---
title: "Detection mode project: Descriptive tables and figures development"
author: "Martina Brayley"
date: last-modified
date-format: long
format: 
  html: 
    embed-resources: true
    toc: true
    toc-location: left
    number-sections: true
    theme: [flatly, style.css]
    page-layout: full
    grid:
      sidebar-width: 300px
      body-width: 2000px
      margin-width: 300px
      gutter-width: 1.5rem
editor: visual
execute:
  warning: false
---

```{r}
#| include: false
#| echo: false

# SET UP ------------------------------------------------------------------------


# execute master file - no need run it as final data now saved
#source("code/master.R")

print(getwd())

library(pacman)
p_load(here)

# set up
setup_path <- here("code", "1.setup_import.r")
source(setup_path)

# import data
# data_path <- here("data", "an_df.rds")
# an_df <- readRDS(data_path)

an_df <- readRDS("Q:/SHARED/USERS/MBrayley/Screening/data/an_df.rds")
```

```{r}
#| include: false
#| echo: false

# DATA PREP ----------------------------------------------------------------------


## Store explanatory variables as an object -------------------------------

# select categorical variables that are labelled

# exploratory vars 
all_expl_variables <- c(
                         # parity related risk factors
                        "d_R1menopause_lab3", 
                        "d_age_menopause_lab", 
                        "d_R1hrtstatus_lab", 
                        "d_parity_lab", 
                        "d_agebirth1_lab",
                        "d_age_menarche_lab", 
                        "d_ocstatus_lab", 
                        # other biological risk factors
                        "d_bbd_lab", 
                        "d_fambrca_lab",
                        # lifestyle risk factors
                        "d_R1alcohol_units_lab", 
                        "d_R1smokingstatus_lab", 
                        "d_bmi_entry_lab", 
                        "d_R1physmet_leis_who_lab")

all_expl_variables_tr <- c(
                            # parity related risk factors
                           "d_R1menopause_lab3", 
                           "d_age_meno_tr_lab", 
                           "d_R1hrt_tr_lab", 
                           "d_parity_lab",
                           "d_age_menarche_lab", 
                           "d_ocstatus_lab",
                            # other biological risk factors
                           "d_bbd_lab", 
                           "d_fambrca_lab",
                           # lifestyle risk factors
                           "d_agebirth1_tr_lab",
                           "d_R1alcohol_units_lab", 
                           "d_R1smokingstatus_lab", 
                           "d_bmi_entry_lab", 
                           "d_R1physmet_leis_who_lab" 
                           )
  

lifestyle <- c("d_R1alcohol_units_lab", 
               "d_R1smokingstatus_lab", 
               "d_bmi_entry_lab", 
               "d_R1physmet_leis_who")
  
reproductive <- c("d_age_menarche_lab", 
                  "d_ocstatus_lab", 
                  "d_R1menopause_lab3", 
                  "d_age_menopause_lab", 
                  "d_R1hrtstatus_lab", 
                  "d_parity_lab", 
                  "d_agebirth1_lab") # do we want to add breastfeeding? If so, it would also have be in the trick model
  
orther_rf <- c("d_bbd", 
               "d_fambrca")

# mammo density excluded from the lists - special variable for subanalyses

# covariates 
all_covariates <- c(# time varying variables
                    "diagage", 
                    "d_R1toBC_y", 
                    "d_R1toMD_y", 
                    "d_MDtoBC_y",
                    # tumour characteristics
                    "d_inv_status", 
                    "d_grade_lab", 
                    "d_morph4", 
                    "d_pos_nodes_lab", 
                    "d_tumour_size", 
                    "d_er_status", 
                    "d_pr_status", 
                    "d_her2_status")  
  
tumour_char <- c("d_inv_status", 
                 "d_grade_lab", 
                 "d_morph4_lab", 
                 "d_pos_nodes_lab", 
                 "d_tumour_size_lab", 
                 "d_er_status_lab", 
                 "d_pr_status_lab", 
                 "d_her2_status_lab")

tumour_char_tr <- c("d_inv_status", 
                    "d_grade_tr", 
                    "d_morph4_tr_lab", 
                    "d_pos_nodes_tr_lab", 
                    "d_tmsize_tr_lab", 
                    "d_er_tr_lab", 
                    "d_pr_tr_lab", 
                    "d_her2_tr_lab")
  
time_vars <- c("diagage", 
               "d_R1toBC_y", 
               "d_R1toMD_y", 
               "d_MDtoBC_y")

# create dataset 
df <- an_df %>% 
  mutate(d_bmi_entry_lab = fct_relevel(d_bmi_entry_lab, "18.5-25", "<18.5", "25-30", "30+")
  )

# set theme for gtsummary tables: 

theme_gtsummary_compact(set_theme = TRUE)
```

# Preliminary descriptive tables

## Tumour characteristics

```{r}
#| echo: false
# tumour characteristics 

# create a dataset with tumour characteristics

tc_df <- an_df %>% 
  select(d_dmode, d_inv_status_lab, diagage, d_grade_lab, d_morph7_lab, d_pos_nodes_lab, d_tumour_size_lab, d_er_status_lab, d_pr_status_lab, d_her2_status_lab) %>% 
  ungroup()

# create tbl summary
tc_df %>% 
  tbl_summary(
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p})"
      ),
    digits = all_continuous() ~ 2,
    label = c(d_inv_status_lab ~ "Invasive status",
              diagage ~ "Age at diagnosis",
              d_grade_lab ~ "Grade",
              d_morph7_lab ~ "Morphology",
              d_pos_nodes_lab ~ "Positive nodes",
              d_tumour_size_lab ~ "Tumour size",
              d_er_status_lab ~ "ER status",
              d_pr_status_lab ~ "PR status",
              d_her2_status_lab ~ "HER2 status")
        
  ) 
```

## Risk factors

```{r}
#| echo: false


rf_df <- df %>% 
  select(d_dmode, d_R1toBC_lab, d_R1toBC_y, d_R1toMD_y, d_MDtoBC_lab, d_MDtoBC_y, 
         d_R1alcohol_status_lab, d_R1alcohol_units_lab, d_R1smokingstatus_lab,
         d_bmi_entry_lab,d_R1physmet_leis_who_lab,
         d_bbd_lab, d_fambrcaN_lab,
         d_age_menarche_lab, d_ocstatus_lab,
         d_R1menopause_lab3, d_age_menopause_lab, d_R1hrtstatus_lab, 
         d_parous_lab, d_parity_lab, d_agebirth1_lab) %>% 
  ungroup()

# create tbl summary
rf_df %>% 
  tbl_summary(
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p})"
      ),
    digits = all_continuous() ~ 2,
    label = c(d_R1toBC_lab ~ "Time from entry to diagnosis (cat)",
              d_R1toBC_y ~ "Time from entry to diagnosis (years)",
              d_R1toMD_y ~ "Time between entry and mammo image (years)",
              d_MDtoBC_lab ~ "Time between mammo and diagnosis (cat)",
              d_MDtoBC_y ~ "Time between mammo and diagnosis (years)",
              d_R1alcohol_status_lab ~ "Alcohol status",
              d_R1alcohol_units_lab ~ "Alcohol units",
              d_R1smokingstatus_lab ~ "Smoking status",
              d_bmi_entry_lab ~ "BMI",
              d_R1physmet_leis_who_lab ~ "Physical Activity (WHO cat)",
              d_bbd_lab ~ "Benign breast disease",
              d_fambrcaN_lab ~ "Family history of BC",
              d_age_menarche_lab ~ "Age at menarche",
              d_ocstatus_lab ~ "Oral contraceptives status",
              d_R1menopause_lab3 ~ "Menopause",
              d_age_menopause_lab ~ "Age at menopause",
              d_R1hrtstatus_lab ~ "HRT status",
              d_parous_lab ~ "Parous",
              d_parity_lab ~ "N of parous pregnancies",
              d_agebirth1_lab ~ "Age at first birth")
        
  ) 
```

## Descriptive tables by mode of detection

## Tumour characteristics

```{r}
#| echo: false

# create a dataset with tumour characteristics

tc_df <- df %>% 
  select(d_dmode, d_inv_status_lab, diagage, d_grade_lab, d_morph7_lab, d_pos_nodes_lab, d_tumour_size_lab, d_er_status_lab, d_pr_status_lab, d_her2_status_lab) %>% 
  ungroup()

# create tbl summary
tc_df %>% 
  tbl_summary(
    by = d_dmode, 
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p})"
      ),
    digits = all_continuous() ~ 2,
    label = c(d_inv_status_lab ~ "Invasive status",
              diagage ~ "Age at diagnosis",
              d_grade_lab ~ "Grade",
              d_morph7_lab ~ "Morphology",
              d_pos_nodes_lab ~ "Positive nodes",
              d_tumour_size_lab ~ "Tumour size",
              d_er_status_lab ~ "ER status",
              d_pr_status_lab ~ "PR status",
              d_her2_status_lab ~ "HER2 status")
        
  ) 



```

## Risk factors

```{r}
#| echo: false


rf_df <- df %>% 
  select(d_dmode, d_R1toBC_lab, d_R1toBC_y, d_R1toMD_y, d_MDtoBC_lab, d_MDtoBC_y, 
         d_R1alcohol_status_lab, d_R1alcohol_units_lab, d_R1smokingstatus_lab,
         d_bmi_entry_lab,d_R1physmet_leis_who_lab,
         d_bbd_lab, d_fambrcaN_lab,
         d_age_menarche_lab, d_ocstatus_lab,
         d_R1menopause_lab3, d_age_menopause_lab, d_R1hrtstatus_lab, 
         d_parous_lab, d_parity_lab, d_agebirth1_lab) %>% 
  ungroup()

# create tbl summary
rf_df %>% 
  tbl_summary(
    by = d_dmode, 
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p})"
      ),
    digits = all_continuous() ~ 2,
    label = c(d_R1toBC_lab ~ "Time from entry to diagnosis (cat)",
              d_R1toBC_y ~ "Time from entry to diagnosis (years)",
              d_R1toMD_y ~ "Time between entry and mammo image (years)",
              d_MDtoBC_lab ~ "Time between mammo and diagnosis (cat)",
              d_MDtoBC_y ~ "Time between mammo and diagnosis (years)",
              d_R1alcohol_status_lab ~ "Alcohol status",
              d_R1alcohol_units_lab ~ "Alcohol units",
              d_R1smokingstatus_lab ~ "Smoking status",
              d_bmi_entry_lab ~ "BMI",
              d_R1physmet_leis_who_lab ~ "Physical Activity (WHO cat)",
              d_bbd_lab ~ "Benign breast disease",
              d_fambrcaN_lab ~ "Family history of BC",
              d_age_menarche_lab ~ "Age at menarche",
              d_ocstatus_lab ~ "Oral contraceptives status",
              d_R1menopause_lab3 ~ "Menopause",
              d_age_menopause_lab ~ "Age at menopause",
              d_R1hrtstatus_lab ~ "HRT status",
              d_parous_lab ~ "Parous",
              d_parity_lab ~ "N of parous pregnancies",
              d_agebirth1_lab ~ "Age at first birth")
        
  ) 
```

## Comparison of trick variables with actual variables

### Tumour characteristics

```{r}
# create a dataset with tumour characteristics

tc_trick_df <- an_df %>% 
  select(d_dmode, d_inv_status_lab, diagage, d_grade_lab, d_grade_tr, d_morph7_lab, d_morph7_tr_lab, d_pos_nodes_lab, d_pos_nodes_tr_lab, d_tumour_size_lab, d_tmsize_tr_lab, d_er_status_lab, d_er_tr_lab, d_pr_status_lab, d_pr_tr_lab, d_her2_status_lab, d_her2_tr_lab) %>% 
  ungroup()

# create tbl summary
tc_trick_df %>% 
  tbl_summary(
    by = d_dmode, 
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p})"
      ),
    digits = all_continuous() ~ 2,
    label = c(d_inv_status_lab ~ "Invasive status",
              diagage ~ "Age at diagnosis",
              d_grade_lab ~ "Grade",
              d_morph7_lab ~ "Morphology",
              d_pos_nodes_lab ~ "Positive nodes",
              d_tumour_size_lab ~ "Tumour size",
              d_er_status_lab ~ "ER status",
              d_pr_status_lab ~ "PR status",
              d_her2_status_lab ~ "HER2 status")
        
  ) 
```

### Parity related

```{r}
#| echo: false


par_df <- df %>% 
  select(d_dmode,
         d_parous_lab, d_parity_lab, d_parity_tr_lab, d_agebirth1_lab, d_agebirth1_tr_lab) %>% 
  ungroup()

# create tbl summary
par_df %>% 
  tbl_summary(
    by = d_dmode, 
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p})"
      ),
    digits = all_continuous() ~ 2,
    label = c(
              d_parous_lab ~ "Parous",
              d_parity_lab ~ "N of parous pregnancies",
              d_agebirth1_lab ~ "Age at first birth")
        
  ) 


```

### Menopause related

```{r}
#| echo: false


meno_df <- an_df %>% 
  select(d_dmode, d_ocstatus_lab, d_oc_tr_lab,
         d_R1menopause_lab3, d_age_menopause_lab, d_R1hrtstatus_lab, 
         ) 

# create tbl summary
meno_df %>% 
  tbl_summary(
    by = d_dmode, 
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p})"
      ),
    digits = all_continuous() ~ 2,
    label = c(
              d_ocstatus_lab ~ "Oral contraceptives status",
              d_R1menopause_lab3 ~ "Menopause",
              d_age_menopause_lab ~ "Age at menopause",
              d_R1hrtstatus_lab ~ "HRT status"
              )
        
  ) 
```

# Tables for publication 

## Table 1 - general study population characteristics by detection mode and overall?

Variables to include:

-   Age at cohort entry

-   age at diagnosis

-   invasive status

-   

# Figures for publication 

## Breast density by risk factors

box plots with jitter

```{r}
# create density subset data 

md_df <- df %>% 
  filter(d_md_avail == 1)
```

### Detection mode

```{r}
# box plots with jitter
md_df %>% 
  ggplot(aes(x = d_dmode, y = d_md, fill = d_dmode)) +
  geom_jitter(position = position_jitter(width = 0.2), 
              alpha = 0.2,
              colour = "grey50") +
  geom_boxplot(alpha = 0.6) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  theme_classic()

```

## Menopausal status

```{r}

# box plots with jitter
md_df %>% 
  ggplot(aes(x = d_R1menopause_lab3, y = d_md, fill = d_R1menopause_lab3)) +
  geom_jitter(position = position_jitter(width = 0.2), 
              alpha = 0.2,
              colour = "grey50") +
  geom_boxplot(alpha = 0.6) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  theme_classic()
```

## Age at menopause

```{r}
# box plots with jitter
md_df %>% 
  ggplot(aes(x = d_age_menopause_lab, y = d_md, fill = d_age_menopause_lab)) +
  geom_jitter(position = position_jitter(width = 0.2), 
              alpha = 0.2,
              colour = "grey50") +
  geom_boxplot(alpha = 0.6) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  theme_classic()
```

## HRT status

```{r}

# box plots with jitter
md_df %>% 
  ggplot(aes(x = d_R1hrtstatus_lab, y = d_md, fill = d_R1hrtstatus_lab)) +
  geom_jitter(position = position_jitter(width = 0.2), 
              alpha = 0.2,
              colour = "grey50") +
  geom_boxplot(alpha = 0.6) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  theme_classic()
```

## Parity

```{r}
# box plots with jitter
md_df %>% 
  ggplot(aes(x = d_parity_lab, y = d_md, fill = d_parity_lab)) +
  geom_jitter(position = position_jitter(width = 0.2), 
              alpha = 0.2,
              colour = "grey50") +
  geom_boxplot(alpha = 0.6) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  theme_classic()
```

## Age at first birth

```{r}

```
