---
title: "Detection mode project: Tables and figures for publication"
author: "Martina Brayley"
date: last-modified
date-format: long
format: 
  html: 
    embed-resources: true
    toc: true
    toc-location: left
    number-sections: true
    theme: [flatly, style.css]
    page-layout: full
    grid:
      sidebar-width: 300px
      body-width: 2000px
      margin-width: 300px
      gutter-width: 1.5rem
editor: visual
execute:
  warning: false
---

```{r}
#| include: false
#| echo: false

# SET UP ------------------------------------------------------------------------


# execute master file - no need run it as final data now saved
#source("code/master.R")

print(getwd())

library(pacman)
p_load(here)

# set up
setup_path <- here("code", "1.setup_import.r")
source(setup_path)

# import data
# data_path <- here("data", "an_df.rds")
# an_df <- readRDS(data_path)

an_df <- readRDS("Q:/SHARED/USERS/MBrayley/Screening/data/an_df.rds")
```

```{r}
#| include: false
#| echo: false

# DATA PREP ----------------------------------------------------------------------


## Store explanatory variables as an object -------------------------------

# select categorical variables that are labelled

# exploratory vars 
all_expl_variables <- c(
                         # parity related risk factors
                        "d_R1menopause_lab3", 
                        "d_age_menopause_lab", 
                        "d_R1hrtstatus_lab", 
                        "d_parity_lab", 
                        "d_agebirth1_lab",
                        "d_age_menarche_lab", 
                        "d_ocstatus_lab", 
                        # other biological risk factors
                        "d_bbd_lab", 
                        "d_fambrca_lab",
                        # lifestyle risk factors
                        "d_R1alcohol_units_lab", 
                        "d_R1smokingstatus_lab", 
                        "d_bmi_entry_lab", 
                        "d_R1physmet_leis_who_lab")

all_expl_variables_tr <- c(
                            # parity related risk factors
                           "d_R1menopause_lab3", 
                           "d_age_meno_tr_lab", 
                           "d_R1hrt_tr_lab", 
                           "d_parity_lab",
                           "d_age_menarche_lab", 
                           "d_ocstatus_lab",
                            # other biological risk factors
                           "d_bbd_lab", 
                           "d_fambrca_lab",
                           # lifestyle risk factors
                           "d_agebirth1_tr_lab",
                           "d_R1alcohol_units_lab", 
                           "d_R1smokingstatus_lab", 
                           "d_bmi_entry_lab", 
                           "d_R1physmet_leis_who_lab" 
                           )
  

lifestyle <- c("d_R1alcohol_units_lab", 
               "d_R1smokingstatus_lab", 
               "d_bmi_entry_lab", 
               "d_R1physmet_leis_who")
  
reproductive <- c("d_age_menarche_lab", 
                  "d_ocstatus_lab", 
                  "d_R1menopause_lab3", 
                  "d_age_menopause_lab", 
                  "d_R1hrtstatus_lab", 
                  "d_parity_lab", 
                  "d_agebirth1_lab") # do we want to add breastfeeding? If so, it would also have be in the trick model
  
orther_rf <- c("d_bbd", 
               "d_fambrca")

# mammo density excluded from the lists - special variable for subanalyses

# covariates 
all_covariates <- c(# time varying variables
                    "diagage", 
                    "d_R1toBC_y", 
                    "d_R1toMD_y", 
                    "d_MDtoBC_y",
                    # tumour characteristics
                    "d_inv_status", 
                    "d_grade_lab", 
                    "d_morph4", 
                    "d_pos_nodes_lab", 
                    "d_tumour_size", 
                    "d_er_status", 
                    "d_pr_status", 
                    "d_her2_status")  
  
tumour_char <- c("d_inv_status", 
                 "d_grade_lab", 
                 "d_morph4_lab", 
                 "d_pos_nodes_lab", 
                 "d_tumour_size_lab", 
                 "d_er_status_lab", 
                 "d_pr_status_lab", 
                 "d_her2_status_lab")

tumour_char_tr <- c("d_inv_status", 
                    "d_grade_tr", 
                    "d_morph4_tr_lab", 
                    "d_pos_nodes_tr_lab", 
                    "d_tmsize_tr_lab", 
                    "d_er_tr_lab", 
                    "d_pr_tr_lab", 
                    "d_her2_tr_lab")
  
time_vars <- c("diagage", 
               "d_R1toBC_y", 
               "d_R1toMD_y", 
               "d_MDtoBC_y")

# create dataset 
df <- an_df %>% 
  mutate(d_bmi_entry_lab = fct_relevel(d_bmi_entry_lab, "18.5-25", "<18.5", "25-30", "30+")
  )

# set theme for gtsummary tables: 

theme_gtsummary_compact(set_theme = TRUE)
```

# Table and figure structure:

**Main tables**

Table 1 - general descriptives (population characteristics)

Table 2 - Risk factor descriptives by detection mode with logistic regression results (unadjusted for TC) - all cases

**Main figures**

Figure 1 - comprised figure of plots describing mamographic density in relation to risk factors (detail in ST 2)

Figure 2 - Forest plot of analysis adjusted vs unadjusted for mammographic density (full table in supplementary material - ST 2)

**Supplementary material**

(ST = Supplementary Table)

ST 1 - Descriptives for MD cases (same as table 1 but subset)

ST 2 - risk factor descriptives by density (complementing figure 1)

ST 3 - subset on cases with MD. Same model as table 3 + adjusted for MD. Include mediation measures (% change)

ST 4 - TC descriptives by detection mode with logistic regression results

ST 5 - Risk factor analysis adjusted for tumour characteristics (sensitivity analysis) - on subset of cases with TC

# Main tables 

## Table 1 - general participant descriptives

## Table 2 - Main analysis

The main analysis is a logistic regression with multiple exposures (risk factors).

This table shows risk factor level descriptives and their associations with the mode of detection (logistic regression) adjusted for all other risk factors, age at diagnosis and time from entry to BC diagnosis.

Model 2 is the final selected model (Default model) to present in the analysis. Model 2 is adjusted for all co-exposures + age at diagnosis and time from entry to BC diagnosis. Model 1 was only adjusted for co-exposures, e.i. all exposures in the model.

Because model 2 was selected as the default model, it is also used on the subanalyses with TC and MD as the comparison against adjusted models for TC and MD.

```{r}
#| echo: false

# summary Logistic Regression of Model 2  -----------------------------------------

# view existing output
#model2

# remove columns with total and interval N - not simple with gtsummary, trying from scratch without adding the columns


# Create clean version of logistic output, without Ns
model <- df %>%
   drop_na(all_of(all_expl_variables_tr), diagage, d_R1toBC_y
           ) %>%
  glm(formula = d_dmode_n ~ d_R1menopause_lab3 + d_age_meno_tr_lab 
      + d_R1hrt_tr_lab
      + d_parity_lab + d_agebirth1_tr_lab + d_age_menarche_lab + d_ocstatus_lab 
      + d_bbd_lab + d_fambrca_lab
      + d_R1alcohol_units_lab + d_R1smokingstatus_lab 
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab
      + diagage + d_R1toBC_y,
            family = binomial(link = "logit"))

# tidy up output
model2 <- tbl_regression(model, 
                         exponentiate = TRUE,
                         label = list(
                          d_R1alcohol_units_lab ~ "Alcohol units",
                          d_R1smokingstatus_lab ~ "Smoking status",
                          d_bmi_entry_lab ~ "BMI",
                          d_R1physmet_leis_who_lab ~ "Physical Activity (WHO cat)",
                          d_bbd_lab ~ "Benign breast disease",
                          d_fambrca_lab ~ "Family history of BC",
                          d_age_menarche_lab ~ "Age at menarche",
                          d_ocstatus_lab ~ "Oral contraceptives status",
                          d_R1menopause_lab3 ~ "Menopause",
                          d_age_meno_tr_lab ~ "Age at menopause",
                          d_R1hrt_tr_lab ~ "HRT status",
                          d_parity_lab ~ "N of parous pregnancies",
                          d_agebirth1_tr_lab ~ "Age at first birth"
                         )
                         ) %>% 
  add_global_p() %>% 
  remove_row_type(variables = c(diagage, d_R1toBC_y), # remove covariates
                  type = "all")
  

#model2




# Create descriptives by mode of detection ------------------------------------
# select(d_dmode_n, d_R1alcohol_units_lab, d_R1smokingstatus_lab, d_bmi_entry_lab,  d_R1physmet_leis_who_lab, d_bbd_lab, d_fambrca_lab, d_age_menarche_lab, d_ocstatus_lab, d_R1menopause_lab3, d_age_meno_tr_lab, d_R1hrt_tr_lab, d_parity_tr_lab, d_agebirth1_tr_lab)

# note: need to rename the non-trick variables according to their trick name to align with the model, which can then allow table merge. 

# include missing 
# create df
  
descr_model2_df <- df %>% 
  select(d_dmode, all_of(all_expl_variables), diagage, d_R1toBC_y 
         ) %>% 
  #tidyr::drop_na(all_of(all_expl_variables), diagage, d_R1toBC_y # filter out all missing to have the same n as in the model
  mutate(d_bmi_entry_lab = fct_relevel(d_bmi_entry_lab, "<18.5", "18.5-25", "25-30", "30+")
              ) %>% 
  mutate(d_age_meno_tr_lab = d_age_menopause_lab,
         d_R1hrt_tr_lab = d_R1hrtstatus_lab,
         d_agebirth1_tr_lab = d_agebirth1_lab
         ) %>% 
  select(-d_age_menopause_lab, -d_R1hrtstatus_lab, -d_agebirth1_lab, -diagage, -d_R1toBC_y)  
  
  

# glimpse(descr_model2)
# str(test)

# create gs summary table
descr_model2 <- descr_model2_df %>% 
  tbl_summary(by = d_dmode,
              statistic = list(
                all_continuous() ~ "{mean} ({sd})",
                all_categorical() ~ "{n} ({p}) "
                ),
              digits = all_continuous() ~ 2,
              label = list(
                          d_R1alcohol_units_lab ~ "Alcohol units",
                          d_R1smokingstatus_lab ~ "Smoking status",
                          d_bmi_entry_lab ~ "BMI",
                          d_R1physmet_leis_who_lab ~ "Physical Activity (WHO cat)",
                          d_bbd_lab ~ "Benign breast disease",
                          d_fambrca_lab ~ "Family history of BC",
                          d_age_menarche_lab ~ "Age at menarche",
                          d_ocstatus_lab ~ "Oral contraceptives status",
                          d_R1menopause_lab3 ~ "Menopause",
                          d_age_meno_tr_lab ~ "Age at menopause",
                          d_R1hrt_tr_lab ~ "HRT status",
                          d_parity_lab ~ "N of parous pregnancies",
                          d_agebirth1_tr_lab ~ "Age at first birth"
                ),
              type = list(d_bbd_lab ~ "categorical",
                          d_fambrca_lab ~ "categorical")
              ) %>% 
  add_overall()

#descr_model2
  
# merge with analysis output -------------------------------------------------


table3 <- tbl_merge(
  tbls = list(descr_model2, model2),
  tab_spanner = c("", "")
   ) 
   
  
   

table3


# this ID is extra in the LR output from model 2 - T276E210
# this is because HRT status was missing as under 50 but it is present in trick hrt as recoded to never. - check what to do with this 
```

# Main figures

## Figure 1 - descriptives of risk factors with MD

## Figure 2 - adjusting for MD

::: callout-important
## Decision needed

Need to decide whether to use quartiles or continuous MD
:::

### Categorical MD (quartiles)

```{r}
#| echo: false

# 1. RUN MODELS AND CREATE REGRESSION DATASETS FOR EACH MODEL 

# Model 6 ----------------------------------------------------------------
 # include parity and meno trick variable instead of original variables

# only included alcohol status - check if we need to do the trick if including both
# including "diagage", "d_R1toBC_y" - MD variables will be included in sub analysis

model <- df %>%
   filter(d_md_avail == 1 # include only those with md
         ) %>%
   drop_na(all_of(all_expl_variables_tr), diagage, d_R1toBC_y
           ) %>% 
  glm(formula = d_dmode_n ~ d_R1menopause_lab3 + d_age_meno_tr_lab 
      + d_R1hrt_tr_lab
      + d_parity_lab + d_agebirth1_tr_lab + d_age_menarche_lab + d_ocstatus_lab 
      + d_bbd_lab + d_fambrca_lab
      + d_R1alcohol_units_lab + d_R1smokingstatus_lab 
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab
      + diagage + d_R1toBC_y,
            family = binomial(link = "logit"))

# create a dataframe with broom::tidy() for forestplots 

tidy_model6 <- model %>%
   tidy(exponentiate = TRUE, conf.int = TRUE) %>%
   mutate(OR = estimate) %>%
   select(-estimate) %>%
   select(term, OR, everything())
#tidy_model6


# Model 8 ----------------------------------------------------------------
 # include parity and meno trick variable instead of original variables

# only included alcohol status - check if we need to do the trick if including both
# including "diagage", "d_R1toBC_y" - MD variables will be included in sub analysis

model <- df %>%
   filter(d_md_avail == 1 # include only those with md
         ) %>%
  drop_na(all_of(all_expl_variables_tr), diagage, d_R1toBC_y
           ) %>% 
  glm(formula = d_dmode_n ~ d_R1menopause_lab3 + d_age_meno_tr_lab 
      + d_R1hrt_tr_lab
      + d_parity_lab + d_agebirth1_tr_lab + d_age_menarche_lab + d_ocstatus_lab 
      + d_bbd_lab + d_fambrca_lab
      + d_R1alcohol_units_lab + d_R1smokingstatus_lab 
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab
      + diagage + d_R1toBC_y + d_MDtoBC_y
      + d_md_qrt,
            family = binomial(link = "logit"))


# create a dataframe with broom::tidy() for forestplots 

tidy_model8 <- model %>%
   tidy(exponentiate = TRUE, conf.int = TRUE) %>%
   mutate(OR = estimate) %>%
   select(-estimate) %>%
   select(term, OR, everything())
 
  
```

```{r}
#| echo: false

# 2. PROCESS DATA FOR FORESTPLOT 

# full join data from both models 
md_forest_df <- tidy_model6 %>% 
  full_join(tidy_model8, by = "term", suffix = c("_mdl6", "_mdl8"))

# RESUME HERE
# create labels and variable names structures 
md_forest_df <- md_forest_df %>% 
  mutate(variable = case_when(
    grepl("menopause", term) ~ "Menopausal status",
    grepl("age_meno", term) ~ "Age at menopause",
    grepl("hrt", term) ~ "HRT status",
    grepl("parity", term) ~ "N of parous pregnancies",
    grepl("agebirth1", term) ~ "Age at first birth",
    grepl("age_menarche", term) ~ "Age at menarche", 
    grepl("ocstatus", term) ~ "Oral contraceptives status",
    grepl("bbd", term) ~ "Benign breast disease",
    grepl("fambrca", term) ~ "Family history of BC",
    grepl("R1alcohol", term) ~ "Alcohol units",
    grepl("R1smokingstatus", term) ~ "Smoking status",
    grepl("bmi_entry", term) ~ "BMI",
    grepl("R1physmet_leis_who", term) ~ "Physical Activity (WHO cat)",
    grepl("md_qrt", term) ~ "Mammographic breast density quartile"
                              ),
    level = case_when(
      grepl("d_R1menopause_lab3", term) ~ sub("d_R1menopause_lab3", "", term),
      grepl("age_meno", term) ~ sub("d_age_meno_tr_lab", "", term),
      grepl("hrt", term) ~ sub("d_R1hrt_tr_lab", "", term),
      grepl("parity", term) ~ sub("d_parity_lab", "", term),
      grepl("agebirth1", term) ~ sub("d_agebirth1_tr_lab", "", term),
      grepl("age_menarche", term) ~ sub("d_age_menarche_lab", "", term), 
      grepl("ocstatus", term) ~ sub("d_ocstatus_lab", "", term),
      grepl("bbd", term) ~ sub("d_bbd_lab", "", term),
      grepl("fambrca", term) ~ sub("d_fambrca_lab", "", term),
      grepl("R1alcohol", term) ~ sub("d_R1alcohol_units_lab", "", term),
      grepl("R1smokingstatus", term) ~ sub("d_R1smokingstatus_lab", "", term),
      grepl("bmi_entry", term) ~ sub("d_bmi_entry_lab", "", term),
      grepl("R1physmet_leis_who", term) ~ sub("d_R1physmet_leis_who_lab", "", term),
      grepl("md_qrt", term) ~ sub("d_md_qrt", "", term)
    )
    )

# add subheading rows 

subheading_data <- md_forest_df %>%
  distinct(variable) %>%
  mutate(level = variable)

md_forest_df <- bind_rows(subheading_data, md_forest_df) %>%
  arrange(variable, factor(level, levels = unique(level)))



# create long format 
md_forest_long <- md_forest_df %>% 
  filter(!term %in% c("(Intercept)", "diagage", "d_R1toBC_y", "d_MDtoBC_y", "d_md_qrt2", "d_md_qrt3", "d_md_qrt4"),
         !is.na(level),
         level != "Mammographic breast density quartile"
         ) %>% 
  select(level, OR_mdl6, conf.low_mdl6, conf.high_mdl6, OR_mdl8, conf.low_mdl8, conf.high_mdl8
         ) %>% 
  mutate(pct_diff = if_else(is.na(OR_mdl6), "-", as.character(round((OR_mdl8 - OR_mdl6) / OR_mdl8 * 100, 1))) 
         ) %>% 
  pivot_longer(
    cols = c(OR_mdl6, conf.low_mdl6, conf.high_mdl6, 
             OR_mdl8, conf.low_mdl8, conf.high_mdl8
             ),
    names_to = c(".value", "mdl"),
    names_sep = "_"
  ) 

```

```{r, fig.width=8, fig.height=10}
#| echo: false

# 3. GENERATE FORESTPLOT

is_summary <- is.na(md_forest_long$OR)


# develop forest

md_forest_long %>% 
  group_by(mdl
           ) %>% 
  mutate(mean = OR,
         lower = conf.low,
         upper = conf.high
         ) %>% 
  forestplot(labeltext = c(level, pct_diff),
             xlog = FALSE,
             xlab = "OR",
             boxsize = 0.3,
             zero = 1,
             line.margin = 0.5,
             clip = c(0, 3),
             ticks = c(0, 0.5, 1, 1.5, 2, 2.5, 3),
             ci.vertices = T,
             ci.vertices.height = 0.2,
             legend = c("Default model", "Adjusted for MD"),
             title = paste("Final model adjusted vs unadjusted for mamographic breast density quartiles"),
             is.summary = is_summary
             ) %>% 
  fp_set_style(box = c("grey50","deeppink2") %>%  lapply(function(x) gpar(fill = x, col = "#666666")),
               default = gpar(vertices = T),
               line = "grey40",
               summary = gpar(col = NA), # Set summary color to NA
               align = "lrrr",
               hrz_lines = gpar(col = NA), # Set line color to NA
               txt_gp = fpTxtGp(label = gpar(cex = 0.8),
                                    ticks = gpar(cex = 0.8),
                                    xlab = gpar(cex = 0.8),
                                )
               ) %>% 
  fp_decorate_graph(graph.pos = 2
                    ) %>% 
  fp_add_header(level = c("", "Risk factor"),
                pct_diff = c("", "% difference")
                ) %>% 
  fp_insert_row(level = "",
                position = 3,
                mean = NULL,
                lower = NULL,
                upper = NULL,
                is.summary = TRUE,
                boxsize = NA)
```

### Continuous MD

```{r}
#| echo: false

# 1. RUN MODEL 8b only AND CREATE REGRESSION DATASET 

# Model 8b - with continuous MD ----------------------------------------------------------------
 # include parity and meno trick variable instead of original variables

# only included alcohol status - check if we need to do the trick if including both
# including "diagage", "d_R1toBC_y" - MD variables will be included in sub analysis

model <- df %>%
   filter(d_md_avail == 1 # include only those with md
         ) %>%
  drop_na(all_of(all_expl_variables_tr), diagage, d_R1toBC_y
           ) %>% 
  glm(formula = d_dmode_n ~ d_R1menopause_lab3 + d_age_meno_tr_lab 
      + d_R1hrt_tr_lab
      + d_parity_lab + d_agebirth1_tr_lab + d_age_menarche_lab + d_ocstatus_lab 
      + d_bbd_lab + d_fambrca_lab
      + d_R1alcohol_units_lab + d_R1smokingstatus_lab 
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab
      + diagage + d_R1toBC_y + d_MDtoBC_y
      + d_md,
            family = binomial(link = "logit"))

# create a dataframe with broom::tidy() for forestplots 

tidy_model8b <- model %>%
   tidy(exponentiate = TRUE, conf.int = TRUE) %>%
   mutate(OR = estimate) %>%
   select(-estimate) %>%
   select(term, OR, everything())
```

```{r}
#| echo: false

# 2. PROCESS DATA FOR FORESTPLOT 

# full join data from both models 
md_forest_df <- tidy_model6 %>% 
  full_join(tidy_model8b, by = "term", suffix = c("_mdl6", "_mdl8"))

# RESUME HERE
# create labels and variable names structures 
md_forest_df <- md_forest_df %>% 
  mutate(variable = case_when(
    grepl("menopause", term) ~ "Menopausal status",
    grepl("age_meno", term) ~ "Age at menopause",
    grepl("hrt", term) ~ "HRT status",
    grepl("parity", term) ~ "N of parous pregnancies",
    grepl("agebirth1", term) ~ "Age at first birth",
    grepl("age_menarche", term) ~ "Age at menarche", 
    grepl("ocstatus", term) ~ "Oral contraceptives status",
    grepl("bbd", term) ~ "Benign breast disease",
    grepl("fambrca", term) ~ "Family history of BC",
    grepl("R1alcohol", term) ~ "Alcohol units",
    grepl("R1smokingstatus", term) ~ "Smoking status",
    grepl("bmi_entry", term) ~ "BMI",
    grepl("R1physmet_leis_who", term) ~ "Physical Activity (WHO cat)",
    grepl("md", term) ~ "Mammographic breast density"
                              ),
    level = case_when(
      grepl("d_R1menopause_lab3", term) ~ sub("d_R1menopause_lab3", "", term),
      grepl("age_meno", term) ~ sub("d_age_meno_tr_lab", "", term),
      grepl("hrt", term) ~ sub("d_R1hrt_tr_lab", "", term),
      grepl("parity", term) ~ sub("d_parity_lab", "", term),
      grepl("agebirth1", term) ~ sub("d_agebirth1_tr_lab", "", term),
      grepl("age_menarche", term) ~ sub("d_age_menarche_lab", "", term), 
      grepl("ocstatus", term) ~ sub("d_ocstatus_lab", "", term),
      grepl("bbd", term) ~ sub("d_bbd_lab", "", term),
      grepl("fambrca", term) ~ sub("d_fambrca_lab", "", term),
      grepl("R1alcohol", term) ~ sub("d_R1alcohol_units_lab", "", term),
      grepl("R1smokingstatus", term) ~ sub("d_R1smokingstatus_lab", "", term),
      grepl("bmi_entry", term) ~ sub("d_bmi_entry_lab", "", term),
      grepl("R1physmet_leis_who", term) ~ sub("d_R1physmet_leis_who_lab", "", term),
      grepl("md", term) ~ sub("d_md", "", term)
    )
    )

# add subheading rows 

subheading_data <- md_forest_df %>%
  distinct(variable) %>%
  mutate(level = variable)

md_forest_df <- bind_rows(subheading_data, md_forest_df) %>%
  arrange(variable, factor(level, levels = unique(level)))



# create long format 
md_forest_long <- md_forest_df %>% 
  filter(!term %in% c("(Intercept)", "diagage", "d_R1toBC_y", "d_MDtoBC_y", "d_md"),
         !is.na(level),
         ) %>% 
  select(level, OR_mdl6, conf.low_mdl6, conf.high_mdl6, OR_mdl8, conf.low_mdl8, conf.high_mdl8
         ) %>% 
  mutate(pct_diff = if_else(is.na(OR_mdl6), "-", as.character(round((OR_mdl8 - OR_mdl6) / OR_mdl8 * 100, 1))) 
         ) %>% 
  pivot_longer(
    cols = c(OR_mdl6, conf.low_mdl6, conf.high_mdl6, 
             OR_mdl8, conf.low_mdl8, conf.high_mdl8
             ),
    names_to = c(".value", "mdl"),
    names_sep = "_"
  ) 
```

```{r, fig.width=8, fig.height=10}
#| echo: false

# 3. GENERATE FORESTPLOT

is_summary <- is.na(md_forest_long$OR)


# develop forest

md_forest_long %>% 
  group_by(mdl
           ) %>% 
  mutate(mean = OR,
         lower = conf.low,
         upper = conf.high
         ) %>% 
  forestplot(labeltext = c(level, pct_diff),
             xlog = FALSE,
             xlab = "OR",
             boxsize = 0.3,
             zero = 1,
             line.margin = 0.5,
             clip = c(0, 3),
             ticks = c(0, 0.5, 1, 1.5, 2, 2.5, 3),
             ci.vertices = T,
             ci.vertices.height = 0.2,
             legend = c("Default model", "Adjusted for MD"),
             title = paste("Final model adjusted vs unadjusted for continuous mamographic breast density"),
             is.summary = is_summary
             ) %>% 
  fp_set_style(box = c("grey50","orangered2") %>%  lapply(function(x) gpar(fill = x, col = "#666666")),
               default = gpar(vertices = T),
               line = "grey40",
               summary = gpar(col = NA), # Set summary color to NA
               align = "lrrr",
               hrz_lines = gpar(col = NA), # Set line color to NA
               txt_gp = fpTxtGp(label = gpar(cex = 0.8),
                                    ticks = gpar(cex = 0.8),
                                    xlab = gpar(cex = 0.8)
                                )
               ) %>% 
  fp_decorate_graph(graph.pos = 2
                    ) %>% 
  fp_add_header(level = c("", "Risk factor"),
                pct_diff = c("", "% difference")
                ) %>% 
  fp_insert_row(level = "",
                position = 3,
                mean = NULL,
                lower = NULL,
                upper = NULL,
                is.summary = TRUE,
                boxsize = NA)
```

# Supplementary material

## ST1 - MD subset general descriptives

## ST 2 - Risk factor descriptives by density 

complementing figure 1

## ST 3 - subset on cases with MD 

Subanalyses comparing default model with model adjusted for MD

### Categorical MD (quartiles)

! need to move MD last

```{r}
#| echo: false

# Model 6 ----------------------------------------------------------------
 # include parity and meno trick variable instead of original variables

# only included alcohol status - check if we need to do the trick if including both
# including "diagage", "d_R1toBC_y" - MD variables will be included in sub analysis

model <- df %>%
   filter(d_md_avail == 1 # include only those with md
         ) %>%
   drop_na(all_of(all_expl_variables_tr), diagage, d_R1toBC_y
           ) %>% 
  glm(formula = d_dmode_n ~ d_R1menopause_lab3 + d_age_meno_tr_lab 
      + d_R1hrt_tr_lab
      + d_parity_lab + d_agebirth1_tr_lab + d_age_menarche_lab + d_ocstatus_lab 
      + d_bbd_lab + d_fambrca_lab
      + d_R1alcohol_units_lab + d_R1smokingstatus_lab 
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab
      + diagage + d_R1toBC_y,
            family = binomial(link = "logit"))


# tidy up output
model6 <- tbl_regression(model, 
                         exponentiate = TRUE,
                         label = list(
                          d_R1alcohol_units_lab ~ "Alcohol units",
                          d_R1smokingstatus_lab ~ "Smoking status",
                          d_bmi_entry_lab ~ "BMI",
                          d_R1physmet_leis_who_lab ~ "Physical Activity (WHO cat)",
                          d_bbd_lab ~ "Benign breast disease",
                          d_fambrca_lab ~ "Family history of BC",
                          d_age_menarche_lab ~ "Age at menarche",
                          d_ocstatus_lab ~ "Oral contraceptives status",
                          d_R1menopause_lab3 ~ "Menopause",
                          d_age_meno_tr_lab ~ "Age at menopause",
                          d_R1hrt_tr_lab ~ "HRT status",
                          d_parity_lab ~ "N of parous pregnancies",
                          d_agebirth1_tr_lab ~ "Age at first birth"
                         )
                         ) %>% 
  add_global_p() %>% 
  remove_row_type(variables = c(diagage, d_R1toBC_y), # remove covariates
                  type = "all")

#model6


# Model 8 ----------------------------------------------------------------
 # include parity and meno trick variable instead of original variables

# only included alcohol status - check if we need to do the trick if including both
# including "diagage", "d_R1toBC_y" - MD variables will be included in sub analysis

model <- df %>%
   filter(d_md_avail == 1 # include only those with md
         ) %>%
  drop_na(all_of(all_expl_variables_tr), diagage, d_R1toBC_y
           ) %>% 
  glm(formula = d_dmode_n ~ d_R1menopause_lab3 + d_age_meno_tr_lab 
      + d_R1hrt_tr_lab
      + d_parity_lab + d_agebirth1_tr_lab + d_age_menarche_lab + d_ocstatus_lab 
      + d_bbd_lab + d_fambrca_lab
      + d_R1alcohol_units_lab + d_R1smokingstatus_lab 
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab
      + diagage + d_R1toBC_y + d_MDtoBC_y
      + d_md_qrt,
            family = binomial(link = "logit"))

model8 <- tbl_regression(model, 
                         exponentiate = TRUE,
                         label = list(
                          d_R1alcohol_units_lab ~ "Alcohol units",
                          d_R1smokingstatus_lab ~ "Smoking status",
                          d_bmi_entry_lab ~ "BMI",
                          d_R1physmet_leis_who_lab ~ "Physical Activity (WHO cat)",
                          d_bbd_lab ~ "Benign breast disease",
                          d_fambrca_lab ~ "Family history of BC",
                          d_age_menarche_lab ~ "Age at menarche",
                          d_ocstatus_lab ~ "Oral contraceptives status",
                          d_R1menopause_lab3 ~ "Menopause",
                          d_age_meno_tr_lab ~ "Age at menopause",
                          d_R1hrt_tr_lab ~ "HRT status",
                          d_parity_lab ~ "N of parous pregnancies",
                          d_agebirth1_tr_lab ~ "Age at first birth",
                          d_md_qrt ~ "Mammographic breast density quartile"
                         )
                         ) %>% 
  add_global_p() %>% 
  remove_row_type(variables = c(diagage, d_R1toBC_y, d_MDtoBC_y), # remove covariates
                  type = "all")

#model8

 
# Descriptives for table 4 -------------------------------------------------
# subset on those with mammo density 


descr_MDsubset_df <- df %>% 
  select(d_dmode, all_of(all_expl_variables), diagage, d_R1toBC_y, d_md_qrt 
         ) %>% 
  tidyr::drop_na(d_md_qrt # filter out all missing to have the same n as in the model
  ) %>% 
  mutate(d_bmi_entry_lab = fct_relevel(d_bmi_entry_lab, "<18.5", "18.5-25", "25-30", "30+")
              ) %>% 
  mutate(d_age_meno_tr_lab = d_age_menopause_lab,
         d_R1hrt_tr_lab = d_R1hrtstatus_lab,
         d_agebirth1_tr_lab = d_agebirth1_lab
         ) %>% 
  select(-d_age_menopause_lab, -d_R1hrtstatus_lab, -d_agebirth1_lab, -diagage, -d_R1toBC_y)  
  
  

# glimpse(descr_MDsubset_df)
# str(descr_MDsubset_df)

# create gs summary table
descr_MDsubset <- descr_MDsubset_df %>% 
  select(d_dmode,
    # parity related risk factors
                        d_R1menopause_lab3, 
                        d_age_meno_tr_lab, 
                        d_R1hrt_tr_lab, 
                        d_parity_lab, 
                        d_agebirth1_tr_lab,
                        d_age_menarche_lab,
                        d_ocstatus_lab, 
                        # other biological risk factors
                        d_bbd_lab, 
                        d_fambrca_lab,
                        # lifestyle risk factors
                        d_R1alcohol_units_lab, 
                        d_R1smokingstatus_lab, 
                        d_bmi_entry_lab, 
                        d_R1physmet_leis_who_lab,
    d_md_qrt) %>% 
  tbl_summary(by = d_dmode,
              statistic = list(
                all_continuous() ~ "{mean} ({sd})",
                all_categorical() ~ "{n} ({p})" 
                ),
              digits = all_continuous() ~ 2,
              label = list(
                          d_R1alcohol_units_lab ~ "Alcohol units",
                          d_R1smokingstatus_lab ~ "Smoking status",
                          d_bmi_entry_lab ~ "BMI",
                          d_R1physmet_leis_who_lab ~ "Physical Activity (WHO cat)",
                          d_bbd_lab ~ "Benign breast disease",
                          d_fambrca_lab ~ "Family history of BC",
                          d_age_menarche_lab ~ "Age at menarche",
                          d_ocstatus_lab ~ "Oral contraceptives status",
                          d_R1menopause_lab3 ~ "Menopause",
                          d_age_meno_tr_lab ~ "Age at menopause",
                          d_R1hrt_tr_lab ~ "HRT status",
                          d_parity_lab ~ "N of parous pregnancies",
                          d_agebirth1_tr_lab ~ "Age at first birth",
                          d_md_qrt ~ "Mammographic breast density quartile"
                ),
              type = list(d_bbd_lab ~ "categorical",
                          d_fambrca_lab ~ "categorical")
              ) #%>% 
 # add_overall()


#descr_MDsubset



# merge tables -------------------------------------------------


table4 <- tbl_merge(
  tbls = list(descr_MDsubset, model6, model8),
  tab_spanner = c("Subset of cases with MD", "Unadjusted for MD", "Adjusted for MD")
   ) 
   
  
#  add difference using gtsummary 
table4_diff <- table4 %>% 
  modify_table_body(
    ~ .x %>% 
      mutate(pct_diff = if_else(
          is.na(estimate_2) | is.na(estimate_3),
          "",
          sprintf("%.1f", (estimate_3 - estimate_2) / estimate_3 * 100))
    )
    ) %>% 
  modify_header(
    pct_diff = "**Difference (%)**"
  ) %>% 
  modify_table_body(
    ~ .x %>% 
      arrange(desc(variable = "d_md_qrt"))
  )

table4_diff
```

### Continuous MD

```{r}
#| echo: false

# Model 8b - with continuous MD ----------------------------------------------------------------
 # include parity and meno trick variable instead of original variables

# only included alcohol status - check if we need to do the trick if including both
# including "diagage", "d_R1toBC_y" - MD variables will be included in sub analysis

model <- df %>%
   filter(d_md_avail == 1 # include only those with md
         ) %>%
  drop_na(all_of(all_expl_variables_tr), diagage, d_R1toBC_y
           ) %>% 
  glm(formula = d_dmode_n ~ d_R1menopause_lab3 + d_age_meno_tr_lab 
      + d_R1hrt_tr_lab
      + d_parity_lab + d_agebirth1_tr_lab + d_age_menarche_lab + d_ocstatus_lab 
      + d_bbd_lab + d_fambrca_lab
      + d_R1alcohol_units_lab + d_R1smokingstatus_lab 
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab
      + diagage + d_R1toBC_y + d_MDtoBC_y
      + d_md,
            family = binomial(link = "logit"))

model8b <- tbl_regression(model, 
                         exponentiate = TRUE,
                         label = list(
                          d_R1alcohol_units_lab ~ "Alcohol units",
                          d_R1smokingstatus_lab ~ "Smoking status",
                          d_bmi_entry_lab ~ "BMI",
                          d_R1physmet_leis_who_lab ~ "Physical Activity (WHO cat)",
                          d_bbd_lab ~ "Benign breast disease",
                          d_fambrca_lab ~ "Family history of BC",
                          d_age_menarche_lab ~ "Age at menarche",
                          d_ocstatus_lab ~ "Oral contraceptives status",
                          d_R1menopause_lab3 ~ "Menopause",
                          d_age_meno_tr_lab ~ "Age at menopause",
                          d_R1hrt_tr_lab ~ "HRT status",
                          d_parity_lab ~ "N of parous pregnancies",
                          d_agebirth1_tr_lab ~ "Age at first birth",
                          d_md ~ "Mammographic breast density"
                         )
                         ) %>% 
  add_global_p() %>% 
  remove_row_type(variables = c(diagage, d_R1toBC_y, d_MDtoBC_y), # remove covariates
                  type = "all")


# Descriptives for table 4 -------------------------------------------------
# subset on those with mammo density 


descr_MDsubset_df <- df %>% 
  select(d_dmode, all_of(all_expl_variables), diagage, d_R1toBC_y, d_md 
         ) %>% 
  tidyr::drop_na(d_md # filter out all missing to have the same n as in the model
  ) %>% 
  mutate(d_bmi_entry_lab = fct_relevel(d_bmi_entry_lab, "<18.5", "18.5-25", "25-30", "30+")
              ) %>% 
  mutate(d_age_meno_tr_lab = d_age_menopause_lab,
         d_R1hrt_tr_lab = d_R1hrtstatus_lab,
         d_agebirth1_tr_lab = d_agebirth1_lab
         ) %>% 
  select(-d_age_menopause_lab, -d_R1hrtstatus_lab, -d_agebirth1_lab, -diagage, -d_R1toBC_y)  
  
  

# glimpse(descr_MDsubset_df)
# str(descr_MDsubset_df)

# create gs summary table
descr_MDsubset <- descr_MDsubset_df %>% 
  select(d_dmode,
    # parity related risk factors
                        d_R1menopause_lab3, 
                        d_age_meno_tr_lab, 
                        d_R1hrt_tr_lab, 
                        d_parity_lab, 
                        d_agebirth1_tr_lab,
                        d_age_menarche_lab,
                        d_ocstatus_lab, 
                        # other biological risk factors
                        d_bbd_lab, 
                        d_fambrca_lab,
                        # lifestyle risk factors
                        d_R1alcohol_units_lab, 
                        d_R1smokingstatus_lab, 
                        d_bmi_entry_lab, 
                        d_R1physmet_leis_who_lab,
    d_md) %>% 
  tbl_summary(by = d_dmode,
              statistic = list(
                all_continuous() ~ "{mean} ({sd})",
                all_categorical() ~ "{n} ({p})" 
                ),
              digits = all_continuous() ~ 2,
              label = list(
                          d_R1alcohol_units_lab ~ "Alcohol units",
                          d_R1smokingstatus_lab ~ "Smoking status",
                          d_bmi_entry_lab ~ "BMI",
                          d_R1physmet_leis_who_lab ~ "Physical Activity (WHO cat)",
                          d_bbd_lab ~ "Benign breast disease",
                          d_fambrca_lab ~ "Family history of BC",
                          d_age_menarche_lab ~ "Age at menarche",
                          d_ocstatus_lab ~ "Oral contraceptives status",
                          d_R1menopause_lab3 ~ "Menopause",
                          d_age_meno_tr_lab ~ "Age at menopause",
                          d_R1hrt_tr_lab ~ "HRT status",
                          d_parity_lab ~ "N of parous pregnancies",
                          d_agebirth1_tr_lab ~ "Age at first birth",
                          d_md ~ "Mammographic breast density"
                ),
              type = list(d_bbd_lab ~ "categorical",
                          d_fambrca_lab ~ "categorical")
              ) #%>% 
 # add_overall()


# descr_MDsubset



# merge tables

table4b <- tbl_merge(
  tbls = list(descr_MDsubset, model6, model8b),
  tab_spanner = c("Subset of cases with MD", "Unadjusted for MD", "Adjusted for cont MD")
   ) 

#table4b   

#  using gtsummary 
table4b_diff <- table4b %>% 
  modify_table_body(
    ~ .x %>% 
      mutate(pct_diff = if_else(
          is.na(estimate_2) | is.na(estimate_3),
          "",
          sprintf("%.1f", (estimate_3 - estimate_2) / estimate_3 * 100))
    )
    ) %>% 
  modify_header(
    pct_diff = "**Difference (%)**"
  )

table4b_diff  
```

## ST 4 - TC descriptives and association with detection mode

TC descriptives by detection mode, logistic regression of TC with detection mode (same principal as risk factor analysis but multiple exposures are TC)

```{r}
#| echo: false

# summary Logistic Regression of Model 10  -----------------------------------------



# remove columns with total and interval N - not simple with gtsummary, trying from scratch without adding the columns


# Create clean version of logistic output, without Ns

model <- df %>%
  drop_na(all_of(tumour_char_tr), diagage, d_R1toBC_y
           ) %>%
  glm(formula = d_dmode_n ~ 
      diagage + d_R1toBC_y
      + d_inv_status + d_grade_tr + d_morph4_tr_lab + d_pos_nodes_tr_lab 
      + d_tmsize_tr_lab + d_er_tr_lab + d_pr_tr_lab + d_her2_tr_lab, 
            family = binomial(link = "logit"))

# tidy up output
model10 <- tbl_regression(model, 
                         exponentiate = TRUE,
                         label = list(
                          d_inv_status ~ "Invasive status",
                          d_grade_tr ~ "Grade",
                          d_morph4_tr_lab ~ "Morphology type",
                          d_pos_nodes_tr_lab ~ "Positive nodes",
                          d_tmsize_tr_lab ~ "Tumour size",
                          d_er_tr_lab ~ "ER status",
                          d_pr_tr_lab ~ "PR status",
                          d_her2_tr_lab ~ "HER2 status"
                         )
                         ) %>% 
  add_global_p() %>% 
  remove_row_type(variables = c(diagage, d_R1toBC_y), # remove covariates
                  type = "all")
  

#model10




# Create descriptives by mode of detection ------------------------------------
# select(d_dmode_n, d_R1alcohol_units_lab, d_R1smokingstatus_lab, d_bmi_entry_lab,  d_R1physmet_leis_who_lab, d_bbd_lab, d_fambrca_lab, d_age_menarche_lab, d_ocstatus_lab, d_R1menopause_lab3, d_age_meno_tr_lab, d_R1hrt_tr_lab, d_parity_tr_lab, d_agebirth1_tr_lab)

# note: need to rename the non-trick variables according to their trick name to align with the model, which can then allow table merge. 

# include missing 
# create df
  
descr_model10_df <- df %>% 
  select(d_dmode, all_of(tumour_char), diagage, d_R1toBC_y 
         ) %>% 
  mutate(d_grade_lab = as.character(if_else(d_grade_lab == "DCIS", NA, d_grade_lab)) # recode dcis to missing
  ) %>% 
  mutate(d_grade_lab = factor(x = d_grade_lab,
                              levels = c(1, 2, 3))
         ) %>% 
  mutate(d_grade_tr = d_grade_lab,
         d_morph4_tr_lab = d_morph4_lab,
         d_pos_nodes_tr_lab = d_pos_nodes_lab,
         d_tmsize_tr_lab = d_tumour_size_lab,
         d_er_tr_lab = d_er_status_lab,
         d_pr_tr_lab = d_pr_status_lab,
         d_her2_tr_lab = d_her2_status_lab
         ) %>% 
  select(-d_grade_lab, -d_morph4_lab, -d_pos_nodes_lab, -d_tumour_size_lab, -d_er_status_lab, -d_pr_status_lab, -d_her2_status_lab, -diagage, -d_R1toBC_y)  
  
  

# glimpse(descr_model2)
# str(test)

# create gs summary table
descr_model10 <- descr_model10_df %>% 
  tbl_summary(by = d_dmode,
              statistic = list(
                all_continuous() ~ "{mean} ({sd})",
                all_categorical() ~ "{n} ({p})"
                ),
              digits = all_continuous() ~ 2,
              label = list(
                          d_inv_status ~ "Invasive status",
                          d_grade_tr ~ "Grade",
                          d_morph4_tr_lab ~ "Morphology type",
                          d_pos_nodes_tr_lab ~ "Positive nodes",
                          d_tmsize_tr_lab ~ "Tumour size",
                          d_er_tr_lab ~ "ER status",
                          d_pr_tr_lab ~ "PR status",
                          d_her2_tr_lab ~ "HER2 status"
                         )
              ) %>% 
  add_overall()

#descr_model10
  
# merge with analysis output -------------------------------------------------


ST2 <- tbl_merge(
  tbls = list(descr_model10, model10),
  tab_spanner = c("", "")
   ) 
   
  
   

ST2
```

## ST 5 - Risk factor analysis adjusted for TC 

Sensitivity analysis - default model with model adjusted for TC

```{r}
#| echo: false

# Model 3 - subset on those with TC----------------------------------------------------------------
 # include parity and meno trick variable instead of original variables

# only included alcohol status - check if we need to do the trick if including both
# including "diagage", "d_R1toBC_y" - MD variables will be included in sub analysis

 model <- df %>%
   tidyr::drop_na(all_of(tumour_char_tr), all_of(all_expl_variables_tr), diagage, d_R1toBC_y
           ) %>% 
  glm(formula = d_dmode_n ~ d_R1menopause_lab3 + d_age_meno_tr_lab 
      + d_R1hrt_tr_lab
      + d_parity_lab + d_agebirth1_tr_lab + d_age_menarche_lab + d_ocstatus_lab 
      + d_bbd_lab + d_fambrca_lab
      + d_R1alcohol_units_lab + d_R1smokingstatus_lab 
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab
      + diagage + d_R1toBC_y,
            family = binomial(link = "logit"))

# tidy up output
model3 <- tbl_regression(model, 
                         exponentiate = TRUE,
                         label = list(
                          d_R1alcohol_units_lab ~ "Alcohol units",
                          d_R1smokingstatus_lab ~ "Smoking status",
                          d_bmi_entry_lab ~ "BMI",
                          d_R1physmet_leis_who_lab ~ "Physical Activity (WHO cat)",
                          d_bbd_lab ~ "Benign breast disease",
                          d_fambrca_lab ~ "Family history of BC",
                          d_age_menarche_lab ~ "Age at menarche",
                          d_ocstatus_lab ~ "Oral contraceptives status",
                          d_R1menopause_lab3 ~ "Menopause",
                          d_age_meno_tr_lab ~ "Age at menopause",
                          d_R1hrt_tr_lab ~ "HRT status",
                          d_parity_lab ~ "N of parous pregnancies",
                          d_agebirth1_tr_lab ~ "Age at first birth"
                         )
                         ) %>% 
  add_global_p() %>% 
  remove_row_type(variables = c(diagage, d_R1toBC_y), # remove covariates
                  type = "all")


model3a <- tbl_regression(model, exponentiate = TRUE) %>%
  add_n(location = "level")  %>%
  add_nevent(location = "level") %>%
  add_global_p() %>%
  modify_header(stat_nevent = "**Interval N**",
                stat_n = "**Total N**")



#model3
#model3a

# create a dataframe with broom::tidy() for forestplots 

tidy_model3 <- model %>%
   tidy(exponentiate = TRUE, conf.int = TRUE) %>%
   mutate(OR = estimate) %>%
   select(-estimate) %>%
   select(term, OR, everything())
#tidy_model3


# Model 4 ----------------------------------------------------------------
 # include parity and meno trick variable instead of original variables

# only included alcohol status - check if we need to do the trick if including both
# including "diagage", "d_R1toBC_y" - MD variables will be included in sub analysis

model <- df %>%
   tidyr::drop_na(all_of(tumour_char_tr), all_of(all_expl_variables_tr), diagage, d_R1toBC_y
           ) %>% 
  glm(formula = d_dmode_n ~ d_R1menopause_lab3 + d_age_meno_tr_lab 
      + d_R1hrt_tr_lab
      + d_parity_lab + d_agebirth1_tr_lab + d_age_menarche_lab + d_ocstatus_lab 
      + d_bbd_lab + d_fambrca_lab
      + d_R1alcohol_units_lab + d_R1smokingstatus_lab 
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab
      + diagage + d_R1toBC_y
      + d_inv_status + d_grade_tr + d_morph4_tr_lab + d_pos_nodes_tr_lab 
      + d_tmsize_tr_lab + d_er_tr_lab + d_pr_tr_lab + d_her2_tr_lab, 
            family = binomial(link = "logit"))

model4 <- tbl_regression(model, 
                         exponentiate = TRUE,
                         label = list(
                          d_R1alcohol_units_lab ~ "Alcohol units",
                          d_R1smokingstatus_lab ~ "Smoking status",
                          d_bmi_entry_lab ~ "BMI",
                          d_R1physmet_leis_who_lab ~ "Physical Activity (WHO cat)",
                          d_bbd_lab ~ "Benign breast disease",
                          d_fambrca_lab ~ "Family history of BC",
                          d_age_menarche_lab ~ "Age at menarche",
                          d_ocstatus_lab ~ "Oral contraceptives status",
                          d_R1menopause_lab3 ~ "Menopause",
                          d_age_meno_tr_lab ~ "Age at menopause",
                          d_R1hrt_tr_lab ~ "HRT status",
                          d_parity_lab ~ "N of parous pregnancies",
                          d_agebirth1_tr_lab ~ "Age at first birth",
                          d_inv_status ~ "Invasive status",
                          d_grade_tr ~ "Grade",
                          d_morph4_tr_lab ~ "Morphology type",
                          d_pos_nodes_tr_lab ~ "Positive nodes",
                          d_tmsize_tr_lab ~ "Tumour size",
                          d_er_tr_lab ~ "ER status",
                          d_pr_tr_lab ~ "PR status",
                          d_her2_tr_lab ~ "HER2 status"
                         )
                         ) %>% 
  add_global_p() %>% 
  remove_row_type(variables = c(diagage, d_R1toBC_y), # remove covariates
                  type = "all")

# model4
# 
 model4a <- tbl_regression(model, exponentiate = TRUE) %>%
  add_n(location = "level")  %>%
   add_nevent(location = "level") %>%
   add_global_p() %>%
   modify_header(stat_nevent = "**Interval N**",
                 stat_n = "**Total N**")

# model4a
# model3a


# create a dataframe with broom::tidy() for forestplots 

tidy_model4 <- model %>%
   tidy(exponentiate = TRUE, conf.int = TRUE) %>%
   mutate(OR = estimate) %>%
   select(-estimate) %>%
   select(term, OR, everything())


 # Descriptives for model 4 -------------------------------------------------
# subset on those with mammo density 


descr_TCsubset_df <- df %>% 
  select(d_dmode,  all_of(all_expl_variables_tr), all_of(all_expl_variables), all_of(tumour_char_tr), all_of(tumour_char), diagage, d_R1toBC_y) %>% 
  tidyr::drop_na(all_of(tumour_char_tr), all_of(all_expl_variables_tr), diagage, d_R1toBC_y
  ) %>% 
  mutate(d_bmi_entry_lab = fct_relevel(d_bmi_entry_lab, "<18.5", "18.5-25", "25-30", "30+")
              ) %>%
  # recode the names of normal variables to trick to allow merge between regression tables and this descriptive table
  mutate(d_age_meno_tr_lab = d_age_menopause_lab,
         d_R1hrt_tr_lab = d_R1hrtstatus_lab,
         d_agebirth1_tr_lab = d_agebirth1_lab,
         d_grade_tr = d_grade_lab,
         d_morph4_tr_lab = d_morph4_lab,
         d_pos_nodes_tr_lab = d_pos_nodes_lab,
         d_tmsize_tr_lab = d_tumour_size_lab,
         d_er_tr_lab = d_er_status_lab,
         d_pr_tr_lab = d_pr_status_lab,
         d_her2_tr_lab = d_her2_status_lab
         ) %>%
  select(-d_age_menopause_lab, -d_R1hrtstatus_lab, -d_agebirth1_lab, -diagage, -d_R1toBC_y, -d_grade_lab, -d_morph4_lab, -d_pos_nodes_lab, -d_tumour_size_lab, -d_er_status_lab, -d_pr_status_lab, -d_her2_status_lab)
  
  

# glimpse(descr_MDsubset_df)
# str(descr_MDsubset_df)

# create gs summary table
descr_TCsubset <- descr_TCsubset_df %>% 
  tbl_summary(by = d_dmode,
              statistic = list(
                all_continuous() ~ "{mean} ({sd})",
                all_categorical() ~ "{n} ({p})"
                ),
              digits = all_continuous() ~ 2,
              label = list(
                          d_R1alcohol_units_lab ~ "Alcohol units",
                          d_R1smokingstatus_lab ~ "Smoking status",
                          d_bmi_entry_lab ~ "BMI",
                          d_R1physmet_leis_who_lab ~ "Physical Activity (WHO cat)",
                          d_bbd_lab ~ "Benign breast disease",
                          d_fambrca_lab ~ "Family history of BC",
                          d_age_menarche_lab ~ "Age at menarche",
                          d_ocstatus_lab ~ "Oral contraceptives status",
                          d_R1menopause_lab3 ~ "Menopause",
                          d_age_meno_tr_lab ~ "Age at menopause",
                          d_R1hrt_tr_lab ~ "HRT status",
                          d_parity_lab ~ "N of parous pregnancies",
                          d_agebirth1_tr_lab ~ "Age at first birth",
                          d_inv_status ~ "Invasive status",
                          d_grade_tr ~ "Grade",
                          d_morph4_tr_lab ~ "Morphology type",
                          d_pos_nodes_tr_lab ~ "Positive nodes",
                          d_tmsize_tr_lab ~ "Tumour size",
                          d_er_tr_lab ~ "ER status",
                          d_pr_tr_lab ~ "PR status",
                          d_her2_tr_lab ~ "HER2 status"
                ),
              type = list(d_bbd_lab ~ "categorical",
                          d_fambrca_lab ~ "categorical")
              ) #%>% 
  #add_overall()


#descr_TCsubset
# check numbers against
# model4a
# model3a
 


# merge tables -------------------------------------------------


ST3 <- tbl_merge(
  tbls = list(descr_TCsubset, model3, model4),
  tab_spanner = c("Subset with TC" ,"Unadjusted for TC", "Adjusted for TC")
   ) 
   

#ST3


# adding pct difference -------------------------------------------

#  using gtsummary 
ST3_diff <- ST3 %>% 
  modify_table_body(
    ~ .x %>% 
      mutate(pct_diff = if_else(
          is.na(estimate_2) | is.na(estimate_3),
          "",
          sprintf("%.1f", (estimate_3 - estimate_2) / estimate_3 * 100))
    )
    ) %>% 
  modify_header(
    pct_diff = "**Difference (%)**"
  )

ST3_diff
```
