---
title: "Detection mode project: Tables and figures for publication"
author: "Martina Brayley, Reuben Frost"
date: last-modified
date-format: long
format: 
  html: 
    embed-resources: true
    toc: true
    toc-location: left
    number-sections: true
    theme: [flatly, style.css]
    page-layout: full
    grid:
      sidebar-width: 300px
      body-width: 2000px
      margin-width: 300px
      gutter-width: 1.5rem
editor: source
execute:
  warning: false
---

## Set-Up

```{r}
#| include: false
#| echo: false

# SET UP ------------------------------------------------------------------------


# execute master file - no need run it as final data now saved
#source("code/master.R")

print(getwd())

library(pacman)
p_load(here,
       broom.helpers,
       paleteer,
       ggthemes,
       nortest,
       cowplot,
       Cairo,
       flexlsx,
       openxlsx2)

# set up
setup_path <- here("code", "1.setup_import.r")
source(setup_path)

# functions are loaded in the setup script


# dataset with missing as separate categories
df <- readRDS("Q:/SHARED/USERS/MBrayley/Screening/data/an_df_missing_cat.rds") |>
  mutate(d_bmi_20_lab = as.factor(if_else(d_bmi_20_lab %in% c("25-30", "30+"), "25+", d_bmi_20_lab)),
         d_bmi_20_lab = relevel(d_bmi_20_lab, ref = "18.5-25"))

# create md_df dataset
md_df <- df %>% 
  filter(d_md_avail == 1)

# complete case data - for sensitivity only
df_compl_case <- readRDS("Q:/SHARED/USERS/MBrayley/Screening/data/an_df_compl_case.rds") |>
  mutate(d_bmi_20_lab = as.factor(if_else(d_bmi_20_lab %in% c("25-30", "30+"), "25+", d_bmi_20_lab)),
         d_bmi_20_lab = relevel(d_bmi_20_lab, ref = "18.5-25"))

```

```{r}
#| include: false
#| echo: false

# DATA PREP ----------------------------------------------------------------------


## Store explanatory variables as an object -------------------------------

# with labels instead of categories --------------------------------------
# select categorical variables that are labelled

# exploratory vars 
all_expl_variables <- c(
                         # parity related risk factors
                        "d_R1menopause_lab3", 
                        "d_age_menopause_lab",
                        "d_age_meno_cont",
                        "d_R1hrtstatus_lab",
                        "d_R1hrt_sttyp_lab",
                        "d_parity_lab", 
                        "d_parity",
                        "d_agebirth1_lab",
                        "d_agebirth1_cont",
                        "d_age_menarche_lab", 
                        "d_age_menarche",
                        "d_ocstatus_lab",
                        "d_ocstatus2_lab",
                        "d_breastfed_lab",
                        "d_bf_dur_lab",
                        "d_bf_dur_month",
                        # other biological risk factors
                        "d_bbd_lab", 
                        "d_fambrca_lab",
                        # lifestyle risk factors
                        "d_R1alcohol_units_lab",
                        "d_R1alcohol_units",
                        "d_R1smokingstatus_lab", 
                        "d_bmi_entry_lab",
                        "d_bmi_entry_per5",
                        "d_bmi_20_lab",
                        "d_bmi_20_per5", 
                        "d_R1physmet_leis_who_lab",
                        "d_R1physmet_leisure")

all_expl_variables_tr <- c(
                            # parity related risk factors
                           "d_R1menopause_lab3", 
                           "d_age_meno_tr_lab",
                           "d_age_meno_cont",
                           "d_R1hrt_tr_lab",
                           "d_R1hrt_sttyp_tr_lab",
                           "d_parity_lab",
                           "d_parity",
                           "d_agebirth1_tr_lab",
                           "d_agebirth1_cont",
                           "d_age_menarche_lab",
                           "d_age_menarche",
                           "d_ocstatus_lab",
                           "d_ocstatus2_lab",
                           "d_breastfed_tr_lab",
                           "d_bf_dur_tr_lab",
                            # other biological risk factors
                           "d_bbd_lab", 
                           "d_fambrca_lab",
                           # lifestyle risk factors
                           
                           "d_R1alcohol_units_lab", 
                           "d_R1alcohol_units",
                           "d_R1smokingstatus_lab", 
                           "d_bmi_entry_lab",
                           "d_bmi_entry_per5",
                           "d_bmi_20_lab",
                           "d_bmi_20_per5",
                           "d_R1physmet_leis_who_lab",
                           "d_R1physmet_leisure"
                           )
  

lifestyle <- c("d_R1alcohol_units_lab", 
               "d_R1smokingstatus_lab", 
               "d_bmi_entry_lab", 
               "d_R1physmet_leis_who")
  
reproductive <- c("d_age_menarche_lab", 
                  "d_ocstatus_lab", 
                  "d_R1menopause_lab3", 
                  "d_age_menopause_lab", 
                  "d_R1hrtstatus_lab", 
                  "d_parity_lab", 
                  "d_agebirth1_lab") # do we want to add breastfeeding? If so, it would also have be in the trick model
  
orther_rf <- c("d_bbd", 
               "d_fambrca")

# mammo density excluded from the lists - special variable for subanalyses

# covariates 
all_covariates <- c(# time varying variables
                    "diagage", 
                    "d_R1toBC_y", 
                    "d_R1toMD_y", 
                    "d_MDtoBC_y",
                    # tumour characteristics
                    "d_inv_status", 
                    "d_grade_lab", 
                    "d_morph4", 
                    "d_pos_nodes_lab", 
                    "d_tumour_size", 
                    "d_er_status", 
                    "d_pr_status", 
                    "d_her2_status")  
  
tumour_char <- c("d_inv_status", 
                 "d_grade_lab", 
                 "d_morph4_lab", 
                 "d_pos_nodes_lab", 
                 "d_tumour_size_lab", 
                 "d_er_status_lab", 
                 "d_pr_status_lab", 
                 "d_her2_status_lab")

tumour_char_tr <- c("d_inv_status", 
                    "d_grade_tr", 
                    "d_morph4_tr_lab", 
                    "d_pos_nodes_tr_lab", 
                    "d_tmsize_tr_lab", 
                    "d_er_tr_lab", 
                    "d_pr_tr_lab", 
                    "d_her2_tr_lab")
  
time_vars <- c("diagage", 
               "d_R1toBC_y", 
               "d_R1toMD_y", 
               "d_MDtoBC_y",
               "d_MDtoBC_lab",
               "yeardiag")

# create dataset 
#df <- an_df %>% 
#  mutate(d_bmi_entry_lab = fct_relevel(d_bmi_entry_lab, "18.5-25", "<18.5", "25-30", "30+")
#  )

#df <- an_df

# set theme for gtsummary tables: 

theme_gtsummary_compact(set_theme = TRUE)


```

```{r}

theme_minimal_custom <- function(){
  
  theme_minimal(base_size = 8) +
    theme(
      plot.title = element_text(size = 10, face = "bold"),
      axis.title = element_text(size = 10, face = "bold"),
      axis.text = element_text(size = 8),
      legend.title = element_text(size = 10, face = "bold"),
      legend.text = element_text(size = 10),
      strip.text = element_text(size = 10, face = "bold"),
      plot.caption = element_text(size = 10)
    )
  
}

```

### Table and figure structure

**Main figures**

**Table 1:** Tumour characteristics of interval and screen-detected breast cancers

**Figure 1:** Flow chart for selection of Generations Study breast cancer cases eligible for analyses (created in drawio (saved into the paper folder) and mermaid (memaid issue with formatting)

**Figure 2:** Mammographic density by detection mode and tumour characteristics

**Figure 3:** Mammographic density by risk factor

**Figure 4:** Forest plot of OR estimates for interval breast cancers based on 3 complexities of models

**Supplementary material**

**Supplementary Table 1:** Comparison of derived detection mode and regsitry detection mode

**Supplementary Table 2:** Multivariable model predicting log(Mammographic Density), both adjusted and non-adjusted for age at mammogram

**Supplementary Table 3:** OR estiamtes for interval breast cancer for 4 complexities of models


## Main figures and tables

### Table 1 - Tumour characteristics of interval and screen-detected breast cancers

by detection mode and with p-value from logistic regression model with tumour characteristics in the model (from supplementary table 1) + adjusted for time variables

```{r}
#| echo: false

# create categorical for year of diagnosis
#df %>% tabyl(yeardiag)

df <- df %>%
  # mutate(across(all_of(c("d_grade_tr", "d_morph4_tr_lab", "d_pos_nodes_tr_lab", "d_tmsize_tr_lab", "d_er_tr_lab", "d_pr_tr_lab", "d_her2_tr_lab")), function(x){
  #   if_else(x == "Missing", "Unknown, N", x)
  # })) |> 
  mutate(yeardiag_cat = case_when(yeardiag < 2009 ~ "2004-2008",
                                  yeardiag >= 2009 & yeardiag < 2014 ~ "2009-2013",
                                  yeardiag >= 2014 & yeardiag < 2019 ~ "2014-2018"
                                  ),
         d_R1toBC_lab = fct_relevel(d_R1toBC_lab, "<3 years", "3-5 years", "6-8 years", "9-11 years", "12+ years")
         )

# compute regression model -----------------------------------
# Adding TC trick variables

model <- df %>%
  glm(formula = d_dmode_n ~ 
       diagage + d_R1toBC_y + yeardiag_cat +
      d_inv_status + d_grade_tr + d_morph4_tr_lab + d_pos_nodes_tr_lab 
      + d_tmsize_tr_lab + d_er_tr_lab + d_pr_tr_lab + d_her2_tr_lab, 
            family = binomial(link = "logit"))


# tidy up output
model_t1 <- tbl_regression(model, 
                         exponentiate = TRUE,
                         label = list(
                          d_inv_status ~ "Invasive status, N (%)",
                          d_grade_tr ~ "Grade, N (%)",
                          d_morph4_tr_lab ~ "Morphology type, N (%)",
                          d_pos_nodes_tr_lab ~ "Node status, N (%)",
                          d_tmsize_tr_lab ~ "Tumour size (mm), N (%)",
                          d_er_tr_lab ~ "ER status, N (%)",
                          d_pr_tr_lab ~ "PR status, N (%)",
                          d_her2_tr_lab ~ "HER2 status, N (%)",
                          diagage ~ "Age at diagnosis, mean (SD)",
                           # AgeatEntry ~ "Age at recruitment, mean (SD)",
                           d_R1toBC_y ~ "Time from recruitment to diagnosis (years), mean (SD)",
                          # d_R1toMD_y ~ "Time from recruitment to mammogram (years), mean (SD)",
                          # mammo_age ~ "Age at mammogram, mean (SD)",
                          yeardiag_cat ~ "Year of diagnosis, mean (SD)"
                          
                         ),
                         missing_text = "Unknown, N",
                         pvalue_fun = custom_pvalue_fun
                         ) |>
  # remove Missing levels from regression output
  modify_table_body(
    ~ .x %>%
      dplyr::filter(!grepl("Missing", .data$label, ignore.case = TRUE))  # Filter 'missing' levels
  )

#model_t1 
# descriptives ------------------------------------

# note: need to rename the non-trick variables according to their trick name to align with the model, which can then allow table merge. 

# include missing 
# create df
  
descr_t1_df <- df %>% 
  mutate(mammo_age = d_R1toMD_y + AgeatEntry) |>
  mutate(d_R1toMD_y = abs(d_R1toMD_y)) |>
  select(d_dmode, all_of(tumour_char), diagage, d_R1toBC_y, d_R1toMD_y, AgeatEntry, mammo_age, yeardiag_cat
  # select(d_dmode, all_of(tumour_char), diagage, d_R1toBC_y, yeardiag_cat
         ) %>% 
  mutate(d_grade_tr = d_grade_lab,
         d_morph4_tr_lab = d_morph4_lab,
         d_pos_nodes_tr_lab = d_pos_nodes_lab,
         d_tmsize_tr_lab = d_tumour_size_lab,
         d_er_tr_lab = d_er_status_lab,
         d_pr_tr_lab = d_pr_status_lab,
         d_her2_tr_lab = d_her2_status_lab
         ) %>% 
  select(-d_grade_lab, -d_morph4_lab, -d_pos_nodes_lab, -d_tumour_size_lab, -d_er_status_lab, -d_pr_status_lab, -d_her2_status_lab)  
  
#str(descr_t1_df)


# create gs summary table
descr_t1 <- descr_t1_df %>% 
  # reselect to ensure the correct order
  mutate(d_dmode = if_else(d_dmode == "I", "Interval", "Screen-Detected")) |>
  select(d_dmode,
         AgeatEntry,
         mammo_age,
         diagage,
         d_R1toMD_y,
         d_R1toBC_y,
         yeardiag_cat,
         d_inv_status, 
         d_grade_tr, 
         d_morph4_tr_lab, 
         d_pos_nodes_tr_lab, 
         d_tmsize_tr_lab, 
         d_er_tr_lab, 
         d_pr_tr_lab, 
         d_her2_tr_lab, 
) %>% 
  tbl_summary(by = d_dmode,
              statistic = list(
                all_continuous() ~ "{mean} ({sd})",
                all_categorical() ~ "{n} ({p})"
                ),
              missing = "ifany",
              missing_text = "Unknown, N",
              digits = all_continuous() ~ 2,
              label = list(
                          d_inv_status ~ "Invasive status, N (%)",
                          d_grade_tr ~ "Grade, N (%)",
                          d_morph4_tr_lab ~ "Morphology type, N (%)",
                          d_pos_nodes_tr_lab ~ "Node status, N (%)",
                          d_tmsize_tr_lab ~ "Tumour size (mm), N (%)",
                          d_er_tr_lab ~ "ER status, N (%)",
                          d_pr_tr_lab ~ "PR status, N (%)",
                          d_her2_tr_lab ~ "HER2 status, N (%)",
                          diagage ~ "Age at diagnosis, mean (SD)",
                           AgeatEntry ~ "Age at recruitment, mean (SD)",
                           d_R1toBC_y ~ "Time from recruitment to diagnosis (years), mean (SD)",
                          d_R1toMD_y ~ "Time from recruitment to mammogram (years), mean (SD)",
                          mammo_age ~ "Age at mammogram, mean (SD)",
                          yeardiag_cat ~ "Year of diagnosis, mean (SD)"
                          
                         )
              ) 


#descr_t1



# join tables 

# merge with analysis output -------------------------------------------------


table1 <- tbl_merge(
  tbls = list(descr_t1, model_t1),
  tab_spanner = FALSE
  )
  
#table1
   
# remove unwanted columns (effect and CI) ----------------------

#show_header_names(table1) 


table1 <- table1 %>% 
  modify_column_hide(columns = c(estimate_2, ci_2, conf.low_2)
                     ) 
# show_header_names(table1)


table1 <- table1 %>%
  modify_footnote(
    p.value_2 ~ "p-value from logistic regression multivariate model of all listed variables"
  )

# Manually changing rows with inviasive/DCIS labels
table1$table_body$label[13] <- "DCIS"
table1$table_body$label[14] <- "Invasive"
table1$table_body$label[23] <- "Mixed"

# Manually adding % for the unknown categories for mammogram age
# table1$table_body$stat_1_1[table1$table_body$label == "Unknown, N"] <- paste0(table1$table_body$stat_1_1[table1$table_body$label == "Unknown, N"], " (",  round(100*as.numeric(table1$table_body$stat_1_1[table1$table_body$label == "Unknown"])/table1$table_styling$header$modify_stat_n[!is.na(table1$table_styling$header$modify_stat_n)][1]), ")")
# 
# 
# 
# table1$table_body$stat_2_1[table1$table_body$label == "Unknown, N"] <- paste0(table1$table_body$stat_2_1[table1$table_body$label == "Unknown, N"], " (",  round(100*as.numeric(table1$table_body$stat_2_1[table1$table_body$label == "Unknown, N"])/table1$table_styling$header$modify_stat_n[!is.na(table1$table_styling$header$modify_stat_n)][2]), ")")


table1

table1 |>
  as_flex_table() |>
  flextable::save_as_docx(path = "../outputs/figures/main/table_1.docx")

```


### Figure 2 - Mammographic density by mode of detection and tumour characteristics

```{r}
#| echo: false
#| include: false
#| warning: false


# rf vars for violin plots
explanatory_vars <- c(
                  "d_dmode",
                 "d_inv_status", 
                 "d_grade_lab", 
                 "d_morph4_lab", 
                 "d_pos_nodes_lab", 
                 "d_tumour_size_lab", 
                 "d_er_status_lab", 
                 "d_pr_status_lab", 
                 "d_her2_status_lab"
                        
                      )


# Summary
table_summary <- df_compl_case |>
  mutate(d_bmi_entry_lab = fct_relevel(d_bmi_entry_lab, "<18.5", "18.5-25", "25-30", "30+"),
        d_bmi_20_lab = fct_relevel(d_bmi_20_lab, "<18.5", "18.5-25", "25-30", "30+")
              ) |>
  filter(d_md_avail == 1) |>
   mutate(d_bf_dur_lab = if_else(d_bf_dur_lab == "Missing", NA, d_bf_dur_lab),
         d_age_menopause_lab = if_else(d_age_menopause_lab == "Pre-menopausal", NA, d_age_menopause_lab),
         d_R1hrtstatus_lab = if_else(d_R1hrtstatus_lab == "Pre-menopausal", NA, d_R1hrtstatus_lab),
         d_agebirth1_lab = if_else(d_agebirth1_lab == "Non-parous", NA, d_agebirth1_lab),
         d_dmode = factor(d_dmode, levels = c("SD", "I"))
         ) |> 
  select(any_of(explanatory_vars)
         ) |> 
    tbl_summary(
              statistic = list(
                all_continuous() ~ "{mean} ({sd})",
                all_categorical() ~ "{n}"
                ),
              digits = all_continuous() ~ 2,
              missing = "no"
    ) |>
    modify_table_body(
      ~ filter(.x, stat_0 != 0 | is.na(stat_0))
    )

# Summary of mammographic
table_summary_md <- df_compl_case |>
  mutate(d_bmi_entry_lab = fct_relevel(d_bmi_entry_lab, "<18.5", "18.5-25", "25-30", "30+"),
        d_bmi_20_lab = fct_relevel(d_bmi_20_lab, "<18.5", "18.5-25", "25-30", "30+") 
              ) |>
  filter(d_md_avail == 1) |>
   mutate(d_bf_dur_lab = if_else(d_bf_dur_lab == "Missing", NA, d_bf_dur_lab),
         d_age_menopause_lab = if_else(d_age_menopause_lab == "Pre-menopausal", NA, d_age_menopause_lab),
         d_R1hrtstatus_lab = if_else(d_R1hrtstatus_lab == "Pre-menopausal", NA, d_R1hrtstatus_lab),
         d_agebirth1_lab = if_else(d_agebirth1_lab == "Non-parous", NA, d_agebirth1_lab),
         d_dmode = factor(d_dmode, levels = c("SD", "I"))
         ) |> 
  select(any_of(explanatory_vars), d_md
         ) |> 
    tbl_custom_summary(
              statistic = list(
                everything() ~ "{mean} ({sd})"
                ),
              stat_fns  = list(
                everything() ~ function(data, full_data, variable, by, type, ...){
                  dplyr::summarize(data, mean = mean(d_md),
                           sd = sd(d_md))
                }
              ),
              digits = all_continuous() ~ 1,
              missing = "no"
    ) |>
    modify_table_body(
      ~ filter(.x, stat_0 != 0 | is.na(stat_0))
    )


# Unadjusted:
table_unadjusted <- df_compl_case |> 
  filter(d_md_avail == 1
  ) |> 
  mutate(d_bf_dur_lab = if_else(d_bf_dur_lab == "Missing", NA, d_bf_dur_lab),
         d_age_menopause_lab = if_else(d_age_menopause_lab == "Pre-menopausal", NA, d_age_menopause_lab),
         d_R1hrtstatus_lab = if_else(d_R1hrtstatus_lab == "Pre-menopausal", NA, d_R1hrtstatus_lab),
         d_agebirth1_lab = if_else(d_agebirth1_lab == "Non-parous", NA, d_agebirth1_lab),
         d_dmode = factor(d_dmode, levels = c("SD", "I"))
         ) |> 
  select(d_md, any_of(explanatory_vars)
         ) |> 
  tbl_uvregression(
    method = lm,
    y = d_md,
    pvalue_fun = custom_pvalue_fun,
    estimate_fun = function(x){
      case_when(
        is.na(x) ~ NA,
        TRUE ~ formatC(x, format = "f", digits = 2)
      )
    },
    # pvalue_fun = ~ .x,
    # estimate_fun = ~ .x
    method.args = list(na.action = na.omit),
    hide_n = T
   ) |> 
  bold_labels() 


#table_unadjusted


# rf vars for violin plots
explanatory_vars <- c(
                  "d_dmode",
                 "d_inv_status", 
                 "d_grade_lab", 
                 "d_morph4_lab", 
                 "d_pos_nodes_lab", 
                 "d_tumour_size_lab", 
                 "d_er_status_lab", 
                 "d_pr_status_lab", 
                 "d_her2_status_lab"
                        
                      )


# Adjusted table
# List to store regression tables
table_list <- list()

# data manipulation to remove redundant levels 

df_temp <- df_compl_case |> 
  mutate(d_bf_dur_lab = if_else(d_bf_dur_lab == "Missing", NA, d_bf_dur_lab),
         d_age_menopause_lab = if_else(d_age_menopause_lab == "Pre-menopausal", NA, d_age_menopause_lab),
         d_R1hrtstatus_lab = if_else(d_R1hrtstatus_lab == "Pre-menopausal", NA, d_R1hrtstatus_lab),
         d_agebirth1_lab = if_else(d_agebirth1_lab == "Non-parous", NA, d_agebirth1_lab),
         d_age_mammo = AgeatEntry + d_R1toMD_y,
         d_dmode = factor(d_dmode, levels = c("SD", "I"))
         )


# Loop through each explanatory variable and create a regression table
for (var in explanatory_vars) {
  # Run the regression model adjusted for age
  model <- lm(as.formula(paste("d_md ~", var, "+ d_age_mammo")), data = df_temp, na.action = na.omit)

  # Create the regression table and store it in the list
  table_list[[var]] <- tbl_regression(
    model,
    pvalue_fun = custom_pvalue_fun,
    estimate_fun = function(x){
      case_when(
        is.na(x) ~ NA,
        TRUE ~ formatC(x, format = "f", digits = 2)
      )
    }
  ) %>%
    remove_row_type(variables = "d_age_mammo", type = "all")  # Remove the age covariate from the table
}

# Stack all the regression tables into one
table_adjusted <- tbl_stack(table_list)

# View the final stacked table
table_adjusted

# Merge the tables

md_desc_table <- tbl_merge(list(table_summary, table_summary_md, table_adjusted),
                           tab_spanner = c(" ", "**Density**", "**Adjusted for age at mammogram**"
                                           )
                         )
                         
# Changing variable names to make sense
tab_labels <- c(
  "d_dmode" = "Mode of detection",
  # "d_age_menarche_lab" = "Age at menarche",
  # "d_ocstatus2_lab" = "Oral contraception status",
  # "d_parity_lab" = "Parity",
  # "d_agebirth1_lab" = "Age at first birth",
  # "d_bf_dur_lab" = "Breastfeeding duration (mths)",
  # "d_R1menopause_lab3" = "Menopausal status",
  # "d_age_menopause_lab" = "Age at menopause",
  # "d_R1hrtstatus_lab" = "MHT status",
  # "d_bbd_lab" = "Benign breast disease",
  # "d_fambrca_lab" = "Family history of breast cancer",
  # "d_bmi_20_lab" = "BMI at age 20 (kg/m²)",
  # "d_bmi_entry_lab" = "BMI at recruitment (kg/m²)",
  # "d_R1physmet_leis_who_lab" = "Physical activity (MET h/wk)",
  # "d_R1alcohol_units_lab" = "Alcohol units per week",
  # "d_R1smokingstatus_lab" = "Smoking status",
  "d_inv_status" = "Invasive status",
  "d_grade_lab" = "Grade",
  "d_morph4_lab" = "Morphology",
  "d_pos_nodes_lab" = "Node status",
  "d_tumour_size_lab" = "Tumour size (mm)",
  "d_er_status_lab" = "ER status",
  "d_pr_status_lab" = "PR status",
  "d_her2_status_lab" = "HER2 status"
)

for(i in 1:length(md_desc_table$table_body$label)){
  
  if(md_desc_table$table_body$label[[i]] %in% names(tab_labels)){
    
    md_desc_table$table_body$label[[i]] <- unname(tab_labels[which(names(tab_labels) == md_desc_table$table_body$label[[i]])])
    
  }
  
}

# Filter unneeded rows out
md_desc_table <- md_desc_table |>
  modify_table_body(
    ~ .x |>
      filter(stat_0_1 != "0" | is.na(stat_0_1)) |>
      filter(stat_0_2 != "NA (NA)" | is.na(stat_0_2)) |>
      filter(variable != "d_md")
  )

# Manually changing the labels of levels in the table
md_desc_table$table_body$label[md_desc_table$table_body$label == "Mixed: ductal and other"] <- "Mixed"

md_desc_table$table_body$label[grepl(md_desc_table$table_body$label, pattern = "months")] <- gsub(" months", "", md_desc_table$table_body$label[grepl(md_desc_table$table_body$label, pattern = "months")])

md_desc_table$table_body$label[md_desc_table$table_body$label == "I"] <- "Interval"
md_desc_table$table_body$label[md_desc_table$table_body$label == "SD"] <- "Screen-Detected"

md_desc_table$table_body$label[(md_desc_table$table_body$term_3) == "d_bbd_labYes"] <- "Ever"
md_desc_table$table_body$label[(md_desc_table$table_body$term_3) == "d_bbd_labNo"] <- "Never"


md_desc_table$table_body$label[(md_desc_table$table_body$term_3) == "d_ocstatus2_labNo"] <- "Never"
md_desc_table$table_body$label[(md_desc_table$table_body$term_3) == "d_ocstatus2_labYes"] <- "Ever"

md_desc_table$table_body$label[(md_desc_table$table_body$term_3) == "d_inv_status1"] <- "Invasive"
md_desc_table$table_body$label[(md_desc_table$table_body$term_3) == "d_inv_status0"] <- "DCIS"


# md_desc_table$table_body

# develop forest

plot <- md_desc_table$table_body |>
  mutate(mean = estimate_3,
         lower = conf.low_3,
         upper = conf.high_3,
         is_summary = is.na(term_3),
         p.value_3 = if_else(p.value_3 < 0.001, "<0.001", sprintf("%.3f", p.value_3)),
         p.value_3 = if_else(!is_summary & is.na(p.value_3), "-", p.value_3)
         ) %>% 
  forestplot(labeltext = c(label, stat_0_1, stat_0_2, p.value_3),
             xlab = "Difference in mammographic density (%), adjusted for age at mammogram",
             boxsize = 0.25,
             zero = 0,
             line.margin = 0.55,
             clip = c(-5, 10),
             xticks = c(-5, -2.5, 0, 2.5, 5, 7.5, 10),
             ci.vertices = T,
             ci.vertices.height = 0.15,
             is.summary = is_summary
             )   %>% 
  fp_set_style(
              box = c("red2") %>%  lapply(function(x) gpar(fill = x, col = "#666666")),
               default = gpar(vertices = T),
               line = "grey40",
               summary = gpar(col = NA), # Set summary color to NA
               # align = "lrrrr",
               hrz_lines = gpar(col = NA), # Set line color to NA
               txt_gp = fpTxtGp(label = gpar(cex = 1.2),
                                    ticks = gpar(cex = 1.2),
                                    xlab = gpar(cex = 1.2),
                                legend = gpar(cex = 1.2)
                                )
               ) %>%
  fp_decorate_graph(graph.pos = 4
                    ) |>
  fp_add_header(label = c("Risk Factor"),
                stat_0_1 = c("N"),
                stat_0_2 = c("Mean (SD)"),
                p.value_3 = c("P Value")
  )


plot

CairoPNG(file=paste0("../outputs/figures/main/figure_2.png"), width = 13, height = 13.5, units = "in", res = 300)
print(plot)
dev.off()

```

### Figure 3 - Mammographic density by risk factors
```{r}


# rf vars for violin plots
explanatory_vars <- c(
                  # "d_dmode",
                      "d_age_menarche_lab",
                  "d_ocstatus2_lab",
                  "d_parity_lab",
                  "d_agebirth1_lab",
                  "d_bf_dur_lab",
                  "d_R1menopause_lab3",
                  "d_age_menopause_lab", 
                  "d_R1hrtstatus_lab",
                  "d_bbd_lab",
                  "d_fambrca_lab",
                  "d_bmi_20_lab",
                  "d_bmi_entry_lab",
                  "d_R1physmet_leis_who_lab",
                  "d_R1alcohol_units_lab",
                  "d_R1smokingstatus_lab"
                        
                 # "d_inv_status", 
                 # "d_grade_lab", 
                 # "d_morph4_lab", 
                 # "d_pos_nodes_lab", 
                 # "d_tumour_size_lab", 
                 # "d_er_status_lab", 
                 # "d_pr_status_lab", 
                 # "d_her2_status_lab"
                        
                      )


# Summary
table_summary <- df_compl_case |>
  mutate(d_bmi_entry_lab = fct_relevel(d_bmi_entry_lab, "<18.5", "18.5-25", "25-30", "30+"),
        d_bmi_20_lab = fct_relevel(d_bmi_20_lab, "<18.5", "18.5-25", "25-30", "30+") 
              ) |>
  filter(d_md_avail == 1) |>
   mutate(d_bf_dur_lab = if_else(d_bf_dur_lab == "Missing", NA, d_bf_dur_lab),
         d_age_menopause_lab = if_else(d_age_menopause_lab == "Pre-menopausal", NA, d_age_menopause_lab),
         d_R1hrtstatus_lab = if_else(d_R1hrtstatus_lab == "Pre-menopausal", NA, d_R1hrtstatus_lab),
         d_agebirth1_lab = if_else(d_agebirth1_lab == "Non-parous", NA, d_agebirth1_lab),
         d_dmode = factor(d_dmode, levels = c("SD", "I"))
         ) |> 
  select(any_of(explanatory_vars)
         ) |> 
    tbl_summary(
              statistic = list(
                all_continuous() ~ "{mean} ({sd})",
                all_categorical() ~ "{n}"
                ),
              digits = all_continuous() ~ 2,,
              type = list(d_ocstatus2_lab ~ "categorical",
                          d_fambrca_lab ~ "categorical",
                          d_bbd_lab ~ "categorical"),
              missing = "no"
    ) |>
    modify_table_body(
      ~ filter(.x, stat_0 != 0 | is.na(stat_0))
    )

# Summary of mammographic
table_summary_md <- df_compl_case |>
  mutate(d_bmi_entry_lab = fct_relevel(d_bmi_entry_lab, "<18.5", "18.5-25", "25-30", "30+"),
        d_bmi_20_lab = fct_relevel(d_bmi_20_lab, "<18.5", "18.5-25", "25-30", "30+") 
              ) |>
  filter(d_md_avail == 1) |>
   mutate(d_bf_dur_lab = if_else(d_bf_dur_lab == "Missing", NA, d_bf_dur_lab),
         d_age_menopause_lab = if_else(d_age_menopause_lab == "Pre-menopausal", NA, d_age_menopause_lab),
         d_R1hrtstatus_lab = if_else(d_R1hrtstatus_lab == "Pre-menopausal", NA, d_R1hrtstatus_lab),
         d_agebirth1_lab = if_else(d_agebirth1_lab == "Non-parous", NA, d_agebirth1_lab),
         d_dmode = factor(d_dmode, levels = c("SD", "I"))
         ) |> 
  select(any_of(explanatory_vars), d_md
         ) |> 
    tbl_custom_summary(
              statistic = list(
                everything() ~ "{mean} ({sd})"
                ),
              stat_fns  = list(
                everything() ~ function(data, full_data, variable, by, type, ...){
                  dplyr::summarize(data, mean = mean(d_md),
                           sd = sd(d_md))
                }
              ),
              digits = all_continuous() ~ 1,
              type = list(d_ocstatus2_lab ~ "categorical",
                          d_fambrca_lab ~ "categorical",
                          d_bbd_lab ~ "categorical"),
              missing = "no"
    ) |>
    modify_table_body(
      ~ filter(.x, stat_0 != 0 | is.na(stat_0))
    )


# rf vars for violin plots
explanatory_vars <- c(
  
                  # "d_dmode",
                      "d_age_menarche_lab",
                  "d_ocstatus2_lab",
                  "d_parity_lab",
                  "d_agebirth1_lab",
                  "d_bf_dur_lab",
                  "d_R1menopause_lab3",
                  "d_age_menopause_lab", 
                  "d_R1hrtstatus_lab",
                  "d_bbd_lab",
                  "d_fambrca_lab",
                  "d_bmi_20_lab",
                  "d_bmi_entry_lab",
                  "d_R1physmet_leis_who_lab",
                  "d_R1alcohol_units_lab",
                  "d_R1smokingstatus_lab"
                        
                 # "d_inv_status", 
                 # "d_grade_lab", 
                 # "d_morph4_lab", 
                 # "d_pos_nodes_lab", 
                 # "d_tumour_size_lab", 
                 # "d_er_status_lab", 
                 # "d_pr_status_lab", 
                 # "d_her2_status_lab"
                        
                      )


# Adjusted table
# List to store regression tables
table_list <- list()

# data manipulation to remove redundant levels 

df_temp <- df_compl_case |> 
  mutate(d_bf_dur_lab = if_else(d_bf_dur_lab == "Missing", NA, d_bf_dur_lab),
         d_age_menopause_lab = if_else(d_age_menopause_lab == "Pre-menopausal", NA, d_age_menopause_lab),
         d_R1hrtstatus_lab = if_else(d_R1hrtstatus_lab == "Pre-menopausal", NA, d_R1hrtstatus_lab),
         d_agebirth1_lab = if_else(d_agebirth1_lab == "Non-parous", NA, d_agebirth1_lab),
         d_age_mammo = AgeatEntry + d_R1toMD_y,
         d_dmode = factor(d_dmode, levels = c("SD", "I"))
         )


# Loop through each explanatory variable and create a regression table
for (var in explanatory_vars) {
  # Run the regression model adjusted for age
  model <- lm(as.formula(paste("d_md ~", var, "+ d_age_mammo")), data = df_temp, na.action = na.omit)

  # Create the regression table and store it in the list
  table_list[[var]] <- tbl_regression(
    model,
    pvalue_fun = custom_pvalue_fun,
    estimate_fun = function(x){
      case_when(
        is.na(x) ~ NA,
        TRUE ~ formatC(x, format = "f", digits = 2)
      )
    },
  ) %>%
    remove_row_type(variables = "d_age_mammo", type = "all")  # Remove the age covariate from the table
}

# Stack all the regression tables into one
table_adjusted <- tbl_stack(table_list)

# View the final stacked table
table_adjusted

# Merge the tables

md_desc_table <- tbl_merge(list(table_summary, table_summary_md, table_adjusted),
                           tab_spanner = c(" ", "**Density**", "**Adjusted for age at mammogram**"
                                           )
                         )
                         
# Changing variable names to make sense
tab_labels <- c(
  # "d_dmode" = "Mode of detection",
  "d_age_menarche_lab" = "Age at menarche",
  "d_ocstatus2_lab" = "Oral contraception status",
  "d_parity_lab" = "Parity",
  "d_agebirth1_lab" = "Age at first birth",
  "d_bf_dur_lab" = "Breastfeeding duration (mths)",
  "d_R1menopause_lab3" = "Menopausal status",
  "d_age_menopause_lab" = "Age at menopause",
  "d_R1hrtstatus_lab" = "MHT status",
  "d_bbd_lab" = "Benign breast disease",
  "d_fambrca_lab" = "Family history of breast cancer",
  "d_bmi_20_lab" = "BMI at age 20 (kg/m²)",
  "d_bmi_entry_lab" = "BMI at recruitment (kg/m²)",
  "d_R1physmet_leis_who_lab" = "Physical activity (MET h/wk)",
  "d_R1alcohol_units_lab" = "Alcohol units per week",
  "d_R1smokingstatus_lab" = "Smoking status"
  # "d_inv_status" = "Invasive status",
  # "d_grade_lab" = "Grade",
  # "d_morph4_lab" = "Morphology",
  # "d_pos_nodes_lab" = "Node status",
  # "d_tumour_size_lab" = "Tumour size (mm)",
  # "d_er_status_lab" = "ER status",
  # "d_pr_status_lab" = "PR status",
  # "d_her2_status_lab" = "HER2 status"
)

for(i in 1:length(md_desc_table$table_body$label)){
  
  if(md_desc_table$table_body$label[[i]] %in% names(tab_labels)){
    
    md_desc_table$table_body$label[[i]] <- unname(tab_labels[which(names(tab_labels) == md_desc_table$table_body$label[[i]])])
    
  }
  
}

# Filter unneeded rows out
md_desc_table <- md_desc_table |>
  modify_table_body(
    ~ .x |>
      filter(stat_0_1 != "0" | is.na(stat_0_1)) |>
      filter(stat_0_2 != "NA (NA)" | is.na(stat_0_2)) |>
      filter(variable != "d_md")
  )

# Manually changing the labels of levels in the table
md_desc_table$table_body$label[md_desc_table$table_body$label == "Mixed: ductal and other"] <- "Mixed"

md_desc_table$table_body$label[grepl(md_desc_table$table_body$label, pattern = "months")] <- gsub(" months", "", md_desc_table$table_body$label[grepl(md_desc_table$table_body$label, pattern = "months")])

md_desc_table$table_body$label[md_desc_table$table_body$label == "I"] <- "Interval"
md_desc_table$table_body$label[md_desc_table$table_body$label == "SD"] <- "Screen-Detected"

md_desc_table$table_body$label[(md_desc_table$table_body$term_3) == "d_bbd_labYes"] <- "Ever"
md_desc_table$table_body$label[(md_desc_table$table_body$term_3) == "d_bbd_labNo"] <- "Never"


md_desc_table$table_body$label[(md_desc_table$table_body$term_3) == "d_ocstatus2_labNo"] <- "Never"
md_desc_table$table_body$label[(md_desc_table$table_body$term_3) == "d_ocstatus2_labYes"] <- "Ever"

md_desc_table$table_body$label[(md_desc_table$table_body$term_3) == "d_inv_status1"] <- "Invasive"
md_desc_table$table_body$label[(md_desc_table$table_body$term_3) == "d_inv_status0"] <- "DCIS"



# develop forest

plot <- md_desc_table$table_body |>
  mutate(mean = estimate_3,
         lower = conf.low_3,
         upper = conf.high_3,
         is_summary = is.na(term_3),
         p.value_3 = if_else(p.value_3 < 0.001, "<0.001", sprintf("%.3f", p.value_3)),
         p.value_3 = if_else(!is_summary & is.na(p.value_3), "-", p.value_3)
         ) %>% 
  forestplot(labeltext = c(label, stat_0_1, stat_0_2, p.value_3),
             xlab = "Difference in mammographic density (%), adjusted for age at mammogram",
             boxsize = 0.25,
             zero = 0,
             line.margin = 0.55,
             clip = c(-20, 15),
             xticks = seq(-20, 15, by = 5),
             # xticks = c(-5, -2.5, 0, 2.5, 5, 7.5, 10),
             ci.vertices = T,
             ci.vertices.height = 0.15,
             is.summary = is_summary
             )   %>% 
  fp_set_style(
              box = c("red2") %>%  lapply(function(x) gpar(fill = x, col = "#666666")),
               default = gpar(vertices = T),
               line = "grey40",
               summary = gpar(col = NA), # Set summary color to NA
               # align = "lrrrr",
               hrz_lines = gpar(col = NA), # Set line color to NA
               txt_gp = fpTxtGp(label = gpar(cex = 1.2),
                                    ticks = gpar(cex = 1.2),
                                    xlab = gpar(cex = 1.2),
                                legend = gpar(cex = 1.2)
                                )
               ) %>%
  fp_decorate_graph(graph.pos = 4
                    ) |>
  fp_add_header(label = c("Risk Factor"),
                stat_0_1 = c("N"),
                stat_0_2 = c("Mean (SD)"),
                p.value_3 = c("P Value")
  )


plot

CairoPNG(file=paste0("../outputs/figures/main/figure_3.png"), width = 13, height = 22, units = "in", res = 300)
print(plot)
dev.off()

```

### Figure 4 - Forest plot

```{r}
#| echo: false
#| warning: false

# summary Logistic Regression of Model 1  -----------------------------------------
model <- df %>%
  glm(formula = d_dmode_n ~ d_R1menopause_lab3 + d_age_meno_tr_lab 
      + d_R1hrt_tr_lab
      + d_parity_lab + d_agebirth1_tr_lab + d_age_menarche_lab + d_ocstatus2_lab 
      + d_bbd_lab + d_fambrca_lab
      + d_R1alcohol_units_lab + d_R1smokingstatus_lab 
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab + d_bmi_20_lab + d_bf_dur_tr_lab
      + diagage + d_R1toBC_y + yeardiag,
            family = binomial(link = "logit"))

tidy_model1 <- model %>%
   tidy(exponentiate = TRUE, conf.int = TRUE) 

# summary Logistic Regression of Model 2  -----------------------------------------

 model <- df %>%
  glm(formula = d_dmode_n ~ d_R1menopause_lab3 + d_age_meno_tr_lab 
      + d_R1hrt_tr_lab
      + d_parity_lab + d_agebirth1_tr_lab + d_age_menarche_lab + d_ocstatus2_lab 
      + d_bbd_lab + d_fambrca_lab
      + d_R1alcohol_units_lab + d_R1smokingstatus_lab 
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab + d_bmi_20_lab + d_bf_dur_tr_lab
      + diagage + d_R1toBC_y + yeardiag
      + d_md_qrt + d_MDtoBC_lab,
            family = binomial(link = "logit"))

tidy_model2 <- model %>%
   tidy(exponentiate = TRUE, conf.int = TRUE)

```

```{r}
#| echo: false

# 2. PROCESS DATA FOR FORESTPLOT 

# full join data from both models 
md_forest_df <- tidy_model1 %>% 
  full_join(tidy_model2, by = "term", suffix = c("_mdl1", "_mdl2")) # |>
  # full_join(tidy_model3 |> rename_with(~ paste0(., "_mdl3"), -term), by = "term", suffix = c("_mdl3", ""))


# Need to add reference levels to the dataframe
for(var in names(model.frame(model))){
  
  # Check for factor
  if(is.factor(model.frame(model)[[var]])){
    
    # Extract reference level
    ref_level <- levels(model.frame(model)[[var]])[1]
    
    # Find current top row for the variable
    top_row <- min(which(grepl(md_forest_df$term, pattern = var)))
    
    # Add in new row
    new_row <- tibble(term = paste0(var, ref_level))
        
    md_forest_df <- bind_rows(
      slice(md_forest_df, 1:(top_row-1)),
      new_row,
      slice(md_forest_df, top_row:n())
    )
    
  }
  
}

# create labels and variable names structures 
md_forest_df <- md_forest_df %>% 
  mutate(variable = case_when(
    grepl("menopause", term) ~ "Menopausal status",
    grepl("age_meno", term) ~ "Age at menopause",
    grepl("hrt", term) ~ "MHT status",
    grepl("parity", term) ~ "Parity",
    grepl("agebirth1", term) ~ "Age at first birth",
    grepl("age_menarche", term) ~ "Age at menarche", 
    grepl("ocstatus2", term) ~ "Oral contraceptive use",
    grepl("bbd", term) ~ "Benign breast disease",
    grepl("fambrca", term) ~ "Family history of breast cancer",
    grepl("R1alcohol", term) ~ "Alcohol units per week",
    grepl("R1smokingstatus", term) ~ "Smoking status",
    grepl("bmi_entry", term) ~ "BMI at recruitment (kg/m²)*",
    grepl("R1physmet_leis_who", term) ~ "Physical activity (MET h/wk)",
    grepl("md_qrt", term) ~ "Mammographic breast density quartile",
    grepl("bmi_20", term) ~ "BMI at age 20 (kg/m²)",
    grepl("bf_dur", term) ~ "Breastfeeding duration (mths)"
                              ),
    level = case_when(
      grepl("d_R1menopause_lab3", term) ~ sub("d_R1menopause_lab3", "", term),
      grepl("age_meno", term) ~ sub("d_age_meno_tr_lab", "", term),
      grepl("hrt", term) ~ sub("d_R1hrt_tr_lab", "", term),
      grepl("parity", term) ~ sub("d_parity_lab", "", term),
      grepl("agebirth1", term) ~ sub("d_agebirth1_tr_lab", "", term),
      grepl("age_menarche", term) ~ sub("d_age_menarche_lab", "", term), 
      grepl("ocstatus2", term) ~ sub("d_ocstatus2_lab", "", term),
      grepl("bbd", term) ~ sub("d_bbd_lab", "", term),
      grepl("fambrca", term) ~ sub("d_fambrca_lab", "", term),
      grepl("R1alcohol", term) ~ sub("d_R1alcohol_units_lab", "", term),
      grepl("R1smokingstatus", term) ~ sub("d_R1smokingstatus_lab", "", term),
      grepl("bmi_entry", term) ~ sub("d_bmi_entry_lab", "", term),
      grepl("R1physmet_leis_who", term) ~ sub("d_R1physmet_leis_who_lab", "", term),
      grepl("md_qrt", term) ~ sub("d_md_qrt", "", term),
      grepl("bmi_20", term) ~ sub("d_bmi_20_lab", "", term),
      grepl("bf_dur", term) ~ sub("d_bf_dur_tr_lab", "", term)
    )
    )

# add subheading rows 

subheading_data <- md_forest_df %>%
  distinct(variable) %>%
  mutate(level = variable)

md_forest_df <- bind_rows(subheading_data, md_forest_df) %>%
  arrange(variable, factor(level, levels = unique(level)))



# create long format 
md_forest_long <- md_forest_df %>% 
  filter(!(variable == "BMI at recruitment (kg/m²)*" & level == "<18.5")) |>
  filter(!term %in% c("(Intercept)", "diagage", "d_R1toBC_y", "d_MDtoBC_y"),
         !is.na(level),
         # level != "Mammographic breast density quartile",
         level != "Missing"
         ) %>% 
  # Make variable for ordering
  mutate(
    
    ordering = case_when(
      !variable %in% c("BMI at recruitment (kg/m²)*") ~ row_number(),
      variable == level ~ 1,
      level == "18.5-25" ~ 2,
      level == "25-30" ~ 3,
      level == "30+" ~ 4,
      TRUE ~ 888
    ),
    
       ordering = case_when(
      !variable %in% c("BMI at age 20 (kg/m²)") ~ ordering,
      variable == level ~ 1,
      level == "18.5-25" ~ 2,
      level == "<18.5" ~ 3,
      level == "25+" ~ 4,
      TRUE ~ 888
    )
    
  ) |>
  group_by(variable) |>
  arrange(ordering, .by_group = T) |>
  ungroup() |>
  # Renaming levels
  mutate(
    
    level = case_when(
      grepl(" months", level) ~ gsub(" months", "", level),
      level == "No" & variable == "Benign breast disease" ~ "Never",
      level == "No" & variable == "Oral contraceptive use" ~ "Never ", # Add in space to prevent copied id row
      level == "Never" & variable == "MHT status" ~ "Never   ",
      level == "Never" & variable == "Smoking status" ~ "Never  ", # Add in space to prevent copied id row
      level == "18.5-25" & variable == "BMI at age 20 (kg/m²)" ~ "18.5-25 ", # Add in space to prevent copied id row
      level == "Yes" & variable == "Benign breast disease" ~ "Ever",
      level == "Yes" & variable == "Oral contraceptive use" ~ "Ever",
      level == "1" & variable == "Mammographic breast density quartile" ~ "1 (<11.45%)",
      level == "2" & variable == "Mammographic breast density quartile" ~ "2 (11.45-22.00%)",
      level == "3" & variable == "Mammographic breast density quartile" ~ "3 (22.00-36.70%)",
      level == "4" & variable == "Mammographic breast density quartile" ~ "4 (>36.70%)",
      TRUE ~ level
    )
    
  ) |>
  # Order by variable
  mutate(
    variable = factor(variable,
                      levels = c("Age at menarche", "Oral contraceptive use", "Parity",
                                 "Age at first birth", "Breastfeeding duration (mths)", "Menopausal status",
                                 "Age at menopause", "MHT status", "Benign breast disease",
                                 "Family history of breast cancer", "BMI at age 20 (kg/m²)", "BMI at recruitment (kg/m²)*",
                                 "Physical activity (MET h/wk)", "Alcohol units per week", "Smoking status"))
  ) |>
  arrange(variable, row_number()) |>
  select(level, estimate_mdl1, conf.low_mdl1, conf.high_mdl1, estimate_mdl2, conf.low_mdl2, conf.high_mdl2, p.value_mdl1, p.value_mdl2
         ) %>% 
  pivot_longer(
    cols = c(estimate_mdl1, conf.low_mdl1, conf.high_mdl1, 
             estimate_mdl2, conf.low_mdl2, conf.high_mdl2
             ),
    names_to = c(".value", "mdl"),
    names_sep = "_"
  ) |>
  mutate(across(starts_with("p.value"), function(x){
    case_when(
      is.na(x) ~ NA,
      x < 0.001 ~ "<0.001",
      x > 0.999 ~ ">0.999",
      is.numeric(x) ~ as.character(formatC(x, format = "f", digits = 3)),
      TRUE ~ NA
    )
  })) |>
  mutate(p.value_mdl1 = if_else(is.na(p.value_mdl1) & !is.na(p.value_mdl2), "-", p.value_mdl1)) |>
  # Find reference categories and convert p value to "-"
  group_by(mdl) |>
  mutate(test = c(diff(cumsum(is.na(p.value_mdl1))), NA)) |>
  ungroup() |>
  mutate(summary = test == 1 & is.na(p.value_mdl1)) |>
  select(-test) |>
  mutate(p.value_mdl1 = if_else(!summary & is.na(p.value_mdl1), "-", p.value_mdl1))|>
  mutate(p.value_mdl2 = if_else(!summary & is.na(p.value_mdl2), "-", p.value_mdl2))


```

```{r, fig.width=8, fig.height=15}
#| echo: false

# 3. GENERATE FORESTPLOT

is_summary <- md_forest_long$summary


# develop forest

plot <- md_forest_long |>
  filter(row_number() <= 70) |>
  mutate(rn = case_when(
    mdl == "mdl1" ~ row_number() + 1,
    TRUE ~ row_number() - 1
  )) |>
  arrange(rn) |>
  select(-rn) |>
  group_by(mdl
           ) %>%
  mutate(mean = estimate,
         lower = conf.low,
         upper = conf.high
         ) %>% 
  forestplot(labeltext = c(level, p.value_mdl1, p.value_mdl2),
             xlog = TRUE,
             xlab = "OR",
             boxsize = 0.25,
             zero = 1,
             line.margin = 0.55,
             clip = c(0.25, 4),
             xticks = log(c(0.25, 0.5, 1, 2, 4)),
             ci.vertices = T,
             ci.vertices.height = 0.15,
             # legend = c("Model 2 (adjusted for MD)", "Model 1"),
             # # legend = c("Model 1", "Model 2 (adjusted for MD)"),
             legend = c("Model 1", "Model 2 (Adjusted for MD)"), # NOTE: This is the incorrect labelling, but the right ORDER. Colour is fixed in post
             # title = paste("Final model adjusted vs unadjusted for mamographic breast density quartiles"),
             is.summary = is_summary[1:70],
             
             )   %>% 
  fp_set_style(
              box = c("red2", "grey50") %>%  lapply(function(x) gpar(fill = x, col = "#666666")),
               default = gpar(vertices = T),
               line = "grey40",
               summary = gpar(col = NA), # Set summary color to NA
               align = "lrrrr",
               hrz_lines = gpar(col = NA), # Set line color to NA
               txt_gp = fpTxtGp(label = gpar(cex = 1.2),
                                    ticks = gpar(cex = 1.2),
                                    xlab = gpar(cex = 1.2),
                                legend = gpar(cex = 1.2)
                                )
               ) %>%
  fp_decorate_graph(graph.pos = 2
                    ) %>%
  fp_add_header(level = c("", "Risk Factor                                              "), # Fixing width to be consistent between tables
                p.value_mdl1 = c("P Value", "Model 1"),
                p.value_mdl2 = c("P Value", "Model 2")
                # p.value_mdl3 = c("P Value", "Model 3")
                # diff = c("Absolute difference", "(Model 4 - Model 2)")
  )


plot

CairoPNG(file=paste0("../outputs/figures/main/figure_4a.png"), width = 13, height = 13.5, units = "in", res = 300)
print(plot)
dev.off()

plot <- md_forest_long |>
  filter(row_number() > 70) |>
  mutate(rn = case_when(
    mdl == "mdl1" ~ row_number() + 1,
    mdl == "mdl2" ~ row_number() - 1
    # TRUE ~ row_number() - 2
  )) |>
  arrange(rn) |>
  select(-rn) |>
  group_by(mdl
           ) %>%
  mutate(mean = estimate,
         lower = conf.low,
         upper = conf.high
         ) %>% 
  forestplot(labeltext = c(level, p.value_mdl1, p.value_mdl2),
             xlog = TRUE,
             xlab = "OR",
             boxsize = 0.25,
             zero = 1,
             line.margin = 0.55,
             clip = c(0.25, 4),
             xticks = log(c(0.25, 0.5, 1, 2, 4)),
             ci.vertices = T,
             ci.vertices.height = 0.15,
             # legend = c("Model 2 (adjusted for MD)", "Model 1"),
             # # legend = c("Model 1", "Model 2 (adjusted for MD)"),
             legend = c("Model 1", "Model 2 (Adjusted for MD)"), # NOTE: This is the incorrect labelling, but the right ORDER. Colour is fixed in post
             # title = paste("Final model adjusted vs unadjusted for mamographic breast density quartiles"),
             is.summary = is_summary[71:length(is_summary)],
             )   %>% 
  fp_set_style(
              box = c("red2", "grey50") %>%  lapply(function(x) gpar(fill = x, col = "#666666")),
               default = gpar(vertices = T),
               line = "grey40",
               summary = gpar(col = NA), # Set summary color to NA
               align = "lrrrr",
               hrz_lines = gpar(col = NA), # Set line color to NA
               txt_gp = fpTxtGp(label = gpar(cex = 1.2),
                                    ticks = gpar(cex = 1.2),
                                    xlab = gpar(cex = 1.2),
                                legend = gpar(cex = 1.2)
                                )
               ) %>%
  fp_decorate_graph(graph.pos = 2
                    ) %>%
  fp_add_header(level = c("", "Risk Factor                                              "), # Fixing width to be consistent between tables
                p.value_mdl1 = c("P Value", "Model 1"),
                p.value_mdl2 = c("P Value", "Model 2")
                # diff = c("Absolute difference", "(Model 4 - Model 2)")
  )


plot

CairoPNG(file=paste0("../outputs/figures/main/figure_4b.png"), width = 13, height = 13.5, units = "in", res = 300)
print(plot)
grid.text("*<18.5 category ommitted due to small sample size",
          x = unit(0.005, "npc"),
          y = unit(0.005, "npc"),  
          just = c("left", "bottom"),
          gp = gpar(cex = 1.2))
dev.off()

```

```{r, fig.width=8, fig.height=15}
#| echo: false

# 3. GENERATE FORESTPLOT

is_summary <- md_forest_long$summary


# develop forest

plot <- md_forest_long |>
  filter(row_number() %in% c(9:14, 55:98, 127:136)) |>
  mutate(rn = case_when(
    mdl == "mdl1" ~ row_number() + 1,
    TRUE ~ row_number() - 1
  )) |>
  arrange(rn) |>
  select(-rn) |>
  group_by(mdl
           ) %>%
  mutate(mean = estimate,
         lower = conf.low,
         upper = conf.high
         ) %>% 
  forestplot(labeltext = c(level, p.value_mdl1, p.value_mdl2),
             xlog = TRUE,
             xlab = "OR",
             boxsize = 0.25,
             zero = 1,
             line.margin = 0.55,
             clip = c(0.25, 4),
             xticks = log(c(0.25, 0.5, 1, 2, 4)),
             ci.vertices = T,
             ci.vertices.height = 0.15,
             # legend = c("Model 2 (adjusted for MD)", "Model 1"),
             # # legend = c("Model 1", "Model 2 (adjusted for MD)"),
             legend = c("Model 1", "Model 2 (Adjusted for MD)"), # NOTE: This is the incorrect labelling, but the right ORDER. Colour is fixed in post
             # title = paste("Final model adjusted vs unadjusted for mamographic breast density quartiles"),
             is.summary = is_summary[c(9:14, 55:98, 127:136)],
             
             )   %>% 
  fp_set_style(
              box = c("red2", "grey50") %>%  lapply(function(x) gpar(fill = x, col = "#666666")),
               default = gpar(vertices = T),
               line = "grey40",
               summary = gpar(col = NA), # Set summary color to NA
               align = "lrrrr",
               hrz_lines = gpar(col = NA), # Set line color to NA
               txt_gp = fpTxtGp(label = gpar(cex = 1.2),
                                    ticks = gpar(cex = 1.2),
                                    xlab = gpar(cex = 1.2),
                                legend = gpar(cex = 1.2)
                                )
               ) %>%
  fp_decorate_graph(graph.pos = 2
                    ) %>%
  fp_add_header(level = c("", "Risk Factor                                              "), # Fixing width to be consistent between tables
                p.value_mdl1 = c("P Value", "Model 1"),
                p.value_mdl2 = c("P Value", "Model 2")
                # p.value_mdl3 = c("P Value", "Model 3")
                # diff = c("Absolute difference", "(Model 4 - Model 2)")
  )


plot

svg(file = "../outputs/figures/poster/forest.svg", width = 13, height = 13.5, bg = "transparent")
print(plot)
dev.off()


```

## Supplementary figures and tables

### Supplementary Figure 1 - Mammographic density by mode of detection and tumour characteristics

```{r}
#| echo: false
#| include: false

# Mode of Detection -----------------------------------------------------------------------
# Define annotation positions
annotation_positions <- list(ref_x = 1, 
                             ref_y = 90, 
                             p_x = 2, 
                             p_y = 90)

# Define category labels
category_labels <- c("I" = "Interval", 
                     "SD" = "Screen detected")

# Call the function
dmode_plot <- md_df |>
  mutate(d_dmode = if_else(d_dmode == "I", "Interval", "Screen-Detected")) |>
  create_violin_plot(
  response_var = "d_md",
  explanatory_var = "d_dmode",
  y_label = "Breast density (%)",
  title = "Mode of Detection",
  category_labels = category_labels,
  annotation_pos = annotation_positions
)

# Print the plot
# dmode_plot

# Save plot
ggsave(dmode_plot + 
         theme(
  panel.background = element_rect(fill = "transparent",
                                  colour = NA_character_), # necessary to avoid drawing panel outline
  panel.grid.major = element_blank(), # get rid of major grid
  panel.grid.minor = element_blank(), # get rid of minor grid
  plot.background = element_rect(fill = "transparent",
                                 colour = NA_character_), # necessary to avoid drawing plot outline
  legend.background = element_rect(fill = "transparent"),
  legend.box.background = element_rect(fill = "transparent"),
  legend.key = element_rect(fill = "transparent")
),
       filename = "../outputs/figures/poster/violin_1.svg",
       height = 4, width = 5, bg = "transparent")

# Invesive status -----------------------------------------------------------------------
#df |> tabyl(d_inv_status_lab)
# Define annotation positions
annotation_positions <- list(ref_x = 1, ref_y = 90, p_x = 2, p_y = 90)
# Define category labels - no labels needed 

# Call the function
invst_plot <- md_df |> 
  filter(d_inv_status_lab != "Missing"
         ) |> 
  create_violin_plot(
  response_var = "d_md",
  explanatory_var = "d_inv_status_lab",
  y_label = "Breast density (%)",
  title = "Invasive status",
  annotation_pos = annotation_positions
)
# Print the plot
#invst_plot


# Grade ----------------------------------------------------------------------------------
#df |> tabyl(d_grade_lab)
# Define annotation positions
annotation_positions <- list(ref_x = 1, ref_y = 90, p_x = 3, p_y = 90)
# Define category labels - no labels needed 

# Call the function
grade_plot <- md_df |> 
  filter(d_grade_lab != "Missing"
         ) |> 
  create_violin_plot(
  response_var = "d_md",
  explanatory_var = "d_grade_lab",
  y_label = "Breast density (%)",
  title = "Grade",
  annotation_pos = annotation_positions
)
# Print the plot
#grade_plot


# Morphology -------------------------------------------------------------------------------
#df |> tabyl(d_morph4_lab)
# Define annotation positions
annotation_positions <- list(ref_x = 1, ref_y = 90, p_x = 4, p_y = 90)
# Define category labels - no labels needed 

# Call the function
morph_plot <- md_df |> 
  mutate(d_morph4_lab = if_else(d_morph4_lab == "Mixed: ductal and other", "Mixed", d_morph4_lab)) |>
  filter(d_morph4_lab != "Missing"
         ) |> 
  create_violin_plot(
  response_var = "d_md",
  explanatory_var = "d_morph4_lab",
  y_label = "Breast density (%)",
  title = "Morphology",
  annotation_pos = annotation_positions
)
# Print the plot
#morph_plot


# Save plot
ggsave(morph_plot + 
         theme(
  panel.background = element_rect(fill = "transparent",
                                  colour = NA_character_), # necessary to avoid drawing panel outline
  panel.grid.major = element_blank(), # get rid of major grid
  panel.grid.minor = element_blank(), # get rid of minor grid
  plot.background = element_rect(fill = "transparent",
                                 colour = NA_character_), # necessary to avoid drawing plot outline
  legend.background = element_rect(fill = "transparent"),
  legend.box.background = element_rect(fill = "transparent"),
  legend.key = element_rect(fill = "transparent")
),
       filename = "../outputs/figures/poster/violin_2.svg",
       height = 4, width = 5, bg = "transparent")


# Node status -------------------------------------------------------------------------------
#df |> tabyl(d_pos_nodes_lab)
# Define annotation positions
annotation_positions <- list(ref_x = 1, ref_y = 90, p_x = 2, p_y = 90)
# Define category labels - no labels needed 

# Call the function
nodes_plot <- md_df |> 
  filter(d_pos_nodes_lab != "Missing"
         ) |> 
  create_violin_plot(
  response_var = "d_md",
  explanatory_var = "d_pos_nodes_lab",
  y_label = "Breast density (%)",
  title = "Node status",
  annotation_pos = annotation_positions
)
# Print the plot
#nodes_plot



# Tumour size -------------------------------------------------------------------------------
#df |> tabyl(d_tumour_size_lab)
# Define annotation positions
annotation_positions <- list(ref_x = 1, ref_y = 90, p_x = 3, p_y = 90)
# Define category labels - no labels needed 

# Call the function
tsize_plot <- md_df |> 
  filter(d_tumour_size_lab != "Missing"
         ) |> 
  create_violin_plot(
  response_var = "d_md",
  explanatory_var = "d_tumour_size_lab",
  y_label = "Breast density (%)",
  title = "Tumour size (mm)",
  annotation_pos = annotation_positions
)
# Print the plot
#tsize_plot



# ER status --------------------------------------------------------------------------------
#df |> tabyl(d_er_status_lab)
# Define annotation positions
annotation_positions <- list(ref_x = 1, ref_y = 90, p_x = 2, p_y = 90)
# Define category labels - no labels needed 

# Call the function
er_plot <- md_df |> 
  filter(d_er_status_lab != "Missing"
         ) |> 
  create_violin_plot(
  response_var = "d_md",
  explanatory_var = "d_er_status_lab",
  y_label = "Breast density (%)",
  title = "ER status",
  annotation_pos = annotation_positions
)
# Print the plot
#er_plot


# PR status ------------------------------------------------------------------------------------
#df |> tabyl(d_pr_status_lab)
# Define annotation positions
annotation_positions <- list(ref_x = 1, ref_y = 90, p_x = 2, p_y = 90)
# Define category labels - no labels needed 

# Call the function
pr_plot <- md_df |> 
  filter(d_pr_status_lab != "Missing"
         ) |> 
  create_violin_plot(
  response_var = "d_md",
  explanatory_var = "d_pr_status_lab",
  y_label = "Breast density (%)",
  title = "PR status",
  annotation_pos = annotation_positions
)
# Print the plot
#pr_plot



# HER2 status -------------------------------------------------------------------------------------
#df |> tabyl(d_her2_status_lab)
# Define annotation positions
annotation_positions <- list(ref_x = 1, ref_y = 90, p_x = 2, p_y = 90)
# Define category labels - no labels needed 

# Call the function
her2_plot <- md_df |> 
  filter(d_her2_status_lab != "Missing"
         ) |> 
  create_violin_plot(
  response_var = "d_md",
  explanatory_var = "d_her2_status_lab",
  y_label = "Breast density (%)",
  title = "HER2 status",
  annotation_pos = annotation_positions
)
# Print the plot
#her2_plot

```

```{r, fig.width=10, fig.height=20}
#| echo: false

# tcomposite_plot <- invst_plot + grade_plot + morph_plot + nodes_plot + tsize_plot + er_plot + pr_plot + her2_plot + plot_layout(ncol = 2)
# 
# tcomposite_plot

plot <- plot_grid(plotlist = list(dmode_plot, invst_plot, grade_plot, morph_plot, nodes_plot, tsize_plot, er_plot, pr_plot, her2_plot), ncol=3)

plot

ggsave(plot, filename = "../outputs/figures/supplement/supplementary_figure_1.png", height = 12, width = 15)

```

### Supplementary Figure 2 - Mammographic density by risk factors

```{r}
#| echo: false
#| include: false


# Age at menarche ------------------------------------------------------------
#df |> tabyl(d_age_menarche_lab)
# Define annotation positions
annotation_positions <- list(ref_x = 1, ref_y = 90, p_x = 3, p_y = 90)
# Define category labels - no labels needed 

# Call the function
agemenarche_plot <- md_df |> 
  filter(d_age_menarche_lab != "Missing"
         ) |> 
  create_violin_plot(
  response_var = "d_md",
  explanatory_var = "d_age_menarche_lab",
  y_label = "Breast density (%)",
  title = "Age at menarche",
  annotation_pos = annotation_positions
)
# Print the plot
#agemenarche_plot


# Save plot
ggsave(agemenarche_plot + 
         theme(
  aspect.ratio = 1,
  panel.background = element_rect(fill = "transparent",
                                  colour = NA_character_), # necessary to avoid drawing panel outline
  panel.grid.major = element_blank(), # get rid of major grid
  panel.grid.minor = element_blank(), # get rid of minor grid
  plot.background = element_rect(fill = "transparent",
                                 colour = NA_character_), # necessary to avoid drawing plot outline
  legend.background = element_rect(fill = "transparent"),
  legend.box.background = element_rect(fill = "transparent"),
  legend.key = element_rect(fill = "transparent")
),
       filename = "../outputs/figures/poster/violin_3.svg",
       height = 4, width = 5, bg = "transparent")


# Oral contraceptives --------------------------------------------------------
#df |> tabyl(d_ocstatus_lab)
# Define annotation positions
annotation_positions <- list(ref_x = 1, ref_y = 90, p_x = 3, p_y = 90)
# Define category labels - no labels needed 

# Call the function
oc_plot <- md_df |> 
  filter(d_ocstatus_lab != "Missing"
         ) |> 
  create_violin_plot(
  response_var = "d_md",
  explanatory_var = "d_ocstatus_lab",
  y_label = "Breast density (%)",
  title = "Oral contraceptive use",
  annotation_pos = annotation_positions
)
# Print the plot
#oc_plot



# Parity --------------------------------------------------------
#df |> tabyl(d_parity_lab)
# Define annotation positions
annotation_positions <- list(ref_x = 1, ref_y = 90, p_x = 4, p_y = 90)
# Define category labels - no labels needed 

# Call the function
parity_plot <- md_df |> 
  filter(d_parity_lab != "Missing"
         ) |> 
  create_violin_plot(
  response_var = "d_md",
  explanatory_var = "d_parity_lab",
  y_label = "Breast density (%)",
  title = "Parity",
  annotation_pos = annotation_positions
)
# Print the plot
#parity_plot


# Save plot
ggsave(parity_plot + 
         theme(
  aspect.ratio = 1,
  panel.background = element_rect(fill = "transparent",
                                  colour = NA_character_), # necessary to avoid drawing panel outline
  panel.grid.major = element_blank(), # get rid of major grid
  panel.grid.minor = element_blank(), # get rid of minor grid
  plot.background = element_rect(fill = "transparent",
                                 colour = NA_character_), # necessary to avoid drawing plot outline
  legend.background = element_rect(fill = "transparent"),
  legend.box.background = element_rect(fill = "transparent"),
  legend.key = element_rect(fill = "transparent")
),
       filename = "../outputs/figures/poster/violin_4.svg",
       height = 4, width = 5, bg = "transparent")


# Age at first birth -------------------------------------------------
#df |> tabyl(d_agebirth1_lab)
# Define annotation positions
annotation_positions <- list(ref_x = 1, ref_y = 90, p_x = 5, p_y = 90)
# Define category labels - no labels needed 

# Call the function
agebirth1_plot <- md_df |> 
  filter(d_agebirth1_lab != "Missing"
         ) |> 
  create_violin_plot(
  response_var = "d_md",
  explanatory_var = "d_agebirth1_lab",
  y_label = "Breast density (%)",
  title = "Age at first birth",
  annotation_pos = annotation_positions
)
# Print the plot
#agebirth1_plot



# Duration of breastfeeding -----------------------------------------------------
#df |> tabyl(d_bf_dur_lab)
# Define annotation positions
annotation_positions <- list(ref_x = 1, ref_y = 90, p_x = 5, p_y = 90)
# Define category labels - no labels needed 

# Call the function
bf_plot <- md_df |> 
  mutate(d_bf_dur_lab = case_when(
    d_bf_dur_lab == "Never breastfed" ~ "Never",
    TRUE ~ gsub(" months", "", d_bf_dur_lab)
  )) |>
  mutate(
    d_bf_dur_lab = factor(d_bf_dur_lab, levels = c("Never", "<6", "6-12", "12-24", "24+"))
  ) |>
  filter(d_bf_dur_lab != "Missing"
         ) |> 
  create_violin_plot(
  response_var = "d_md",
  explanatory_var = "d_bf_dur_lab",
  y_label = "Breast density (%)",
  title = "Breastfeeding duration (mths)",
  annotation_pos = annotation_positions
)

# bf_plot <- bf_plot +
#   theme(axis.text.x = element_text(angle = 20, hjust = 1))
# Print the plot
#bf_plot


# Save plot

ggsave(bf_plot + 
         theme(
  aspect.ratio = 1,
  panel.background = element_rect(fill = "transparent",
                                  colour = NA_character_), # necessary to avoid drawing panel outline
  panel.grid.major = element_blank(), # get rid of major grid
  panel.grid.minor = element_blank(), # get rid of minor grid
  plot.background = element_rect(fill = "transparent",
                                 colour = NA_character_), # necessary to avoid drawing plot outline
  legend.background = element_rect(fill = "transparent"),
  legend.box.background = element_rect(fill = "transparent"),
  legend.key = element_rect(fill = "transparent")
),
       filename = "../outputs/figures/poster/violin_5.svg",
       height = 4, width = 5, bg = "transparent")


# Menopausal status ------------------------------------------------------------------------
#df |> tabyl(d_R1menopause_lab3)
# Define annotation positions
annotation_positions <- list(ref_x = 1, ref_y = 90, p_x = 2, p_y = 90)
# Define category labels - no labels needed 

# Call the function
menost_plot <- create_violin_plot(
  data = md_df,
  response_var = "d_md",
  explanatory_var = "d_R1menopause_lab3",
  y_label = "Breast density (%)",
  title = "Menopausal status",
  annotation_pos = annotation_positions
)
# Print the plot
menost_plot



# Age at Menopause ------------------------------------------------------------------------
#df |> tabyl(d_age_menopause_lab2)
# Define annotation positions
annotation_positions <- list(ref_x = 1, ref_y = 90, p_x = 3, p_y = 90)
# Define category labels - no labels needed 

# Call the function
menoage_plot <- md_df |> 
  filter(!d_age_menopause_lab2 %in% c("Missing", "Pre-menopausal")
         ) |> 
  mutate(d_age_menopause_lab2 = fct_relevel(d_age_menopause_lab2, "<50", "50-54", "55+")
         ) |> 
  create_violin_plot(
  response_var = "d_md",
  explanatory_var = "d_age_menopause_lab2",
  y_label = "Breast density (%)",
  title = "Age at menopause",
  annotation_pos = annotation_positions
)

# Print the plot
#menoage_plot





# HRT status (simpler HRT variable) --------------------------------------------------------------------------------
# Define annotation positions
annotation_positions <- list(ref_x = 1, ref_y = 90, p_x = 3, p_y = 90)
# Define category labels - no labels needed 

# Call the function
hrtst_plot <- md_df |> 
  filter(d_R1hrtstatus_lab != "Missing"
         ) |> 
  create_violin_plot(
  response_var = "d_md",
  explanatory_var = "d_R1hrtstatus_lab",
  y_label = "Breast density (%)",
  title = "MHT status",
  annotation_pos = annotation_positions
)



# Save plot
ggsave(hrtst_plot + 
         theme(
  aspect.ratio = 1,
  panel.background = element_rect(fill = "transparent",
                                  colour = NA_character_), # necessary to avoid drawing panel outline
  panel.grid.major = element_blank(), # get rid of major grid
  panel.grid.minor = element_blank(), # get rid of minor grid
  plot.background = element_rect(fill = "transparent",
                                 colour = NA_character_), # necessary to avoid drawing plot outline
  legend.background = element_rect(fill = "transparent"),
  legend.box.background = element_rect(fill = "transparent"),
  legend.key = element_rect(fill = "transparent")
),
       filename = "../outputs/figures/poster/violin_6.svg",
       height = 4, width = 5, bg = "transparent")

# Benign breast disease ---------------------------------------------------------
#df |> tabyl(d_bbd_lab)
# Define annotation positions
annotation_positions <- list(ref_x = 1, ref_y = 90, p_x = 2, p_y = 90)
# Define category labels - no labels needed 

# Call the function
bbd_plot <- md_df |> 
  mutate(d_bbd_lab = case_when(
    d_bbd_lab == "Yes" ~ "Ever",
    d_bbd_lab == "No" ~ "Never",
    TRUE ~ d_bbd_lab
  )) |>
  filter(d_bbd_lab != "Missing"
         ) |> 
  create_violin_plot(
  response_var = "d_md",
  explanatory_var = "d_bbd_lab",
  y_label = "Breast density (%)",
  title = "Benign breast disease",
  annotation_pos = annotation_positions
)
# Print the plot
#bbd_plot

# Save plot
ggsave(bbd_plot + 
         theme(
  aspect.ratio = 1,
  panel.background = element_rect(fill = "transparent",
                                  colour = NA_character_), # necessary to avoid drawing panel outline
  panel.grid.major = element_blank(), # get rid of major grid
  panel.grid.minor = element_blank(), # get rid of minor grid
  plot.background = element_rect(fill = "transparent",
                                 colour = NA_character_), # necessary to avoid drawing plot outline
  legend.background = element_rect(fill = "transparent"),
  legend.box.background = element_rect(fill = "transparent"),
  legend.key = element_rect(fill = "transparent")
),
       filename = "../outputs/figures/poster/violin_9.svg",
       height = 4, width = 5, bg = "transparent")


# Family history of breast cancer -----------------------------------------------
#df |> tabyl(d_fambrca_lab)
# Define annotation positions
annotation_positions <- list(ref_x = 1, ref_y = 90, p_x = 2, p_y = 90)
# Define category labels - no labels needed 

# Call the function
fambrca_plot <- md_df |> 
  filter(d_fambrca_lab != "Missing"
         ) |> 
  create_violin_plot(
  response_var = "d_md",
  explanatory_var = "d_fambrca_lab",
  y_label = "Breast density (%)",
  title = "Family history of breast cancer",
  annotation_pos = annotation_positions
)
# Print the plot
#fambrca_plot



# BMI at age 20 (kg/m²) ------------------------------------------------------------------
#df |> tabyl(d_bmi_20_lab)
# Define annotation positions
annotation_positions <- list(ref_x = 2, ref_y = 90, p_x = 4, p_y = 90)
# Define category labels - no labels needed 

# Call the function
bmi20_plot <- md_df |>
  mutate(d_bmi_20_lab = factor(d_bmi_20_lab, levels = c("<18.5", "18.5-25", "25+", "Missing"))) |>
  filter(d_bmi_20_lab != "Missing"
         ) |> 
  create_violin_plot(
  response_var = "d_md",
  explanatory_var = "d_bmi_20_lab",
  y_label = "Breast density (%)",
  title = "BMI at age 20 (kg/m²)",
  annotation_pos = annotation_positions
)
# Print the plot
# bmi20_plot

# Save plot
ggsave(bmi20_plot + 
         theme(
           aspect.ratio = 1,
  panel.background = element_rect(fill = "transparent",
                                  colour = NA_character_), # necessary to avoid drawing panel outline
  panel.grid.major = element_blank(), # get rid of major grid
  panel.grid.minor = element_blank(), # get rid of minor grid
  plot.background = element_rect(fill = "transparent",
                                 colour = NA_character_), # necessary to avoid drawing plot outline
  legend.background = element_rect(fill = "transparent"),
  legend.box.background = element_rect(fill = "transparent"),
  legend.key = element_rect(fill = "transparent")
),
       filename = "../outputs/figures/poster/violin_10.svg",
       height = 4, width = 5, bg = "transparent")

# BMI -----------------------------------------------------------------------------
#df |> tabyl(d_bmi_entry_lab)
# Define annotation positions
annotation_positions <- list(ref_x = 2, ref_y = 90, p_x = 4, p_y = 90)
# Define category labels - no labels needed 

# Call the function
bmi_plot <- md_df |>
  mutate(d_bmi_entry_lab = factor(d_bmi_entry_lab, levels = c("<18.5", "18.5-25", "25-30", "30+", "Missing"))) |>
  filter(d_bmi_entry_lab != "Missing"
         ) |> 
  create_violin_plot(
  response_var = "d_md",
  explanatory_var = "d_bmi_entry_lab",
  y_label = "Breast density (%)",
  title = "BMI at recruitment (kg/m²)",
  annotation_pos = annotation_positions
)
# Print the plot
#bmi_plot



# Save plot
ggsave(bmi_plot + 
         theme(
  aspect.ratio = 1,
  panel.background = element_rect(fill = "transparent",
                                  colour = NA_character_), # necessary to avoid drawing panel outline
  panel.grid.major = element_blank(), # get rid of major grid
  panel.grid.minor = element_blank(), # get rid of minor grid
  plot.background = element_rect(fill = "transparent",
                                 colour = NA_character_), # necessary to avoid drawing plot outline
  legend.background = element_rect(fill = "transparent"),
  legend.box.background = element_rect(fill = "transparent"),
  legend.key = element_rect(fill = "transparent")
),
       filename = "../outputs/figures/poster/violin_8.svg",
       height = 4, width = 5, bg = "transparent")



# Physical activity ---------------------------------------------------------------
#df |> tabyl(d_R1physmet_leis_who_lab)
# Define annotation positions
annotation_positions <- list(ref_x = 1, ref_y = 90, p_x = 3, p_y = 90)
# Define category labels - no labels needed 

# Call the function
pa_plot <- md_df |> 
  filter(d_R1physmet_leis_who_lab != "Missing"
         ) |> 
  create_violin_plot(
  response_var = "d_md",
  explanatory_var = "d_R1physmet_leis_who_lab",
  y_label = "Breast density (%)",
  title = "Physical activity (MET h/wk)",
  annotation_pos = annotation_positions
)
# Print the plot
#pa_plot




# Alcohol ------------------------------------------------------------------------
#df |> tabyl(d_R1alcohol_units_lab)
# Define annotation positions
annotation_positions <- list(ref_x = 1, ref_y = 90, p_x = 5, p_y = 90)
# Define category labels - no labels needed 

# Call the function
alc_plot <- md_df |> 
  filter(d_R1alcohol_units_lab != "Missing"
         ) |> 
  create_violin_plot(
  response_var = "d_md",
  explanatory_var = "d_R1alcohol_units_lab",
  y_label = "Breast density (%)",
  title = "Alcohol units per week",
  annotation_pos = annotation_positions
)
# Print the plot
#alc_plot


# Save plot
# Save plot
ggsave(alc_plot + 
         theme(
  aspect.ratio = 1,
  panel.background = element_rect(fill = "transparent",
                                  colour = NA_character_), # necessary to avoid drawing panel outline
  panel.grid.major = element_blank(), # get rid of major grid
  panel.grid.minor = element_blank(), # get rid of minor grid
  plot.background = element_rect(fill = "transparent",
                                 colour = NA_character_), # necessary to avoid drawing plot outline
  legend.background = element_rect(fill = "transparent"),
  legend.box.background = element_rect(fill = "transparent"),
  legend.key = element_rect(fill = "transparent")
),
       filename = "../outputs/figures/poster/violin_7.svg",
       height = 4, width = 5, bg = "transparent")

# Smoking -----------------------------------------------------------------------
#df |> tabyl(d_R1smokingstatus_lab)
# Define annotation positions
annotation_positions <- list(ref_x = 1, ref_y = 90, p_x = 3, p_y = 90)
# Define category labels - no labels needed 

# Call the function
smok_plot <- md_df |> 
  filter(d_R1smokingstatus_lab != "Missing"
         ) |> 
  create_violin_plot(
  response_var = "d_md",
  explanatory_var = "d_R1smokingstatus_lab",
  y_label = "Breast density (%)",
  title = "Smoking status",
  annotation_pos = annotation_positions
)
# Print the plot
smok_plot


```


```{r, fig.width=15, fig.height=20}
#| echo: false

# All plots combined
plot <- plot_grid(plotlist = list(agemenarche_plot, oc_plot, parity_plot, agebirth1_plot, bf_plot, menost_plot, menoage_plot, hrtst_plot, bbd_plot, fambrca_plot, bmi20_plot, bmi_plot, pa_plot, alc_plot, smok_plot),
          nrow = 5)

plot

ggsave(plot, filename = "../outputs/figures/supplement/supplementary_figure_2.png", height = 20, width = 15)

```

### Supplementary Table 1

```{r}

#| echo: false

# Read in data
df_mod <- readRDS("Q:/SHARED/USERS/MBrayley/Screening/data/dm_df_include_unknown.rds")

# Read in casummary data
df_casummary <- jsonlite::fromJSON("//rds/data/DGE/DUDGE/Shared/GENERATIONS/_JSON_Data/v20260123/CancerSummary.json") 

df_casummary_clean <- df_casummary |>
  filter(substr(ICD_CODE, 1, 3) == "C50" | ICD_CODE %in% c("D051", "D059")) |>
  mutate(DIAGNOSIS_DATE = as.Date(DIAGNOSIS_DATE_SHIFTED, format = "%Y-%m-%d")) |>
  arrange(DIAGNOSIS_DATE) |>
  group_by(TCode) |>
  filter(row_number() == 1) |>
  ungroup() |>
  select(TCode, DIAGNOSIS_DATE, SCREENINGSTATUSCOSD_CODE)

# use row percentages and rename reg_sd levels to align with dmode

dmode_crosstab_tb <- df_mod |>
  merge(df_casummary_clean, by.x = "tcode", by.y = "TCode", all.x = T) |>
  filter(source_dm2 != "reg_sd") |>
  mutate(
    
    reg_sd = if_else(is.na(SCREENINGSTATUSCOSD_CODE), "NA", SCREENINGSTATUSCOSD_CODE),
    reg_sd = if_else(reg_sd %in% c("1", "2", "NA"), reg_sd, "3"),
    
    reg_sd = factor(reg_sd,
                    levels = c("2","1", "3", "NA"),
                    labels = c("Interval", "Screen-Detected", "Other", "No Data")),
    
    d_dmode = case_when(
      ancat_dmode_v2 == "I" ~ "Interval",
      ancat_dmode_v2 == "SD" ~ "Screen-Detected",
      TRUE ~ "Other"
    ),
    
    d_dmode = factor(d_dmode, levels = c("Interval", "Screen-Detected", "Other"))
    
  ) |> 
  tbl_cross(
    row = reg_sd,
    col = d_dmode,
    percent = "cell"
  ) |> 
  bold_labels()

#dmode_crosstab_tb |> show_header_names()


dmode_crosstab_tb <- dmode_crosstab_tb |> 
  modify_header(
    label ~ "**Registry Detection Mode**"
  ) |> 
  modify_spanning_header(all_stat_cols() ~ "**Derived Detection Mode**" # Rename column variable
  ) 


dmode_crosstab_tb <- dmode_crosstab_tb |> 
  modify_table_body(
    ~ .x |> 
      filter(label != "reg_sd") |>  # Remove rows where the label matches the old variable name
      mutate(stat_1 = if_else(row_number() == n(), stat_1, gsub("\\(.*", "", stat_1)),
             stat_2 = if_else(row_number() == n(), stat_2, gsub("\\(.*", "", stat_2)),
             stat_3 = if_else(row_number() == n(), stat_3, gsub("\\(.*", "", stat_3)),
             )
  )

dmode_crosstab_tb

dmode_crosstab_tb |>
  as_flex_table() |>
  flextable::save_as_docx(path = "../outputs/figures/supplement/supplementary_table_1.docx")

```

### Supplementary Table 2

```{r}
#| echo: false
#| include: false
#| warning: false


# rf vars for violin plots
explanatory_vars <- c(
                  "d_dmode",
                  #     "d_age_menarche_lab",
                  # "d_ocstatus2_lab",
                  # "d_parity_lab",
                  # "d_agebirth1_lab",
                  # "d_bf_dur_lab",
                  # "d_R1menopause_lab3",
                  # "d_age_menopause_lab", 
                  # "d_R1hrtstatus_lab",
                  # "d_bbd_lab",
                  # "d_fambrca_lab",
                  # "d_bmi_20_lab",
                  # "d_bmi_entry_lab",
                  # "d_R1physmet_leis_who_lab",
                  # "d_R1alcohol_units_lab",
                  # "d_R1smokingstatus_lab",
                        
                 "d_inv_status", 
                 "d_grade_lab", 
                 "d_morph4_lab", 
                 "d_pos_nodes_lab", 
                 "d_tumour_size_lab", 
                 "d_er_status_lab", 
                 "d_pr_status_lab", 
                 "d_her2_status_lab"
                        
                      )


# Summary
table_summary <- df_compl_case |>
  mutate(d_bmi_entry_lab = fct_relevel(d_bmi_entry_lab, "<18.5", "18.5-25", "25-30", "30+"),
        d_bmi_20_lab = fct_relevel(d_bmi_20_lab, "<18.5", "18.5-25", "25-30", "30+")
              ) |>
  filter(d_md_avail == 1) |>
   mutate(d_bf_dur_lab = if_else(d_bf_dur_lab == "Missing", NA, d_bf_dur_lab),
         d_age_menopause_lab = if_else(d_age_menopause_lab == "Pre-menopausal", NA, d_age_menopause_lab),
         d_R1hrtstatus_lab = if_else(d_R1hrtstatus_lab == "Pre-menopausal", NA, d_R1hrtstatus_lab),
         d_agebirth1_lab = if_else(d_agebirth1_lab == "Non-parous", NA, d_agebirth1_lab),
         d_dmode = factor(d_dmode, levels = c("SD", "I"))
         ) |> 
  select(any_of(explanatory_vars)
         ) |> 
    tbl_summary(
              statistic = list(
                all_continuous() ~ "{mean} ({sd})",
                all_categorical() ~ "{n}"
                ),
              digits = all_continuous() ~ 2,
              # type = list(d_ocstatus2_lab ~ "categorical",
              #             d_fambrca_lab ~ "categorical",
              #             d_bbd_lab ~ "categorical"),
              missing = "no"
    ) |>
    modify_table_body(
      ~ filter(.x, stat_0 != 0 | is.na(stat_0))
    )

# Summary of mammographic
table_summary_md <- df_compl_case |>
  mutate(d_bmi_entry_lab = fct_relevel(d_bmi_entry_lab, "<18.5", "18.5-25", "25-30", "30+"),
        d_bmi_20_lab = fct_relevel(d_bmi_20_lab, "<18.5", "18.5-25", "25-30", "30+") 
              ) |>
  filter(d_md_avail == 1) |>
   mutate(d_bf_dur_lab = if_else(d_bf_dur_lab == "Missing", NA, d_bf_dur_lab),
         d_age_menopause_lab = if_else(d_age_menopause_lab == "Pre-menopausal", NA, d_age_menopause_lab),
         d_R1hrtstatus_lab = if_else(d_R1hrtstatus_lab == "Pre-menopausal", NA, d_R1hrtstatus_lab),
         d_agebirth1_lab = if_else(d_agebirth1_lab == "Non-parous", NA, d_agebirth1_lab),
         d_dmode = factor(d_dmode, levels = c("SD", "I"))
         ) |> 
  select(any_of(explanatory_vars), d_md
         ) |> 
    tbl_custom_summary(
              statistic = list(
                everything() ~ "{mean} ({sd})"
                ),
              stat_fns  = list(
                everything() ~ function(data, full_data, variable, by, type, ...){
                  dplyr::summarize(data, mean = mean(d_md),
                           sd = sd(d_md))
                }
              ),
              digits = all_continuous() ~ 2,
              # type = list(d_ocstatus2_lab ~ "categorical",
              #             d_fambrca_lab ~ "categorical",
              #             d_bbd_lab ~ "categorical"),
              missing = "no"
    ) |>
    modify_table_body(
      ~ filter(.x, stat_0 != 0 | is.na(stat_0))
    )


# Unadjusted:
table_unadjusted <- df_compl_case |> 
  filter(d_md_avail == 1
  ) |> 
  mutate(d_bf_dur_lab = if_else(d_bf_dur_lab == "Missing", NA, d_bf_dur_lab),
         d_age_menopause_lab = if_else(d_age_menopause_lab == "Pre-menopausal", NA, d_age_menopause_lab),
         d_R1hrtstatus_lab = if_else(d_R1hrtstatus_lab == "Pre-menopausal", NA, d_R1hrtstatus_lab),
         d_agebirth1_lab = if_else(d_agebirth1_lab == "Non-parous", NA, d_agebirth1_lab),
         d_dmode = factor(d_dmode, levels = c("SD", "I"))
         ) |> 
  select(d_md, any_of(explanatory_vars)
         ) |> 
  tbl_uvregression(
    method = lm,
    y = d_md,
    pvalue_fun = custom_pvalue_fun,
    estimate_fun = function(x){
      case_when(
        is.na(x) ~ NA,
        TRUE ~ formatC(x, format = "f", digits = 2)
      )
    },
    # pvalue_fun = ~ .x,
    # estimate_fun = ~ .x
    method.args = list(na.action = na.omit),
    hide_n = T
   ) |> 
  # modify_table_body(~ .x |>
  #   mutate(across(c(estimate, conf.low, conf.high), exp))
  # ) |>
  # modify_header(
  #   estimate = "**exp(Beta)**"
  # ) |>
  # add_global_p() |>  # add global p-value
  bold_labels() 


#table_unadjusted


# rf vars for violin plots
explanatory_vars <- c(
                  "d_dmode",
                  #     "d_age_menarche_lab",
                  # "d_ocstatus2_lab",
                  # "d_parity_lab",
                  # "d_agebirth1_lab",
                  # "d_bf_dur_lab",
                  # "d_R1menopause_lab3",
                  # "d_age_menopause_lab", 
                  # "d_R1hrtstatus_lab",
                  # "d_bbd_lab",
                  # "d_fambrca_lab",
                  # "d_bmi_20_lab",
                  # "d_bmi_entry_lab",
                  # "d_R1physmet_leis_who_lab",
                  # "d_R1alcohol_units_lab",
                  # "d_R1smokingstatus_lab",
                        
                 "d_inv_status", 
                 "d_grade_lab", 
                 "d_morph4_lab", 
                 "d_pos_nodes_lab", 
                 "d_tumour_size_lab", 
                 "d_er_status_lab", 
                 "d_pr_status_lab", 
                 "d_her2_status_lab"
                        
                      )


# Adjusted table
# List to store regression tables
table_list <- list()

# data manipulation to remove redundant levels 

df_temp <- df_compl_case |> 
  mutate(d_bf_dur_lab = if_else(d_bf_dur_lab == "Missing", NA, d_bf_dur_lab),
         d_age_menopause_lab = if_else(d_age_menopause_lab == "Pre-menopausal", NA, d_age_menopause_lab),
         d_R1hrtstatus_lab = if_else(d_R1hrtstatus_lab == "Pre-menopausal", NA, d_R1hrtstatus_lab),
         d_agebirth1_lab = if_else(d_agebirth1_lab == "Non-parous", NA, d_agebirth1_lab),
         d_age_mammo = AgeatEntry + d_R1toMD_y,
         d_dmode = factor(d_dmode, levels = c("SD", "I"))
         )


# Loop through each explanatory variable and create a regression table
for (var in explanatory_vars) {
  # Run the regression model adjusted for age
  model <- lm(as.formula(paste("d_md ~", var, "+ d_age_mammo")), data = df_temp, na.action = na.omit)

  # Create the regression table and store it in the list
  table_list[[var]] <- tbl_regression(
    model,
    pvalue_fun = custom_pvalue_fun,
    estimate_fun = function(x){
      case_when(
        is.na(x) ~ NA,
        TRUE ~ formatC(x, format = "f", digits = 2)
      )
    }
    # pvalue_fun = ~ .x,
    # estimate_fun = ~ .x
  ) %>%
  # modify_table_body(~ .x |>
  #   mutate(across(c(estimate, conf.low, conf.high), exp))
  # ) |>
  # modify_header(
  #   estimate = "**exp(Beta)**"
  # ) |>
    # add_global_p(
    # ) |> 
    #add_n() |> # checked with unadjusted and n is the same so no need here
    remove_row_type(variables = "d_age_mammo", type = "all")  # Remove the age covariate from the table
}

# Stack all the regression tables into one
table_adjusted <- tbl_stack(table_list)

# View the final stacked table
table_adjusted

# Merge the tables

md_desc_table <- tbl_merge(list(table_summary, table_summary_md, table_unadjusted, table_adjusted),
                           tab_spanner = c(" ", "**Density**", "**Unadjusted**", "**Adjusted for age at mammogram**"
                                           )
                         )
                         
# Changing variable names to make sense
tab_labels <- c(
  "d_dmode" = "Mode of detection",
  # "d_age_menarche_lab" = "Age at menarche",
  # "d_ocstatus2_lab" = "Oral contraception status",
  # "d_parity_lab" = "Parity",
  # "d_agebirth1_lab" = "Age at first birth",
  # "d_bf_dur_lab" = "Breastfeeding duration (mths)",
  # "d_R1menopause_lab3" = "Menopausal status",
  # "d_age_menopause_lab" = "Age at menopause",
  # "d_R1hrtstatus_lab" = "MHT status",
  # "d_bbd_lab" = "Benign breast disease",
  # "d_fambrca_lab" = "Family history of breast cancer",
  # "d_bmi_20_lab" = "BMI at age 20 (kg/m²)",
  # "d_bmi_entry_lab" = "BMI at recruitment (kg/m²)",
  # "d_R1physmet_leis_who_lab" = "Physical activity (MET h/wk)",
  # "d_R1alcohol_units_lab" = "Alcohol units per week",
  # "d_R1smokingstatus_lab" = "Smoking status",
  "d_inv_status" = "Invasive status",
  "d_grade_lab" = "Grade",
  "d_morph4_lab" = "Morphology",
  "d_pos_nodes_lab" = "Node status",
  "d_tumour_size_lab" = "Tumour size (mm)",
  "d_er_status_lab" = "ER status",
  "d_pr_status_lab" = "PR status",
  "d_her2_status_lab" = "HER2 status"
)

for(i in 1:length(md_desc_table$table_body$label)){
  
  if(md_desc_table$table_body$label[[i]] %in% names(tab_labels)){
    
    md_desc_table$table_body$label[[i]] <- unname(tab_labels[which(names(tab_labels) == md_desc_table$table_body$label[[i]])])
    
  }
  
}

# Filter unneeded rows out
md_desc_table <- md_desc_table |>
  modify_table_body(
    ~ .x |>
      filter(stat_0_1 != "0" | is.na(stat_0_1)) |>
      filter(stat_0_2 != "NA (NA)" | is.na(stat_0_2)) |>
      filter(variable != "d_md")
  )

# Manually changing the labels of levels in the table
md_desc_table$table_body$label[md_desc_table$table_body$label == "Mixed: ductal and other"] <- "Mixed"

md_desc_table$table_body$label[grepl(md_desc_table$table_body$label, pattern = "months")] <- gsub(" months", "", md_desc_table$table_body$label[grepl(md_desc_table$table_body$label, pattern = "months")])

md_desc_table$table_body$label[md_desc_table$table_body$label == "I"] <- "Interval"
md_desc_table$table_body$label[md_desc_table$table_body$label == "SD"] <- "Screen-Detected"

md_desc_table$table_body$label[(md_desc_table$table_body$term_3) == "d_bbd_labYes"] <- "Ever"
md_desc_table$table_body$label[(md_desc_table$table_body$term_3) == "d_bbd_labNo"] <- "Never"


md_desc_table$table_body$label[(md_desc_table$table_body$term_3) == "d_ocstatus2_labNo"] <- "Never"
md_desc_table$table_body$label[(md_desc_table$table_body$term_3) == "d_ocstatus2_labYes"] <- "Ever"

md_desc_table$table_body$label[(md_desc_table$table_body$term_3) == "d_inv_status1"] <- "Invasive"
md_desc_table$table_body$label[(md_desc_table$table_body$term_3) == "d_inv_status0"] <- "DCIS"





md_desc_table

wb <- wb_workbook()$add_worksheet("Supplementary_Table_2")
wb <- wb_add_flextable(wb, "Supplementary_Table_2", as_flex_table(md_desc_table))

```

### Supplementary Table 3

```{r}
#| echo: false
#| include: false
#| warning: false


# rf vars for violin plots
explanatory_vars <- c(
                  # "d_dmode",
                      "d_age_menarche_lab",
                  "d_ocstatus2_lab",
                  "d_parity_lab",
                  "d_agebirth1_lab",
                  "d_bf_dur_lab",
                  "d_R1menopause_lab3",
                  "d_age_menopause_lab", 
                  "d_R1hrtstatus_lab",
                  "d_bbd_lab",
                  "d_fambrca_lab",
                  "d_bmi_20_lab",
                  "d_bmi_entry_lab",
                  "d_R1physmet_leis_who_lab",
                  "d_R1alcohol_units_lab",
                  "d_R1smokingstatus_lab"
                        
                 # "d_inv_status", 
                 # "d_grade_lab", 
                 # "d_morph4_lab", 
                 # "d_pos_nodes_lab", 
                 # "d_tumour_size_lab", 
                 # "d_er_status_lab", 
                 # "d_pr_status_lab", 
                 # "d_her2_status_lab"
                        
                      )


# Summary
table_summary <- df_compl_case |>
  mutate(d_bmi_entry_lab = fct_relevel(d_bmi_entry_lab, "<18.5", "18.5-25", "25-30", "30+"),
        d_bmi_20_lab = fct_relevel(d_bmi_20_lab, "<18.5", "18.5-25", "25-30", "30+") 
              ) |>
  filter(d_md_avail == 1) |>
   mutate(d_bf_dur_lab = if_else(d_bf_dur_lab == "Missing", NA, d_bf_dur_lab),
         d_age_menopause_lab = if_else(d_age_menopause_lab == "Pre-menopausal", NA, d_age_menopause_lab),
         d_R1hrtstatus_lab = if_else(d_R1hrtstatus_lab == "Pre-menopausal", NA, d_R1hrtstatus_lab),
         d_agebirth1_lab = if_else(d_agebirth1_lab == "Non-parous", NA, d_agebirth1_lab),
         d_dmode = factor(d_dmode, levels = c("SD", "I"))
         ) |> 
  select(any_of(explanatory_vars)
         ) |> 
    tbl_summary(
              statistic = list(
                all_continuous() ~ "{mean} ({sd})",
                all_categorical() ~ "{n}"
                ),
              digits = all_continuous() ~ 2,,
              type = list(d_ocstatus2_lab ~ "categorical",
                          d_fambrca_lab ~ "categorical",
                          d_bbd_lab ~ "categorical"),
              missing = "no"
    ) |>
    modify_table_body(
      ~ filter(.x, stat_0 != 0 | is.na(stat_0))
    )

# Summary of mammographic
table_summary_md <- df_compl_case |>
  mutate(d_bmi_entry_lab = fct_relevel(d_bmi_entry_lab, "<18.5", "18.5-25", "25-30", "30+"),
        d_bmi_20_lab = fct_relevel(d_bmi_20_lab, "<18.5", "18.5-25", "25-30", "30+") 
              ) |>
  filter(d_md_avail == 1) |>
   mutate(d_bf_dur_lab = if_else(d_bf_dur_lab == "Missing", NA, d_bf_dur_lab),
         d_age_menopause_lab = if_else(d_age_menopause_lab == "Pre-menopausal", NA, d_age_menopause_lab),
         d_R1hrtstatus_lab = if_else(d_R1hrtstatus_lab == "Pre-menopausal", NA, d_R1hrtstatus_lab),
         d_agebirth1_lab = if_else(d_agebirth1_lab == "Non-parous", NA, d_agebirth1_lab),
         d_dmode = factor(d_dmode, levels = c("SD", "I"))
         ) |> 
  select(any_of(explanatory_vars), d_md
         ) |> 
    tbl_custom_summary(
              statistic = list(
                everything() ~ "{mean} ({sd})"
                ),
              stat_fns  = list(
                everything() ~ function(data, full_data, variable, by, type, ...){
                  dplyr::summarize(data, mean = mean(d_md),
                           sd = sd(d_md))
                }
              ),
              digits = all_continuous() ~ 2,,
              type = list(d_ocstatus2_lab ~ "categorical",
                          d_fambrca_lab ~ "categorical",
                          d_bbd_lab ~ "categorical"),
              missing = "no"
    ) |>
    modify_table_body(
      ~ filter(.x, stat_0 != 0 | is.na(stat_0))
    )


# Unadjusted:
table_unadjusted <- df_compl_case |> 
  filter(d_md_avail == 1
  ) |> 
  mutate(d_bf_dur_lab = if_else(d_bf_dur_lab == "Missing", NA, d_bf_dur_lab),
         d_age_menopause_lab = if_else(d_age_menopause_lab == "Pre-menopausal", NA, d_age_menopause_lab),
         d_R1hrtstatus_lab = if_else(d_R1hrtstatus_lab == "Pre-menopausal", NA, d_R1hrtstatus_lab),
         d_agebirth1_lab = if_else(d_agebirth1_lab == "Non-parous", NA, d_agebirth1_lab),
         d_dmode = factor(d_dmode, levels = c("SD", "I"))
         ) |> 
  select(d_md, any_of(explanatory_vars)
         ) |> 
  tbl_uvregression(
    method = lm,
    y = d_md,
    pvalue_fun = custom_pvalue_fun,
    estimate_fun = function(x){
      case_when(
        is.na(x) ~ NA,
        TRUE ~ formatC(x, format = "f", digits = 2)
      )
    },
    # pvalue_fun = ~ .x,
    # estimate_fun = ~ .x
    method.args = list(na.action = na.omit),
    hide_n = T
   ) |> 
  # modify_table_body(~ .x |>
  #   mutate(across(c(estimate, conf.low, conf.high), exp))
  # ) |>
  # modify_header(
  #   estimate = "**exp(Beta)**"
  # ) |>
  # add_global_p() |>  # add global p-value
  bold_labels() 


#table_unadjusted


# rf vars for violin plots
explanatory_vars <- c(
  
                  # "d_dmode",
                      "d_age_menarche_lab",
                  "d_ocstatus2_lab",
                  "d_parity_lab",
                  "d_agebirth1_lab",
                  "d_bf_dur_lab",
                  "d_R1menopause_lab3",
                  "d_age_menopause_lab", 
                  "d_R1hrtstatus_lab",
                  "d_bbd_lab",
                  "d_fambrca_lab",
                  "d_bmi_20_lab",
                  "d_bmi_entry_lab",
                  "d_R1physmet_leis_who_lab",
                  "d_R1alcohol_units_lab",
                  "d_R1smokingstatus_lab"
                        
                 # "d_inv_status", 
                 # "d_grade_lab", 
                 # "d_morph4_lab", 
                 # "d_pos_nodes_lab", 
                 # "d_tumour_size_lab", 
                 # "d_er_status_lab", 
                 # "d_pr_status_lab", 
                 # "d_her2_status_lab"
                        
                      )


# Adjusted table
# List to store regression tables
table_list <- list()

# data manipulation to remove redundant levels 

df_temp <- df_compl_case |> 
  mutate(d_bf_dur_lab = if_else(d_bf_dur_lab == "Missing", NA, d_bf_dur_lab),
         d_age_menopause_lab = if_else(d_age_menopause_lab == "Pre-menopausal", NA, d_age_menopause_lab),
         d_R1hrtstatus_lab = if_else(d_R1hrtstatus_lab == "Pre-menopausal", NA, d_R1hrtstatus_lab),
         d_agebirth1_lab = if_else(d_agebirth1_lab == "Non-parous", NA, d_agebirth1_lab),
         d_age_mammo = AgeatEntry + d_R1toMD_y,
         d_dmode = factor(d_dmode, levels = c("SD", "I"))
         )


# Loop through each explanatory variable and create a regression table
for (var in explanatory_vars) {
  # Run the regression model adjusted for age
  model <- lm(as.formula(paste("d_md ~", var, "+ d_age_mammo")), data = df_temp, na.action = na.omit)

  # Create the regression table and store it in the list
  table_list[[var]] <- tbl_regression(
    model,
    pvalue_fun = custom_pvalue_fun,
    estimate_fun = function(x){
      case_when(
        is.na(x) ~ NA,
        TRUE ~ formatC(x, format = "f", digits = 2)
      )
    },
    # pvalue_fun = ~ .x,
    # estimate_fun = ~ .x
  ) %>%
  # modify_table_body(~ .x |>
  #   mutate(across(c(estimate, conf.low, conf.high), exp))
  # ) |>
  # modify_header(
  #   estimate = "**exp(Beta)**"
  # ) |>
    # add_global_p(
    # ) |> 
    #add_n() |> # checked with unadjusted and n is the same so no need here
    remove_row_type(variables = "d_age_mammo", type = "all")  # Remove the age covariate from the table
}

# Stack all the regression tables into one
table_adjusted <- tbl_stack(table_list)

# View the final stacked table
table_adjusted

# Merge the tables

md_desc_table <- tbl_merge(list(table_summary, table_summary_md, table_unadjusted, table_adjusted),
                           tab_spanner = c(" ", "**Density**", "**Unadjusted**", "**Adjusted for age at mammogram**"
                                           )
                         )
                         
# Changing variable names to make sense
tab_labels <- c(
  # "d_dmode" = "Mode of detection",
  "d_age_menarche_lab" = "Age at menarche",
  "d_ocstatus2_lab" = "Oral contraception status",
  "d_parity_lab" = "Parity",
  "d_agebirth1_lab" = "Age at first birth",
  "d_bf_dur_lab" = "Breastfeeding duration (mths)",
  "d_R1menopause_lab3" = "Menopausal status",
  "d_age_menopause_lab" = "Age at menopause",
  "d_R1hrtstatus_lab" = "MHT status",
  "d_bbd_lab" = "Benign breast disease",
  "d_fambrca_lab" = "Family history of breast cancer",
  "d_bmi_20_lab" = "BMI at age 20 (kg/m²)",
  "d_bmi_entry_lab" = "BMI at recruitment (kg/m²)",
  "d_R1physmet_leis_who_lab" = "Physical activity (MET h/wk)",
  "d_R1alcohol_units_lab" = "Alcohol units per week",
  "d_R1smokingstatus_lab" = "Smoking status"
  # "d_inv_status" = "Invasive status",
  # "d_grade_lab" = "Grade",
  # "d_morph4_lab" = "Morphology",
  # "d_pos_nodes_lab" = "Node status",
  # "d_tumour_size_lab" = "Tumour size (mm)",
  # "d_er_status_lab" = "ER status",
  # "d_pr_status_lab" = "PR status",
  # "d_her2_status_lab" = "HER2 status"
)

for(i in 1:length(md_desc_table$table_body$label)){
  
  if(md_desc_table$table_body$label[[i]] %in% names(tab_labels)){
    
    md_desc_table$table_body$label[[i]] <- unname(tab_labels[which(names(tab_labels) == md_desc_table$table_body$label[[i]])])
    
  }
  
}

# Filter unneeded rows out
md_desc_table <- md_desc_table |>
  modify_table_body(
    ~ .x |>
      filter(stat_0_1 != "0" | is.na(stat_0_1)) |>
      filter(stat_0_2 != "NA (NA)" | is.na(stat_0_2)) |>
      filter(variable != "d_md")
  )

# Manually changing the labels of levels in the table
md_desc_table$table_body$label[md_desc_table$table_body$label == "Mixed: ductal and other"] <- "Mixed"

md_desc_table$table_body$label[grepl(md_desc_table$table_body$label, pattern = "months")] <- gsub(" months", "", md_desc_table$table_body$label[grepl(md_desc_table$table_body$label, pattern = "months")])

md_desc_table$table_body$label[md_desc_table$table_body$label == "I"] <- "Interval"
md_desc_table$table_body$label[md_desc_table$table_body$label == "SD"] <- "Screen-Detected"

md_desc_table$table_body$label[(md_desc_table$table_body$term_3) == "d_bbd_labYes"] <- "Ever"
md_desc_table$table_body$label[(md_desc_table$table_body$term_3) == "d_bbd_labNo"] <- "Never"


md_desc_table$table_body$label[(md_desc_table$table_body$term_3) == "d_ocstatus2_labNo"] <- "Never"
md_desc_table$table_body$label[(md_desc_table$table_body$term_3) == "d_ocstatus2_labYes"] <- "Ever"

md_desc_table$table_body$label[(md_desc_table$table_body$term_3) == "d_inv_status1"] <- "Invasive"
md_desc_table$table_body$label[(md_desc_table$table_body$term_3) == "d_inv_status0"] <- "DCIS"





md_desc_table

wb <- wb$add_worksheet("Supplementary_Table_3")
wb <- wb_add_flextable(wb, "Supplementary_Table_3", as_flex_table(md_desc_table))

```


### Supplementary Table 4

```{r}
#| echo: false
#| warning: false

# summary Logistic Regression of Model 1  -----------------------------------------

# view existing output
#model2

# remove columns with total and interval N - not simple with gtsummary, trying from scratch without adding the columns


# Create clean version of logistic output, without Ns
model <- df %>%
  glm(formula = d_dmode_n ~ d_R1menopause_lab3 + d_age_meno_tr_lab
      + d_R1hrt_tr_lab
      + d_parity_lab + d_agebirth1_tr_lab + d_age_menarche_lab + d_ocstatus2_lab
      + d_bbd_lab + d_fambrca_lab
      + d_R1alcohol_units_lab + d_R1smokingstatus_lab
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab + d_bmi_20_lab + d_bf_dur_tr_lab
      + diagage + d_R1toBC_y + yeardiag,
            family = binomial(link = "logit"))

# tidy up output
model1 <- tbl_regression(model, 
                         exponentiate = TRUE,
                         label = list(
                          d_R1alcohol_units_lab ~ "Alcohol units per week",
                          d_R1smokingstatus_lab ~ "Smoking status",
                          d_bmi_entry_lab ~ "BMI at recruitment (kg/m²)",
                          d_R1physmet_leis_who_lab ~ "Physical activity (MET h/wk)",
                          d_bbd_lab ~ "Benign breast disease",
                          d_fambrca_lab ~ "Family history of breast cancer",
                          d_age_menarche_lab ~ "Age at menarche",
                          d_ocstatus2_lab ~ "Oral contraceptive use",
                          d_R1menopause_lab3 ~ "Menopausal status",
                          d_age_meno_tr_lab ~ "Age at menopause",
                          d_R1hrt_tr_lab ~ "MHT status",
                          d_parity_lab ~ "Parity",
                          d_agebirth1_tr_lab ~ "Age at first birth",
                          d_bmi_20_lab ~ "BMI at age 20 (kg/m²)",
                          d_bf_dur_tr_lab ~ "Breastfeeding duration (mths)"
                         ),
    pvalue_fun = custom_pvalue_fun
    # pvalue_fun = ~ .x,
    # estimate_fun = ~ .x
                         ) %>% 
  #add_global_p() %>% 
  remove_row_type(variables = c(diagage, d_R1toBC_y, yeardiag), # remove covariates
                  type = "all") %>%
  # add_global_p() |> 
  # remove Missing levels from regression output
  modify_table_body(
    ~ .x %>%
      dplyr::filter(!grepl("Missing", .data$label, ignore.case = TRUE))  # Filter 'missing' levels
  ) |>
  bold_labels() 
  

#model2

# summary Logistic Regression of Model 2  -----------------------------------------

 model <- df %>%
  glm(formula = d_dmode_n ~ d_R1menopause_lab3 + d_age_meno_tr_lab
      + d_R1hrt_tr_lab
      + d_parity_lab + d_agebirth1_tr_lab + d_age_menarche_lab + d_ocstatus2_lab
      + d_bbd_lab + d_fambrca_lab
      + d_R1alcohol_units_lab + d_R1smokingstatus_lab
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab + d_bmi_20_lab + d_bf_dur_tr_lab
      + diagage + d_R1toBC_y + yeardiag
      + d_md_qrt + d_MDtoBC_lab,
            family = binomial(link = "logit"))


# tidy up output
model2 <- tbl_regression(model, 
                         exponentiate = TRUE,
                         label = list(
                          d_R1alcohol_units_lab ~ "Alcohol units per week",
                          d_R1smokingstatus_lab ~ "Smoking status",
                          d_bmi_entry_lab ~ "BMI at recruitment (kg/m²)",
                          d_R1physmet_leis_who_lab ~ "Physical activity (MET h/wk)",
                          d_bbd_lab ~ "Benign breast disease",
                          d_fambrca_lab ~ "Family history of breast cancer",
                          d_age_menarche_lab ~ "Age at menarche",
                          d_ocstatus2_lab ~ "Oral contraceptive use",
                          d_R1menopause_lab3 ~ "Menopausal status",
                          d_age_meno_tr_lab ~ "Age at menopause",
                          d_R1hrt_tr_lab ~ "MHT status",
                          d_parity_lab ~ "Parity",
                          d_agebirth1_tr_lab ~ "Age at first birth",
                          d_bmi_20_lab ~ "BMI at age 20 (kg/m²)",
                          d_bf_dur_tr_lab ~ "Breastfeeding duration (mths)",
                          d_md_qrt ~ "Mammographic density (quartiles)"
                         ),
    pvalue_fun = custom_pvalue_fun
    # pvalue_fun = ~ .x,
    # estimate_fun = ~ .x
                         ) %>% 
  #add_global_p() %>% 
  remove_row_type(variables = c(diagage, d_R1toBC_y, yeardiag, d_MDtoBC_lab), # remove covariates
                  type = "all") %>%
  # add_global_p()|> 
  # remove Missing levels from regression output
  modify_table_body(
    ~ .x %>%
      dplyr::filter(!grepl("Missing", .data$label, ignore.case = TRUE))  # Filter 'missing' levels
  )


# summary Logistic Regression of Model 3  -----------------------------------------

model <- df %>%
  glm(formula = d_dmode_n ~ d_R1menopause_lab3 + d_age_meno_tr_lab
      + d_R1hrt_tr_lab
      + d_parity_lab + d_agebirth1_tr_lab + d_age_menarche_lab + d_ocstatus2_lab
      + d_bbd_lab + d_fambrca_lab
      + d_R1alcohol_units_lab + d_R1smokingstatus_lab
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab + d_bmi_20_lab + d_bf_dur_tr_lab
      + diagage + d_R1toBC_y + yeardiag
      + d_inv_status + d_grade_tr + d_morph4_tr_lab + d_pos_nodes_tr_lab
      + d_tmsize_tr_lab + d_er_tr_lab + d_pr_tr_lab + d_her2_tr_lab,
            family = binomial(link = "logit"))

# tidy up output
model3 <- tbl_regression(model,
                         exponentiate = TRUE,
                         label = list(
                          d_R1alcohol_units_lab ~ "Alcohol units per week",
                          d_R1smokingstatus_lab ~ "Smoking status",
                          d_bmi_entry_lab ~ "BMI at recruitment (kg/m²)",
                          d_R1physmet_leis_who_lab ~ "Physical activity (MET h/wk)",
                          d_bbd_lab ~ "Benign breast disease",
                          d_fambrca_lab ~ "Family history of breast cancer",
                          d_age_menarche_lab ~ "Age at menarche",
                          d_ocstatus2_lab ~ "Oral contraceptive use",
                          d_R1menopause_lab3 ~ "Menopausal status",
                          d_age_meno_tr_lab ~ "Age at menopause",
                          d_R1hrt_tr_lab ~ "MHT status",
                          d_parity_lab ~ "Parity",
                          d_agebirth1_tr_lab ~ "Age at first birth",
                          d_bmi_20_lab ~ "BMI at age 20 (kg/m²)",
                          d_bf_dur_tr_lab ~ "Breastfeeding duration (mths)",
                          d_inv_status ~ "Invasive status",
                          d_grade_tr ~ "Grade",
                          d_morph4_tr_lab ~ "Morphology type",
                          d_pos_nodes_tr_lab ~ "Node status",
                          d_tmsize_tr_lab ~ "Tumour size (mm)",
                          d_er_tr_lab ~ "ER status",
                          d_pr_tr_lab ~ "PR status",
                          d_her2_tr_lab ~ "HER2 status"
                         ),
    pvalue_fun = custom_pvalue_fun
    # pvalue_fun = ~ .x,
    # estimate_fun = ~ .x
                         ) %>%
  #add_global_p() %>%
  remove_row_type(variables = c(diagage, d_R1toBC_y, yeardiag), # remove covariates
                  type = "all") %>%
  # add_global_p()|>
  # remove Missing levels from regression output
  modify_table_body(
    ~ .x %>%
      dplyr::filter(!grepl("Missing", .data$label, ignore.case = TRUE))  # Filter 'missing' levels
  )


# summary Logistic Regression of Model 4  -----------------------------------------

model <- df %>%
  glm(formula = d_dmode_n ~ d_R1menopause_lab3 + d_age_meno_tr_lab
      + d_R1hrt_tr_lab
      + d_parity_lab + d_agebirth1_tr_lab + d_age_menarche_lab + d_ocstatus2_lab
      + d_bbd_lab + d_fambrca_lab
      + d_R1alcohol_units_lab + d_R1smokingstatus_lab
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab + d_bmi_20_lab + d_bf_dur_tr_lab
      + diagage + d_R1toBC_y + yeardiag
      + d_md_qrt + d_MDtoBC_lab
      + d_inv_status + d_grade_tr + d_morph4_tr_lab + d_pos_nodes_tr_lab 
      + d_tmsize_tr_lab + d_er_tr_lab + d_pr_tr_lab + d_her2_tr_lab,
            family = binomial(link = "logit"))

# tidy up output
model4 <- tbl_regression(model, 
                         exponentiate = TRUE,
                         label = list(
                          d_R1alcohol_units_lab ~ "Alcohol units per week",
                          d_R1smokingstatus_lab ~ "Smoking status",
                          d_bmi_entry_lab ~ "BMI at recruitment (kg/m²)",
                          d_R1physmet_leis_who_lab ~ "Physical activity (MET h/wk)",
                          d_bbd_lab ~ "Benign breast disease",
                          d_fambrca_lab ~ "Family history of breast cancer",
                          d_age_menarche_lab ~ "Age at menarche",
                          d_ocstatus2_lab ~ "Oral contraceptive use",
                          d_R1menopause_lab3 ~ "Menopausal status",
                          d_age_meno_tr_lab ~ "Age at menopause",
                          d_R1hrt_tr_lab ~ "MHT status",
                          d_parity_lab ~ "Parity",
                          d_agebirth1_tr_lab ~ "Age at first birth",
                          d_bmi_20_lab ~ "BMI at age 20 (kg/m²)",
                          d_bf_dur_tr_lab ~ "Breastfeeding duration (mths)",
                          d_md_qrt ~ "Mammographic density (quartiles)",
                          d_inv_status ~ "Invasive status",
                          d_grade_tr ~ "Grade",
                          d_morph4_tr_lab ~ "Morphology type",
                          d_pos_nodes_tr_lab ~ "Node status",
                          d_tmsize_tr_lab ~ "Tumour size (mm)",
                          d_er_tr_lab ~ "ER status",
                          d_pr_tr_lab ~ "PR status",
                          d_her2_tr_lab ~ "HER2 status"
                         ),
    pvalue_fun = custom_pvalue_fun
    # estimate_fun = ~ .x
                         ) %>% 
  #add_global_p() %>% 
  remove_row_type(variables = c(diagage, d_R1toBC_y, yeardiag, d_MDtoBC_lab), # remove covariates
                  type = "all") %>%
  # add_global_p()|> 
  # remove Missing levels from regression output
  modify_table_body(
    ~ .x %>%
      dplyr::filter(!grepl("Missing", .data$label, ignore.case = TRUE))  # Filter 'missing' levels
  )


# Create descriptives by mode of detection ------------------------------------
# select(d_dmode_n, d_R1alcohol_units_lab, d_R1smokingstatus_lab, d_bmi_entry_lab,  d_R1physmet_leis_who_lab, d_bbd_lab, d_fambrca_lab, d_age_menarche_lab, d_ocstatus_lab, d_R1menopause_lab3, d_age_meno_tr_lab, d_R1hrt_sttyp_tr_lab, d_parity_tr_lab, d_agebirth1_tr_lab)

# note: need to rename the non-trick variables according to their trick name to align with the model, which can then allow table merge. 

# include missing 
# create df
  
descr_tab2_df <- df  %>%
  select(d_dmode, all_of(all_expl_variables), all_of(tumour_char), diagage, d_R1toBC_y, d_md_qrt
         ) %>% 
  mutate(d_bmi_entry_lab = fct_relevel(d_bmi_entry_lab, "<18.5", "18.5-25", "25-30", "30+"),
        d_bmi_20_lab = fct_relevel(d_bmi_20_lab, "<18.5", "18.5-25", "25-30", "30+") 
              ) %>% 
  mutate(d_age_meno_tr_lab = d_age_menopause_lab,
         d_R1hrt_tr_lab = d_R1hrtstatus_lab,
         d_agebirth1_tr_lab = d_agebirth1_lab,
         d_bf_dur_tr_lab = d_bf_dur_lab
         ) %>% 
  select(-d_age_menopause_lab, -d_R1hrtstatus_lab, -d_agebirth1_lab, -d_bf_dur_lab, -diagage, -d_R1toBC_y)  
  
  

# glimpse(descr_model2)
# str(test)

# create gs summary table
descr_tab2 <- descr_tab2_df %>% 
  mutate(d_dmode = if_else(d_dmode == "I", "Interval", "Screen-Detected")) |>
  select(d_dmode,
         d_age_menarche_lab,
         d_ocstatus2_lab,
         d_parity_lab,
         d_agebirth1_tr_lab,
         d_bf_dur_tr_lab,
         d_R1menopause_lab3,
         d_age_meno_tr_lab,
         d_R1hrt_tr_lab,
         d_bbd_lab,
         d_fambrca_lab,
         d_bmi_entry_lab,
         d_bmi_20_lab,
         d_R1physmet_leis_who_lab,
         d_R1alcohol_units_lab, 
         d_R1smokingstatus_lab,
         d_md_qrt,
         d_inv_status, 
         d_grade_tr = d_grade_lab, 
         d_morph4_tr_lab = d_morph4_lab, 
         d_pos_nodes_tr_lab = d_pos_nodes_lab, 
         d_tmsize_tr_lab = d_tumour_size_lab, 
         d_er_tr_lab = d_er_status_lab, 
         d_pr_tr_lab = d_pr_status_lab, 
         d_her2_tr_lab = d_her2_status_lab) %>% 
  tbl_summary(by = d_dmode,
              statistic = list(
                all_continuous() ~ "{mean} ({sd})",
                all_categorical() ~ "{n} ({p})"
                ),
              digits = all_continuous() ~ 2,
              label = list(
                          d_R1alcohol_units_lab ~ "Alcohol units per week",
                          d_R1smokingstatus_lab ~ "Smoking status",
                          d_bmi_entry_lab ~ "BMI at recruitment (kg/m²)",
                          d_R1physmet_leis_who_lab ~ "Physical activity (MET h/wk)",
                          d_bbd_lab ~ "Benign breast disease",
                          d_fambrca_lab ~ "Family history of breast cancer",
                          d_age_menarche_lab ~ "Age at menarche",
                          d_ocstatus2_lab ~ "Oral contraceptive use",
                          d_R1menopause_lab3 ~ "Menopausal status",
                          d_age_meno_tr_lab ~ "Age at menopause",
                          d_R1hrt_tr_lab ~ "MHT status",
                          d_parity_lab ~ "Parity",
                          d_agebirth1_tr_lab ~ "Age at first birth",
                          d_bmi_20_lab ~ "BMI at age 20 (kg/m²)",
                          d_bf_dur_tr_lab ~ "Breastfeeding duration (mths)",
                          d_md_qrt ~ "Mammographic density (quartiles)",
                          d_inv_status ~ "Invasive status",
                          d_grade_tr ~ "Grade",
                          d_morph4_tr_lab ~ "Morphology type",
                          d_pos_nodes_tr_lab ~ "Node status",
                          d_tmsize_tr_lab ~ "Tumour size (mm)",
                          d_er_tr_lab ~ "ER status",
                          d_pr_tr_lab ~ "PR status",
                          d_her2_tr_lab ~ "HER2 status"
                ),
              type = list(d_bbd_lab ~ "categorical",
                          d_fambrca_lab ~ "categorical")
              )

#descr_model2
  
# merge with analysis output -------------------------------------------------


table2 <- tbl_merge(
  tbls = list(descr_tab2, model1, model2, model3, model4),
  tab_spanner = c("**Distribution by detection mode**", "**Model 1**", "**Model 2**", "**Model 3**", "**Model 4**")
   ) 

#show_header_names(table2)

table2 <- table2 %>% 
  modify_footnote(
    estimate_2 ~ "Model 1: Age at menarche, Menopause, Age at menopause, MHT status, Oral contraceptive use, Parity, Age at first birth, Breast feeding duration (mths), Benign breast disease, Family history of breast cancer, BMI at recruitment (kg/m²), BMI at age 20 (kg/m²), Physical activity (MET h/wk), Alcohol units per week, Smoking status, Age at Diagnosis, Time from Recruitment to BC diagnosis, Year of diagnosis",
    estimate_3 ~ "Model 2: Age at menarche, Menopause, Age at menopause, MHT status, Oral contraceptive use, Parity, Age at first birth, Breast feeding duration (mths), Benign breast disease, Family history of breast cancer, BMI at recruitment (kg/m²), BMI at age 20 (kg/m²), Physical activity (MET h/wk), Alcohol units per week, Smoking status, Age at diagnosis, Time from recruitment to BC diagnosis, Year of diagnosis, Percent mammographic density (quartiles), Time from mammographic density measurement to BRCA diagnosis",
    estimate_4 ~ "Model 3: Age at menarche, Menopause, Age at menopause, MHT status, Oral contraceptive use, Parity, Age at first birth, Breast feeding duration (mths), Benign breast disease, Family history of breast cancer, BMI at recruitment (kg/m²), BMI at age 20 (kg/m²), Physical activity (MET h/wk), Alcohol units per week, Smoking status, Age at diagnosis, Time from recruitment to BC diagnosis, Year of diagnosis, Invasive status, Morphology, Node status, Tumour size, ER status, PR status, HER2 status",
    estimate_5 ~ "Model 4: Age at menarche, Menopause, Age at menopause, MHT status, Oral contraceptive use, Parity, Age at first birth, Breast feeding duration (mths), Benign breast disease, Family history of breast cancer, BMI at recruitment (kg/m²), BMI at age 20 (kg/m²), Physical activity (MET h/wk), Alcohol units per week, Smoking status, Age at diagnosis, Time from recruitment to BC diagnosis, Year of diagnosis, Percent mammographic density (quartiles), Time from mammographic density measurement to BRCA diagnosis, Invasive status, Morphology, Node status, Tumour size, ER status, PR status, HER2 status"
  )

# Manually changing labels for the levels
table2$table_body$label[grepl(" months", table2$table_body$label)] <- gsub(" months", "", table2$table_body$label[grepl(" months", table2$table_body$label)])

table2$table_body$label[43] <- "Never"
table2$table_body$label[44] <- "Ever"

table2$table_body$label[7] <- "Never"
table2$table_body$label[8] <- "Ever"

table2$table_body$label[82] <- "DCIS"
table2$table_body$label[83] <- "Invasive"
table2$table_body$label[table2$table_body$label == "Mixed: ductal and other"] <- "Mixed"

table2$table_body$label[table2$table_body$term_3 == "d_md_qrt1"] <- "1 (<11.45%)"
table2$table_body$label[table2$table_body$term_3 == "d_md_qrt2"] <- "2 (11.45-22.00%)"
table2$table_body$label[table2$table_body$term_3 == "d_md_qrt3"] <- "3 (22.00-26.70%)"
table2$table_body$label[table2$table_body$term_3 == "d_md_qrt4"] <- "4 (>36.70%)"

table2

table2_df <- as_tibble(table2)

write_xlsx(table2_df, path = "../outputs/results.xlsx")


wb <- wb$add_worksheet("Supplementary_Table_4")
wb <- wb_add_flextable(wb, "Supplementary_Table_4", as_flex_table(table2))

wb$save("../outputs/figures/supplement/supplementary_tables.xlsx")

```


## Miscellaneous

### Associations of BBD with TC and MOD

WARNINGS?
```{r, warning = F, results = "asis"}

# Try and find associations with tumour characteristics


model <- df |>
  glm(formula = as.formula(paste0("d_bbd_lab ~", paste0(tumour_char_tr,collapse = "+"))), family = "binomial")


tbl_regression(x = model,
                 exponentiate = T) |>
  add_global_p() |>
  print()

for(i in 1:3){
  
  mean <- mean(df$d_bbd_lab[df$d_grade_lab == i]=="Yes")
  sd <- sd(df$d_bbd_lab[df$d_grade_lab == i]=="Yes")
  
  print(paste0(mean, " (", mean - 1.96*sd, ", ", mean + 1.96*sd, ")"))
  
}

# Testing BBD History association with MOD with/without removing Grade

# Full model
model <- df %>%
  glm(formula = d_dmode_n ~ d_R1menopause_lab3 + d_age_meno_tr_lab
      + d_R1hrt_tr_lab
      + d_parity_lab + d_agebirth1_tr_lab + d_age_menarche_lab + d_ocstatus2_lab
      + d_bbd_lab + d_fambrca_lab
      + d_R1alcohol_units_lab + d_R1smokingstatus_lab
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab + d_bmi_20_lab + d_bf_dur_tr_lab
      + diagage + d_R1toBC_y + yeardiag
      + d_md_qrt + d_MDtoBC_lab
      + d_inv_status + d_grade_tr + d_morph4_tr_lab + d_pos_nodes_tr_lab 
      + d_tmsize_tr_lab + d_er_tr_lab + d_pr_tr_lab + d_her2_tr_lab,
            family = binomial(link = "logit"))

tbl_regression(x = model, exponentiate = T) |>
  modify_table_body(
    ~ .x %>%
      dplyr::filter(.data$variable == "d_bbd_lab")
  )

# Model without TC
model <- df %>%
  glm(formula = d_dmode_n ~ d_R1menopause_lab3 + d_age_meno_tr_lab
      + d_R1hrt_tr_lab
      + d_parity_lab + d_agebirth1_tr_lab + d_age_menarche_lab + d_ocstatus2_lab
      + d_bbd_lab + d_fambrca_lab
      + d_R1alcohol_units_lab + d_R1smokingstatus_lab
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab + d_bmi_20_lab + d_bf_dur_tr_lab
      + diagage + d_R1toBC_y + yeardiag
      + d_md_qrt + d_MDtoBC_lab,
            family = binomial(link = "logit"))

tbl_regression(x = model, exponentiate = T) |>
  modify_table_body(
    ~ .x %>%
      dplyr::filter(.data$variable == "d_bbd_lab")
  )

# Including Grade
model <- df %>%
  glm(formula = d_dmode_n ~ d_R1menopause_lab3 + d_age_meno_tr_lab
      + d_R1hrt_tr_lab
      + d_parity_lab + d_agebirth1_tr_lab + d_age_menarche_lab + d_ocstatus2_lab
      + d_bbd_lab + d_fambrca_lab
      + d_R1alcohol_units_lab + d_R1smokingstatus_lab
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab + d_bmi_20_lab + d_bf_dur_tr_lab
      + diagage + d_R1toBC_y + yeardiag
      + d_md_qrt + d_MDtoBC_lab
      + d_grade_tr,
            family = binomial(link = "logit"))

tbl_regression(x = model, exponentiate = T) |>
  modify_table_body(
    ~ .x %>%
      dplyr::filter(.data$variable == "d_bbd_lab")
  )



```
### Associations of Family History of BC with TC and MOD

```{r, warning = F, results = "asis"}

# Try and find associations with tumour characteristics


model <- df |>
  glm(formula = as.formula(paste0("d_fambrca_lab ~", paste0(tumour_char_tr,collapse = "+"))), family = "binomial")


tbl_regression(x = model,
                 exponentiate = T) |>
  print()

# Testing BBD History association with MOD with/without removing Grade

# Full model
model <- df %>%
  glm(formula = d_dmode_n ~ d_R1menopause_lab3 + d_age_meno_tr_lab
      + d_R1hrt_tr_lab
      + d_parity_lab + d_agebirth1_tr_lab + d_age_menarche_lab + d_ocstatus2_lab
      + d_bbd_lab + d_fambrca_lab
      + d_R1alcohol_units_lab + d_R1smokingstatus_lab
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab + d_bmi_20_lab + d_bf_dur_tr_lab
      + diagage + d_R1toBC_y + yeardiag
      + d_md_qrt + d_MDtoBC_lab
      + d_inv_status + d_grade_tr + d_morph4_tr_lab + d_pos_nodes_tr_lab 
      + d_tmsize_tr_lab + d_er_tr_lab + d_pr_tr_lab + d_her2_tr_lab,
            family = binomial(link = "logit"))

tbl_regression(x = model, exponentiate = T) |>
  modify_table_body(
    ~ .x %>%
      dplyr::filter(.data$variable == "d_fambrca_lab")
  )

# Model without TC
model <- df %>%
  glm(formula = d_dmode_n ~ d_R1menopause_lab3 + d_age_meno_tr_lab
      + d_R1hrt_tr_lab
      + d_parity_lab + d_agebirth1_tr_lab + d_age_menarche_lab + d_ocstatus2_lab
      + d_bbd_lab + d_fambrca_lab
      + d_R1alcohol_units_lab + d_R1smokingstatus_lab
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab + d_bmi_20_lab + d_bf_dur_tr_lab
      + diagage + d_R1toBC_y + yeardiag
      + d_md_qrt + d_MDtoBC_lab,
            family = binomial(link = "logit"))

tbl_regression(x = model, exponentiate = T) |>
  modify_table_body(
    ~ .x %>%
      dplyr::filter(.data$variable == "d_fambrca_lab")
  )

# Adding grade
model <- df %>%
  glm(formula = d_dmode_n ~ d_R1menopause_lab3 + d_age_meno_tr_lab
      + d_R1hrt_tr_lab
      + d_parity_lab + d_agebirth1_tr_lab + d_age_menarche_lab + d_ocstatus2_lab
      + d_bbd_lab + d_fambrca_lab
      + d_R1alcohol_units_lab + d_R1smokingstatus_lab
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab + d_bmi_20_lab + d_bf_dur_tr_lab
      + diagage + d_R1toBC_y + yeardiag
      + d_md_qrt + d_MDtoBC_lab
      + d_grade_tr,
            family = binomial(link = "logit"))

tbl_regression(x = model, exponentiate = T) |>
  modify_table_body(
    ~ .x %>%
      dplyr::filter(.data$variable == "d_fambrca_lab")
  )


```



### Counting individuals by ordering of events in MOD

```{r}
#| echo: false

df |>
  mutate(
    
    MOD_order = case_when(
      d_dmode == "I" & is.na(d_MDtoBC_y) ~ "Interval: no MD",
      d_dmode == "I" & d_R1toMD_y <= 0 ~ "Interval: MD -> R1",
      d_dmode == "I" & d_R1toMD_y > 0 ~ "Interval: R1 -> MD",
      d_dmode == "SD" & is.na(d_MDtoBC_y) ~ "Screen-Detected: no MD",
      d_dmode == "SD" & d_R1toMD_y <= 0 ~ "Screen-Detected: MD -> R1",
      d_dmode == "SD" & d_R1toMD_y > 0 & d_MDtoBC_y == 0 ~ "Screen-Detected: R1 -> MD == BC",
      d_dmode == "SD" & d_R1toMD_y > 0 & d_MDtoBC_y > 0.25 ~ "Screen-Detected: R1 -> MD -> BC",
      d_dmode == "SD" & d_R1toMD_y > 0 & d_MDtoBC_y > 0 ~ "Screen-Detected: R1 -> MD == BC",
      d_dmode == "SD" & d_R1toMD_y > 0 & d_MDtoBC_y < 0 ~ "Screen-Detected: R1 -> MD == BC",
      TRUE ~ "NA"
    )
    
  ) |>
  xtabs(formula = ~ MOD_order)

```

### Summary of time differences

```{r}
#| echo: false

for(mod in unique(df$d_dmode)){
  
  label <- if_else(mod == "I", "Interval", "Screen-Detected")
  
  cat("Summary of time from R1 to breast cancer for", label, "cancers\n")
  with(df |> filter(d_dmode == mod), print(summary(d_R1toBC_y)))
  cat("\n")
  
  cat("Summary of time from R1 to mammogram for", label, "cancers\n")
  with(df |> filter(d_dmode == mod), print(summary(d_R1toMD_y)))
  cat("\n")

  cat("Summary of time from mammogram to breast cancer for", label, "cancers\n")
  with(df |> filter(d_dmode == mod), print(summary(d_MDtoBC_y)))
  cat("\n")

  
  
}

```

### Statistics for paper
```{r}
#| echo: false

# Summary of density for IBC vs SDBC
cat("Summary of mammographic density for IBC\n")
summary(df$d_md[df$d_dmode == "I"], na.rm = T)


cat("Summary of mammographic density for SDBC\n")
summary(df$d_md[df$d_dmode == "SD"], na.rm = T)

# Percentage of missing densities
cat("Percentage of missing densities\n")
summary(1-df$d_md_avail)


```

### Choice of transform for the mammographic density
```{r}

# Normal (no transform)
model <- df %>%
  filter(d_md_avail == 1) |>
  lm(formula = d_md ~ d_dmode + d_R1menopause_lab3 + d_age_meno_tr_lab
      + d_R1hrt_tr_lab
      + d_parity_lab + d_agebirth1_tr_lab + d_age_menarche_lab + d_ocstatus2_lab
      + d_bbd_lab + d_fambrca_lab
      + d_R1alcohol_units_lab + d_R1smokingstatus_lab
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab + d_bmi_20_lab + d_bf_dur_tr_lab
      + diagage + d_R1toBC_y + yeardiag + d_MDtoBC_lab
      + d_inv_status + d_grade_tr + d_morph4_tr_lab + d_pos_nodes_tr_lab 
      + d_tmsize_tr_lab + d_er_tr_lab + d_pr_tr_lab + d_her2_tr_lab)

shapiro.test(model$residuals)
qqnorm(model$residuals)
qqline(model$residuals)

# Log transform
model <- df %>%
  filter(d_md_avail == 1) |>
  lm(formula = log(d_md) ~ d_dmode +  d_R1menopause_lab3 + d_age_meno_tr_lab
      + d_R1hrt_tr_lab
      + d_parity_lab + d_agebirth1_tr_lab + d_age_menarche_lab + d_ocstatus2_lab
      + d_bbd_lab + d_fambrca_lab
      + d_R1alcohol_units_lab + d_R1smokingstatus_lab
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab + d_bmi_20_lab + d_bf_dur_tr_lab
      + diagage + d_R1toBC_y + yeardiag + d_MDtoBC_lab
      + d_inv_status + d_grade_tr + d_morph4_tr_lab + d_pos_nodes_tr_lab 
      + d_tmsize_tr_lab + d_er_tr_lab + d_pr_tr_lab + d_her2_tr_lab)

shapiro.test(model$residuals)
qqnorm(model$residuals)
qqline(model$residuals)

# Logistic transform
model <- df %>%
  filter(d_md_avail == 1) |>
  lm(formula = logit(d_md/100) ~ d_dmode + d_R1menopause_lab3 + d_age_meno_tr_lab
      + d_R1hrt_tr_lab
      + d_parity_lab + d_agebirth1_tr_lab + d_age_menarche_lab + d_ocstatus2_lab
      + d_bbd_lab + d_fambrca_lab
      + d_R1alcohol_units_lab + d_R1smokingstatus_lab
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab + d_bmi_20_lab + d_bf_dur_tr_lab
      + diagage + d_R1toBC_y + yeardiag + d_MDtoBC_lab
      + d_inv_status + d_grade_tr + d_morph4_tr_lab + d_pos_nodes_tr_lab 
      + d_tmsize_tr_lab + d_er_tr_lab + d_pr_tr_lab + d_her2_tr_lab)

shapiro.test(model$residuals)
qqnorm(model$residuals)
qqline(model$residuals)


# Square root transform
model <- df %>%
  filter(d_md_avail == 1) |>
  lm(formula = sqrt(d_md) ~ d_dmode + d_R1menopause_lab3 + d_age_meno_tr_lab
      + d_R1hrt_tr_lab
      + d_parity_lab + d_agebirth1_tr_lab + d_age_menarche_lab + d_ocstatus2_lab
      + d_bbd_lab + d_fambrca_lab
      + d_R1alcohol_units_lab + d_R1smokingstatus_lab
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab + d_bmi_20_lab + d_bf_dur_tr_lab
      + diagage + d_R1toBC_y + yeardiag + d_MDtoBC_lab
      + d_inv_status + d_grade_tr + d_morph4_tr_lab + d_pos_nodes_tr_lab 
      + d_tmsize_tr_lab + d_er_tr_lab + d_pr_tr_lab + d_her2_tr_lab)

shapiro.test(model$residuals)
qqnorm(model$residuals)
qqline(model$residuals)



```

### Fitting a model to see if any variables are associated with no mammographic density measured
```{r}

model <- glm(data = df, formula = d_md_avail ~ d_dmode + d_R1menopause_lab3 + d_age_meno_tr_lab
      + d_R1hrt_tr_lab
      + d_parity_lab + d_agebirth1_tr_lab + d_age_menarche_lab + d_ocstatus2_lab
      + d_bbd_lab + d_fambrca_lab
      + d_R1alcohol_units_lab + d_R1smokingstatus_lab
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab + d_bmi_20_lab + d_bf_dur_tr_lab
      + diagage + d_R1toBC_y + yeardiag
      + d_inv_status + d_grade_tr + d_morph4_tr_lab + d_pos_nodes_tr_lab 
      + d_tmsize_tr_lab + d_er_tr_lab + d_pr_tr_lab + d_her2_tr_lab,
      family = binomial)

summary(model)

```

### Complete-case analysis
```{r}
#| echo: false
#| warning: false

# summary Logistic Regression of Model 1  -----------------------------------------

# view existing output
#model2

# remove columns with total and interval N - not simple with gtsummary, trying from scratch without adding the columns


# Create clean version of logistic output, without Ns
model <- df %>%
  filter(d_md_avail == 1) |>
  glm(formula = d_dmode_n ~ d_R1menopause_lab3 + d_age_meno_tr_lab
      + d_R1hrt_tr_lab
      + d_parity_lab + d_agebirth1_tr_lab + d_age_menarche_lab + d_ocstatus2_lab
      + d_bbd_lab + d_fambrca_lab
      + d_R1alcohol_units_lab + d_R1smokingstatus_lab
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab + d_bmi_20_lab + d_bf_dur_tr_lab
      + diagage + d_R1toBC_y + yeardiag,
            family = binomial(link = "logit"))

# tidy up output
model1 <- tbl_regression(model, 
                         exponentiate = TRUE,
                         label = list(
                          d_R1alcohol_units_lab ~ "Alcohol units per week",
                          d_R1smokingstatus_lab ~ "Smoking status",
                          d_bmi_entry_lab ~ "BMI at recruitment (kg/m²)",
                          d_R1physmet_leis_who_lab ~ "Physical activity (MET h/wk)",
                          d_bbd_lab ~ "Benign breast disease",
                          d_fambrca_lab ~ "Family history of breast cancer",
                          d_age_menarche_lab ~ "Age at menarche",
                          d_ocstatus2_lab ~ "Oral contraceptive use",
                          d_R1menopause_lab3 ~ "Menopausal status",
                          d_age_meno_tr_lab ~ "Age at menopause",
                          d_R1hrt_tr_lab ~ "MHT status",
                          d_parity_lab ~ "Parity",
                          d_agebirth1_tr_lab ~ "Age at first birth",
                          d_bmi_20_lab ~ "BMI at age 20 (kg/m²)",
                          d_bf_dur_tr_lab ~ "Breastfeeding duration (mths)"
                         ),
    pvalue_fun = custom_pvalue_fun
    # pvalue_fun = ~ .x,
    # estimate_fun = ~ .x
                         ) %>% 
  #add_global_p() %>% 
  remove_row_type(variables = c(diagage, d_R1toBC_y, yeardiag), # remove covariates
                  type = "all") %>%
  # add_global_p() |> 
  # remove Missing levels from regression output
  modify_table_body(
    ~ .x %>%
      dplyr::filter(!grepl("Missing", .data$label, ignore.case = TRUE))  # Filter 'missing' levels
  ) |>
  bold_labels() 
  

#model2

# summary Logistic Regression of Model 2  -----------------------------------------

 model <- df %>%
  filter(d_md_avail == 1) |>
  glm(formula = d_dmode_n ~ d_R1menopause_lab3 + d_age_meno_tr_lab
      + d_R1hrt_tr_lab
      + d_parity_lab + d_agebirth1_tr_lab + d_age_menarche_lab + d_ocstatus2_lab
      + d_bbd_lab + d_fambrca_lab
      + d_R1alcohol_units_lab + d_R1smokingstatus_lab
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab + d_bmi_20_lab + d_bf_dur_tr_lab
      + diagage + d_R1toBC_y + yeardiag
      + d_md_qrt + d_MDtoBC_lab,
            family = binomial(link = "logit"))


# tidy up output
model2 <- tbl_regression(model, 
                         exponentiate = TRUE,
                         label = list(
                          d_R1alcohol_units_lab ~ "Alcohol units per week",
                          d_R1smokingstatus_lab ~ "Smoking status",
                          d_bmi_entry_lab ~ "BMI at recruitment (kg/m²)",
                          d_R1physmet_leis_who_lab ~ "Physical activity (MET h/wk)",
                          d_bbd_lab ~ "Benign breast disease",
                          d_fambrca_lab ~ "Family history of breast cancer",
                          d_age_menarche_lab ~ "Age at menarche",
                          d_ocstatus2_lab ~ "Oral contraceptive use",
                          d_R1menopause_lab3 ~ "Menopausal status",
                          d_age_meno_tr_lab ~ "Age at menopause",
                          d_R1hrt_tr_lab ~ "MHT status",
                          d_parity_lab ~ "Parity",
                          d_agebirth1_tr_lab ~ "Age at first birth",
                          d_bmi_20_lab ~ "BMI at age 20 (kg/m²)",
                          d_bf_dur_tr_lab ~ "Breastfeeding duration (mths)",
                          d_md_qrt ~ "Mammographic density (quartiles)"
                         ),
    pvalue_fun = custom_pvalue_fun
    # pvalue_fun = ~ .x,
    # estimate_fun = ~ .x
                         ) %>% 
  #add_global_p() %>% 
  remove_row_type(variables = c(diagage, d_R1toBC_y, yeardiag, d_MDtoBC_lab), # remove covariates
                  type = "all") %>%
  # add_global_p()|> 
  # remove Missing levels from regression output
  modify_table_body(
    ~ .x %>%
      dplyr::filter(!grepl("Missing", .data$label, ignore.case = TRUE))  # Filter 'missing' levels
  )


# summary Logistic Regression of Model 3  -----------------------------------------

model <- df %>%
  filter(d_md_avail == 1) |>
  glm(formula = d_dmode_n ~ d_R1menopause_lab3 + d_age_meno_tr_lab
      + d_R1hrt_tr_lab
      + d_parity_lab + d_agebirth1_tr_lab + d_age_menarche_lab + d_ocstatus2_lab
      + d_bbd_lab + d_fambrca_lab
      + d_R1alcohol_units_lab + d_R1smokingstatus_lab
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab + d_bmi_20_lab + d_bf_dur_tr_lab
      + diagage + d_R1toBC_y + yeardiag
      + d_inv_status + d_grade_tr + d_morph4_tr_lab + d_pos_nodes_tr_lab
      + d_tmsize_tr_lab + d_er_tr_lab + d_pr_tr_lab + d_her2_tr_lab,
            family = binomial(link = "logit"))

# tidy up output
model3 <- tbl_regression(model,
                         exponentiate = TRUE,
                         label = list(
                          d_R1alcohol_units_lab ~ "Alcohol units per week",
                          d_R1smokingstatus_lab ~ "Smoking status",
                          d_bmi_entry_lab ~ "BMI at recruitment (kg/m²)",
                          d_R1physmet_leis_who_lab ~ "Physical activity (MET h/wk)",
                          d_bbd_lab ~ "Benign breast disease",
                          d_fambrca_lab ~ "Family history of breast cancer",
                          d_age_menarche_lab ~ "Age at menarche",
                          d_ocstatus2_lab ~ "Oral contraceptive use",
                          d_R1menopause_lab3 ~ "Menopausal status",
                          d_age_meno_tr_lab ~ "Age at menopause",
                          d_R1hrt_tr_lab ~ "MHT status",
                          d_parity_lab ~ "Parity",
                          d_agebirth1_tr_lab ~ "Age at first birth",
                          d_bmi_20_lab ~ "BMI at age 20 (kg/m²)",
                          d_bf_dur_tr_lab ~ "Breastfeeding duration (mths)",
                          d_inv_status ~ "Invasive status",
                          d_grade_tr ~ "Grade",
                          d_morph4_tr_lab ~ "Morphology type",
                          d_pos_nodes_tr_lab ~ "Node status",
                          d_tmsize_tr_lab ~ "Tumour size (mm)",
                          d_er_tr_lab ~ "ER status",
                          d_pr_tr_lab ~ "PR status",
                          d_her2_tr_lab ~ "HER2 status"
                         ),
    pvalue_fun = custom_pvalue_fun
    # pvalue_fun = ~ .x,
    # estimate_fun = ~ .x
                         ) %>%
  #add_global_p() %>%
  remove_row_type(variables = c(diagage, d_R1toBC_y, yeardiag), # remove covariates
                  type = "all") %>%
  # add_global_p()|>
  # remove Missing levels from regression output
  modify_table_body(
    ~ .x %>%
      dplyr::filter(!grepl("Missing", .data$label, ignore.case = TRUE))  # Filter 'missing' levels
  )


# summary Logistic Regression of Model 4  -----------------------------------------

model <- df %>%
  filter(d_md_avail == 1) |>
  glm(formula = d_dmode_n ~ d_R1menopause_lab3 + d_age_meno_tr_lab
      + d_R1hrt_tr_lab
      + d_parity_lab + d_agebirth1_tr_lab + d_age_menarche_lab + d_ocstatus2_lab
      + d_bbd_lab + d_fambrca_lab
      + d_R1alcohol_units_lab + d_R1smokingstatus_lab
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab + d_bmi_20_lab + d_bf_dur_tr_lab
      + diagage + d_R1toBC_y + yeardiag
      + d_md_qrt + d_MDtoBC_lab
      + d_inv_status + d_grade_tr + d_morph4_tr_lab + d_pos_nodes_tr_lab 
      + d_tmsize_tr_lab + d_er_tr_lab + d_pr_tr_lab + d_her2_tr_lab,
            family = binomial(link = "logit"))

# tidy up output
model4 <- tbl_regression(model, 
                         exponentiate = TRUE,
                         label = list(
                          d_R1alcohol_units_lab ~ "Alcohol units per week",
                          d_R1smokingstatus_lab ~ "Smoking status",
                          d_bmi_entry_lab ~ "BMI at recruitment (kg/m²)",
                          d_R1physmet_leis_who_lab ~ "Physical activity (MET h/wk)",
                          d_bbd_lab ~ "Benign breast disease",
                          d_fambrca_lab ~ "Family history of breast cancer",
                          d_age_menarche_lab ~ "Age at menarche",
                          d_ocstatus2_lab ~ "Oral contraceptive use",
                          d_R1menopause_lab3 ~ "Menopausal status",
                          d_age_meno_tr_lab ~ "Age at menopause",
                          d_R1hrt_tr_lab ~ "MHT status",
                          d_parity_lab ~ "Parity",
                          d_agebirth1_tr_lab ~ "Age at first birth",
                          d_bmi_20_lab ~ "BMI at age 20 (kg/m²)",
                          d_bf_dur_tr_lab ~ "Breastfeeding duration (mths)",
                          d_md_qrt ~ "Mammographic density (quartiles)",
                          d_inv_status ~ "Invasive status",
                          d_grade_tr ~ "Grade",
                          d_morph4_tr_lab ~ "Morphology type",
                          d_pos_nodes_tr_lab ~ "Node status",
                          d_tmsize_tr_lab ~ "Tumour size (mm)",
                          d_er_tr_lab ~ "ER status",
                          d_pr_tr_lab ~ "PR status",
                          d_her2_tr_lab ~ "HER2 status"
                         ),
    pvalue_fun = custom_pvalue_fun
    # estimate_fun = ~ .x
                         ) %>% 
  #add_global_p() %>% 
  remove_row_type(variables = c(diagage, d_R1toBC_y, yeardiag, d_MDtoBC_lab), # remove covariates
                  type = "all") %>%
  # add_global_p()|> 
  # remove Missing levels from regression output
  modify_table_body(
    ~ .x %>%
      dplyr::filter(!grepl("Missing", .data$label, ignore.case = TRUE))  # Filter 'missing' levels
  )


# Create descriptives by mode of detection ------------------------------------
# select(d_dmode_n, d_R1alcohol_units_lab, d_R1smokingstatus_lab, d_bmi_entry_lab,  d_R1physmet_leis_who_lab, d_bbd_lab, d_fambrca_lab, d_age_menarche_lab, d_ocstatus_lab, d_R1menopause_lab3, d_age_meno_tr_lab, d_R1hrt_sttyp_tr_lab, d_parity_tr_lab, d_agebirth1_tr_lab)

# note: need to rename the non-trick variables according to their trick name to align with the model, which can then allow table merge. 

# include missing 
# create df
  
descr_tab2_df <- df  %>%
  filter(d_md_avail == 1) |>
  select(d_dmode, all_of(all_expl_variables), all_of(tumour_char), diagage, d_R1toBC_y, d_md_qrt
         ) %>% 
  mutate(d_bmi_entry_lab = fct_relevel(d_bmi_entry_lab, "<18.5", "18.5-25", "25-30", "30+"),
        d_bmi_20_lab = fct_relevel(d_bmi_20_lab, "<18.5", "18.5-25", "25-30", "30+") 
              ) %>% 
  mutate(d_age_meno_tr_lab = d_age_menopause_lab,
         d_R1hrt_tr_lab = d_R1hrtstatus_lab,
         d_agebirth1_tr_lab = d_agebirth1_lab,
         d_bf_dur_tr_lab = d_bf_dur_lab
         ) %>% 
  select(-d_age_menopause_lab, -d_R1hrtstatus_lab, -d_agebirth1_lab, -d_bf_dur_lab, -diagage, -d_R1toBC_y)  
  
  

# glimpse(descr_model2)
# str(test)

# create gs summary table
descr_tab2 <- descr_tab2_df %>%
  mutate(d_dmode = if_else(d_dmode == "I", "Interval", "Screen-Detected")) |>
  select(d_dmode,
         d_age_menarche_lab,
         d_ocstatus2_lab,
         d_parity_lab,
         d_agebirth1_tr_lab,
         d_bf_dur_tr_lab,
         d_R1menopause_lab3,
         d_age_meno_tr_lab,
         d_R1hrt_tr_lab,
         d_bbd_lab,
         d_fambrca_lab,
         d_bmi_entry_lab,
         d_bmi_20_lab,
         d_R1physmet_leis_who_lab,
         d_R1alcohol_units_lab, 
         d_R1smokingstatus_lab,
         d_md_qrt,
         d_inv_status, 
         d_grade_tr = d_grade_lab, 
         d_morph4_tr_lab = d_morph4_lab, 
         d_pos_nodes_tr_lab = d_pos_nodes_lab, 
         d_tmsize_tr_lab = d_tumour_size_lab, 
         d_er_tr_lab = d_er_status_lab, 
         d_pr_tr_lab = d_pr_status_lab, 
         d_her2_tr_lab = d_her2_status_lab) %>% 
  tbl_summary(by = d_dmode,
              statistic = list(
                all_continuous() ~ "{mean} ({sd})",
                all_categorical() ~ "{n} ({p})"
                ),
              digits = all_continuous() ~ 2,
              label = list(
                          d_R1alcohol_units_lab ~ "Alcohol units per week",
                          d_R1smokingstatus_lab ~ "Smoking status",
                          d_bmi_entry_lab ~ "BMI at recruitment (kg/m²)",
                          d_R1physmet_leis_who_lab ~ "Physical activity (MET h/wk)",
                          d_bbd_lab ~ "Benign breast disease",
                          d_fambrca_lab ~ "Family history of breast cancer",
                          d_age_menarche_lab ~ "Age at menarche",
                          d_ocstatus2_lab ~ "Oral contraceptive use",
                          d_R1menopause_lab3 ~ "Menopausal status",
                          d_age_meno_tr_lab ~ "Age at menopause",
                          d_R1hrt_tr_lab ~ "MHT status",
                          d_parity_lab ~ "Parity",
                          d_agebirth1_tr_lab ~ "Age at first birth",
                          d_bmi_20_lab ~ "BMI at age 20 (kg/m²)",
                          d_bf_dur_tr_lab ~ "Breastfeeding duration (mths)",
                          d_md_qrt ~ "Mammographic density (quartiles)",
                          d_inv_status ~ "Invasive status",
                          d_grade_tr ~ "Grade",
                          d_morph4_tr_lab ~ "Morphology type",
                          d_pos_nodes_tr_lab ~ "Node status",
                          d_tmsize_tr_lab ~ "Tumour size (mm)",
                          d_er_tr_lab ~ "ER status",
                          d_pr_tr_lab ~ "PR status",
                          d_her2_tr_lab ~ "HER2 status"
                ),
              type = list(d_bbd_lab ~ "categorical",
                          d_fambrca_lab ~ "categorical")
              )

#descr_model2
  
# merge with analysis output -------------------------------------------------


table2 <- tbl_merge(
  tbls = list(descr_tab2, model1, model2, model3, model4),
  tab_spanner = c("**Distribution by detection mode**", "**Model 1**", "**Model 2**", "**Model 3**", "**Model 4**")
   ) 

#show_header_names(table2)

table2 <- table2 %>% 
  modify_footnote(
    estimate_2 ~ "Model 1: Age at menarche, Menopause, Age at menopause, MHT status, Oral contraceptive use, Parity, Age at first birth, Breast feeding duration (mths), Benign breast disease, Family history of breast cancer, BMI at recruitment (kg/m²), BMI at age 20 (kg/m²), Physical activity (MET h/wk), Alcohol units per week, Smoking status, Age at Diagnosis, Time from Recruitment to BC diagnosis, Year of diagnosis",
    estimate_3 ~ "Model 2: Age at menarche, Menopause, Age at menopause, MHT status, Oral contraceptive use, Parity, Age at first birth, Breast feeding duration (mths), Benign breast disease, Family history of breast cancer, BMI at recruitment (kg/m²), BMI at age 20 (kg/m²), Physical activity (MET h/wk), Alcohol units per week, Smoking status, Age at diagnosis, Time from recruitment to BC diagnosis, Year of diagnosis, Percent mammographic density (quartiles), Time from mammographic density measurement to BRCA diagnosis",
    estimate_4 ~ "Model 3: Age at menarche, Menopause, Age at menopause, MHT status, Oral contraceptive use, Parity, Age at first birth, Breast feeding duration (mths), Benign breast disease, Family history of breast cancer, BMI at recruitment (kg/m²), BMI at age 20 (kg/m²), Physical activity (MET h/wk), Alcohol units per week, Smoking status, Age at diagnosis, Time from recruitment to BC diagnosis, Year of diagnosis, Invasive status, Morphology, Node status, Tumour size, ER status, PR status, HER2 status",
    estimate_5 ~ "Model 4: Age at menarche, Menopause, Age at menopause, MHT status, Oral contraceptive use, Parity, Age at first birth, Breast feeding duration (mths), Benign breast disease, Family history of breast cancer, BMI at recruitment (kg/m²), BMI at age 20 (kg/m²), Physical activity (MET h/wk), Alcohol units per week, Smoking status, Age at diagnosis, Time from recruitment to BC diagnosis, Year of diagnosis, Percent mammographic density (quartiles), Time from mammographic density measurement to BRCA diagnosis, Invasive status, Morphology, Node status, Tumour size, ER status, PR status, HER2 status"
  )

# Manually changing labels for the levels
table2$table_body$label[grepl(" months", table2$table_body$label)] <- gsub(" months", "", table2$table_body$label[grepl(" months", table2$table_body$label)])

table2$table_body$label[43] <- "Never"
table2$table_body$label[44] <- "Ever"

table2$table_body$label[7] <- "Never"
table2$table_body$label[8] <- "Ever"

table2$table_body$label[82] <- "DCIS"
table2$table_body$label[83] <- "Invasive"
table2$table_body$label[table2$table_body$label == "Mixed: ductal and other"] <- "Mixed"

table2$table_body$label[table2$table_body$term_3 == "d_md_qrt1"] <- "1 (<11.45%)"
table2$table_body$label[table2$table_body$term_3 == "d_md_qrt2"] <- "2 (11.45-22.00%)"
table2$table_body$label[table2$table_body$term_3 == "d_md_qrt3"] <- "3 (22.00-26.70%)"
table2$table_body$label[table2$table_body$term_3 == "d_md_qrt4"] <- "4 (>36.70%)"

table2

table2_df <- as_tibble(table2)

write_xlsx(table2_df, path = "../outputs/cc_results.xlsx")

```

### Testing for effect of SES on BMI association
```{r}

#| echo: false
#| warning: false

# summary Logistic Regression of Model 1  -----------------------------------------

# view existing output
#model2

# Read in SES data
data_ses <- read.csv("//dge/DGE/AETEPID/SHARED/USERS/ZRichards/ForMartina_Nov25_GS_carstairs2001.csv")

# remove columns with total and interval N - not simple with gtsummary, trying from scratch without adding the columns
df_ses <- df |>
  merge(data_ses, by.x = "tcode", by.y = "TCode", all.x = T) |>
  mutate(missing_ses = if_else(is.na(carstairs), 1, 0)) |>
  mutate(carstairs = replace_values(carstairs, NA ~ 0))

# Create clean version of logistic output, without Ns
model <- df %>%
  glm(formula = d_dmode_n ~ d_R1menopause_lab3 + d_age_meno_tr_lab
      + d_R1hrt_tr_lab
      + d_parity_lab + d_agebirth1_tr_lab + d_age_menarche_lab + d_ocstatus2_lab
      + d_bbd_lab + d_fambrca_lab
      + d_R1alcohol_units_lab + d_R1smokingstatus_lab
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab + d_bmi_20_lab + d_bf_dur_tr_lab
      + diagage + d_R1toBC_y + yeardiag,
            family = binomial(link = "logit"))

# tidy up output
model1 <- tbl_regression(model, 
                         exponentiate = TRUE,
                         label = list(
                          d_R1alcohol_units_lab ~ "Alcohol units per week",
                          d_R1smokingstatus_lab ~ "Smoking status",
                          d_bmi_entry_lab ~ "BMI at recruitment (kg/m²)",
                          d_R1physmet_leis_who_lab ~ "Physical activity (MET h/wk)",
                          d_bbd_lab ~ "Benign breast disease",
                          d_fambrca_lab ~ "Family history of breast cancer",
                          d_age_menarche_lab ~ "Age at menarche",
                          d_ocstatus2_lab ~ "Oral contraceptive use",
                          d_R1menopause_lab3 ~ "Menopausal status",
                          d_age_meno_tr_lab ~ "Age at menopause",
                          d_R1hrt_tr_lab ~ "MHT status",
                          d_parity_lab ~ "Parity",
                          d_agebirth1_tr_lab ~ "Age at first birth",
                          d_bmi_20_lab ~ "BMI at age 20 (kg/m²)",
                          d_bf_dur_tr_lab ~ "Breastfeeding duration (mths)"
                         ),
    pvalue_fun = custom_pvalue_fun
    # pvalue_fun = ~ .x,
    # estimate_fun = ~ .x
                         ) %>% 
  #add_global_p() %>% 
  remove_row_type(variables = c(diagage, d_R1toBC_y, yeardiag), # remove covariates
                  type = "all") %>%
  # add_global_p() |> 
  # remove Missing levels from regression output
  modify_table_body(
    ~ .x %>%
      dplyr::filter(!grepl("Missing", .data$label, ignore.case = TRUE))  # Filter 'missing' levels
  ) |>
  bold_labels() 
  

#model2

# summary Logistic Regression of Model 2  -----------------------------------------

 model <- df %>%
  glm(formula = d_dmode_n ~ d_R1menopause_lab3 + d_age_meno_tr_lab
      + d_R1hrt_tr_lab
      + d_parity_lab + d_agebirth1_tr_lab + d_age_menarche_lab + d_ocstatus2_lab
      + d_bbd_lab + d_fambrca_lab
      + d_R1alcohol_units_lab + d_R1smokingstatus_lab
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab + d_bmi_20_lab + d_bf_dur_tr_lab
      + diagage + d_R1toBC_y + yeardiag
      + d_md_qrt + d_MDtoBC_lab,
            family = binomial(link = "logit"))


# tidy up output
model2 <- tbl_regression(model, 
                         exponentiate = TRUE,
                         label = list(
                          d_R1alcohol_units_lab ~ "Alcohol units per week",
                          d_R1smokingstatus_lab ~ "Smoking status",
                          d_bmi_entry_lab ~ "BMI at recruitment (kg/m²)",
                          d_R1physmet_leis_who_lab ~ "Physical activity (MET h/wk)",
                          d_bbd_lab ~ "Benign breast disease",
                          d_fambrca_lab ~ "Family history of breast cancer",
                          d_age_menarche_lab ~ "Age at menarche",
                          d_ocstatus2_lab ~ "Oral contraceptive use",
                          d_R1menopause_lab3 ~ "Menopausal status",
                          d_age_meno_tr_lab ~ "Age at menopause",
                          d_R1hrt_tr_lab ~ "MHT status",
                          d_parity_lab ~ "Parity",
                          d_agebirth1_tr_lab ~ "Age at first birth",
                          d_bmi_20_lab ~ "BMI at age 20 (kg/m²)",
                          d_bf_dur_tr_lab ~ "Breastfeeding duration (mths)",
                          d_md_qrt ~ "Mammographic density (quartiles)"
                         ),
    pvalue_fun = custom_pvalue_fun
    # pvalue_fun = ~ .x,
    # estimate_fun = ~ .x
                         ) %>% 
  #add_global_p() %>% 
  remove_row_type(variables = c(diagage, d_R1toBC_y, yeardiag, d_MDtoBC_lab), # remove covariates
                  type = "all") %>%
  # add_global_p()|> 
  # remove Missing levels from regression output
  modify_table_body(
    ~ .x %>%
      dplyr::filter(!grepl("Missing", .data$label, ignore.case = TRUE))  # Filter 'missing' levels
  )


# summary Logistic Regression of Model 3  -----------------------------------------

model <- df %>%
  glm(formula = d_dmode_n ~ d_R1menopause_lab3 + d_age_meno_tr_lab
      + d_R1hrt_tr_lab
      + d_parity_lab + d_agebirth1_tr_lab + d_age_menarche_lab + d_ocstatus2_lab
      + d_bbd_lab + d_fambrca_lab
      + d_R1alcohol_units_lab + d_R1smokingstatus_lab
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab + d_bmi_20_lab + d_bf_dur_tr_lab
      + diagage + d_R1toBC_y + yeardiag
      + d_inv_status + d_grade_tr + d_morph4_tr_lab + d_pos_nodes_tr_lab
      + d_tmsize_tr_lab + d_er_tr_lab + d_pr_tr_lab + d_her2_tr_lab,
            family = binomial(link = "logit"))

# tidy up output
model3 <- tbl_regression(model,
                         exponentiate = TRUE,
                         label = list(
                          d_R1alcohol_units_lab ~ "Alcohol units per week",
                          d_R1smokingstatus_lab ~ "Smoking status",
                          d_bmi_entry_lab ~ "BMI at recruitment (kg/m²)",
                          d_R1physmet_leis_who_lab ~ "Physical activity (MET h/wk)",
                          d_bbd_lab ~ "Benign breast disease",
                          d_fambrca_lab ~ "Family history of breast cancer",
                          d_age_menarche_lab ~ "Age at menarche",
                          d_ocstatus2_lab ~ "Oral contraceptive use",
                          d_R1menopause_lab3 ~ "Menopausal status",
                          d_age_meno_tr_lab ~ "Age at menopause",
                          d_R1hrt_tr_lab ~ "MHT status",
                          d_parity_lab ~ "Parity",
                          d_agebirth1_tr_lab ~ "Age at first birth",
                          d_bmi_20_lab ~ "BMI at age 20 (kg/m²)",
                          d_bf_dur_tr_lab ~ "Breastfeeding duration (mths)",
                          d_inv_status ~ "Invasive status",
                          d_grade_tr ~ "Grade",
                          d_morph4_tr_lab ~ "Morphology type",
                          d_pos_nodes_tr_lab ~ "Node status",
                          d_tmsize_tr_lab ~ "Tumour size (mm)",
                          d_er_tr_lab ~ "ER status",
                          d_pr_tr_lab ~ "PR status",
                          d_her2_tr_lab ~ "HER2 status"
                         ),
    pvalue_fun = custom_pvalue_fun
    # pvalue_fun = ~ .x,
    # estimate_fun = ~ .x
                         ) %>%
  #add_global_p() %>%
  remove_row_type(variables = c(diagage, d_R1toBC_y, yeardiag), # remove covariates
                  type = "all") %>%
  # add_global_p()|>
  # remove Missing levels from regression output
  modify_table_body(
    ~ .x %>%
      dplyr::filter(!grepl("Missing", .data$label, ignore.case = TRUE))  # Filter 'missing' levels
  )


# summary Logistic Regression of Model 4  -----------------------------------------

model <- df %>%
  glm(formula = d_dmode_n ~ d_R1menopause_lab3 + d_age_meno_tr_lab
      + d_R1hrt_tr_lab
      + d_parity_lab + d_agebirth1_tr_lab + d_age_menarche_lab + d_ocstatus2_lab
      + d_bbd_lab + d_fambrca_lab
      + d_R1alcohol_units_lab + d_R1smokingstatus_lab
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab + d_bmi_20_lab + d_bf_dur_tr_lab
      + diagage + d_R1toBC_y + yeardiag
      + d_md_qrt + d_MDtoBC_lab
      + d_inv_status + d_grade_tr + d_morph4_tr_lab + d_pos_nodes_tr_lab 
      + d_tmsize_tr_lab + d_er_tr_lab + d_pr_tr_lab + d_her2_tr_lab,
            family = binomial(link = "logit"))

# tidy up output
model4 <- tbl_regression(model, 
                         exponentiate = TRUE,
                         label = list(
                          d_R1alcohol_units_lab ~ "Alcohol units per week",
                          d_R1smokingstatus_lab ~ "Smoking status",
                          d_bmi_entry_lab ~ "BMI at recruitment (kg/m²)",
                          d_R1physmet_leis_who_lab ~ "Physical activity (MET h/wk)",
                          d_bbd_lab ~ "Benign breast disease",
                          d_fambrca_lab ~ "Family history of breast cancer",
                          d_age_menarche_lab ~ "Age at menarche",
                          d_ocstatus2_lab ~ "Oral contraceptive use",
                          d_R1menopause_lab3 ~ "Menopausal status",
                          d_age_meno_tr_lab ~ "Age at menopause",
                          d_R1hrt_tr_lab ~ "MHT status",
                          d_parity_lab ~ "Parity",
                          d_agebirth1_tr_lab ~ "Age at first birth",
                          d_bmi_20_lab ~ "BMI at age 20 (kg/m²)",
                          d_bf_dur_tr_lab ~ "Breastfeeding duration (mths)",
                          d_md_qrt ~ "Mammographic density (quartiles)",
                          d_inv_status ~ "Invasive status",
                          d_grade_tr ~ "Grade",
                          d_morph4_tr_lab ~ "Morphology type",
                          d_pos_nodes_tr_lab ~ "Node status",
                          d_tmsize_tr_lab ~ "Tumour size (mm)",
                          d_er_tr_lab ~ "ER status",
                          d_pr_tr_lab ~ "PR status",
                          d_her2_tr_lab ~ "HER2 status"
                         ),
    pvalue_fun = custom_pvalue_fun
    # estimate_fun = ~ .x
                         ) %>% 
  #add_global_p() %>% 
  remove_row_type(variables = c(diagage, d_R1toBC_y, yeardiag, d_MDtoBC_lab), # remove covariates
                  type = "all") %>%
  # add_global_p()|> 
  # remove Missing levels from regression output
  modify_table_body(
    ~ .x %>%
      dplyr::filter(!grepl("Missing", .data$label, ignore.case = TRUE))  # Filter 'missing' levels
  )

# summary Logistic Regression of Model 4  -----------------------------------------

model <- df_ses %>%
  glm(formula = d_dmode_n ~ d_R1menopause_lab3 + d_age_meno_tr_lab
      + d_R1hrt_tr_lab
      + d_parity_lab + d_agebirth1_tr_lab + d_age_menarche_lab + d_ocstatus2_lab
      + d_bbd_lab + d_fambrca_lab
      + d_R1alcohol_units_lab + d_R1smokingstatus_lab
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab + d_bmi_20_lab + d_bf_dur_tr_lab
      + diagage + d_R1toBC_y + yeardiag
      + d_md_qrt + d_MDtoBC_lab
      + d_inv_status + d_grade_tr + d_morph4_tr_lab + d_pos_nodes_tr_lab 
      + d_tmsize_tr_lab + d_er_tr_lab + d_pr_tr_lab + d_her2_tr_lab + carstairs + missing_ses,
            family = binomial(link = "logit"))

# tidy up output
model5 <- tbl_regression(model, 
                         exponentiate = TRUE,
                         label = list(
                          d_R1alcohol_units_lab ~ "Alcohol units per week",
                          d_R1smokingstatus_lab ~ "Smoking status",
                          d_bmi_entry_lab ~ "BMI at recruitment (kg/m²)",
                          d_R1physmet_leis_who_lab ~ "Physical activity (MET h/wk)",
                          d_bbd_lab ~ "Benign breast disease",
                          d_fambrca_lab ~ "Family history of breast cancer",
                          d_age_menarche_lab ~ "Age at menarche",
                          d_ocstatus2_lab ~ "Oral contraceptive use",
                          d_R1menopause_lab3 ~ "Menopausal status",
                          d_age_meno_tr_lab ~ "Age at menopause",
                          d_R1hrt_tr_lab ~ "MHT status",
                          d_parity_lab ~ "Parity",
                          d_agebirth1_tr_lab ~ "Age at first birth",
                          d_bmi_20_lab ~ "BMI at age 20 (kg/m²)",
                          d_bf_dur_tr_lab ~ "Breastfeeding duration (mths)",
                          d_md_qrt ~ "Mammographic density (quartiles)",
                          d_inv_status ~ "Invasive status",
                          d_grade_tr ~ "Grade",
                          d_morph4_tr_lab ~ "Morphology type",
                          d_pos_nodes_tr_lab ~ "Node status",
                          d_tmsize_tr_lab ~ "Tumour size (mm)",
                          d_er_tr_lab ~ "ER status",
                          d_pr_tr_lab ~ "PR status",
                          d_her2_tr_lab ~ "HER2 status"
                         ),
    pvalue_fun = custom_pvalue_fun
    # estimate_fun = ~ .x
                         ) %>% 
  #add_global_p() %>% 
  remove_row_type(variables = c(diagage, d_R1toBC_y, yeardiag, d_MDtoBC_lab), # remove covariates
                  type = "all") %>%
  # add_global_p()|> 
  # remove Missing levels from regression output
  modify_table_body(
    ~ .x %>%
      dplyr::filter(!grepl("Missing", .data$label, ignore.case = TRUE))  # Filter 'missing' levels
  )


# Create descriptives by mode of detection ------------------------------------
# select(d_dmode_n, d_R1alcohol_units_lab, d_R1smokingstatus_lab, d_bmi_entry_lab,  d_R1physmet_leis_who_lab, d_bbd_lab, d_fambrca_lab, d_age_menarche_lab, d_ocstatus_lab, d_R1menopause_lab3, d_age_meno_tr_lab, d_R1hrt_sttyp_tr_lab, d_parity_tr_lab, d_agebirth1_tr_lab)

# note: need to rename the non-trick variables according to their trick name to align with the model, which can then allow table merge. 

# include missing 
# create df
  
descr_tab2_df <- df  %>%
  select(d_dmode, all_of(all_expl_variables), all_of(tumour_char), diagage, d_R1toBC_y, d_md_qrt
         ) %>% 
  mutate(d_bmi_entry_lab = fct_relevel(d_bmi_entry_lab, "<18.5", "18.5-25", "25-30", "30+"),
        d_bmi_20_lab = fct_relevel(d_bmi_20_lab, "<18.5", "18.5-25", "25-30", "30+") 
              ) %>% 
  mutate(d_age_meno_tr_lab = d_age_menopause_lab,
         d_R1hrt_tr_lab = d_R1hrtstatus_lab,
         d_agebirth1_tr_lab = d_agebirth1_lab,
         d_bf_dur_tr_lab = d_bf_dur_lab
         ) %>% 
  select(-d_age_menopause_lab, -d_R1hrtstatus_lab, -d_agebirth1_lab, -d_bf_dur_lab, -diagage, -d_R1toBC_y)  
  
  

# glimpse(descr_model2)
# str(test)

# create gs summary table
descr_tab2 <- descr_tab2_df %>% 
  mutate(d_dmode = if_else(d_dmode == "I", "Interval", "Screen-Detected")) |>
  select(d_dmode,
         d_age_menarche_lab,
         d_ocstatus2_lab,
         d_parity_lab,
         d_agebirth1_tr_lab,
         d_bf_dur_tr_lab,
         d_R1menopause_lab3,
         d_age_meno_tr_lab,
         d_R1hrt_tr_lab,
         d_bbd_lab,
         d_fambrca_lab,
         d_bmi_entry_lab,
         d_bmi_20_lab,
         d_R1physmet_leis_who_lab,
         d_R1alcohol_units_lab, 
         d_R1smokingstatus_lab,
         d_md_qrt,
         d_inv_status, 
         d_grade_tr = d_grade_lab, 
         d_morph4_tr_lab = d_morph4_lab, 
         d_pos_nodes_tr_lab = d_pos_nodes_lab, 
         d_tmsize_tr_lab = d_tumour_size_lab, 
         d_er_tr_lab = d_er_status_lab, 
         d_pr_tr_lab = d_pr_status_lab, 
         d_her2_tr_lab = d_her2_status_lab) %>% 
  tbl_summary(by = d_dmode,
              statistic = list(
                all_continuous() ~ "{mean} ({sd})",
                all_categorical() ~ "{n} ({p})"
                ),
              digits = all_continuous() ~ 2,
              label = list(
                          d_R1alcohol_units_lab ~ "Alcohol units per week",
                          d_R1smokingstatus_lab ~ "Smoking status",
                          d_bmi_entry_lab ~ "BMI at recruitment (kg/m²)",
                          d_R1physmet_leis_who_lab ~ "Physical activity (MET h/wk)",
                          d_bbd_lab ~ "Benign breast disease",
                          d_fambrca_lab ~ "Family history of breast cancer",
                          d_age_menarche_lab ~ "Age at menarche",
                          d_ocstatus2_lab ~ "Oral contraceptive use",
                          d_R1menopause_lab3 ~ "Menopausal status",
                          d_age_meno_tr_lab ~ "Age at menopause",
                          d_R1hrt_tr_lab ~ "MHT status",
                          d_parity_lab ~ "Parity",
                          d_agebirth1_tr_lab ~ "Age at first birth",
                          d_bmi_20_lab ~ "BMI at age 20 (kg/m²)",
                          d_bf_dur_tr_lab ~ "Breastfeeding duration (mths)",
                          d_md_qrt ~ "Mammographic density (quartiles)",
                          d_inv_status ~ "Invasive status",
                          d_grade_tr ~ "Grade",
                          d_morph4_tr_lab ~ "Morphology type",
                          d_pos_nodes_tr_lab ~ "Node status",
                          d_tmsize_tr_lab ~ "Tumour size (mm)",
                          d_er_tr_lab ~ "ER status",
                          d_pr_tr_lab ~ "PR status",
                          d_her2_tr_lab ~ "HER2 status"
                ),
              type = list(d_bbd_lab ~ "categorical",
                          d_fambrca_lab ~ "categorical")
              )

#descr_model2
  
# merge with analysis output -------------------------------------------------


table2 <- tbl_merge(
  tbls = list(descr_tab2, model1, model2, model3, model4, model5),
  tab_spanner = c("**Distribution by detection mode**", "**Model 1**", "**Model 2**", "**Model 3**", "**Model 4**", "**Model 5**")
   ) 

#show_header_names(table2)

table2 <- table2 %>% 
  modify_footnote(
    estimate_2 ~ "Model 1: Age at menarche, Menopause, Age at menopause, MHT status, Oral contraceptive use, Parity, Age at first birth, Breast feeding duration (mths), Benign breast disease, Family history of breast cancer, BMI at recruitment (kg/m²), BMI at age 20 (kg/m²), Physical activity (MET h/wk), Alcohol units per week, Smoking status, Age at Diagnosis, Time from Recruitment to BC diagnosis, Year of diagnosis",
    estimate_3 ~ "Model 2: Age at menarche, Menopause, Age at menopause, MHT status, Oral contraceptive use, Parity, Age at first birth, Breast feeding duration (mths), Benign breast disease, Family history of breast cancer, BMI at recruitment (kg/m²), BMI at age 20 (kg/m²), Physical activity (MET h/wk), Alcohol units per week, Smoking status, Age at diagnosis, Time from recruitment to BC diagnosis, Year of diagnosis, Percent mammographic density (quartiles), Time from mammographic density measurement to BRCA diagnosis",
    estimate_4 ~ "Model 3: Age at menarche, Menopause, Age at menopause, MHT status, Oral contraceptive use, Parity, Age at first birth, Breast feeding duration (mths), Benign breast disease, Family history of breast cancer, BMI at recruitment (kg/m²), BMI at age 20 (kg/m²), Physical activity (MET h/wk), Alcohol units per week, Smoking status, Age at diagnosis, Time from recruitment to BC diagnosis, Year of diagnosis, Invasive status, Morphology, Node status, Tumour size, ER status, PR status, HER2 status",
    estimate_5 ~ "Model 4: Age at menarche, Menopause, Age at menopause, MHT status, Oral contraceptive use, Parity, Age at first birth, Breast feeding duration (mths), Benign breast disease, Family history of breast cancer, BMI at recruitment (kg/m²), BMI at age 20 (kg/m²), Physical activity (MET h/wk), Alcohol units per week, Smoking status, Age at diagnosis, Time from recruitment to BC diagnosis, Year of diagnosis, Percent mammographic density (quartiles), Time from mammographic density measurement to BRCA diagnosis, Invasive status, Morphology, Node status, Tumour size, ER status, PR status, HER2 status"
  )

# Manually changing labels for the levels
table2$table_body$label[grepl(" months", table2$table_body$label)] <- gsub(" months", "", table2$table_body$label[grepl(" months", table2$table_body$label)])

table2$table_body$label[43] <- "Never"
table2$table_body$label[44] <- "Ever"

table2$table_body$label[7] <- "Never"
table2$table_body$label[8] <- "Ever"

table2$table_body$label[82] <- "DCIS"
table2$table_body$label[83] <- "Invasive"
table2$table_body$label[table2$table_body$label == "Mixed: ductal and other"] <- "Mixed"

table2$table_body$label[table2$table_body$term_3 == "d_md_qrt1"] <- "1 (<11.45%)"
table2$table_body$label[table2$table_body$term_3 == "d_md_qrt2"] <- "2 (11.45-22.00%)"
table2$table_body$label[table2$table_body$term_3 == "d_md_qrt3"] <- "3 (22.00-26.70%)"
table2$table_body$label[table2$table_body$term_3 == "d_md_qrt4"] <- "4 (>36.70%)"

table2

table2_df <- as_tibble(table2)

write_xlsx(table2_df, path = "../outputs/ses_results.xlsx")



```
