---
title: "Detection mode project: Tables and figures for publication"
author: "Martina Brayley"
date: last-modified
date-format: long
format: 
  html: 
    embed-resources: true
    toc: true
    toc-location: left
    number-sections: true
    theme: [flatly, style.css]
    page-layout: full
    grid:
      sidebar-width: 300px
      body-width: 2000px
      margin-width: 300px
      gutter-width: 1.5rem
editor: visual
execute:
  warning: false
---

version 5 - Using data missing as a separate category in analysis

```{r}
#| include: false
#| echo: false

# SET UP ------------------------------------------------------------------------


# execute master file - no need run it as final data now saved
#source("code/master.R")

print(getwd())

library(pacman)
p_load(here,
       broom.helpers,
       paleteer,
       ggthemes,
       nortest)

# set up
setup_path <- here("code", "1.setup_import.r")
source(setup_path)

# import data
# data_path <- here("data", "an_df.rds")
# an_df <- readRDS(data_path)

#an_df <- readRDS("Q:/SHARED/USERS/MBrayley/Screening/data/an_df.rds")


# dataset processed in analysis script v3 - adjusting missing levels for TC
an_df <- readRDS("Q:/SHARED/USERS/MBrayley/Screening/data/df_from_analysis_script.rds")

```

```{r}
#| include: false
#| echo: false

# DATA PREP ----------------------------------------------------------------------


## Store explanatory variables as an object -------------------------------

# select categorical variables that are labelled

# exploratory vars 
all_expl_variables <- c(
                         # parity related risk factors
                        "d_R1menopause_lab3", 
                        "d_age_menopause_lab", 
                        "d_R1hrtstatus_lab", 
                        "d_parity_lab", 
                        "d_agebirth1_lab",
                        "d_age_menarche_lab", 
                        "d_ocstatus_lab", 
                        # other biological risk factors
                        "d_bbd_lab", 
                        "d_fambrca_lab",
                        # lifestyle risk factors
                        "d_R1alcohol_units_lab", 
                        "d_R1smokingstatus_lab", 
                        "d_bmi_entry_lab", 
                        "d_R1physmet_leis_who_lab")

all_expl_variables_tr <- c(
                            # parity related risk factors
                           "d_R1menopause_lab3", 
                           "d_age_meno_tr_lab", 
                           "d_R1hrt_tr_lab", 
                           "d_parity_lab",
                           "d_age_menarche_lab", 
                           "d_ocstatus_lab",
                            # other biological risk factors
                           "d_bbd_lab", 
                           "d_fambrca_lab",
                           # lifestyle risk factors
                           "d_agebirth1_tr_lab",
                           "d_R1alcohol_units_lab", 
                           "d_R1smokingstatus_lab", 
                           "d_bmi_entry_lab", 
                           "d_R1physmet_leis_who_lab" 
                           )
  

lifestyle <- c("d_R1alcohol_units_lab", 
               "d_R1smokingstatus_lab", 
               "d_bmi_entry_lab", 
               "d_R1physmet_leis_who")
  
reproductive <- c("d_age_menarche_lab", 
                  "d_ocstatus_lab", 
                  "d_R1menopause_lab3", 
                  "d_age_menopause_lab", 
                  "d_R1hrtstatus_lab", 
                  "d_parity_lab", 
                  "d_agebirth1_lab") # do we want to add breastfeeding? If so, it would also have be in the trick model
  
orther_rf <- c("d_bbd", 
               "d_fambrca")

# mammo density excluded from the lists - special variable for subanalyses

# covariates 
all_covariates <- c(# time varying variables
                    "diagage", 
                    "d_R1toBC_y", 
                    "d_R1toMD_y", 
                    "d_MDtoBC_y",
                    # tumour characteristics
                    "d_inv_status", 
                    "d_grade_lab", 
                    "d_morph4", 
                    "d_pos_nodes_lab", 
                    "d_tumour_size", 
                    "d_er_status", 
                    "d_pr_status", 
                    "d_her2_status")  
  
tumour_char <- c("d_inv_status", 
                 "d_grade_lab", 
                 "d_morph4_lab", 
                 "d_pos_nodes_lab", 
                 "d_tumour_size_lab", 
                 "d_er_status_lab", 
                 "d_pr_status_lab", 
                 "d_her2_status_lab")

tumour_char_tr <- c("d_inv_status", 
                    "d_grade_tr", 
                    "d_morph4_tr_lab", 
                    "d_pos_nodes_tr_lab", 
                    "d_tmsize_tr_lab", 
                    "d_er_tr_lab", 
                    "d_pr_tr_lab", 
                    "d_her2_tr_lab")
  
time_vars <- c("diagage", 
               "d_R1toBC_y", 
               "d_R1toMD_y", 
               "d_MDtoBC_y")

# create dataset 
df <- an_df %>% 
  mutate(d_bmi_entry_lab = fct_relevel(d_bmi_entry_lab, "18.5-25", "<18.5", "25-30", "30+")
  )

# set theme for gtsummary tables: 

theme_gtsummary_compact(set_theme = TRUE)


```

# Table and figure structure:

**Main tables**

Table 1 - general descriptives (population characteristics)

Table 2 - Risk factor descriptives by detection mode with logistic regression results from default model ( adjusted for multiple exposures and selected covariates) + model additionally adjusted for mammo density (add mediation measure)

**Main figures**

Figures 1 to 3 - comprised figure of plots describing mamographic density in relation to risk factors (detail in ST 2)

**Supplementary material**

(ST = Supplementary Table)

ST 2 - risk factor descriptives by density (complementing figures 1 to 3)

ST 3 - TC descriptives by detection mode with logistic regression results for TC only + full model adjusting for TC + adjusting for density

Suplemntar Figure 1 - Forest plot of analysis adjusted vs unadjusted for mammographic density (full table in supplementary material - ST 2)

# Main tables

## Table 1 - general participant descriptives

stage coding - DCIS coded as stage 0 - should we keep it that way or code them as unknown? or DCIS?

time from recruitement to diagnosis categorical or continuous?

```{r}
#| echo: false

df %>% 
  select(diagage, AgeatEntry, d_R1toBC_lab, d_R1toBC_y, d_inv_status_lab, d_grade_lab, d_stage_cat, d_morph4_lab, d_pos_nodes_lab, d_tumour_size_lab, d_er_status_lab, d_pr_status_lab, d_her2_status_lab
         ) %>% 
  tbl_summary(statistic = list(
                all_continuous() ~ "{mean} ({sd})",
                all_categorical() ~ "{n} ({p}) "),
              digits = all_continuous() ~ 2,
              label = list(diagage ~ "Age at diagnosis",
                           AgeatEntry ~ "Age at recruitment",
                           d_R1toBC_y ~ "Time from recruitment to diagnosis",
                           d_R1toBC_lab ~ "Time from recruitment to diagnosis",
                          d_inv_status_lab ~ "Invasive status",
                          d_grade_lab ~ "Grade",
                          d_stage_cat ~ "Stage",
                          d_morph4_lab ~ "Morphology type",
                          d_pos_nodes_lab ~ "Positive nodes",
                          d_tumour_size_lab ~ "Tumour size",
                          d_er_status_lab ~ "ER status",
                          d_pr_status_lab ~ "PR status",
                          d_her2_status_lab ~ "HER2 status"
                          )
  )

```

## Table 2 - Main analysis

The main analysis is a logistic regression with multiple exposures (risk factors).

This table shows risk factor level descriptives and their associations with the mode of detection (logistic regression) adjusted for all other risk factors, age at diagnosis and time from entry to BC diagnosis.

Model 2 is the final selected model (Default model) to present in the analysis. Model 2 is adjusted for all co-exposures + age at diagnosis and time from entry to BC diagnosis. Model 1 was only adjusted for co-exposures, e.i. all exposures in the model.

Because model 2 was selected as the default model, it is also used on the subanalyses with TC and MD as the comparison against adjusted models for TC and MD.

**Model 1**: “d_R1alcohol_status_cat”, “d_R1alcohol_units_cat”, “d_R1smokingstatus_cat”, “d_bmi_entry_cat”, “d_R1physmet_leis_who”, “d_bbd”, “d_fambrca”, “d_age_menarche_cat”, “d_ocstatus_cat”, “d_R1menopause_cat3”, “d_age_menopause_cat”, “d_R1hrtstatus_cat”, “d_parous_cat”, “d_parity_cat”, “d_agebirth1_cat”

**Model 2**: Model 1 + “diagage”, “d_R1toBC_y”, “yeardiag”

**Model 3:** Model 2 + “d_md_qrt”

**Model 4:** Model 3 + d_MDtoBC_lab (categorical)

```{r}
#| echo: false
#| warning: false

# summary Logistic Regression of Model 2  -----------------------------------------

# view existing output
#model2

# remove columns with total and interval N - not simple with gtsummary, trying from scratch without adding the columns


# Create clean version of logistic output, without Ns
model <- df %>%
  glm(formula = d_dmode_n ~ d_R1menopause_lab3 + d_age_meno_tr_lab 
      + d_R1hrt_tr_lab
      + d_parity_lab + d_agebirth1_tr_lab + d_age_menarche_lab + d_ocstatus_lab 
      + d_bbd_lab + d_fambrca_lab
      + d_R1alcohol_units_lab + d_R1smokingstatus_lab 
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab
      + diagage + d_R1toBC_y + yeardiag,
            family = binomial(link = "logit"))




# tidy up output
model2 <- tbl_regression(model, 
                         exponentiate = TRUE,
                         label = list(
                          d_R1alcohol_units_lab ~ "Alcohol units",
                          d_R1smokingstatus_lab ~ "Smoking status",
                          d_bmi_entry_lab ~ "BMI",
                          d_R1physmet_leis_who_lab ~ "Physical Activity (WHO cat)",
                          d_bbd_lab ~ "Benign breast disease",
                          d_fambrca_lab ~ "Family history of BC",
                          d_age_menarche_lab ~ "Age at menarche",
                          d_ocstatus_lab ~ "Oral contraceptives status",
                          d_R1menopause_lab3 ~ "Menopause",
                          d_age_meno_tr_lab ~ "Age at menopause",
                          d_R1hrt_tr_lab ~ "HRT status",
                          d_parity_lab ~ "N of parous pregnancies",
                          d_agebirth1_tr_lab ~ "Age at first birth"
                         )
                         ) %>% 
  #add_global_p() %>% 
  remove_row_type(variables = c(diagage, d_R1toBC_y, yeardiag), # remove covariates
                  type = "all")
  

#model2

# summary Logistic Regression of Model 4  -----------------------------------------

 model <- df %>%
  glm(formula = d_dmode_n ~ d_R1menopause_lab3 + d_age_meno_tr_lab 
      + d_R1hrt_tr_lab
      + d_parity_lab + d_agebirth1_tr_lab + d_age_menarche_lab + d_ocstatus_lab 
      + d_bbd_lab + d_fambrca_lab
      + d_R1alcohol_units_lab + d_R1smokingstatus_lab 
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab
      + diagage + d_R1toBC_y + yeardiag
      + d_md_qrt + d_MDtoBC_lab,
            family = binomial(link = "logit"))


# tidy up output
model4 <- tbl_regression(model, 
                         exponentiate = TRUE,
                         label = list(
                          d_R1alcohol_units_lab ~ "Alcohol units",
                          d_R1smokingstatus_lab ~ "Smoking status",
                          d_bmi_entry_lab ~ "BMI",
                          d_R1physmet_leis_who_lab ~ "Physical Activity (WHO cat)",
                          d_bbd_lab ~ "Benign breast disease",
                          d_fambrca_lab ~ "Family history of BC",
                          d_age_menarche_lab ~ "Age at menarche",
                          d_ocstatus_lab ~ "Oral contraceptives status",
                          d_R1menopause_lab3 ~ "Menopause",
                          d_age_meno_tr_lab ~ "Age at menopause",
                          d_R1hrt_tr_lab ~ "HRT status",
                          d_parity_lab ~ "N of parous pregnancies",
                          d_agebirth1_tr_lab ~ "Age at first birth"
                         )
                         ) %>% 
  #add_global_p() %>% 
  remove_row_type(variables = c(diagage, d_R1toBC_y, yeardiag), # remove covariates
                  type = "all")



# Create descriptives by mode of detection ------------------------------------
# select(d_dmode_n, d_R1alcohol_units_lab, d_R1smokingstatus_lab, d_bmi_entry_lab,  d_R1physmet_leis_who_lab, d_bbd_lab, d_fambrca_lab, d_age_menarche_lab, d_ocstatus_lab, d_R1menopause_lab3, d_age_meno_tr_lab, d_R1hrt_tr_lab, d_parity_tr_lab, d_agebirth1_tr_lab)

# note: need to rename the non-trick variables according to their trick name to align with the model, which can then allow table merge. 

# include missing 
# create df
  
descr_tab2_df <- df %>% 
  select(d_dmode, all_of(all_expl_variables), diagage, d_R1toBC_y 
         ) %>% 
  #tidyr::drop_na(all_of(all_expl_variables), diagage, d_R1toBC_y # filter out all missing to have the same n as in the model
  mutate(d_bmi_entry_lab = fct_relevel(d_bmi_entry_lab, "<18.5", "18.5-25", "25-30", "30+")
              ) %>% 
  mutate(d_age_meno_tr_lab = d_age_menopause_lab,
         d_R1hrt_tr_lab = d_R1hrtstatus_lab,
         d_agebirth1_tr_lab = d_agebirth1_lab
         ) %>% 
  select(-d_age_menopause_lab, -d_R1hrtstatus_lab, -d_agebirth1_lab, -diagage, -d_R1toBC_y)  
  
  

# glimpse(descr_model2)
# str(test)

# create gs summary table
descr_tab2 <- descr_tab2_df %>% 
  tbl_summary(by = d_dmode,
              statistic = list(
                all_continuous() ~ "{mean} ({sd})",
                all_categorical() ~ "{n} ({p}) "
                ),
              digits = all_continuous() ~ 2,
              label = list(
                          d_R1alcohol_units_lab ~ "Alcohol units",
                          d_R1smokingstatus_lab ~ "Smoking status",
                          d_bmi_entry_lab ~ "BMI",
                          d_R1physmet_leis_who_lab ~ "Physical Activity (WHO cat)",
                          d_bbd_lab ~ "Benign breast disease",
                          d_fambrca_lab ~ "Family history of BC",
                          d_age_menarche_lab ~ "Age at menarche",
                          d_ocstatus_lab ~ "Oral contraceptives status",
                          d_R1menopause_lab3 ~ "Menopause",
                          d_age_meno_tr_lab ~ "Age at menopause",
                          d_R1hrt_tr_lab ~ "HRT status",
                          d_parity_lab ~ "N of parous pregnancies",
                          d_agebirth1_tr_lab ~ "Age at first birth"
                ),
              type = list(d_bbd_lab ~ "categorical",
                          d_fambrca_lab ~ "categorical")
              )

#descr_model2
  
# merge with analysis output -------------------------------------------------


table2 <- tbl_merge(
  tbls = list(descr_tab2, model2, model4),
  tab_spanner = c("**Distribution by detection mode**", "**Model 2**", "**Model 4**")
   ) 
   
  
   

table2


# this ID is extra in the LR output from model 2 - T276E210
# this is because HRT status was missing as under 50 but it is present in trick hrt as recoded to never. - check what to do with this 
```

# Main figures

## Figures 1 to 3 - descriptives of risk factors with MD

#### Data transformation - formal statistical test of differences between categories

29/07/2024 decided to use square root transformation as data not normally distributed. Linear regression is then used to and p-values (what about estimate?) will be then added to the figures.

::: {.callout-note style="color: orange"}
## Question

Should the transformed estimate also be presented in the figures? Not sure how useful that will be as interpretation may be difficult due to the transformation. To solve this:

Option1: backtransform the estimate but complicated as need to back transform the CI or variance as well, claculations would have to be done for each chart individually. Problems with interpretation and explanations potentially

Option 2: present p-values on the figures, explain in footnote that it was done on transformed data and present linear regression results in the supplement - if anyone wants to backgtransform they can then do it themselves - more transparent
:::

#### Distribution before and after transformation

Figure for internal info only - data approximately normally distributed with more symmetry (checked with Michael and agreed this is appropriate for linear regression)

Notes: full code with more formal tests and distribution descriptives is in general_descriptives.qmd file

```{r, fig.width=15, fig.height=6}
#| echo: false

# DATA TRANSFORMATION TO TEST DIFFERENCE BETWEEN GROUPS 
# to be added to the figures
# notes full code with more formal tests and distribution descriptives is in general_descriptives.qmd file

# create md_df dataset
md_df <- df %>% 
  filter(d_md_avail == 1)

# distribution of untransformed density
# polygon
original <- md_df %>% 
  ggplot(aes(x = d_md)) +
  geom_density(fill = "skyblue", color = "black", alpha = 0.5)


# transform data
md_df <- md_df %>% 
  mutate(d_md_sqrt = sqrt(d_md))

# check distribution again 
# polygon
transformed <- md_df %>% 
  ggplot(aes(x = d_md_sqrt)) +
  geom_density(fill = "skyblue", color = "black", alpha = 0.5)

original + transformed

```

::: {.callout-note style="color: orange"}
## Note

Missing categories left in for now for info but will probably need to be removed
:::

### Reproductive, lifestyle and other risk factors

```{r}
#| echo: false
#| include: false
# code dev for plots with p-values

# ttest - to check is similar to LR
t.test(d_md_sqrt ~ d_dmode, data = md_df)
 

# linear regression 
lr <- md_df %>% 
  filter(!is.na(d_dmode) 
         ) %>% 
  lm(d_md_sqrt ~ d_dmode, data = .
     ) %>% tidy() 

str(lr)

# format p-values 
lr <- lr %>% 
  format_p_values(p.value)

lr

SD_p <- lr %>% 
  filter(term == "d_dmodeSD") %>% 
  pull(p_val1)

print(SD_p)

  
# plot

plot1 <- md_df %>% 
  ggplot(aes(x = d_dmode, y = d_md, fill = d_dmode)) +
  geom_violin(alpha = 0.4) +
  geom_boxplot(width = 0.1)+
  ylim(0, 100) +
  labs(title = "Mode of detection",
       x = "",
       y = "Breast density (%)") +
  scale_x_discrete(labels = c("I" = "Interval", "SD" = "Screen detected")) +
  theme_base() +
  theme(legend.position = "none") +
 scale_fill_brewer(palette="Set2") +
  annotate("text",
           x = 1,
           y = 90,
           label = "reference",
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 2,
           y = 90,
           label = paste("p", SD_p),
           col = "black",
           size = 3.5)

#plot1
```

```{r}
#| echo: false
#| include: false
# age at menopause
# code dev for plots with p-values


# linear regression 
lr <- md_df %>% 
 filter(!is.na(d_age_menopause_lab)) %>% 
  mutate(d_age_menopause_lab = fct_relevel(d_age_menopause_lab, "Pre-menopausal", "<50", "50-54", "55+")) %>% 
  lm(d_md_sqrt ~ d_age_menopause_lab, data = .
     ) %>% tidy() 

str(lr)

# format p-values 
lr <- lr %>% 
  format_p_values(p.value)

lr

# pull p-values
meno_under50_p <- lr %>% 
  filter(term == "d_age_menopause_lab<50") %>% 
  pull(p_val1)


meno_50to54_p <- lr %>% 
  filter(term == "d_age_menopause_lab50-54") %>% 
  pull(p_val1)

meno_over55_p <- lr %>% 
  filter(term == "d_age_menopause_lab55+") %>% 
  pull(p_val1)

# Age at menopause
plot2 <- md_df %>% 
  filter(!is.na(d_age_menopause_lab)) %>% 
  mutate(d_age_menopause_lab = fct_relevel(d_age_menopause_lab, "Pre-menopausal", "<50", "50-54", "55+")) %>% 
  ggplot(aes(x = d_age_menopause_lab, y = d_md, fill = d_age_menopause_lab)) +
  geom_violin(alpha = 0.4) +
  geom_boxplot(width = 0.1)+
  ylim(0, 100) +
  labs(title = "Age at menopause",
       x = "",
       y = "Breast density (%)") +
  theme_base() +
  theme(legend.position = "none") +
  scale_fill_brewer(palette="Set2") +
  annotate("text",
           x = 1,
           y = 90,
           label = "reference",
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 2,
           y = 90,
           label = paste("p =", meno_under50_p),
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 3,
           y = 90,
           label = paste("p =", meno_50to54_p),
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 4,
           y = 90,
           label = paste("p", meno_over55_p),
           col = "black",
           size = 3.5) 

#plot2
```

```{r}
#| echo: false
#| include: false
# HRT 
# code dev for plots with p-values


# linear regression 
lr <- md_df %>% 
 filter(!is.na(d_R1hrtstatus_lab)) %>% 
  mutate(d_R1hrtstatus_lab = fct_relevel(d_R1hrtstatus_lab, "Pre-menopausal", "Never", "Former", "Current")) %>% 
  lm(d_md_sqrt ~ d_R1hrtstatus_lab, data = .
     ) %>% tidy() 

str(lr)

# format p-values 
lr <- lr %>% 
  format_p_values(p.value)

lr

# pull p-values
hrt_never_p <- lr %>% 
  filter(term == "d_R1hrtstatus_labNever") %>% 
  pull(p_val1)


hrt_former_p <- lr %>% 
  filter(term == "d_R1hrtstatus_labFormer") %>% 
  pull(p_val1)

hrt_current_p <- lr %>% 
  filter(term == "d_R1hrtstatus_labCurrent") %>% 
  pull(p_val1)


# plot

# HRT 
plot3 <- md_df %>% 
  filter(!is.na(d_R1hrtstatus_lab)) %>% 
  mutate(d_R1hrtstatus_lab = fct_relevel(d_R1hrtstatus_lab, "Pre-menopausal", "Never", "Former", "Current")) %>% 
  ggplot(aes(x = d_R1hrtstatus_lab, y = d_md, fill = d_R1hrtstatus_lab)) +
  geom_violin(alpha = 0.4) +
  geom_boxplot(width = 0.1)+
  ylim(0, 100) +
  labs(title = "HRT status",
       x = "",
       y = "Breast density (%)") +
  theme_base() +
  theme(legend.position = "none") +
  scale_fill_brewer(palette="Set2") +
  annotate("text",
           x = 1,
           y = 90,
           label = "reference",
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 2,
           y = 90,
           label = paste("p", hrt_never_p),
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 3,
           y = 90,
           label = paste("p", hrt_former_p),
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 4,
           y = 90,
           label = paste("p =", hrt_current_p),
           col = "black",
           size = 3.5) 

#plot3
```

```{r}
#| echo: false
#| include: false

# parity 

# code dev for plots with p-values


# linear regression 
lr <- md_df %>% 
 filter(!is.na(d_parity_lab)) %>% 
  lm(d_md_sqrt ~ d_parity_lab, data = .
     ) %>% tidy() 

str(lr)

# format p-values 
lr <- lr %>% 
  format_p_values(p.value)

lr

# pull p-values
parity1_p <- lr %>% 
  filter(term == "d_parity_lab1") %>% 
  pull(p_val1)


parity2_p <- lr %>% 
  filter(term == "d_parity_lab2") %>% 
  pull(p_val1)

parity3_p <- lr %>% 
  filter(term == "d_parity_lab3+") %>% 
  pull(p_val1)


# plot
plot4 <- md_df %>% 
  filter(!is.na(d_parity_lab)) %>% 
  ggplot(aes(x = d_parity_lab, y = d_md, fill = d_parity_lab)) +
  geom_violin(alpha = 0.4) +
  geom_boxplot(width = 0.1)+
  ylim(0, 100) +
  labs(title = "Number of parous pregnancies",
       x = "",
       y = "Breast density (%)") +
  theme_base() +
  theme(legend.position = "none") +
  scale_fill_brewer(palette="Set2") +
  annotate("text",
           x = 1,
           y = 90,
           label = "reference",
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 2,
           y = 90,
           label = paste("p =", parity1_p),
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 3,
           y = 90,
           label = paste("p", parity2_p),
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 4,
           y = 90,
           label = paste("p =", parity3_p),
           col = "black",
           size = 3.5) 

#plot4

```

```{r}
#| echo: false
#| include: false

# Age at first birth

# code dev for plots with p-values


# linear regression 
lr <- md_df %>% 
 filter(!is.na(d_agebirth1_lab) & d_agebirth1_lab != "Non-parous") %>% 
  lm(d_md_sqrt ~ d_agebirth1_lab, data = .
     ) %>% tidy() 

str(lr)

# format p-values 
lr <- lr %>% 
  format_p_values(p.value)

lr

# pull p-values
agebirth1_20to24_p <- lr %>% 
  filter(term == "d_agebirth1_lab20-24") %>% 
  pull(p_val1)


agebirth1_lab25to29_p <- lr %>% 
  filter(term == "d_agebirth1_lab25-29") %>% 
  pull(p_val1)

agebirth1_lab30to34_p <- lr %>% 
  filter(term == "d_agebirth1_lab30-34") %>% 
  pull(p_val1)

agebirth1_lab35_p <- lr %>% 
  filter(term == "d_agebirth1_lab35+") %>% 
  pull(p_val1)


# plot 
plot5 <- md_df %>% 
  filter(!is.na(d_agebirth1_lab) & d_agebirth1_lab != "Non-parous") %>% 
  ggplot(aes(x = d_agebirth1_lab, y = d_md, fill = d_agebirth1_lab)) +
  geom_violin(alpha = 0.4) +
  geom_boxplot(width = 0.1)+
  ylim(0, 100) +
  labs(title = "Age at first birth",
       x = "",
       y = "Breast density (%)") +
  theme_base() +
  theme(legend.position = "none") +
  scale_fill_brewer(palette="Set2") +
  annotate("text",
           x = 1,
           y = 90,
           label = "reference",
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 2,
           y = 90,
           label = paste("p =", agebirth1_20to24_p),
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 3,
           y = 90,
           label = paste("p =", agebirth1_lab25to29_p),
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 4,
           y = 90,
           label = paste("p =", agebirth1_lab30to34_p),
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 5,
           y = 90,
           label = paste("p =", agebirth1_lab35_p),
           col = "black",
           size = 3.5) 

plot5
```

```{r}
#| echo: false
#| include: false

# age at menarche

# code dev for plots with p-values


# linear regression 
lr <- md_df %>% 
 filter(!is.na(d_age_menarche_lab)) %>% 
  lm(d_md_sqrt ~ d_age_menarche_lab, data = .
     ) %>% tidy() 

str(lr)

# format p-values 
lr <- lr %>% 
  format_p_values(p.value)

lr

# pull p-values
menarche_12to13_p <- lr %>% 
  filter(term == "d_age_menarche_lab12-13") %>% 
  pull(p_val1)


menarche_14_p <- lr %>% 
  filter(term == "d_age_menarche_lab14+") %>% 
  pull(p_val1)


# plot
plot6 <- md_df %>% 
  filter(!is.na(d_age_menarche_lab)) %>% 
  ggplot(aes(x = d_age_menarche_lab, y = d_md, fill = d_age_menarche_lab)) +
  geom_violin(alpha = 0.4) +
  geom_boxplot(width = 0.1)+
  ylim(0, 100) +
  labs(title = "Age at menarche",
       x = "",
       y = "Breast density (%)") +
  theme_base() +
  theme(legend.position = "none") +
  scale_fill_brewer(palette="Set2") +
  annotate("text",
           x = 1,
           y = 90,
           label = "reference",
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 2,
           y = 90,
           label = paste("p", menarche_12to13_p),
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 3,
           y = 90,
           label = paste("p", menarche_14_p),
           col = "black",
           size = 3.5)

#plot6
```

```{r}
#| echo: false
#| include: false

# oral contraceptives 

# code dev for plots with p-values


# linear regression 
lr <- md_df %>% 
  filter(!is.na(d_ocstatus_lab)) %>% 
  lm(d_md_sqrt ~ d_ocstatus_lab, data = .
     ) %>% tidy() 

str(lr)

# format p-values 
lr <- lr %>% 
  format_p_values(p.value)

lr

# pull p-values
oc_former_p <- lr %>% 
  filter(term == "d_ocstatus_labFormer") %>% 
  pull(p_val1)


oc_current_p <- lr %>% 
  filter(term == "d_ocstatus_labCurrent") %>% 
  pull(p_val1)


# plot
plot7 <- md_df %>% 
  filter(!is.na(d_ocstatus_lab)) %>% 
  ggplot(aes(x = d_ocstatus_lab, y = d_md, fill = d_ocstatus_lab)) +
  geom_violin(alpha = 0.4) +
  geom_boxplot(width = 0.1)+
  ylim(0, 100) +
  labs(title = "Oral contraceptive status",
       x = "",
       y = "Breast density (%)") +
  theme_base() +
  theme(legend.position = "none") +
  scale_fill_brewer(palette="Set2") +
  annotate("text",
           x = 1,
           y = 90,
           label = "reference",
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 2,
           y = 90,
           label = paste("p =", oc_former_p),
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 3,
           y = 90,
           label = paste("p =", oc_current_p),
           col = "black",
           size = 3.5)

plot7
```

```{r}
#| echo: false
#| include: false

# benign breast disease

# code dev for plots with p-values


# linear regression 
lr <- md_df %>% 
  filter(!is.na(d_bbd_lab)) %>%
  lm(d_md_sqrt ~ d_bbd_lab, data = .
     ) %>% tidy() 

# format p-values 
lr <- lr %>% 
  format_p_values(p.value)

lr

# pull p-values
bbd_yes_p <- lr %>% 
  filter(term == "d_bbd_labYes") %>% 
  pull(p_val1)


# plot
plot8 <- md_df %>% 
  filter(!is.na(d_bbd_lab)) %>% 
  ggplot(aes(x = d_bbd_lab, y = d_md, fill = d_bbd_lab)) +
  geom_violin(alpha = 0.4) +
  geom_boxplot(width = 0.1)+
  ylim(0, 100) +
  labs(title = "Benign Breast Disease",
       x = "",
       y = "Breast density (%)") +
  theme_base() +
  theme(legend.position = "none") +
  scale_fill_brewer(palette="Set2") +
  annotate("text",
           x = 1,
           y = 90,
           label = "reference",
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 2,
           y = 90,
           label = paste("p", bbd_yes_p),
           col = "black",
           size = 3.5)

#plot8
```

```{r}
#| echo: false
#| include: false


# family history of breast cancer

# code dev for plots with p-values


# linear regression 
lr <- md_df %>% 
  filter(!is.na(d_fambrca_lab)) %>% 
  lm(d_md_sqrt ~ d_fambrca_lab, data = .
     ) %>% tidy() 

# format p-values 
lr <- lr %>% 
  format_p_values(p.value)

lr

# pull p-values
fambrca_yes_p <- lr %>% 
  filter(term == "d_fambrca_labYes") %>% 
  pull(p_val1)


# plot
plot9 <- md_df %>% 
  filter(!is.na(d_fambrca_lab)) %>% 
  ggplot(aes(x = d_fambrca_lab, y = d_md, fill = d_fambrca_lab)) +
  geom_violin(alpha = 0.4) +
  geom_boxplot(width = 0.1)+
  ylim(0, 100) +
  labs(title = "Family history of breast cancer",
       x = "",
       y = "Breast density (%)") +
  theme_base() +
  theme(legend.position = "none") +
  scale_fill_brewer(palette="Set2") +
  annotate("text",
           x = 1,
           y = 90,
           label = "reference",
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 2,
           y = 90,
           label = paste("p =", fambrca_yes_p),
           col = "black",
           size = 3.5)

#plot9

```

```{r}
#| echo: false
#| include: false


# alcohol 

# code dev for plots with p-values


# linear regression 
lr <- md_df %>% 
  filter(!is.na(d_R1alcohol_units_lab)) %>% 
  lm(d_md_sqrt ~ d_R1alcohol_units_lab, data = .
     ) %>% tidy() 

# format p-values 
lr <- lr %>% 
  format_p_values(p.value)

lr

# pull p-values
alc_1to9_p <- lr %>% 
  filter(term == "d_R1alcohol_units_lab1-9") %>% 
  pull(p_val1)

alc_10to19_p <- lr %>% 
  filter(term == "d_R1alcohol_units_lab10-19") %>% 
  pull(p_val1)

alc_20to29_p <- lr %>% 
  filter(term == "d_R1alcohol_units_lab20-29") %>% 
  pull(p_val1)

alc_30_p <- lr %>% 
  filter(term == "d_R1alcohol_units_lab30+") %>% 
  pull(p_val1)


# plot
plot10 <- md_df %>% 
  filter(!is.na(d_R1alcohol_units_lab)) %>% 
  ggplot(aes(x = d_R1alcohol_units_lab, y = d_md, fill = d_R1alcohol_units_lab)) +
  geom_violin(alpha = 0.4) +
  geom_boxplot(width = 0.1)+
  ylim(0, 100) +
  labs(title = "Alcohol units",
       x = "",
       y = "Breast density (%)") +
  theme_base() +
  theme(legend.position = "none") +
  scale_fill_brewer(palette="Set2") +
  annotate("text",
           x = 1,
           y = 90,
           label = "reference",
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 2,
           y = 90,
           label = paste("p =", alc_1to9_p ),
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 3,
           y = 90,
           label = paste("p =", alc_10to19_p),
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 4,
           y = 90,
           label = paste("p =", alc_20to29_p),
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 5,
           y = 90,
           label = paste("p =", alc_30_p),
           col = "black",
           size = 3.5)


#plot10
```

```{r}
#| echo: false
#| include: false

#smoking
# code dev for plots with p-values


# linear regression 
lr <- md_df %>% 
   filter(!is.na(d_R1smokingstatus_lab)) %>% 
  lm(d_md_sqrt ~ d_R1smokingstatus_lab, data = .
     ) %>% tidy() 

# format p-values 
lr <- lr %>% 
  format_p_values(p.value)

lr

# pull p-values
smok_former_p <- lr %>% 
  filter(term == "d_R1smokingstatus_labFormer") %>% 
  pull(p_val1)

smok_current_p <- lr %>% 
  filter(term == "d_R1smokingstatus_labCurrent") %>% 
  pull(p_val1)



# plot
plot11 <- md_df %>% 
  filter(!is.na(d_R1smokingstatus_lab)) %>% 
  ggplot(aes(x = d_R1smokingstatus_lab, y = d_md, fill = d_R1smokingstatus_lab)) +
  geom_violin(alpha = 0.4) +
  geom_boxplot(width = 0.1)+
  ylim(0, 100) +
  labs(title = "Smoking status",
       x = "",
       y = "Breast density (%)") +
  theme_base() +
  theme(legend.position = "none") +
  scale_fill_brewer(palette="Set2") +
  annotate("text",
           x = 1,
           y = 90,
           label = "reference",
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 2,
           y = 90,
           label = paste("p =", smok_former_p),
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 3,
           y = 90,
           label = paste("p =", smok_current_p),
           col = "black",
           size = 3.5)

plot11

```

```{r}
#| echo: false
#| include: false

# BMI
# code dev for plots with p-values


# linear regression 
lr <- md_df %>% 
   filter(!is.na(d_bmi_entry_lab)) %>% 
  mutate(d_bmi_entry_lab= fct_relevel(d_bmi_entry_lab, "<18.5", "18.5-25", "25-30", "30+")) %>% 
  lm(d_md_sqrt ~ d_bmi_entry_lab, data = .
     ) %>% tidy() 

# format p-values 
lr <- lr %>% 
  format_p_values(p.value)

lr

# pull p-values
bmi_18.5to25 <- lr %>% 
  filter(term == "d_bmi_entry_lab18.5-25") %>% 
  pull(p_val1)

bmi_25to30 <- lr %>% 
  filter(term == "d_bmi_entry_lab25-30") %>% 
  pull(p_val1)

bmi_30 <- lr %>% 
  filter(term == "d_bmi_entry_lab30+") %>% 
  pull(p_val1)


# plot
plot12 <- md_df %>% 
  filter(!is.na(d_bmi_entry_lab)) %>% 
  mutate(d_bmi_entry_lab= fct_relevel(d_bmi_entry_lab, "<18.5", "18.5-25", "25-30", "30+")) %>%
  ggplot(aes(x = d_bmi_entry_lab, y = d_md, fill = d_bmi_entry_lab)) +
  geom_violin(alpha = 0.4) +
  geom_boxplot(width = 0.1)+
  ylim(0, 100) +
  labs(title = "Body Mass Index",
       x = "",
       y = "Breast density (%)") +
  theme_base() +
  theme(legend.position = "none") +
  scale_fill_brewer(palette="Set2") +
  annotate("text",
           x = 1,
           y = 90,
           label = "reference",
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 2,
           y = 90,
           label = paste("p =", bmi_18.5to25),
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 3,
           y = 90,
           label = paste("p =", bmi_25to30),
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 4,
           y = 90,
           label = paste("p", bmi_30),
           col = "black",
           size = 3.5)

plot12
```

```{r}
#| echo: false
#| include: false

# Physical activity 

# code dev for plots with p-values


# linear regression 
lr <- md_df %>% 
   filter(!is.na(d_R1physmet_leis_who_lab)) %>%
  lm(d_md_sqrt ~ d_R1physmet_leis_who_lab, data = .
     ) %>% tidy() 

# format p-values 
lr <- lr %>% 
  format_p_values(p.value)

lr

# pull p-values
pa_9to18 <- lr %>% 
  filter(term == "d_R1physmet_leis_who_lab9-18") %>% 
  pull(p_val1)

pa_18 <- lr %>% 
  filter(term == "d_R1physmet_leis_who_lab18+") %>% 
  pull(p_val1)



# plot
plot13 <- md_df %>% 
  filter(!is.na(d_R1physmet_leis_who_lab)) %>% 
  ggplot(aes(x = d_R1physmet_leis_who_lab, y = d_md, fill = d_R1physmet_leis_who_lab)) +
  geom_violin(alpha = 0.4) +
  geom_boxplot(width = 0.1)+
  ylim(0, 100) +
  labs(title = "Physical Activity",
       x = "",
       y = "Breast density (%)") +
  theme_base() +
  theme(legend.position = "none") +
  scale_fill_brewer(palette="Set2") +
  annotate("text",
           x = 1,
           y = 90,
           label = "reference",
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 2,
           y = 90,
           label = paste("p =", pa_9to18),
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 3,
           y = 90,
           label = paste("p =", pa_18),
           col = "black",
           size = 3.5)


plot13
```

```{r, fig.width=15, fig.height=9}
#| echo: false

# reproductive factors
reproductive_plots <- (plot2 + plot3 + plot4) / (plot5 + plot6 + plot7)
reproductive_plots

```

```{r, fig.width=15, fig.height=9}
#| echo: false

# lifestyle factors
lifestyle_plots <- (plot8 + plot9 + plot10) /  (plot11 + plot12 + plot13)
lifestyle_plots
```

### Tumour characteristics

```{r}
#| echo: false
#| include: false

# Invasive status 

# code dev for plots with p-values


# linear regression 
lr <- md_df %>% 
   filter(!is.na(d_inv_status_lab)) %>% 
  lm(d_md_sqrt ~ d_inv_status_lab, data = .
     ) %>% tidy() 

# format p-values 
lr <- lr %>% 
  format_p_values(p.value)

lr

# pull p-values
inv_yes <- lr %>% 
  filter(term == "d_inv_status_labInvasive") %>% 
  pull(p_val1)




# plot
tplot1 <- md_df %>% 
  filter(!is.na(d_inv_status_lab)) %>% 
  ggplot(aes(x = d_inv_status_lab, y = d_md, fill = d_inv_status_lab)) +
  geom_violin(alpha = 0.4) +
  geom_boxplot(width = 0.1)+
  ylim(0, 100) +
  labs(title = "Invasive status",
       x = "",
       y = "Breast density (%)") +
  theme_base() +
  theme(legend.position = "none") +
  scale_fill_brewer(palette="Set2") +
  annotate("text",
           x = 1,
           y = 90,
           label = "reference",
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 2,
           y = 90,
           label = paste("p =", inv_yes),
           col = "black",
           size = 3.5)

tplot1
```

```{r}
#| echo: false
#| include: false

# Grade 

# code dev for plots with p-values


# linear regression 
lr <- md_df %>% 
   filter(!is.na(d_grade_lab)) %>% 
  lm(d_md_sqrt ~ d_grade_lab, data = .
     ) %>% tidy() 

# format p-values 
lr <- lr %>% 
  format_p_values(p.value)

lr

# pull p-values
grade2_p <- lr %>% 
  filter(term == "d_grade_lab2") %>% 
  pull(p_val1)

grade3_p <- lr %>% 
  filter(term == "d_grade_lab3") %>% 
  pull(p_val1)


# plot
tplot2 <- md_df %>% 
  filter(!is.na(d_grade_lab)) %>% 
  ggplot(aes(x = d_grade_lab, y = d_md, fill = d_grade_lab)) +
  geom_violin(alpha = 0.4) +
  geom_boxplot(width = 0.1)+
  ylim(0, 100) +
  labs(title = "Grade",
       x = "",
       y = "Breast density (%)") +
  theme_base() +
  theme(legend.position = "none") +
  scale_fill_brewer(palette="Set2") +
  annotate("text",
           x = 1,
           y = 90,
           label = "reference",
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 2,
           y = 90,
           label = paste("p =", grade2_p),
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 3,
           y = 90,
           label = paste("p =", grade3_p),
           col = "black",
           size = 3.5)

tplot2
```

```{r}
#| echo: false
#| include: false

# Morphology

# code dev for plots with p-values


# linear regression 
lr <- md_df %>% 
   filter(!is.na(d_morph4_lab)) %>%  
  lm(d_md_sqrt ~ d_morph4_lab, data = .
     ) %>% tidy() 

# format p-values 
lr <- lr %>% 
  format_p_values(p.value)

lr

# pull p-values
morph_lob_p <- lr %>% 
  filter(term == "d_morph4_labLobular") %>% 
  pull(p_val1)

morph_mixed_p <- lr %>% 
  filter(term == "d_morph4_labMixed: ductal and other") %>% 
  pull(p_val1)

morph_other_p <- lr %>% 
  filter(term == "d_morph4_labOther") %>% 
  pull(p_val1)


# plot
tplot3 <- md_df %>% 
  filter(!is.na(d_morph4_lab)) %>% 
  ggplot(aes(x = d_morph4_lab, y = d_md, fill = d_morph4_lab)) +
  geom_violin(alpha = 0.4) +
  geom_boxplot(width = 0.1)+
  ylim(0, 100) +
  labs(title = "Morphology",
       x = "",
       y = "Breast density (%)") +
  theme_base() +
  theme(legend.position = "none") +
  scale_fill_brewer(palette="Set2")  +
  annotate("text",
           x = 1,
           y = 90,
           label = "reference",
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 2,
           y = 90,
           label = paste("p =", morph_lob_p),
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 3,
           y = 90,
           label = paste("p =", morph_mixed_p),
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 4,
           y = 90,
           label = paste("p =", morph_other_p),
           col = "black",
           size = 3.5)

tplot3
```

```{r}
#| echo: false
#| include: false

# positive nodes 

# code dev for plots with p-values


# linear regression 
lr <- md_df %>% 
   filter(!is.na(d_pos_nodes_lab)) %>%   
  lm(d_md_sqrt ~ d_pos_nodes_lab, data = .
     ) %>% tidy() 

# format p-values 
lr <- lr %>% 
  format_p_values(p.value)

lr

# pull p-values
pos_nodes_p <- lr %>% 
  filter(term == "d_pos_nodes_labPositive") %>% 
  pull(p_val1)


# plot
tplot4 <- md_df %>% 
  filter(!is.na(d_pos_nodes_lab)) %>% 
  ggplot(aes(x = d_pos_nodes_lab, y = d_md, fill = d_pos_nodes_lab)) +
  geom_violin(alpha = 0.4) +
  geom_boxplot(width = 0.1)+
  ylim(0, 100) +
  labs(title = "Positive nodes",
       x = "",
       y = "Breast density (%)") +
  theme_base() +
  theme(legend.position = "none") +
  scale_fill_brewer(palette="Set2") +
  annotate("text",
           x = 1,
           y = 90,
           label = "reference",
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 2,
           y = 90,
           label = paste("p =", pos_nodes_p),
           col = "black",
           size = 3.5)

tplot4
```

```{r}
#| echo: false
#| include: false

# tumour size 

# code dev for plots with p-values


# linear regression 
lr <- md_df %>% 
   filter(!is.na(d_tumour_size_lab)) %>%    
  lm(d_md_sqrt ~ d_tumour_size_lab, data = .
     ) %>% tidy() 

# format p-values 
lr <- lr %>% 
  format_p_values(p.value)

lr

# pull p-values
tms_21to50_p <- lr %>% 
  filter(term == "d_tumour_size_lab21-50") %>% 
  pull(p_val1)

tms_over50_p <- lr %>% 
  filter(term == "d_tumour_size_lab50+") %>% 
  pull(p_val1)


# plot
tplot5 <- md_df %>% 
  filter(!is.na(d_tumour_size_lab)) %>% 
  ggplot(aes(x = d_tumour_size_lab, y = d_md, fill = d_tumour_size_lab)) +
  geom_violin(alpha = 0.4) +
  geom_boxplot(width = 0.1)+
  ylim(0, 100) +
  labs(title = "Tumour size",
       x = "",
       y = "Breast density (%)") +
  theme_base() +
  theme(legend.position = "none") +
  scale_fill_brewer(palette="Set2") +
  annotate("text",
           x = 1,
           y = 90,
           label = "reference",
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 2,
           y = 90,
           label = paste("p =", tms_21to50_p),
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 3,
           y = 90,
           label = paste("p =", tms_over50_p),
           col = "black",
           size = 3.5)

tplot5
  
```

```{r}
#| echo: false
#| include: false

# er status

# code dev for plots with p-values


# linear regression 
lr <- md_df %>% 
   filter(!is.na(d_er_status_lab)) %>%     
  lm(d_md_sqrt ~ d_er_status_lab, data = .
     ) %>% tidy() 

# format p-values 
lr <- lr %>% 
  format_p_values(p.value)

lr

# pull p-values
er_neg_p <- lr %>% 
  filter(term == "d_er_status_labNegative") %>% 
  pull(p_val1)


# plot
tplot6 <- md_df %>% 
  filter(!is.na(d_er_status_lab)) %>% 
  ggplot(aes(x = d_er_status_lab, y = d_md, fill = d_er_status_lab)) +
  geom_violin(alpha = 0.4) +
  geom_boxplot(width = 0.1)+
  ylim(0, 100) +
  labs(title = "ER status",
       x = "",
       y = "Breast density (%)") +
  theme_base() +
  theme(legend.position = "none") +
  scale_fill_brewer(palette="Set2") +
  annotate("text",
           x = 1,
           y = 90,
           label = "reference",
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 2,
           y = 90,
           label = paste("p =", er_neg_p),
           col = "black",
           size = 3.5)

tplot6
```

```{r}
#| echo: false
#| include: false

# pr status

# code dev for plots with p-values


# linear regression 
lr <- md_df %>% 
   filter(!is.na(d_pr_status_lab)) %>%   
  lm(d_md_sqrt ~ d_pr_status_lab, data = .
     ) %>% tidy() 

# format p-values 
lr <- lr %>% 
  format_p_values(p.value)

lr

# pull p-values
pr_neg_p <- lr %>% 
  filter(term == "d_pr_status_labNegative") %>% 
  pull(p_val1)


# plot
tplot7 <- md_df %>% 
  filter(!is.na(d_pr_status_lab)) %>% 
  ggplot(aes(x = d_pr_status_lab, y = d_md, fill = d_pr_status_lab)) +
  geom_violin(alpha = 0.4) +
  geom_boxplot(width = 0.1)+
  ylim(0, 100) +
  labs(title = "PR status",
       x = "",
       y = "Breast density (%)") +
  theme_base() +
  theme(legend.position = "none") +
  scale_fill_brewer(palette="Set2") +
  annotate("text",
           x = 1,
           y = 90,
           label = "reference",
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 2,
           y = 90,
           label = paste("p =", pr_neg_p),
           col = "black",
           size = 3.5)

tplot7

```

```{r}
#| echo: false
#| include: false

# her2 status

# code dev for plots with p-values


# linear regression 
lr <- md_df %>% 
   filter(!is.na(d_her2_status_lab)) %>%   
  lm(d_md_sqrt ~ d_her2_status_lab, data = .
     ) %>% tidy() 

# format p-values 
lr <- lr %>% 
  format_p_values(p.value)

lr

# pull p-values
her2_neg_p <- lr %>% 
  filter(term == "d_her2_status_labNegative") %>% 
  pull(p_val1)


# plot
tplot8 <- md_df %>% 
  filter(!is.na(d_her2_status_lab)) %>% 
  ggplot(aes(x = d_her2_status_lab, y = d_md, fill = d_her2_status_lab)) +
  geom_violin(alpha = 0.4) +
  geom_boxplot(width = 0.1)+
  ylim(0, 100) +
  labs(title = "HER2 status",
       x = "",
       y = "Breast density (%)") +
  theme_base() +
  theme(legend.position = "none") +
  scale_fill_brewer(palette="Set2") +
  annotate("text",
           x = 1,
           y = 90,
           label = "reference",
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 2,
           y = 90,
           label = paste("p =", her2_neg_p),
           col = "black",
           size = 3.5) 

tplot8
```

```{r, fig.width=19, fig.height=13}
#| echo: false

# composite plot
tcomposite_plot <- tplot1 + tplot2 + tplot3 + tplot4 + tplot5 + tplot6 + tplot7 + tplot8 + plot1 + plot_layout(ncol = 3)

tcomposite_plot


```

# Supplementary material

## ST1 - MD subset general descriptives

::: {.callout-note style="color: red"}
## Question

Do we need this table now that we have MD in the main analysis?
:::

```{r}
#| echo: false

df %>% 
  filter(d_md_avail == 1) %>% 
  select(diagage, AgeatEntry, d_R1toBC_lab, d_R1toBC_y, d_inv_status_lab, d_grade_lab, d_stage_cat, d_morph4_lab, d_pos_nodes_lab, d_tumour_size_lab, d_er_status_lab, d_pr_status_lab, d_her2_status_lab
         ) %>% 
  tbl_summary(statistic = list(
                all_continuous() ~ "{mean} ({sd})",
                all_categorical() ~ "{n} ({p}) "),
              digits = all_continuous() ~ 2,
              label = list(diagage ~ "Age at diagnosis",
                           AgeatEntry ~ "Age at recruitment",
                           d_R1toBC_y ~ "Time from recruitment to diagnosis",
                           d_R1toBC_lab ~ "Time from recruitment to diagnosis",
                          d_inv_status_lab ~ "Invasive status",
                          d_grade_lab ~ "Grade",
                          d_stage_cat ~ "Stage",
                          d_morph4_lab ~ "Morphology type",
                          d_pos_nodes_lab ~ "Positive nodes",
                          d_tumour_size_lab ~ "Tumour size",
                          d_er_status_lab ~ "ER status",
                          d_pr_status_lab ~ "PR status",
                          d_her2_status_lab ~ "HER2 status"
                          )
  )
```

## ST 2 - Density by risk factors

complementing figure 1 to 3

::: {.callout-note style="color: orange"}
## Question

Do we need this table? As mentioned in the question box under Figure 1-3 section, presenting the linear regression results might be better than replicating the chart data into the table
:::

## ST 4 - TC descriptives and association with detection mode

TC descriptives by detection mode, logistic regression of TC with detection mode (same principal as risk factor analysis but multiple exposures are TC)

::: {.callout-note style="color: orange"}
## Question

Probably better to split the descriptives of TC and the logistic regression outputs from the 3 models

headers TBC

previously we had table with TC descriptives + TC in the model only and a second table adjusting for risk factors
:::

```{r}
#| echo: false
#| warning: false

# summary Logistic Regression of Model 5  -----------------------------------------


# Adding TC trick variables
 
model <- df %>%
  glm(formula = d_dmode_n ~ 
      d_inv_status + d_grade_tr + d_morph4_tr_lab + d_pos_nodes_tr_lab 
      + d_tmsize_tr_lab + d_er_tr_lab + d_pr_tr_lab + d_her2_tr_lab, 
            family = binomial(link = "logit"))

# tidy up output
model5 <- tbl_regression(model, 
                         exponentiate = TRUE,
                         label = list(
                          d_inv_status ~ "Invasive status",
                          d_grade_tr ~ "Grade",
                          d_morph4_tr_lab ~ "Morphology type",
                          d_pos_nodes_tr_lab ~ "Positive nodes",
                          d_tmsize_tr_lab ~ "Tumour size",
                          d_er_tr_lab ~ "ER status",
                          d_pr_tr_lab ~ "PR status",
                          d_her2_tr_lab ~ "HER2 status"
                         )
                         ) 
  


# summary Logistic Regression of Model 6  -----------------------------------------



# remove columns with total and interval N - not simple with gtsummary, trying from scratch without adding the columns


# Create clean version of logistic output, without Ns

model <- df %>%
  glm(formula = d_dmode_n ~ 
      d_R1menopause_lab3 + d_age_meno_tr_lab 
      + d_R1hrt_tr_lab
      + d_parity_lab + d_agebirth1_tr_lab + d_age_menarche_lab + d_ocstatus_lab 
      + d_bbd_lab + d_fambrca_lab
      + d_R1alcohol_units_lab + d_R1smokingstatus_lab 
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab
      + diagage + d_R1toBC_y + yeardiag
      + d_inv_status + d_grade_tr + d_morph4_tr_lab + d_pos_nodes_tr_lab 
      + d_tmsize_tr_lab + d_er_tr_lab + d_pr_tr_lab + d_her2_tr_lab, 
            family = binomial(link = "logit"))

# tidy up output
model6 <- tbl_regression(model, 
                         exponentiate = TRUE,
                         label = list(
                          d_inv_status ~ "Invasive status",
                          d_grade_tr ~ "Grade",
                          d_morph4_tr_lab ~ "Morphology type",
                          d_pos_nodes_tr_lab ~ "Positive nodes",
                          d_tmsize_tr_lab ~ "Tumour size",
                          d_er_tr_lab ~ "ER status",
                          d_pr_tr_lab ~ "PR status",
                          d_her2_tr_lab ~ "HER2 status"
                         )
                         ) 


# summary Logistic Regression of Model 7  -----------------------------------------



# remove columns with total and interval N - not simple with gtsummary, trying from scratch without adding the columns


# Create clean version of logistic output, without Ns

model <- df %>%
  glm(formula = d_dmode_n ~ d_R1menopause_lab3 + d_age_meno_tr_lab 
      + d_R1hrt_tr_lab
      + d_parity_lab + d_agebirth1_tr_lab + d_age_menarche_lab + d_ocstatus_lab 
      + d_bbd_lab + d_fambrca_lab
      + d_R1alcohol_units_lab + d_R1smokingstatus_lab 
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab
      + diagage + d_R1toBC_y + yeardiag
      + d_md_qrt + d_MDtoBC_lab
      + d_inv_status + d_grade_tr + d_morph4_tr_lab + d_pos_nodes_tr_lab 
      + d_tmsize_tr_lab + d_er_tr_lab + d_pr_tr_lab + d_her2_tr_lab,
            family = binomial(link = "logit"))

# tidy up output
model7 <- tbl_regression(model, 
                         exponentiate = TRUE,
                         label = list(
                          d_inv_status ~ "Invasive status",
                          d_grade_tr ~ "Grade",
                          d_morph4_tr_lab ~ "Morphology type",
                          d_pos_nodes_tr_lab ~ "Positive nodes",
                          d_tmsize_tr_lab ~ "Tumour size",
                          d_er_tr_lab ~ "ER status",
                          d_pr_tr_lab ~ "PR status",
                          d_her2_tr_lab ~ "HER2 status"
                         )
                         ) 

# Create descriptives by mode of detection ------------------------------------
# select(d_dmode_n, d_R1alcohol_units_lab, d_R1smokingstatus_lab, d_bmi_entry_lab,  d_R1physmet_leis_who_lab, d_bbd_lab, d_fambrca_lab, d_age_menarche_lab, d_ocstatus_lab, d_R1menopause_lab3, d_age_meno_tr_lab, d_R1hrt_tr_lab, d_parity_tr_lab, d_agebirth1_tr_lab)

# note: need to rename the non-trick variables according to their trick name to align with the model, which can then allow table merge. 

# include missing 
# create df
  
descr_tc_df <- df %>% 
  select(d_dmode, all_of(tumour_char), diagage, d_R1toBC_y 
         ) %>% 
  mutate(d_grade_lab = as.character(if_else(d_grade_lab == "DCIS", NA, d_grade_lab)) # recode dcis to missing
  ) %>% 
  mutate(d_grade_lab = factor(x = d_grade_lab,
                              levels = c(1, 2, 3))
         ) %>% 
  mutate(d_grade_tr = d_grade_lab,
         d_morph4_tr_lab = d_morph4_lab,
         d_pos_nodes_tr_lab = d_pos_nodes_lab,
         d_tmsize_tr_lab = d_tumour_size_lab,
         d_er_tr_lab = d_er_status_lab,
         d_pr_tr_lab = d_pr_status_lab,
         d_her2_tr_lab = d_her2_status_lab
         ) %>% 
  select(-d_grade_lab, -d_morph4_lab, -d_pos_nodes_lab, -d_tumour_size_lab, -d_er_status_lab, -d_pr_status_lab, -d_her2_status_lab, -diagage, -d_R1toBC_y)  
  
  

# glimpse(descr_model2)
# str(test)

# create gs summary table
descr_tc <- descr_tc_df %>% 
  tbl_summary(by = d_dmode,
              statistic = list(
                all_continuous() ~ "{mean} ({sd})",
                all_categorical() ~ "{n} ({p})"
                ),
              digits = all_continuous() ~ 2,
              label = list(
                          d_inv_status ~ "Invasive status",
                          d_grade_tr ~ "Grade",
                          d_morph4_tr_lab ~ "Morphology type",
                          d_pos_nodes_tr_lab ~ "Positive nodes",
                          d_tmsize_tr_lab ~ "Tumour size",
                          d_er_tr_lab ~ "ER status",
                          d_pr_tr_lab ~ "PR status",
                          d_her2_tr_lab ~ "HER2 status"
                         )
              ) #%>% 
  #add_overall()

#descr_model10
  
# merge with analysis output -------------------------------------------------


ST3 <- tbl_merge(
  tbls = list(descr_tc, model5, model6, model7),
  tab_spanner = c("**Distribution by detection mode**", "**Model with TC only**", "**Model with risk factors**", "**MOdel with risk factors and MD**")
   ) 
   
  
   

ST3
```

```{r}
#| echo: false
#| include: false


#ST 5 - Risk factor analysis adjusted for TC

#Sensitivity analysis - default model with model adjusted for TC

# Model 3 - subset on those with TC----------------------------------------------------------------
 # include parity and meno trick variable instead of original variables

# only included alcohol status - check if we need to do the trick if including both
# including "diagage", "d_R1toBC_y" - MD variables will be included in sub analysis

#  model <- df %>%
#    tidyr::drop_na(all_of(tumour_char_tr), all_of(all_expl_variables_tr), diagage, d_R1toBC_y
#            ) %>% 
#   glm(formula = d_dmode_n ~ d_R1menopause_lab3 + d_age_meno_tr_lab 
#       + d_R1hrt_tr_lab
#       + d_parity_lab + d_agebirth1_tr_lab + d_age_menarche_lab + d_ocstatus_lab 
#       + d_bbd_lab + d_fambrca_lab
#       + d_R1alcohol_units_lab + d_R1smokingstatus_lab 
#       + d_bmi_entry_lab + d_R1physmet_leis_who_lab
#       + diagage + d_R1toBC_y,
#             family = binomial(link = "logit"))
# 
# # tidy up output
# model3 <- tbl_regression(model, 
#                          exponentiate = TRUE,
#                          label = list(
#                           d_R1alcohol_units_lab ~ "Alcohol units",
#                           d_R1smokingstatus_lab ~ "Smoking status",
#                           d_bmi_entry_lab ~ "BMI",
#                           d_R1physmet_leis_who_lab ~ "Physical Activity (WHO cat)",
#                           d_bbd_lab ~ "Benign breast disease",
#                           d_fambrca_lab ~ "Family history of BC",
#                           d_age_menarche_lab ~ "Age at menarche",
#                           d_ocstatus_lab ~ "Oral contraceptives status",
#                           d_R1menopause_lab3 ~ "Menopause",
#                           d_age_meno_tr_lab ~ "Age at menopause",
#                           d_R1hrt_tr_lab ~ "HRT status",
#                           d_parity_lab ~ "N of parous pregnancies",
#                           d_agebirth1_tr_lab ~ "Age at first birth"
#                          )
#                          ) %>% 
#   add_global_p() %>% 
#   remove_row_type(variables = c(diagage, d_R1toBC_y), # remove covariates
#                   type = "all")
# 
# 
# model3a <- tbl_regression(model, exponentiate = TRUE) %>%
#   add_n(location = "level")  %>%
#   add_nevent(location = "level") %>%
#   add_global_p() %>%
#   modify_header(stat_nevent = "**Interval N**",
#                 stat_n = "**Total N**")
# 
# 
# 
# #model3
# #model3a
# 
# # create a dataframe with broom::tidy() for forestplots 
# 
# tidy_model3 <- model %>%
#    tidy(exponentiate = TRUE, conf.int = TRUE) %>%
#    mutate(OR = estimate) %>%
#    select(-estimate) %>%
#    select(term, OR, everything())
# #tidy_model3
# 
# 
# # Model 4 ----------------------------------------------------------------
#  # include parity and meno trick variable instead of original variables
# 
# # only included alcohol status - check if we need to do the trick if including both
# # including "diagage", "d_R1toBC_y" - MD variables will be included in sub analysis
# 
# model <- df %>%
#    tidyr::drop_na(all_of(tumour_char_tr), all_of(all_expl_variables_tr), diagage, d_R1toBC_y
#            ) %>% 
#   glm(formula = d_dmode_n ~ d_R1menopause_lab3 + d_age_meno_tr_lab 
#       + d_R1hrt_tr_lab
#       + d_parity_lab + d_agebirth1_tr_lab + d_age_menarche_lab + d_ocstatus_lab 
#       + d_bbd_lab + d_fambrca_lab
#       + d_R1alcohol_units_lab + d_R1smokingstatus_lab 
#       + d_bmi_entry_lab + d_R1physmet_leis_who_lab
#       + diagage + d_R1toBC_y
#       + d_inv_status + d_grade_tr + d_morph4_tr_lab + d_pos_nodes_tr_lab 
#       + d_tmsize_tr_lab + d_er_tr_lab + d_pr_tr_lab + d_her2_tr_lab, 
#             family = binomial(link = "logit"))
# 
# model4 <- tbl_regression(model, 
#                          exponentiate = TRUE,
#                          label = list(
#                           d_R1alcohol_units_lab ~ "Alcohol units",
#                           d_R1smokingstatus_lab ~ "Smoking status",
#                           d_bmi_entry_lab ~ "BMI",
#                           d_R1physmet_leis_who_lab ~ "Physical Activity (WHO cat)",
#                           d_bbd_lab ~ "Benign breast disease",
#                           d_fambrca_lab ~ "Family history of BC",
#                           d_age_menarche_lab ~ "Age at menarche",
#                           d_ocstatus_lab ~ "Oral contraceptives status",
#                           d_R1menopause_lab3 ~ "Menopause",
#                           d_age_meno_tr_lab ~ "Age at menopause",
#                           d_R1hrt_tr_lab ~ "HRT status",
#                           d_parity_lab ~ "N of parous pregnancies",
#                           d_agebirth1_tr_lab ~ "Age at first birth",
#                           d_inv_status ~ "Invasive status",
#                           d_grade_tr ~ "Grade",
#                           d_morph4_tr_lab ~ "Morphology type",
#                           d_pos_nodes_tr_lab ~ "Positive nodes",
#                           d_tmsize_tr_lab ~ "Tumour size",
#                           d_er_tr_lab ~ "ER status",
#                           d_pr_tr_lab ~ "PR status",
#                           d_her2_tr_lab ~ "HER2 status"
#                          )
#                          ) %>% 
#   add_global_p() %>% 
#   remove_row_type(variables = c(diagage, d_R1toBC_y), # remove covariates
#                   type = "all")
# 
# # model4
# # 
#  model4a <- tbl_regression(model, exponentiate = TRUE) %>%
#   add_n(location = "level")  %>%
#    add_nevent(location = "level") %>%
#    add_global_p() %>%
#    modify_header(stat_nevent = "**Interval N**",
#                  stat_n = "**Total N**")
# 
# # model4a
# # model3a
# 
# 
# # create a dataframe with broom::tidy() for forestplots 
# 
# tidy_model4 <- model %>%
#    tidy(exponentiate = TRUE, conf.int = TRUE) %>%
#    mutate(OR = estimate) %>%
#    select(-estimate) %>%
#    select(term, OR, everything())
# 
# 
#  # Descriptives for model 4 -------------------------------------------------
# # subset on those with mammo density 
# 
# 
# descr_TCsubset_df <- df %>% 
#   select(d_dmode,  all_of(all_expl_variables_tr), all_of(all_expl_variables), all_of(tumour_char_tr), all_of(tumour_char), diagage, d_R1toBC_y) %>% 
#   tidyr::drop_na(all_of(tumour_char_tr), all_of(all_expl_variables_tr), diagage, d_R1toBC_y
#   ) %>% 
#   mutate(d_bmi_entry_lab = fct_relevel(d_bmi_entry_lab, "<18.5", "18.5-25", "25-30", "30+")
#               ) %>%
#   # recode the names of normal variables to trick to allow merge between regression tables and this descriptive table
#   mutate(d_age_meno_tr_lab = d_age_menopause_lab,
#          d_R1hrt_tr_lab = d_R1hrtstatus_lab,
#          d_agebirth1_tr_lab = d_agebirth1_lab,
#          d_grade_tr = d_grade_lab,
#          d_morph4_tr_lab = d_morph4_lab,
#          d_pos_nodes_tr_lab = d_pos_nodes_lab,
#          d_tmsize_tr_lab = d_tumour_size_lab,
#          d_er_tr_lab = d_er_status_lab,
#          d_pr_tr_lab = d_pr_status_lab,
#          d_her2_tr_lab = d_her2_status_lab
#          ) %>%
#   select(-d_age_menopause_lab, -d_R1hrtstatus_lab, -d_agebirth1_lab, -diagage, -d_R1toBC_y, -d_grade_lab, -d_morph4_lab, -d_pos_nodes_lab, -d_tumour_size_lab, -d_er_status_lab, -d_pr_status_lab, -d_her2_status_lab)
#   
#   
# 
# # glimpse(descr_MDsubset_df)
# # str(descr_MDsubset_df)
# 
# # create gs summary table
# descr_TCsubset <- descr_TCsubset_df %>% 
#   tbl_summary(by = d_dmode,
#               statistic = list(
#                 all_continuous() ~ "{mean} ({sd})",
#                 all_categorical() ~ "{n} ({p})"
#                 ),
#               digits = all_continuous() ~ 2,
#               label = list(
#                           d_R1alcohol_units_lab ~ "Alcohol units",
#                           d_R1smokingstatus_lab ~ "Smoking status",
#                           d_bmi_entry_lab ~ "BMI",
#                           d_R1physmet_leis_who_lab ~ "Physical Activity (WHO cat)",
#                           d_bbd_lab ~ "Benign breast disease",
#                           d_fambrca_lab ~ "Family history of BC",
#                           d_age_menarche_lab ~ "Age at menarche",
#                           d_ocstatus_lab ~ "Oral contraceptives status",
#                           d_R1menopause_lab3 ~ "Menopause",
#                           d_age_meno_tr_lab ~ "Age at menopause",
#                           d_R1hrt_tr_lab ~ "HRT status",
#                           d_parity_lab ~ "N of parous pregnancies",
#                           d_agebirth1_tr_lab ~ "Age at first birth",
#                           d_inv_status ~ "Invasive status",
#                           d_grade_tr ~ "Grade",
#                           d_morph4_tr_lab ~ "Morphology type",
#                           d_pos_nodes_tr_lab ~ "Positive nodes",
#                           d_tmsize_tr_lab ~ "Tumour size",
#                           d_er_tr_lab ~ "ER status",
#                           d_pr_tr_lab ~ "PR status",
#                           d_her2_tr_lab ~ "HER2 status"
#                 ),
#               type = list(d_bbd_lab ~ "categorical",
#                           d_fambrca_lab ~ "categorical")
#               ) #%>% 
#   #add_overall()
# 
# 
# #descr_TCsubset
# # check numbers against
# # model4a
# # model3a
#  
# 
# 
# # merge tables -------------------------------------------------
# 
# 
# ST3 <- tbl_merge(
#   tbls = list(descr_TCsubset, model3, model4),
#   tab_spanner = c("Subset with TC" ,"Unadjusted for TC", "Adjusted for TC")
#    ) 
#    
# 
# #ST3
# 
# 
# # adding pct difference -------------------------------------------
# 
# #  using gtsummary 
# ST3_diff <- ST3 %>% 
#   modify_table_body(
#     ~ .x %>% 
#       mutate(pct_diff = if_else(
#           is.na(estimate_2) | is.na(estimate_3),
#           "",
#           sprintf("%.1f", (estimate_3 - estimate_2) / estimate_3 * 100))
#     )
#     ) %>% 
#   modify_header(
#     pct_diff = "**Difference (%)**"
#   )
# 
# ST3_diff
```

## ST 6 - Differences in characteristics between those with and without mammographic density

this table needs additional formatting changes, labels, add p-value for difference between groups, add time from entry to dg and age at diagnosis

any other variables to add or remove?

::: {.callout-note style="color: red"}
## Question

Do we still need this table?
:::

```{r}
#| echo: false

descr_mammo_diff <- df %>% 
  select(d_dmode, d_md_avail_lab, all_of(all_expl_variables), all_of(tumour_char), diagage, d_R1toBC_y
         ) %>% 
  mutate(d_bmi_entry_lab = fct_relevel(d_bmi_entry_lab, "<18.5", "18.5-25", "25-30", "30+")
              ) %>%
  tbl_summary(by = d_md_avail_lab,
              statistic = list(
                all_continuous() ~ "{mean} ({sd})",
                all_categorical() ~ "{n} ({p})"
                ),
              digits = all_continuous() ~ 2,
              label = list(
                          # d_R1alcohol_units_lab ~ "Alcohol units",
                          # d_R1smokingstatus_lab ~ "Smoking status",
                          # d_bmi_entry_lab ~ "BMI",
                          # d_R1physmet_leis_who_lab ~ "Physical Activity (WHO cat)",
                          # d_bbd_lab ~ "Benign breast disease",
                          # d_fambrca_lab ~ "Family history of BC",
                          # d_age_menarche_lab ~ "Age at menarche",
                          # d_ocstatus_lab ~ "Oral contraceptives status",
                          # d_R1menopause_lab3 ~ "Menopause",
                          # d_age_meno_tr_lab ~ "Age at menopause",
                          #d_R1hrt_tr_lab ~ "HRT status",
                          #d_parity_lab ~ "N of parous pregnancies",
                          #d_agebirth1_tr_lab ~ "Age at first birth",
                          #d_inv_status ~ "Invasive status",
                          #d_grade_tr ~ "Grade",
                          #d_morph4_tr_lab ~ "Morphology type",
                          #d_pos_nodes_tr_lab ~ "Positive nodes",
                          #d_tmsize_tr_lab ~ "Tumour size",
                          # d_er_tr_lab ~ "ER status",
                          # d_pr_tr_lab ~ "PR status",
                          # d_her2_tr_lab ~ "HER2 status",
                          # d_inv_status ~ "Invasive status",
                          # d_grade_tr ~ "Grade",
                          # d_morph4_tr_lab ~ "Morphology type",
                          # d_pos_nodes_tr_lab ~ "Positive nodes",
                          # d_tmsize_tr_lab ~ "Tumour size",
                          # d_er_tr_lab ~ "ER status",
                          # d_pr_tr_lab ~ "PR status",
                          # d_her2_tr_lab ~ "HER2 status",
                          #d_dmode ~ "Mode of detection"
                ),
              type = list(d_bbd_lab ~ "categorical",
                          d_fambrca_lab ~ "categorical")
              )

descr_mammo_diff
```
