---
title: "Detection mode project: Tables and figures for publication"
author: "Martina Brayley"
date: last-modified
date-format: long
format: 
  html: 
    embed-resources: true
    toc: true
    toc-location: left
    number-sections: true
    theme: [flatly, style.css]
    page-layout: full
    grid:
      sidebar-width: 300px
      body-width: 2000px
      margin-width: 300px
      gutter-width: 1.5rem
editor: visual
execute:
  warning: false
---

version 5 - Using data missing as a separate category in analysis

version 6 - swapping HRT variable for HRT subtypes variable

version 7 -

version 8 - 21/01/2025 - publication edits

### Changes agreed with Montse 14/01/2025

-   Violin plots - MD by riskfactors keep unadjusted but swap individual p-value for p-trend

-   create a table for MD by riskfactors and add adjusted for age at mammo reading

-   create a crosstab for SD registry & derived dmode (expl done in analysis script - create a nice table)

-   formatting table 1 - remove missing rows estimates, but keep percentages

-   update flow chart (remove exclusions, add another level for other dmode categories, removed MD numbers)

```{r}
#| include: false
#| echo: false

# SET UP ------------------------------------------------------------------------


# execute master file - no need run it as final data now saved
#source("code/master.R")

print(getwd())

library(pacman)
p_load(here,
       broom.helpers,
       paleteer,
       ggthemes,
       nortest)

# set up
setup_path <- here("code", "1.setup_import.r")
source(setup_path)

# import data
# data_path <- here("data", "an_df.rds")
# an_df <- readRDS(data_path)

#an_df <- readRDS("Q:/SHARED/USERS/MBrayley/Screening/data/an_df.rds")


# dataset with missing as separate categories
df <- readRDS("Q:/SHARED/USERS/MBrayley/Screening/data/an_df_missing_cat.rds")

```

```{r}
#| include: false
#| echo: false

# DATA PREP ----------------------------------------------------------------------


## Store explanatory variables as an object -------------------------------

# with labels instead of categories --------------------------------------
# select categorical variables that are labelled

# exploratory vars 
all_expl_variables <- c(
                         # parity related risk factors
                        "d_R1menopause_lab3", 
                        "d_age_menopause_lab",
                        "d_age_meno_cont",
                        "d_R1hrtstatus_lab",
                        "d_R1hrt_sttyp_lab",
                        "d_parity_lab", 
                        "d_parity",
                        "d_agebirth1_lab",
                        "d_agebirth1_cont",
                        "d_age_menarche_lab", 
                        "d_age_menarche",
                        "d_ocstatus_lab",
                        "d_ocstatus2_lab",
                        "d_breastfed_lab",
                        "d_bf_dur_lab",
                        "d_bf_dur_month",
                        # other biological risk factors
                        "d_bbd_lab", 
                        "d_fambrca_lab",
                        # lifestyle risk factors
                        "d_R1alcohol_units_lab",
                        "d_R1alcohol_units",
                        "d_R1smokingstatus_lab", 
                        "d_bmi_entry_lab",
                        "d_bmi_entry_per5",
                        "d_bmi_20_lab",
                        "d_bmi_20_per5", 
                        "d_R1physmet_leis_who_lab",
                        "d_R1physmet_leisure")

all_expl_variables_tr <- c(
                            # parity related risk factors
                           "d_R1menopause_lab3", 
                           "d_age_meno_tr_lab",
                           "d_age_meno_cont",
                           "d_R1hrt_tr_lab",
                           "d_R1hrt_sttyp_tr_lab",
                           "d_parity_lab",
                           "d_parity",
                           "d_agebirth1_tr_lab",
                           "d_agebirth1_cont",
                           "d_age_menarche_lab",
                           "d_age_menarche",
                           "d_ocstatus_lab",
                           "d_ocstatus2_lab",
                           "d_breastfed_tr_lab",
                           "d_bf_dur_tr_lab",
                            # other biological risk factors
                           "d_bbd_lab", 
                           "d_fambrca_lab",
                           # lifestyle risk factors
                           
                           "d_R1alcohol_units_lab", 
                           "d_R1alcohol_units",
                           "d_R1smokingstatus_lab", 
                           "d_bmi_entry_lab",
                           "d_bmi_entry_per5",
                           "d_bmi_20_lab",
                           "d_bmi_20_per5",
                           "d_R1physmet_leis_who_lab",
                           "d_R1physmet_leisure"
                           )
  

lifestyle <- c("d_R1alcohol_units_lab", 
               "d_R1smokingstatus_lab", 
               "d_bmi_entry_lab", 
               "d_R1physmet_leis_who")
  
reproductive <- c("d_age_menarche_lab", 
                  "d_ocstatus_lab", 
                  "d_R1menopause_lab3", 
                  "d_age_menopause_lab", 
                  "d_R1hrtstatus_lab", 
                  "d_parity_lab", 
                  "d_agebirth1_lab") # do we want to add breastfeeding? If so, it would also have be in the trick model
  
orther_rf <- c("d_bbd", 
               "d_fambrca")

# mammo density excluded from the lists - special variable for subanalyses

# covariates 
all_covariates <- c(# time varying variables
                    "diagage", 
                    "d_R1toBC_y", 
                    "d_R1toMD_y", 
                    "d_MDtoBC_y",
                    # tumour characteristics
                    "d_inv_status", 
                    "d_grade_lab", 
                    "d_morph4", 
                    "d_pos_nodes_lab", 
                    "d_tumour_size", 
                    "d_er_status", 
                    "d_pr_status", 
                    "d_her2_status")  
  
tumour_char <- c("d_inv_status", 
                 "d_grade_lab", 
                 "d_morph4_lab", 
                 "d_pos_nodes_lab", 
                 "d_tumour_size_lab", 
                 "d_er_status_lab", 
                 "d_pr_status_lab", 
                 "d_her2_status_lab")

tumour_char_tr <- c("d_inv_status", 
                    "d_grade_tr", 
                    "d_morph4_tr_lab", 
                    "d_pos_nodes_tr_lab", 
                    "d_tmsize_tr_lab", 
                    "d_er_tr_lab", 
                    "d_pr_tr_lab", 
                    "d_her2_tr_lab")
  
time_vars <- c("diagage", 
               "d_R1toBC_y", 
               "d_R1toMD_y", 
               "d_MDtoBC_y",
               "d_MDtoBC_lab",
               "yeardiag")

# create dataset 
#df <- an_df %>% 
#  mutate(d_bmi_entry_lab = fct_relevel(d_bmi_entry_lab, "18.5-25", "<18.5", "25-30", "30+")
#  )

#df <- an_df

# set theme for gtsummary tables: 

theme_gtsummary_compact(set_theme = TRUE)


```

# Table and figure structure:

**Main tables**

Table 1 - general descriptives (population characteristics)

Table 2 - Risk factor descriptives by detection mode with logistic regression results from default model ( adjusted for multiple exposures and selected covariates) + model additionally adjusted for mammo density (add mediation measure)

**Main figures**

Figures 1 to 3 - comprised figure of plots describing mamographic density in relation to risk factors (detail in ST 2)

**Supplementary material**

(ST = Supplementary Table)

ST 1 - TC descriptives by detection mode with logistic regression results for TC only + full model RF adjusting for TC + fulll RFmodel adjusting for TC and density

ST 2 - difference in characteristics between those with and without mammographic density

### 22/01/2024 Updated table and figure structure

based on draft paper

**Figure 1:** Flow chart for selection of Generations Study breast cancer cases eligible for analyses (created in drawio (saved into the paper folder) and mermaid (memaid issue with formatting)

**Figure 2:** Mammographic density by detection mode (TBC)

**Figure 3:** Mammographic density across risk factor categories

**Table 1:** Tumour characteristics of interval and screen-detected breast cancers

**Table 1b:** Tumour characteristics of interval and screen-detected breast cancers with stratification by mammographic density

**Table 2:** Risk factors of interval and screen-detected breast cancers

**Table 3:** Tumour characteristics and risk factors, by detection mode - multivar models\* including a) all cases, b) cases with mammographic density data, c) adjustment for mammographic density

# Main figures and tables

## **Table 1:** Tumour characteristics of interval and screen-detected breast cancers

by detection mode and with p-value from logistic regression model with tumour characteristics in the model (from supplementary table 1) + adjusted for time variables

```{r}
#| echo: false

# create categorical for year of diagnosis
#df %>% tabyl(yeardiag)

df <- df %>% 
  mutate(yeardiag_cat = case_when(yeardiag < 2009 ~ "2004-2008",
                                  yeardiag >= 2009 & yeardiag < 2014 ~ "2009-2013",
                                  yeardiag >= 2014 & yeardiag < 2019 ~ "2014-2018"
                                  ),
         d_R1toBC_lab = fct_relevel(d_R1toBC_lab, "<3 years", "3-5 years", "6-8 years", "9-11 years", "12+ years")
         )

# compute regression model -----------------------------------
# Adding TC trick variables
 
model <- df %>%
  glm(formula = d_dmode_n ~ 
       diagage + AgeatEntry +  d_R1toBC_y + yeardiag_cat +
      d_inv_status + d_grade_tr + d_morph4_tr_lab + d_pos_nodes_tr_lab 
      + d_tmsize_tr_lab + d_er_tr_lab + d_pr_tr_lab + d_her2_tr_lab, 
            family = binomial(link = "logit"))


# tidy up output
model_t1 <- tbl_regression(model, 
                         exponentiate = TRUE,
                         label = list(
                          d_inv_status ~ "Invasive status",
                          d_grade_tr ~ "Grade",
                          d_morph4_tr_lab ~ "Morphology type",
                          d_pos_nodes_tr_lab ~ "Positive nodes",
                          d_tmsize_tr_lab ~ "Tumour size",
                          d_er_tr_lab ~ "ER status",
                          d_pr_tr_lab ~ "PR status",
                          d_her2_tr_lab ~ "HER2 status",
                          diagage ~ "Age at diagnosis",
                           AgeatEntry ~ "Age at recruitment",
                           d_R1toBC_y ~ "Time from recruitment to diagnosis",
                          yeardiag_cat ~ "Year of diagnosis"
                         )
                         ) 

#model_t1 
# descriptives ------------------------------------

# note: need to rename the non-trick variables according to their trick name to align with the model, which can then allow table merge. 

# include missing 
# create df
  
descr_t1_df <- df %>% 
  select(d_dmode, all_of(tumour_char), diagage, d_R1toBC_y, AgeatEntry, yeardiag_cat
         ) %>% 
  mutate(d_grade_tr = d_grade_lab,
         d_morph4_tr_lab = d_morph4_lab,
         d_pos_nodes_tr_lab = d_pos_nodes_lab,
         d_tmsize_tr_lab = d_tumour_size_lab,
         d_er_tr_lab = d_er_status_lab,
         d_pr_tr_lab = d_pr_status_lab,
         d_her2_tr_lab = d_her2_status_lab
         ) %>% 
  select(-d_grade_lab, -d_morph4_lab, -d_pos_nodes_lab, -d_tumour_size_lab, -d_er_status_lab, -d_pr_status_lab, -d_her2_status_lab)  
  
#str(descr_t1_df)


# create gs summary table
descr_t1 <- descr_t1_df %>% 
  # reselect to ensure the correct order
  select(d_dmode,
         diagage,
         AgeatEntry,
         d_R1toBC_y,
         yeardiag_cat,
         d_inv_status, 
         d_grade_tr, 
         d_morph4_tr_lab, 
         d_pos_nodes_tr_lab, 
         d_tmsize_tr_lab, 
         d_er_tr_lab, 
         d_pr_tr_lab, 
         d_her2_tr_lab, 
) %>% 
  tbl_summary(by = d_dmode,
              statistic = list(
                all_continuous() ~ "{mean} ({sd})",
                all_categorical() ~ "{n} ({p})"
                ),
              digits = all_continuous() ~ 2,
              label = list(
                          d_inv_status ~ "Invasive status",
                          d_grade_tr ~ "Grade",
                          d_morph4_tr_lab ~ "Morphology type",
                          d_pos_nodes_tr_lab ~ "Positive nodes",
                          d_tmsize_tr_lab ~ "Tumour size",
                          d_er_tr_lab ~ "ER status",
                          d_pr_tr_lab ~ "PR status",
                          d_her2_tr_lab ~ "HER2 status",
                          diagage ~ "Age at diagnosis",
                           AgeatEntry ~ "Age at recruitment",
                           d_R1toBC_y ~ "Time from recruitment to diagnosis",
                          yeardiag_cat ~ "Year of diagnosis"
                          
                         )
              ) 


#descr_t1



# join tables 

# merge with analysis output -------------------------------------------------


table1 <- tbl_merge(
  tbls = list(descr_t1, model_t1),
  tab_spanner = FALSE
  )
  
#table1
   
# remove unwanted columns (effect and CI) ----------------------

#show_header_names(table1) 


table1 <- table1 %>% 
  modify_column_hide(columns = c(estimate_2, ci_2)
                     ) 
#show_header_names(table1)


table1 <- table1 %>%
  modify_footnote(
    p.value_2 ~ "p-value from logistic regression multivariate model of all listed variables"
  )

table1

```

## Figures 1 to 2 - descriptives of risk factors with MD

#### Data transformation - formal statistical test of differences between categories

29/07/2024 decided to use square root transformation as data not normally distributed. Linear regression is then used to and p-values (what about estimate?) will be then added to the figures.

#### Distribution before and after transformation

Figure for internal info only - data approximately normally distributed with more symmetry (checked with Michael and agreed this is appropriate for linear regression)

Notes: full code with more formal tests and distribution descriptives is in general_descriptives.qmd file

```{r, fig.width=15, fig.height=6}
#| echo: false

# DATA TRANSFORMATION TO TEST DIFFERENCE BETWEEN GROUPS 
# to be added to the figures
# notes full code with more formal tests and distribution descriptives is in general_descriptives.qmd file

# create md_df dataset
md_df <- df %>% 
  filter(d_md_avail == 1)

# distribution of untransformed density
# polygon
original <- md_df %>% 
  ggplot(aes(x = d_md)) +
  geom_density(fill = "skyblue", color = "black", alpha = 0.5)


# transform data
md_df <- md_df %>% 
  mutate(d_md_sqrt = sqrt(d_md))

# check distribution again 
# polygon
transformed <- md_df %>% 
  ggplot(aes(x = d_md_sqrt)) +
  geom_density(fill = "skyblue", color = "black", alpha = 0.5)

original + transformed

```

::: {.callout-note style="color: orange"}
## Note

Missing categories left in for now for info but will probably need to be removed
:::

```{r}
# function to extract global p-value from linear regression 

#define function to extract overall p-value of model
extract_overall_p_from_lm <- function(my_model) {
    f <- summary(my_model)$fstatistic
    p <- pf(f[1],f[2],f[3],lower.tail=F)
    attributes(p) <- NULL
    return(p)
}

#extract overall p-value of model
# e.g. extract_overall_p_from_lm(model)


# function to format single p value (not a dataframe)
format_single_p_value <- function(p_value) {
  formatted_p <- case_when(
    p_value < 0.001 ~ "<0.001",
    p_value >= 0.001 & p_value < 0.01 ~ as.character(round(p_value, 3)),
    round(p_value, 2) == 1 ~ "0.99",
    TRUE ~ as.character(format(round(p_value, 2), nsmall = 1))
  )
  return(formatted_p)
}

```

### Reproductive, lifestyle and other risk factors

```{r}
#| echo: false
#| include: false
# code dev for plots with p-values

# ADAPTED - TEST -----------------------------------------------------

# ttest - to check is similar to LR
t.test(d_md_sqrt ~ d_dmode, data = md_df)
 

# linear regression without tidy output
lr <- md_df %>% 
  filter(!is.na(d_dmode) 
         ) %>% 
  lm(d_md_sqrt ~ d_dmode, data = .
     )
summary(lr)

global_p <- extract_overall_p_from_lm(lr)

global_p

global_p <- format_single_p_value(global_p)

global_p


# plot

plot1 <- md_df %>% 
  ggplot(aes(x = d_dmode, y = d_md, fill = d_dmode)) +
  geom_violin(alpha = 0.4) +
  geom_boxplot(width = 0.1)+
  ylim(0, 100) +
  labs(title = "Mode of detection",
       x = "",
       y = "Breast density (%)") +
  scale_x_discrete(labels = c("I" = "Interval", "SD" = "Screen detected")) +
  theme_base() +
  theme(legend.position = "none") +
 scale_fill_brewer(palette="Set2") +
  annotate("text",
           x = 1,
           y = 90,
           label = "reference",
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 2,
           y = 90,
           label = paste("p-trend", global_p),
           col = "black",
           size = 3.5)

plot1
```

```{r}
#| echo: false
#| include: false

# WRAP INTO A FUNCTION 

create_violin_plot <- function(data, response_var, explanatory_var, 
                               y_label, title, category_labels, 
                               annotation_pos) {
  # Fit the linear model
  formula <- as.formula(paste(response_var, "~", explanatory_var))
  lm_model <- lm(formula, data = data)
  
  # Extract and format the overall p-value
  global_p <- extract_overall_p_from_lm(lm_model)
  formatted_p <- format_single_p_value(global_p)
  
  # Create the plot
  plot <- data %>%
    ggplot(aes(x = .data[[explanatory_var]], y = .data[[response_var]], fill = .data[[explanatory_var]])) +
    geom_violin(alpha = 0.4) +
    geom_boxplot(width = 0.1) +
    ylim(0, 100) +
    labs(title = title, x = "", y = y_label) +
    scale_x_discrete(labels = category_labels) +
    theme_base() +
    theme(legend.position = "none") +
    scale_fill_brewer(palette = "Set2") +
    annotate("text",
             x = annotation_pos$ref_x,
             y = annotation_pos$ref_y,
             label = "reference",
             col = "black",
             size = 3.5) +
    annotate("text",
             x = annotation_pos$p_x,
             y = annotation_pos$p_y,
             label = paste("p-trend", formatted_p),
             col = "black",
             size = 3.5)
  
  return(plot)
}


# TEST FUNCTION 

# Define annotation positions
annotation_positions <- list(ref_x = 1, ref_y = 90, p_x = 2, p_y = 90)

# Define category labels
category_labels <- c("I" = "Interval", "SD" = "Screen detected")

# Call the function
plot1 <- create_violin_plot(
  data = md_df,
  response_var = "d_md",
  explanatory_var = "d_dmode",
  y_label = "Breast density (%)",
  title = "Mode of Detection",
  category_labels = category_labels,
  annotation_pos = annotation_positions
)

# Print the plot
plot1

```

```{r}
#| echo: false
#| include: false
# code dev for plots with p-values

# ttest - to check is similar to LR
t.test(d_md_sqrt ~ d_dmode, data = md_df)

# linear regression 
lr <- md_df %>% 
  filter(!is.na(d_dmode) 
         ) %>% 
  lm(d_md_sqrt ~ d_dmode, data = .
     ) %>% tidy() 

lr

str(lr)

# format p-values 
lr <- lr %>% 
  format_p_values(p.value)

summary(lr)

SD_p <- lr %>% 
  filter(term == "d_dmodeSD") %>% 
  pull(p_val1)

print(SD_p)

  
# plot

plot1 <- md_df %>% 
  ggplot(aes(x = d_dmode, y = d_md, fill = d_dmode)) +
  geom_violin(alpha = 0.4) +
  geom_boxplot(width = 0.1)+
  ylim(0, 100) +
  labs(title = "Mode of detection",
       x = "",
       y = "Breast density (%)") +
  scale_x_discrete(labels = c("I" = "Interval", "SD" = "Screen detected")) +
  theme_base() +
  theme(legend.position = "none") +
 scale_fill_brewer(palette="Set2") +
  annotate("text",
           x = 1,
           y = 90,
           label = "reference",
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 2,
           y = 90,
           label = paste("p", SD_p),
           col = "black",
           size = 3.5)

plot1
```

```{r}
#| echo: false
#| include: false
# code dev for plots with p-values


# linear regression 
lr <- md_df %>% 
  filter(d_R1menopause_lab3 != "Missing"
         ) %>% 
  lm(d_md_sqrt ~ d_R1menopause_lab3, data = .
     ) %>% tidy() 

str(lr)

# format p-values 
lr <- lr %>% 
  format_p_values(p.value)

lr

# pull p values
post_meno_p <- lr %>% 
  filter(term == "d_R1menopause_lab3Postmenopausal") %>% 
  pull(p_val1)

print(post_meno_p)

  
# plot

plot2 <- md_df %>% 
  ggplot(aes(x = d_R1menopause_lab3, y = d_md, fill = d_R1menopause_lab3)) +
  geom_violin(alpha = 0.4) +
  geom_boxplot(width = 0.1)+
  ylim(0, 100) +
  labs(title = "Menopausal status",
       x = "",
       y = "Breast density (%)") +
  theme_base() +
  theme(legend.position = "none") +
 scale_fill_brewer(palette="Set2") +
  annotate("text",
           x = 1,
           y = 90,
           label = "reference",
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 2,
           y = 90,
           label = paste("p", post_meno_p),
           col = "black",
           size = 3.5)

#plot2
```

```{r}
#| echo: false
#| include: false
# age at menopause
# code dev for plots with p-values

#md_df %>% tabyl(d_age_menopause_lab)

# linear regression 
lr <- md_df %>% 
 filter(d_age_menopause_lab2 != "Missing") %>% 
  mutate(d_age_menopause_lab2 = fct_relevel(d_age_menopause_lab2,  "Pre-menopausal","<50", "50-54", "55+")) %>% 
  lm(d_md_sqrt ~ d_age_menopause_lab2, data = .
     ) %>% tidy() 

lr

summary(lr)

str(lr)

# format p-values 
lr <- lr %>% 
  format_p_values(p.value)

lr

# pull p-values

meno_premeno_p <- lr %>% 
  filter(term == "d_age_menopause_lab2Pre-menopausal") %>% 
  pull(p_val1)

meno_under50_p <- lr %>% 
  filter(term == "d_age_menopause_lab2<50") %>% 
  pull(p_val1)


meno_50to54_p <- lr %>% 
  filter(term == "d_age_menopause_lab250-54") %>% 
  pull(p_val1)

meno_over55_p <- lr %>% 
  filter(term == "d_age_menopause_lab255+") %>% 
  pull(p_val1)

# Age at menopause
plot3 <- md_df %>% 
  filter(d_age_menopause_lab2 != "Missing") %>%  
  mutate(d_age_menopause_lab2 = fct_relevel(d_age_menopause_lab2, "Pre-menopausal", "<50", "50-54", "55+")) %>% 
  ggplot(aes(x = d_age_menopause_lab2, y = d_md, fill = d_age_menopause_lab2)) +
  geom_violin(alpha = 0.4) +
  geom_boxplot(width = 0.1)+
  ylim(0, 100) +
  labs(title = "Age at menopause",
       x = "",
       y = "Breast density (%)") +
  theme_base() +
  theme(legend.position = "none") +
  scale_fill_brewer(palette="Set2") +
  annotate("text",
           x = 1,
           y = 90,
           label = "reference",
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 2,
           y = 90,
           label = paste("p =", meno_under50_p),
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 3,
           y = 90,
           label = paste("p =", meno_50to54_p),
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 4,
           y = 90,
           label = paste("p =", meno_over55_p),
           col = "black",
           size = 3.5) 

#plot3
```

```{r}
#| echo: false
#| include: false
# HRT 
# code dev for plots with p-values


# linear regression 
lr <- md_df %>% 
 filter(d_R1hrt_sttyp_lab != "Missing") %>% 
  mutate(d_R1hrt_sttyp_lab = fct_relevel(d_R1hrt_sttyp_lab, "Never", "Former", "Current: Estrogen only", "Current: Estrogen and progestogen", "Current: Other HRT")) %>% 
  lm(d_md_sqrt ~ d_R1hrt_sttyp_lab, data = .
     ) %>% tidy() 

str(lr)

# format p-values 
lr <- lr %>% 
  format_p_values(p.value)

lr

# pull p-values
hrt_never_p <- lr %>% 
  filter(term == "d_R1hrt_sttyp_labNever") %>% 
  pull(p_val1)


hrt_former_p <- lr %>% 
  filter(term == "d_R1hrt_sttyp_labFormer") %>% 
  pull(p_val1)


hrt_current_e_p <- lr %>% 
  filter(term == "d_R1hrt_sttyp_labCurrent: Estrogen only") %>% 
  pull(p_val1)

hrt_current_ep_p <- lr %>% 
  filter(term == "d_R1hrt_sttyp_labCurrent: Estrogen and progestogen") %>% 
  pull(p_val1)

hrt_current_o_p <- lr %>% 
  filter(term == "d_R1hrt_sttyp_labCurrent: Other HRT") %>% 
  pull(p_val1)


# plot

# HRT 
plot4 <- md_df %>% 
  filter(d_R1hrt_sttyp_lab != "Missing") %>% 
  mutate(d_R1hrt_sttyp_lab = fct_relevel(d_R1hrt_sttyp_lab, "Never", "Former", "Current: Estrogen only", "Current: Estrogen and progestogen", "Current: Other HRT")) %>% 
  ggplot(aes(x = d_R1hrt_sttyp_lab, y = d_md, fill = d_R1hrt_sttyp_lab)) +
  geom_violin(alpha = 0.4) +
  geom_boxplot(width = 0.1)+
  ylim(0, 100) +
  labs(title = "HRT status",
       x = "",
       y = "Breast density (%)") +
  theme_base() +
  theme(legend.position = "none") +
  scale_fill_brewer(palette="Set2") +
  annotate("text",
           x = 1,
           y = 90,
           label = "reference",
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 2,
           y = 90,
           label = paste("p", hrt_former_p),
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 3,
           y = 90,
           label = paste("p =", hrt_current_e_p),
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 4,
           y = 90,
           label = paste("p =", hrt_current_ep_p),
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 5,
           y = 90,
           label = paste("p =", hrt_current_o_p),
           col = "black",
           size = 3.5) +
  theme(axis.text.x = element_text(angle = 20, hjust = 1))
  

#plot4
```

```{r}
#| echo: false
#| include: false

# parity 

# code dev for plots with p-values


# linear regression 
lr <- md_df %>% 
 filter(d_parity_lab != "Missing") %>% 
  lm(d_md_sqrt ~ d_parity_lab, data = .
     ) %>% tidy() 

str(lr)

# format p-values 
lr <- lr %>% 
  format_p_values(p.value)

lr

# pull p-values
parity1_p <- lr %>% 
  filter(term == "d_parity_lab1") %>% 
  pull(p_val1)


parity2_p <- lr %>% 
  filter(term == "d_parity_lab2") %>% 
  pull(p_val1)

parity3_p <- lr %>% 
  filter(term == "d_parity_lab3+") %>% 
  pull(p_val1)


# plot
plot5 <- md_df %>% 
  filter(d_parity_lab != "Missing") %>% 
  ggplot(aes(x = d_parity_lab, y = d_md, fill = d_parity_lab)) +
  geom_violin(alpha = 0.4) +
  geom_boxplot(width = 0.1)+
  ylim(0, 100) +
  labs(title = "Number of parous pregnancies",
       x = "",
       y = "Breast density (%)") +
  theme_base() +
  theme(legend.position = "none") +
  scale_fill_brewer(palette="Set2") +
  annotate("text",
           x = 1,
           y = 90,
           label = "reference",
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 2,
           y = 90,
           label = paste("p =", parity1_p),
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 3,
           y = 90,
           label = paste("p", parity2_p),
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 4,
           y = 90,
           label = paste("p =", parity3_p),
           col = "black",
           size = 3.5) 

#plot5

```

```{r}
#| echo: false
#| include: false

# Age at first birth

# code dev for plots with p-values


# linear regression 
lr <- md_df %>% 
 filter(d_agebirth1_lab != "Missing") %>% 
  lm(d_md_sqrt ~ d_agebirth1_lab, data = .
     ) %>% tidy() 

str(lr)

# format p-values 
lr <- lr %>% 
  format_p_values(p.value)

lr

# pull p-values
agebirth1_20to24_p <- lr %>% 
  filter(term == "d_agebirth1_lab20-24") %>% 
  pull(p_val1)


agebirth1_lab25to29_p <- lr %>% 
  filter(term == "d_agebirth1_lab25-29") %>% 
  pull(p_val1)

agebirth1_lab30to34_p <- lr %>% 
  filter(term == "d_agebirth1_lab30-34") %>% 
  pull(p_val1)

agebirth1_lab35_p <- lr %>% 
  filter(term == "d_agebirth1_lab35+") %>% 
  pull(p_val1)


# plot 
plot6 <- md_df %>% 
  filter(d_agebirth1_lab != "Missing") %>% 
  ggplot(aes(x = d_agebirth1_lab, y = d_md, fill = d_agebirth1_lab)) +
  geom_violin(alpha = 0.4) +
  geom_boxplot(width = 0.1)+
  ylim(0, 100) +
  labs(title = "Age at first birth",
       x = "",
       y = "Breast density (%)") +
  theme_base() +
  theme(legend.position = "none") +
  scale_fill_brewer(palette="Set2") +
  annotate("text",
           x = 1,
           y = 90,
           label = "reference",
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 2,
           y = 90,
           label = paste("p =", agebirth1_20to24_p),
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 3,
           y = 90,
           label = paste("p =", agebirth1_lab25to29_p),
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 4,
           y = 90,
           label = paste("p =", agebirth1_lab30to34_p),
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 5,
           y = 90,
           label = paste("p =", agebirth1_lab35_p),
           col = "black",
           size = 3.5) 

#plot6
```

```{r}
#| echo: false
#| include: false

# age at menarche

# code dev for plots with p-values


# linear regression 
lr <- md_df %>% 
 filter(d_age_menarche_lab != "Missing") %>% 
  lm(d_md_sqrt ~ d_age_menarche_lab, data = .
     ) %>% tidy() 

str(lr)

# format p-values 
lr <- lr %>% 
  format_p_values(p.value)

lr

# pull p-values
menarche_12to13_p <- lr %>% 
  filter(term == "d_age_menarche_lab12-13") %>% 
  pull(p_val1)


menarche_14_p <- lr %>% 
  filter(term == "d_age_menarche_lab14+") %>% 
  pull(p_val1)


# plot
plot7 <- md_df %>% 
  filter(d_age_menarche_lab != "Missing") %>% 
  ggplot(aes(x = d_age_menarche_lab, y = d_md, fill = d_age_menarche_lab)) +
  geom_violin(alpha = 0.4) +
  geom_boxplot(width = 0.1)+
  ylim(0, 100) +
  labs(title = "Age at menarche",
       x = "",
       y = "Breast density (%)") +
  theme_base() +
  theme(legend.position = "none") +
  scale_fill_brewer(palette="Set2") +
  annotate("text",
           x = 1,
           y = 90,
           label = "reference",
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 2,
           y = 90,
           label = paste("p", menarche_12to13_p),
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 3,
           y = 90,
           label = paste("p", menarche_14_p),
           col = "black",
           size = 3.5)

#plot7


```

```{r}
#| echo: false
#| include: false

# oral contraceptives 

# code dev for plots with p-values


# linear regression 
lr <- md_df %>% 
  filter(d_ocstatus_lab != "Missing") %>% 
  lm(d_md_sqrt ~ d_ocstatus_lab, data = .
     ) %>% tidy() 

str(lr)

# format p-values 
lr <- lr %>% 
  format_p_values(p.value)

lr

# pull p-values
oc_former_p <- lr %>% 
  filter(term == "d_ocstatus_labFormer") %>% 
  pull(p_val1)


oc_current_p <- lr %>% 
  filter(term == "d_ocstatus_labCurrent") %>% 
  pull(p_val1)


# plot
plot8 <- md_df %>% 
  filter(d_ocstatus_lab != "Missing") %>% 
  ggplot(aes(x = d_ocstatus_lab, y = d_md, fill = d_ocstatus_lab)) +
  geom_violin(alpha = 0.4) +
  geom_boxplot(width = 0.1)+
  ylim(0, 100) +
  labs(title = "Oral contraceptive status",
       x = "",
       y = "Breast density (%)") +
  theme_base() +
  theme(legend.position = "none") +
  scale_fill_brewer(palette="Set2") +
  annotate("text",
           x = 1,
           y = 90,
           label = "reference",
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 2,
           y = 90,
           label = paste("p =", oc_former_p),
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 3,
           y = 90,
           label = paste("p =", oc_current_p),
           col = "black",
           size = 3.5)

#plot8
```

```{r}
#| echo: false
#| include: false

# benign breast disease

# code dev for plots with p-values


# linear regression 
lr <- md_df %>% 
  filter(d_bbd_lab != "Missing") %>%
  lm(d_md_sqrt ~ d_bbd_lab, data = .
     ) %>% tidy() 

# format p-values 
lr <- lr %>% 
  format_p_values(p.value)

lr

# pull p-values
bbd_yes_p <- lr %>% 
  filter(term == "d_bbd_labYes") %>% 
  pull(p_val1)


# plot
plot9 <- md_df %>% 
  filter(d_bbd_lab != "Missing") %>% 
  ggplot(aes(x = d_bbd_lab, y = d_md, fill = d_bbd_lab)) +
  geom_violin(alpha = 0.4) +
  geom_boxplot(width = 0.1)+
  ylim(0, 100) +
  labs(title = "Benign Breast Disease",
       x = "",
       y = "Breast density (%)") +
  theme_base() +
  theme(legend.position = "none") +
  scale_fill_brewer(palette="Set2") +
  annotate("text",
           x = 1,
           y = 90,
           label = "reference",
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 2,
           y = 90,
           label = paste("p", bbd_yes_p),
           col = "black",
           size = 3.5)

#plot9
```

```{r}
#| echo: false
#| include: false


# family history of breast cancer

# code dev for plots with p-values


# linear regression 
lr <- md_df %>% 
  filter(d_fambrca_lab != "Missing") %>% 
  lm(d_md_sqrt ~ d_fambrca_lab, data = .
     ) %>% tidy() 

# format p-values 
lr <- lr %>% 
  format_p_values(p.value)

lr

# pull p-values
fambrca_yes_p <- lr %>% 
  filter(term == "d_fambrca_labYes") %>% 
  pull(p_val1)


# plot
plot10 <- md_df %>% 
   filter(d_fambrca_lab != "Missing") %>%  
  ggplot(aes(x = d_fambrca_lab, y = d_md, fill = d_fambrca_lab)) +
  geom_violin(alpha = 0.4) +
  geom_boxplot(width = 0.1)+
  ylim(0, 100) +
  labs(title = "Family history of breast cancer",
       x = "",
       y = "Breast density (%)") +
  theme_base() +
  theme(legend.position = "none") +
  scale_fill_brewer(palette="Set2") +
  annotate("text",
           x = 1,
           y = 90,
           label = "reference",
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 2,
           y = 90,
           label = paste("p =", fambrca_yes_p),
           col = "black",
           size = 3.5)

#plot10

```

```{r}
#| echo: false
#| include: false


# alcohol 

# code dev for plots with p-values


# linear regression 
lr <- md_df %>% 
  filter(d_R1alcohol_units_lab != "Missing") %>% 
  lm(d_md_sqrt ~ d_R1alcohol_units_lab, data = .
     ) %>% tidy() 

# format p-values 
lr <- lr %>% 
  format_p_values(p.value)

lr

# pull p-values
alc_1to9_p <- lr %>% 
  filter(term == "d_R1alcohol_units_lab1-9") %>% 
  pull(p_val1)

alc_10to19_p <- lr %>% 
  filter(term == "d_R1alcohol_units_lab10-19") %>% 
  pull(p_val1)

alc_20to29_p <- lr %>% 
  filter(term == "d_R1alcohol_units_lab20-29") %>% 
  pull(p_val1)

alc_30_p <- lr %>% 
  filter(term == "d_R1alcohol_units_lab30+") %>% 
  pull(p_val1)


# plot
plot11 <- md_df %>% 
  filter(d_R1alcohol_units_lab != "Missing") %>% 
  ggplot(aes(x = d_R1alcohol_units_lab, y = d_md, fill = d_R1alcohol_units_lab)) +
  geom_violin(alpha = 0.4) +
  geom_boxplot(width = 0.1)+
  ylim(0, 100) +
  labs(title = "Alcohol units",
       x = "",
       y = "Breast density (%)") +
  theme_base() +
  theme(legend.position = "none") +
  scale_fill_brewer(palette="Set2") +
  annotate("text",
           x = 1,
           y = 90,
           label = "reference",
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 2,
           y = 90,
           label = paste("p =", alc_1to9_p ),
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 3,
           y = 90,
           label = paste("p =", alc_10to19_p),
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 4,
           y = 90,
           label = paste("p =", alc_20to29_p),
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 5,
           y = 90,
           label = paste("p =", alc_30_p),
           col = "black",
           size = 3.5)


#plot11
```

```{r}
#| echo: false
#| include: false

#smoking
# code dev for plots with p-values


# linear regression 
lr <- md_df %>% 
   filter(d_R1smokingstatus_lab != "Missing") %>% 
  lm(d_md_sqrt ~ d_R1smokingstatus_lab, data = .
     ) %>% tidy() 

# format p-values 
lr <- lr %>% 
  format_p_values(p.value)

lr

# pull p-values
smok_former_p <- lr %>% 
  filter(term == "d_R1smokingstatus_labFormer") %>% 
  pull(p_val1)

smok_current_p <- lr %>% 
  filter(term == "d_R1smokingstatus_labCurrent") %>% 
  pull(p_val1)



# plot
plot12 <- md_df %>% 
  filter(d_R1smokingstatus_lab != "Missing") %>% 
  ggplot(aes(x = d_R1smokingstatus_lab, y = d_md, fill = d_R1smokingstatus_lab)) +
  geom_violin(alpha = 0.4) +
  geom_boxplot(width = 0.1)+
  ylim(0, 100) +
  labs(title = "Smoking status",
       x = "",
       y = "Breast density (%)") +
  theme_base() +
  theme(legend.position = "none") +
  scale_fill_brewer(palette="Set2") +
  annotate("text",
           x = 1,
           y = 90,
           label = "reference",
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 2,
           y = 90,
           label = paste("p =", smok_former_p),
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 3,
           y = 90,
           label = paste("p =", smok_current_p),
           col = "black",
           size = 3.5)

#plot12

```

```{r}
#| echo: false
#| include: false

# BMI
# code dev for plots with p-values


# linear regression 
lr <- md_df %>% 
   filter(d_bmi_entry_lab != "Missing") %>% 
  mutate(d_bmi_entry_lab= fct_relevel(d_bmi_entry_lab, "<18.5", "18.5-25", "25-30", "30+")) %>% 
  lm(d_md_sqrt ~ d_bmi_entry_lab, data = .
     ) %>% tidy() 

# format p-values 
lr <- lr %>% 
  format_p_values(p.value)

lr

# pull p-values
bmi_18.5to25 <- lr %>% 
  filter(term == "d_bmi_entry_lab18.5-25") %>% 
  pull(p_val1)

bmi_25to30 <- lr %>% 
  filter(term == "d_bmi_entry_lab25-30") %>% 
  pull(p_val1)

bmi_30 <- lr %>% 
  filter(term == "d_bmi_entry_lab30+") %>% 
  pull(p_val1)


# plot
plot13 <- md_df %>% 
   filter(d_bmi_entry_lab != "Missing") %>% 
  mutate(d_bmi_entry_lab= fct_relevel(d_bmi_entry_lab, "<18.5", "18.5-25", "25-30", "30+")) %>%
  ggplot(aes(x = d_bmi_entry_lab, y = d_md, fill = d_bmi_entry_lab)) +
  geom_violin(alpha = 0.4) +
  geom_boxplot(width = 0.1)+
  ylim(0, 100) +
  labs(title = "Body Mass Index at entry",
       x = "",
       y = "Breast density (%)") +
  theme_base() +
  theme(legend.position = "none") +
  scale_fill_brewer(palette="Set2") +
  annotate("text",
           x = 1,
           y = 90,
           label = "reference",
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 2,
           y = 90,
           label = paste("p =", bmi_18.5to25),
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 3,
           y = 90,
           label = paste("p =", bmi_25to30),
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 4,
           y = 90,
           label = paste("p", bmi_30),
           col = "black",
           size = 3.5)

#plot13
```

```{r}
#| echo: false
#| include: false

# Physical activity 

# code dev for plots with p-values


# linear regression 
lr <- md_df %>% 
   filter(d_R1physmet_leis_who_lab != "Missing") %>%
  lm(d_md_sqrt ~ d_R1physmet_leis_who_lab, data = .
     ) %>% tidy() 

# format p-values 
lr <- lr %>% 
  format_p_values(p.value)

lr

# pull p-values
pa_9to18 <- lr %>% 
  filter(term == "d_R1physmet_leis_who_lab9-18") %>% 
  pull(p_val1)

pa_18 <- lr %>% 
  filter(term == "d_R1physmet_leis_who_lab18+") %>% 
  pull(p_val1)



# plot
plot14 <- md_df %>% 
  filter(d_R1physmet_leis_who_lab != "Missing") %>%
  ggplot(aes(x = d_R1physmet_leis_who_lab, y = d_md, fill = d_R1physmet_leis_who_lab)) +
  geom_violin(alpha = 0.4) +
  geom_boxplot(width = 0.1)+
  ylim(0, 100) +
  labs(title = "Physical Activity (MET-h/w)",
       x = "",
       y = "Breast density (%)") +
  theme_base() +
  theme(legend.position = "none") +
  scale_fill_brewer(palette="Set2") +
  annotate("text",
           x = 1,
           y = 90,
           label = "reference",
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 2,
           y = 90,
           label = paste("p =", pa_9to18),
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 3,
           y = 90,
           label = paste("p =", pa_18),
           col = "black",
           size = 3.5)


#plot14
```

```{r}
#| echo: false
#| include: false

# Duration of breastfeeding

# code dev for plots with p-values


# linear regression 
lr <- md_df %>% 
 filter(d_bf_dur_lab != "Missing") %>% 
  lm(d_md_sqrt ~ d_bf_dur_lab, data = .
     ) %>% tidy() 

str(lr)

# format p-values 
lr <- lr %>% 
  format_p_values(p.value)

lr

# pull p-values
bf_dur_6m_p <- lr %>% 
  filter(term == "d_bf_dur_lab<6 months") %>% 
  pull(p_val1)

bf_dur_6to12_p <- lr %>% 
  filter(term == "d_bf_dur_lab6-12 months") %>% 
  pull(p_val1)

bf_dur_12to24_p <- lr %>% 
  filter(term == "d_bf_dur_lab12-24 months") %>% 
  pull(p_val1)

bf_dur_24m_p <- lr %>% 
  filter(term == "d_bf_dur_lab24+ months") %>% 
  pull(p_val1)


# plot 
plot15 <- md_df %>% 
  filter(d_bf_dur_lab != "Missing") %>% 
  ggplot(aes(x = d_bf_dur_lab, y = d_md, fill = d_bf_dur_lab)) +
  geom_violin(alpha = 0.4) +
  geom_boxplot(width = 0.1)+
  ylim(0, 100) +
  labs(title = "Breastfeeding duration",
       x = "",
       y = "Breast density (%)") +
  theme_base() +
  theme(legend.position = "none") +
  scale_fill_brewer(palette="Set2") +
  annotate("text",
           x = 1,
           y = 90,
           label = "reference",
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 2,
           y = 90,
           label = paste("p =", bf_dur_6m_p),
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 3,
           y = 90,
           label = paste("p =", bf_dur_6to12_p),
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 4,
           y = 90,
           label = paste("p =", bf_dur_12to24_p),
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 5,
           y = 90,
           label = paste("p =", bf_dur_24m_p),
           col = "black",
           size = 3.5) +
  theme(axis.text.x = element_text(angle = 20, hjust = 1))

#plot15
```

```{r}
#| echo: false
#| include: false

# BMI age 20
# code dev for plots with p-values


# linear regression 
lr <- md_df %>% 
   filter(d_bmi_20_lab != "Missing") %>% 
  mutate(d_bmi_20_lab= fct_relevel(d_bmi_20_lab, "<18.5", "18.5-25", "25-30", "30+")) %>% 
  lm(d_md_sqrt ~ d_bmi_20_lab, data = .
     ) %>% tidy() 

# format p-values 
lr <- lr %>% 
  format_p_values(p.value)

lr

# pull p-values
bmi20_18.5to25 <- lr %>% 
  filter(term == "d_bmi_20_lab18.5-25") %>% 
  pull(p_val1)

bmi20_25to30 <- lr %>% 
  filter(term == "d_bmi_20_lab25-30") %>% 
  pull(p_val1)

bmi20_30 <- lr %>% 
  filter(term == "d_bmi_20_lab30+") %>% 
  pull(p_val1)


# plot
plot16 <- md_df %>% 
   filter(d_bmi_20_lab != "Missing") %>% 
  mutate(d_bmi_20_lab= fct_relevel(d_bmi_20_lab, "<18.5", "18.5-25", "25-30", "30+")) %>%
  ggplot(aes(x = d_bmi_20_lab, y = d_md, fill = d_bmi_20_lab)) +
  geom_violin(alpha = 0.4) +
  geom_boxplot(width = 0.1)+
  ylim(0, 100) +
  labs(title = "Body Mass Index at age 20",
       x = "",
       y = "Breast density (%)") +
  theme_base() +
  theme(legend.position = "none") +
  scale_fill_brewer(palette="Set2") +
  annotate("text",
           x = 1,
           y = 90,
           label = "reference",
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 2,
           y = 90,
           label = paste("p =", bmi20_18.5to25),
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 3,
           y = 90,
           label = paste("p =", bmi20_25to30),
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 4,
           y = 90,
           label = paste("p", bmi20_30),
           col = "black",
           size = 3.5) 

# plot16
```

plot list:

plot 1 - detection mode

plot 2 - Meno status

plot 3 - age meno

plot 4 - HRT

plot 5 - n of parous preg

plot 6 - age at first birth

plot 7 - Age at menarche

plot 8 - Oral contraceptives

plot 9 - BBD

plot 10 - Family history of BC

plot 11 - Alcohol

plot 12 - Smoking status

plot 13 - BMI entry

plot 14 - physical activity

plot 15 - Breastfeeding duration

plot 16 - BMI age 20

#### Ordering for RF plots

(chronological order)

Composite plot 1

plot 7 - Age at menarche

plot 8 - Oral contraceptives

Composite plot 2

plot 5 - n of parous preg

plot 6 - age at first birth

plot 15 - Breastfeeding duration

Composite plot 3

plot 2 - Meno status

plot 3 - age meno

plot 4 - HRT

Composite plot 4

plot 9 - BBD

plot 10 - Family history of BC

Composite plot 5

plot 16 - BMI age 20

plot 13 - BMI entry

plot 14 - physical activity

Composite plot 6

plot 11 - Alcohol

plot 12 - Smoking status

```{r, fig.width=10, fig.height=4}
#| echo: false

cp_1 <- plot7 + plot8
cp_1

```

```{r, fig.width=15, fig.height=4}
#| echo: false

cp_2 <- plot5 + plot6 + plot15

cp_2

```

```{r, fig.width=15, fig.height=4}
#| echo: false

cp_3 <- plot2 + plot3 + plot4
cp_3

```

```{r, fig.width=10, fig.height=4}
#| echo: false

cp_4 <- plot9 + plot10
cp_4
```

```{r, fig.width=15, fig.height=4}
#| echo: false

cp_5 <- plot16 + plot13 + plot14
cp_5
```

```{r, fig.width=10, fig.height=4}
#| echo: false

cp_6 <- plot11 + plot12
cp_6

```

### old plots

```{r, fig.width=15, fig.height=9}
#| echo: false

# # reproductive factors
# reproductive_plots <- (plot5 + plot6 + plot7) / (plot3 + plot4 + plot8)
# reproductive_plots

```

```{r, fig.width=15, fig.height=9}
#| echo: false

# # lifestyle factors
# lifestyle_plots <- (plot9 + plot10 + plot11) /  (plot12 + plot13 + plot14)
# lifestyle_plots
```

### Tumour characteristics

```{r}
#| echo: false
#| include: false

# Invasive status 

# code dev for plots with p-values


# linear regression 
lr <- md_df %>% 
   filter(d_inv_status_lab != "Missing") %>% 
  lm(d_md_sqrt ~ d_inv_status_lab, data = .
     ) %>% tidy() 

# format p-values 
lr <- lr %>% 
  format_p_values(p.value)

lr

# pull p-values
inv_yes <- lr %>% 
  filter(term == "d_inv_status_labInvasive") %>% 
  pull(p_val1)




# plot
tplot1 <- md_df %>% 
  filter(d_inv_status_lab != "Missing") %>%  
  ggplot(aes(x = d_inv_status_lab, y = d_md, fill = d_inv_status_lab)) +
  geom_violin(alpha = 0.4) +
  geom_boxplot(width = 0.1)+
  ylim(0, 100) +
  labs(title = "Invasive status",
       x = "",
       y = "Breast density (%)") +
  theme_base() +
  theme(legend.position = "none") +
  scale_fill_brewer(palette="Set2") +
  annotate("text",
           x = 1,
           y = 90,
           label = "reference",
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 2,
           y = 90,
           label = paste("p =", inv_yes),
           col = "black",
           size = 3.5)

#tplot1
```

```{r}
#| echo: false
#| include: false

# Grade 

# code dev for plots with p-values


# linear regression 
lr <- md_df %>% 
   filter(d_grade_lab != "Missing") %>% 
  lm(d_md_sqrt ~ d_grade_lab, data = .
     ) %>% tidy() 

# format p-values 
lr <- lr %>% 
  format_p_values(p.value)

lr

# pull p-values
grade2_p <- lr %>% 
  filter(term == "d_grade_lab2") %>% 
  pull(p_val1)

grade3_p <- lr %>% 
  filter(term == "d_grade_lab3") %>% 
  pull(p_val1)


# plot
tplot2 <- md_df %>% 
  filter(d_grade_lab != "Missing") %>%  
  ggplot(aes(x = d_grade_lab, y = d_md, fill = d_grade_lab)) +
  geom_violin(alpha = 0.4) +
  geom_boxplot(width = 0.1)+
  ylim(0, 100) +
  labs(title = "Grade",
       x = "",
       y = "Breast density (%)") +
  theme_base() +
  theme(legend.position = "none") +
  scale_fill_brewer(palette="Set2") +
  annotate("text",
           x = 1,
           y = 90,
           label = "reference",
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 2,
           y = 90,
           label = paste("p =", grade2_p),
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 3,
           y = 90,
           label = paste("p =", grade3_p),
           col = "black",
           size = 3.5)

#tplot2
```

```{r}
#| echo: false
#| include: false

# Morphology

# code dev for plots with p-values


# linear regression 
lr <- md_df %>% 
   filter(d_morph4_lab != "Missing") %>%  
  lm(d_md_sqrt ~ d_morph4_lab, data = .
     ) %>% tidy() 

# format p-values 
lr <- lr %>% 
  format_p_values(p.value)

lr

# pull p-values
morph_lob_p <- lr %>% 
  filter(term == "d_morph4_labLobular") %>% 
  pull(p_val1)

morph_mixed_p <- lr %>% 
  filter(term == "d_morph4_labMixed: ductal and other") %>% 
  pull(p_val1)

morph_other_p <- lr %>% 
  filter(term == "d_morph4_labOther") %>% 
  pull(p_val1)


# plot
tplot3 <- md_df %>% 
  filter(d_morph4_lab != "Missing") %>% 
  ggplot(aes(x = d_morph4_lab, y = d_md, fill = d_morph4_lab)) +
  geom_violin(alpha = 0.4) +
  geom_boxplot(width = 0.1)+
  ylim(0, 100) +
  labs(title = "Morphology",
       x = "",
       y = "Breast density (%)") +
  theme_base() +
  theme(legend.position = "none") +
  scale_fill_brewer(palette="Set2")  +
  annotate("text",
           x = 1,
           y = 90,
           label = "reference",
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 2,
           y = 90,
           label = paste("p =", morph_lob_p),
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 3,
           y = 90,
           label = paste("p =", morph_mixed_p),
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 4,
           y = 90,
           label = paste("p =", morph_other_p),
           col = "black",
           size = 3.5) +
  theme(axis.text.x = element_text(angle = 20, hjust = 1))

#tplot3
```

```{r}
#| echo: false
#| include: false

# positive nodes 

# code dev for plots with p-values


# linear regression 
lr <- md_df %>% 
   filter(d_pos_nodes_lab != "Missing") %>%   
  lm(d_md_sqrt ~ d_pos_nodes_lab, data = .
     ) %>% tidy() 

# format p-values 
lr <- lr %>% 
  format_p_values(p.value)

lr

# pull p-values
pos_nodes_p <- lr %>% 
  filter(term == "d_pos_nodes_labPositive") %>% 
  pull(p_val1)


# plot
tplot4 <- md_df %>% 
  filter(d_pos_nodes_lab != "Missing") %>% 
  ggplot(aes(x = d_pos_nodes_lab, y = d_md, fill = d_pos_nodes_lab)) +
  geom_violin(alpha = 0.4) +
  geom_boxplot(width = 0.1)+
  ylim(0, 100) +
  labs(title = "Positive nodes",
       x = "",
       y = "Breast density (%)") +
  theme_base() +
  theme(legend.position = "none") +
  scale_fill_brewer(palette="Set2") +
  annotate("text",
           x = 1,
           y = 90,
           label = "reference",
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 2,
           y = 90,
           label = paste("p =", pos_nodes_p),
           col = "black",
           size = 3.5)

#tplot4
```

```{r}
#| echo: false
#| include: false

# tumour size 

# code dev for plots with p-values


# linear regression 
lr <- md_df %>% 
   filter(d_tumour_size_lab != "Missing") %>%    
  lm(d_md_sqrt ~ d_tumour_size_lab, data = .
     ) %>% tidy() 

# format p-values 
lr <- lr %>% 
  format_p_values(p.value)

lr

# pull p-values
tms_21to50_p <- lr %>% 
  filter(term == "d_tumour_size_lab21-50") %>% 
  pull(p_val1)

tms_over50_p <- lr %>% 
  filter(term == "d_tumour_size_lab50+") %>% 
  pull(p_val1)


# plot
tplot5 <- md_df %>% 
  filter(d_tumour_size_lab != "Missing") %>% 
  ggplot(aes(x = d_tumour_size_lab, y = d_md, fill = d_tumour_size_lab)) +
  geom_violin(alpha = 0.4) +
  geom_boxplot(width = 0.1)+
  ylim(0, 100) +
  labs(title = "Tumour size",
       x = "",
       y = "Breast density (%)") +
  theme_base() +
  theme(legend.position = "none") +
  scale_fill_brewer(palette="Set2") +
  annotate("text",
           x = 1,
           y = 90,
           label = "reference",
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 2,
           y = 90,
           label = paste("p =", tms_21to50_p),
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 3,
           y = 90,
           label = paste("p =", tms_over50_p),
           col = "black",
           size = 3.5)

#tplot5
  
```

```{r}
#| echo: false
#| include: false

# er status

# code dev for plots with p-values


# linear regression 
lr <- md_df %>% 
   filter(d_er_status_lab != "Missing") %>%     
  lm(d_md_sqrt ~ d_er_status_lab, data = .
     ) %>% tidy() 

# format p-values 
lr <- lr %>% 
  format_p_values(p.value)

lr

# pull p-values
er_neg_p <- lr %>% 
  filter(term == "d_er_status_labNegative") %>% 
  pull(p_val1)


# plot
tplot6 <- md_df %>% 
  filter(d_er_status_lab != "Missing") %>%
  ggplot(aes(x = d_er_status_lab, y = d_md, fill = d_er_status_lab)) +
  geom_violin(alpha = 0.4) +
  geom_boxplot(width = 0.1)+
  ylim(0, 100) +
  labs(title = "ER status",
       x = "",
       y = "Breast density (%)") +
  theme_base() +
  theme(legend.position = "none") +
  scale_fill_brewer(palette="Set2") +
  annotate("text",
           x = 1,
           y = 90,
           label = "reference",
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 2,
           y = 90,
           label = paste("p =", er_neg_p),
           col = "black",
           size = 3.5)

#tplot6
```

```{r}
#| echo: false
#| include: false

# pr status

# code dev for plots with p-values


# linear regression 
lr <- md_df %>% 
   filter(d_pr_status_lab != "Missing") %>%   
  lm(d_md_sqrt ~ d_pr_status_lab, data = .
     ) %>% tidy() 

# format p-values 
lr <- lr %>% 
  format_p_values(p.value)

lr

# pull p-values
pr_neg_p <- lr %>% 
  filter(term == "d_pr_status_labNegative") %>% 
  pull(p_val1)


# plot
tplot7 <- md_df %>% 
  filter(d_pr_status_lab != "Missing") %>%
  ggplot(aes(x = d_pr_status_lab, y = d_md, fill = d_pr_status_lab)) +
  geom_violin(alpha = 0.4) +
  geom_boxplot(width = 0.1)+
  ylim(0, 100) +
  labs(title = "PR status",
       x = "",
       y = "Breast density (%)") +
  theme_base() +
  theme(legend.position = "none") +
  scale_fill_brewer(palette="Set2") +
  annotate("text",
           x = 1,
           y = 90,
           label = "reference",
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 2,
           y = 90,
           label = paste("p =", pr_neg_p),
           col = "black",
           size = 3.5)

#tplot7

```

```{r}
#| echo: false
#| include: false

# her2 status

# code dev for plots with p-values


# linear regression 
lr <- md_df %>% 
   filter(d_her2_status_lab != "Missing") %>%   
  lm(d_md_sqrt ~ d_her2_status_lab, data = .
     ) %>% tidy() 

# format p-values 
lr <- lr %>% 
  format_p_values(p.value)

lr

# pull p-values
her2_neg_p <- lr %>% 
  filter(term == "d_her2_status_labNegative") %>% 
  pull(p_val1)


# plot
tplot8 <- md_df %>% 
  filter(d_her2_status_lab != "Missing") %>% 
  ggplot(aes(x = d_her2_status_lab, y = d_md, fill = d_her2_status_lab)) +
  geom_violin(alpha = 0.4) +
  geom_boxplot(width = 0.1)+
  ylim(0, 100) +
  labs(title = "HER2 status",
       x = "",
       y = "Breast density (%)") +
  theme_base() +
  theme(legend.position = "none") +
  scale_fill_brewer(palette="Set2") +
  annotate("text",
           x = 1,
           y = 90,
           label = "reference",
           col = "black",
           size = 3.5) +
  annotate("text",
           x = 2,
           y = 90,
           label = paste("p =", her2_neg_p),
           col = "black",
           size = 3.5) 

#tplot8
```

```{r, fig.width=19, fig.height=13}
#| echo: false

# composite plot
tcomposite_plot <-plot1 + tplot1 + tplot2 + tplot4 + tplot5 + tplot6 + tplot7 + tplot8 + tplot3 + plot_layout(ncol = 3)

tcomposite_plot


```

## Table 2 - Main analysis

The main analysis is a logistic regression with multiple exposures (risk factors).

This table shows risk factor level descriptives and their associations with the mode of detection (logistic regression) adjusted for all other risk factors, age at diagnosis and time from entry to BC diagnosis.

Model 2 is the final selected model (Default model) to present in the analysis. Model 2 is adjusted for all co-exposures + age at diagnosis and time from entry to BC diagnosis. Model 1 was only adjusted for co-exposures, e.i. all exposures in the model.

Because model 2 was selected as the default model, it is also used on the subanalyses with TC and MD as the comparison against adjusted models for TC and MD.

03/10/2024 - adding breastfeeding and BMI age 20 to the models, changed OC status to never/ever

**Model 1**: “d_R1alcohol_status_cat”, “d_R1alcohol_units_cat”, “d_R1smokingstatus_cat”, “d_bmi_entry_cat”, “d_R1physmet_leis_who”, “d_bbd”, “d_fambrca”, “d_age_menarche_cat”, “d_ocstatus_cat”, “d_R1menopause_cat3”, “d_age_menopause_cat”, “d_R1hrtstatus_cat”, “d_parous_cat”, “d_parity_cat”, “d_agebirth1_cat” + breastfeeding duration + bmi at 20

**Model 2**: Model 1 + “diagage”, “d_R1toBC_y”, “yeardiag”

**Model 3:** Model 2 + “d_md_qrt”

**Model 4:** Model 3 + d_MDtoBC_lab (categorical)

### Detailed table - internal use only

```{r}
#| echo: false
#| warning: false

# summary Logistic Regression of Model 2  -----------------------------------------

# view existing output
#model2

# remove columns with total and interval N - not simple with gtsummary, trying from scratch without adding the columns


# Create clean version of logistic output, without Ns
model <- df %>%
  glm(formula = d_dmode_n ~ d_R1menopause_lab3 + d_age_meno_tr_lab 
      + d_R1hrt_sttyp_tr_lab
      + d_parity_lab + d_agebirth1_tr_lab + d_age_menarche_lab + d_ocstatus_lab 
      + d_bbd_lab + d_fambrca_lab
      + d_R1alcohol_units_lab + d_R1smokingstatus_lab 
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab + d_bmi_20_lab + d_bf_dur_tr_lab
      + diagage + d_R1toBC_y + yeardiag,
            family = binomial(link = "logit"))




# tidy up output
model2 <- tbl_regression(model, 
                         exponentiate = TRUE,
                         label = list(
                          d_R1alcohol_units_lab ~ "Alcohol units",
                          d_R1smokingstatus_lab ~ "Smoking status",
                          d_bmi_entry_lab ~ "BMI at entry",
                          d_R1physmet_leis_who_lab ~ "Physical Activity (WHO cat)",
                          d_bbd_lab ~ "Benign breast disease",
                          d_fambrca_lab ~ "Family history of BC",
                          d_age_menarche_lab ~ "Age at menarche",
                          d_ocstatus_lab ~ "Oral contraceptives status",
                          d_R1menopause_lab3 ~ "Menopause",
                          d_age_meno_tr_lab ~ "Age at menopause",
                          d_R1hrt_sttyp_tr_lab ~ "HRT status",
                          d_parity_lab ~ "N of parous pregnancies",
                          d_agebirth1_tr_lab ~ "Age at first birth",
                          d_bmi_20_lab ~ "BMI at age 20y",
                          d_bf_dur_tr_lab ~ "Cummulative breastfeeding duration"
                         )
                         ) %>% 
  #add_global_p() %>% 
  remove_row_type(variables = c(diagage, d_R1toBC_y, yeardiag), # remove covariates
                  type = "all")
  

#model2

# summary Logistic Regression of Model 4  -----------------------------------------

 model <- df %>%
  glm(formula = d_dmode_n ~ d_R1menopause_lab3 + d_age_meno_tr_lab 
      + d_R1hrt_sttyp_tr_lab
      + d_parity_lab + d_agebirth1_tr_lab + d_age_menarche_lab + d_ocstatus_lab 
      + d_bbd_lab + d_fambrca_lab
      + d_R1alcohol_units_lab + d_R1smokingstatus_lab 
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab + d_bmi_20_lab + d_bf_dur_tr_lab
      + diagage + d_R1toBC_y + yeardiag
      + d_md_qrt + d_MDtoBC_lab,
            family = binomial(link = "logit"))


# tidy up output
model4 <- tbl_regression(model, 
                         exponentiate = TRUE,
                         label = list(
                          d_R1alcohol_units_lab ~ "Alcohol units",
                          d_R1smokingstatus_lab ~ "Smoking status",
                          d_bmi_entry_lab ~ "BMI at entry",
                          d_R1physmet_leis_who_lab ~ "Physical Activity (WHO cat)",
                          d_bbd_lab ~ "Benign breast disease",
                          d_fambrca_lab ~ "Family history of BC",
                          d_age_menarche_lab ~ "Age at menarche",
                          d_ocstatus_lab ~ "Oral contraceptives status",
                          d_R1menopause_lab3 ~ "Menopause",
                          d_age_meno_tr_lab ~ "Age at menopause",
                          d_R1hrt_sttyp_tr_lab ~ "HRT status",
                          d_parity_lab ~ "N of parous pregnancies",
                          d_agebirth1_tr_lab ~ "Age at first birth",
                          d_bmi_20_lab ~ "BMI at age 20y",
                          d_bf_dur_tr_lab ~ "Cummulative breastfeeding duration",
                          d_md_qrt ~ "Mammographic density (quartiles)"
                         )
                         ) %>% 
  #add_global_p() %>% 
  remove_row_type(variables = c(diagage, d_R1toBC_y, yeardiag, d_MDtoBC_lab), # remove covariates
                  type = "all")



# Create descriptives by mode of detection ------------------------------------
# select(d_dmode_n, d_R1alcohol_units_lab, d_R1smokingstatus_lab, d_bmi_entry_lab,  d_R1physmet_leis_who_lab, d_bbd_lab, d_fambrca_lab, d_age_menarche_lab, d_ocstatus_lab, d_R1menopause_lab3, d_age_meno_tr_lab, d_R1hrt_sttyp_tr_lab, d_parity_tr_lab, d_agebirth1_tr_lab)

# note: need to rename the non-trick variables according to their trick name to align with the model, which can then allow table merge. 

# include missing 
# create df
  
descr_tab2_df <- df %>% 
  select(d_dmode, all_of(all_expl_variables), diagage, d_R1toBC_y, d_md_qrt
         ) %>% 
  #tidyr::drop_na(all_of(all_expl_variables), diagage, d_R1toBC_y # filter out all missing to have the same n as in the model
  mutate(d_bmi_entry_lab = fct_relevel(d_bmi_entry_lab, "<18.5", "18.5-25", "25-30", "30+"),
        d_bmi_20_lab = fct_relevel(d_bmi_entry_lab, "<18.5", "18.5-25", "25-30", "30+") 
              ) %>% 
  mutate(d_age_meno_tr_lab = d_age_menopause_lab,
         d_R1hrt_sttyp_tr_lab = d_R1hrt_sttyp_lab,
         d_agebirth1_tr_lab = d_agebirth1_lab,
         d_bf_dur_tr_lab = d_bf_dur_lab
         ) %>% 
  select(-d_age_menopause_lab, -d_R1hrt_sttyp_lab, -d_agebirth1_lab, -d_bf_dur_lab, -diagage, -d_R1toBC_y)  
  
  

# glimpse(descr_model2)
# str(test)

# create gs summary table
descr_tab2 <- descr_tab2_df %>% 
  select(d_dmode,
         d_age_menarche_lab,
         d_ocstatus_lab,
         d_parity_lab,
         d_agebirth1_tr_lab,
         d_bf_dur_tr_lab,
         d_R1menopause_lab3,
         d_age_meno_tr_lab,
         d_R1hrt_sttyp_tr_lab,
         d_bbd_lab,
         d_fambrca_lab,
         d_bmi_entry_lab,
         d_bmi_20_lab,
         d_R1physmet_leis_who_lab,
         d_R1alcohol_units_lab, 
         d_R1smokingstatus_lab,
         d_md_qrt) %>% 
  tbl_summary(by = d_dmode,
              statistic = list(
                all_continuous() ~ "{mean} ({sd})",
                all_categorical() ~ "{n} ({p}) "
                ),
              digits = all_continuous() ~ 2,
              label = list(
                          d_R1alcohol_units_lab ~ "Alcohol units",
                          d_R1smokingstatus_lab ~ "Smoking status",
                          d_bmi_entry_lab ~ "BMI at entry",
                          d_R1physmet_leis_who_lab ~ "Physical Activity (WHO cat)",
                          d_bbd_lab ~ "Benign breast disease",
                          d_fambrca_lab ~ "Family history of BC",
                          d_age_menarche_lab ~ "Age at menarche",
                          d_ocstatus_lab ~ "Oral contraceptives status",
                          d_R1menopause_lab3 ~ "Menopause",
                          d_age_meno_tr_lab ~ "Age at menopause",
                          d_R1hrt_sttyp_tr_lab ~ "HRT status",
                          d_parity_lab ~ "N of parous pregnancies",
                          d_agebirth1_tr_lab ~ "Age at first birth",
                          d_bmi_20_lab ~ "BMI at age 20y",
                          d_bf_dur_tr_lab ~ "Cummulative breastfeeding duration",
                          d_md_qrt ~ "Mammographic density (quartiles)"
                ),
              type = list(d_bbd_lab ~ "categorical",
                          d_fambrca_lab ~ "categorical")
              )

#descr_model2
  
# merge with analysis output -------------------------------------------------


table2 <- tbl_merge(
  tbls = list(descr_tab2, model2, model4),
  tab_spanner = c("**Distribution by detection mode**", "**Model 2**", "**Model 4**")
   ) 

#show_header_names(table2)

table2 <- table2 %>% 
  modify_footnote(
    estimate_2 ~ "Model 2: Age at menarche, Menopause, Age at menopause, HRT status, Oral contraceptives status, N of parous pregnancies, Age at first birth, Cummulative breastfeeding duration, Benign breast disease, Family history of BC, BMI at entry, BMI at age 20y, Leisure Physical Activity (MET-h/w), Alcohol units per week, Smoking status, Age at Diagnosis, Time from Recruitment to BC diagnosis, Year of diagnosis",
    estimate_3 ~ "Model 4: Age at menarche, Menopause, Age at menopause, HRT status, Oral contraceptives status, N of parous pregnancies, Age at first birth, Cummulative breastfeeding duration, Benign breast disease, Family history of BC, BMI at entry, BMI at age 20y, Leisure Physical Activity (MET-h/w), Alcohol units per week, Smoking status, Age at Diagnosis, Time from Recruitment to BC diagnosis, Year of diagnosis Percent Mammographic density (quartiles), Time from Mammographic density measurement to BRCA diagnosis."
  )

table2

```

### Simplified table

on 01/10/2024 agreed to:

-   remove percentages

-   use HRT simple variable (never/former/current)

-   Use OC status simple variable (ever/never)

-   remove p for categories

-   add overall p-value (only for model 4?)

Forest plot will be also presented

```{r}
#| echo: false
#| warning: false

# summary Logistic Regression of Model 2  -----------------------------------------

# view existing output
#model2

# remove columns with total and interval N - not simple with gtsummary, trying from scratch without adding the columns


# Create clean version of logistic output, without Ns
model <- df %>%
  glm(formula = d_dmode_n ~ d_R1menopause_lab3 + d_age_meno_tr_lab 
      + d_R1hrt_tr_lab
      + d_parity_lab + d_agebirth1_tr_lab + d_age_menarche_lab + d_ocstatus2_lab 
      + d_bbd_lab + d_fambrca_lab
      + d_R1alcohol_units_lab + d_R1smokingstatus_lab 
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab + d_bmi_20_lab + d_bf_dur_tr_lab
      + diagage + d_R1toBC_y + yeardiag,
            family = binomial(link = "logit"))




# tidy up output
model2 <- tbl_regression(model, 
                         exponentiate = TRUE,
                         label = list(
                          d_R1alcohol_units_lab ~ "Alcohol units",
                          d_R1smokingstatus_lab ~ "Smoking status",
                          d_bmi_entry_lab ~ "BMI at entry",
                          d_R1physmet_leis_who_lab ~ "Physical Activity (WHO cat)",
                          d_bbd_lab ~ "Benign breast disease",
                          d_fambrca_lab ~ "Family history of BC",
                          d_age_menarche_lab ~ "Age at menarche",
                          d_ocstatus2_lab ~ "Oral contraceptives status",
                          d_R1menopause_lab3 ~ "Menopause",
                          d_age_meno_tr_lab ~ "Age at menopause",
                          d_R1hrt_tr_lab ~ "HRT status",
                          d_parity_lab ~ "N of parous pregnancies",
                          d_agebirth1_tr_lab ~ "Age at first birth",
                          d_bmi_20_lab ~ "BMI at age 20y",
                          d_bf_dur_tr_lab ~ "Cummulative breastfeeding duration"
                         )
                         ) %>% 
  #add_global_p() %>% 
  remove_row_type(variables = c(diagage, d_R1toBC_y, yeardiag), # remove covariates
                  type = "all") %>% 
  add_global_p()
  

#model2

# summary Logistic Regression of Model 4  -----------------------------------------

 model <- df %>%
  filter(d_MDtoBC_lab != "Missing") %>% 
  glm(formula = d_dmode_n ~ d_R1menopause_lab3 + d_age_meno_tr_lab 
      + d_R1hrt_tr_lab
      + d_parity_lab + d_agebirth1_tr_lab + d_age_menarche_lab + d_ocstatus2_lab 
      + d_bbd_lab + d_fambrca_lab
      + d_R1alcohol_units_lab + d_R1smokingstatus_lab 
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab + d_bmi_20_lab + d_bf_dur_tr_lab
      + diagage + d_R1toBC_y + yeardiag
      + d_md_qrt + d_MDtoBC_lab,
            family = binomial(link = "logit"))

#summary(model)


# tidy up output
model4 <- tbl_regression(model, 
                         exponentiate = TRUE,
                         label = list(
                          d_R1alcohol_units_lab ~ "Alcohol units",
                          d_R1smokingstatus_lab ~ "Smoking status",
                          d_bmi_entry_lab ~ "BMI at entry",
                          d_R1physmet_leis_who_lab ~ "Physical Activity (WHO cat)",
                          d_bbd_lab ~ "Benign breast disease",
                          d_fambrca_lab ~ "Family history of BC",
                          d_age_menarche_lab ~ "Age at menarche",
                          d_ocstatus2_lab ~ "Oral contraceptives status",
                          d_R1menopause_lab3 ~ "Menopause",
                          d_age_meno_tr_lab ~ "Age at menopause",
                          d_R1hrt_tr_lab ~ "HRT status",
                          d_parity_lab ~ "N of parous pregnancies",
                          d_agebirth1_tr_lab ~ "Age at first birth",
                          d_bmi_20_lab ~ "BMI at age 20y",
                          d_bf_dur_tr_lab ~ "Cummulative breastfeeding duration",
                          d_md_qrt ~ "Mammographic density (quartiles)"
                         )
                         ) %>% 
  #add_global_p() %>% 
  remove_row_type(variables = c(diagage, d_R1toBC_y, yeardiag, d_MDtoBC_lab), # remove covariates
                  type = "all") %>% 
  add_global_p()



# Create descriptives by mode of detection ------------------------------------
# select(d_dmode_n, d_R1alcohol_units_lab, d_R1smokingstatus_lab, d_bmi_entry_lab,  d_R1physmet_leis_who_lab, d_bbd_lab, d_fambrca_lab, d_age_menarche_lab, d_ocstatus_lab, d_R1menopause_lab3, d_age_meno_tr_lab, d_R1hrt_sttyp_tr_lab, d_parity_tr_lab, d_agebirth1_tr_lab)

# note: need to rename the non-trick variables according to their trick name to align with the model, which can then allow table merge. 

# include missing 
# create df
  
descr_tab2_df <- df %>% 
  select(d_dmode, all_of(all_expl_variables), diagage, d_R1toBC_y, d_md_qrt
         ) %>% 
  #tidyr::drop_na(all_of(all_expl_variables), diagage, d_R1toBC_y # filter out all missing to have the same n as in the model
  mutate(d_bmi_entry_lab = fct_relevel(d_bmi_entry_lab, "<18.5", "18.5-25", "25-30", "30+"),
        d_bmi_20_lab = fct_relevel(d_bmi_entry_lab, "<18.5", "18.5-25", "25-30", "30+") 
              ) %>% 
  mutate(d_age_meno_tr_lab = d_age_menopause_lab,
         d_R1hrt_tr_lab = d_R1hrtstatus_lab,
         d_agebirth1_tr_lab = d_agebirth1_lab,
         d_bf_dur_tr_lab = d_bf_dur_lab
         ) %>% 
  select(-d_age_menopause_lab, -d_R1hrtstatus_lab, -d_agebirth1_lab, -d_bf_dur_lab, -diagage, -d_R1toBC_y)  
  
  

# glimpse(descr_model2)
# str(test)

# create gs summary table
descr_tab2 <- descr_tab2_df %>% 
  select(d_dmode,
         d_age_menarche_lab,
         d_ocstatus2_lab,
         d_parity_lab,
         d_agebirth1_tr_lab,
         d_bf_dur_tr_lab,
         d_R1menopause_lab3,
         d_age_meno_tr_lab,
         d_R1hrt_tr_lab,
         d_bbd_lab,
         d_fambrca_lab,
         d_bmi_entry_lab,
         d_bmi_20_lab,
         d_R1physmet_leis_who_lab,
         d_R1alcohol_units_lab, 
         d_R1smokingstatus_lab,
         d_md_qrt) %>% 
  tbl_summary(by = d_dmode,
              statistic = list(
                all_continuous() ~ "{mean}",
                all_categorical() ~ "{n}"
                ),
              digits = all_continuous() ~ 2,
              label = list(
                          d_R1alcohol_units_lab ~ "Alcohol units",
                          d_R1smokingstatus_lab ~ "Smoking status",
                          d_bmi_entry_lab ~ "BMI at entry",
                          d_R1physmet_leis_who_lab ~ "Physical Activity (WHO cat)",
                          d_bbd_lab ~ "Benign breast disease",
                          d_fambrca_lab ~ "Family history of BC",
                          d_age_menarche_lab ~ "Age at menarche",
                          d_ocstatus2_lab ~ "Oral contraceptives status",
                          d_R1menopause_lab3 ~ "Menopause",
                          d_age_meno_tr_lab ~ "Age at menopause",
                          d_R1hrt_tr_lab ~ "HRT status",
                          d_parity_lab ~ "N of parous pregnancies",
                          d_agebirth1_tr_lab ~ "Age at first birth",
                          d_bmi_20_lab ~ "BMI at age 20y",
                          d_bf_dur_tr_lab ~ "Cummulative breastfeeding duration",
                          d_md_qrt ~ "Mammographic density (quartiles)"
                ),
              type = list(d_bbd_lab ~ "categorical",
                          d_fambrca_lab ~ "categorical")
              )

#descr_model2
  
# merge with analysis output -------------------------------------------------


table2 <- tbl_merge(
  tbls = list(descr_tab2, model2, model4),
  tab_spanner = c("**Distribution by detection mode**", "**Model 2**", "**Model 4**")
   ) 

#show_header_names(table2)

table2 <- table2 %>% 
  modify_footnote(
    estimate_2 ~ "Model 2: Age at menarche, Menopause, Age at menopause, HRT status, Oral contraceptives status, N of parous pregnancies, Age at first birth, Cummulative breastfeeding duration, Benign breast disease, Family history of BC, BMI at entry, BMI at age 20y, Leisure Physical Activity (MET-h/w), Alcohol units per week, Smoking status, Age at Diagnosis, Time from Recruitment to BC diagnosis, Year of diagnosis",
    estimate_3 ~ "Model 4: Age at menarche, Menopause, Age at menopause, HRT status, Oral contraceptives status, N of parous pregnancies, Age at first birth, Cummulative breastfeeding duration, Benign breast disease, Family history of BC, BMI at entry, BMI at age 20y, Leisure Physical Activity (MET-h/w), Alcohol units per week, Smoking status, Age at Diagnosis, Time from Recruitment to BC diagnosis, Year of diagnosis Percent Mammographic density (quartiles), Time from Mammographic density measurement to BRCA diagnosis."
  )

table2


```

## Figure X - forest plot

```{r}
#| echo: false
#| warning: false


# run models 

# summary Logistic Regression of Model 2  -----------------------------------------

# view existing output
#model2

# remove columns with total and interval N - not simple with gtsummary, trying from scratch without adding the columns


# Create clean version of logistic output, without Ns
model <- df %>%
  glm(formula = d_dmode_n ~ d_R1menopause_lab3 + d_age_meno_tr_lab 
      + d_R1hrt_tr_lab
      + d_parity_lab + d_agebirth1_tr_lab + d_age_menarche_lab + d_ocstatus2_lab 
      + d_bbd_lab + d_fambrca_lab
      + d_R1alcohol_units_lab + d_R1smokingstatus_lab 
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab + d_bmi_20_lab + d_bf_dur_tr_lab
      + diagage + d_R1toBC_y + yeardiag,
            family = binomial(link = "logit"))

tidy_model2 <- model %>%
   tidy(exponentiate = TRUE, conf.int = TRUE) 

# %>% 
#   mutate(OR = estimate) %>%
#    select(-estimate) %>%
#    select(term, estimate, everything())


#tidy_model2

# summary Logistic Regression of Model 4  -----------------------------------------

 model <- df %>%
  filter(d_MDtoBC_lab != "Missing") %>% 
  glm(formula = d_dmode_n ~ d_R1menopause_lab3 + d_age_meno_tr_lab 
      + d_R1hrt_tr_lab
      + d_parity_lab + d_agebirth1_tr_lab + d_age_menarche_lab + d_ocstatus2_lab 
      + d_bbd_lab + d_fambrca_lab
      + d_R1alcohol_units_lab + d_R1smokingstatus_lab 
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab + d_bmi_20_lab + d_bf_dur_tr_lab
      + diagage + d_R1toBC_y + yeardiag
      + d_md_qrt + d_MDtoBC_lab,
            family = binomial(link = "logit"))

tidy_model4 <- model %>%
   tidy(exponentiate = TRUE, conf.int = TRUE) 


#tidy_model4
```

```{r}
#| echo: false

# 2. PROCESS DATA FOR FORESTPLOT 

# full join data from both models 
md_forest_df <- tidy_model2 %>% 
  full_join(tidy_model4, by = "term", suffix = c("_mdl2", "_mdl4"))


# create labels and variable names structures 
md_forest_df <- md_forest_df %>% 
  mutate(variable = case_when(
    grepl("menopause", term) ~ "Menopausal status",
    grepl("age_meno", term) ~ "Age at menopause",
    grepl("hrt", term) ~ "HRT status",
    grepl("parity", term) ~ "N of parous pregnancies",
    grepl("agebirth1", term) ~ "Age at first birth",
    grepl("age_menarche", term) ~ "Age at menarche", 
    grepl("ocstatus2", term) ~ "Oral contraceptives status",
    grepl("bbd", term) ~ "Benign breast disease",
    grepl("fambrca", term) ~ "Family history of BC",
    grepl("R1alcohol", term) ~ "Alcohol units",
    grepl("R1smokingstatus", term) ~ "Smoking status",
    grepl("bmi_entry", term) ~ "BMI at entry",
    grepl("R1physmet_leis_who", term) ~ "Physical Activity (WHO cat)",
    grepl("md_qrt", term) ~ "Mammographic breast density quartile",
    grepl("bmi_20", term) ~ "BMI at age 20y",
    grepl("bf_dur", term) ~ "Cummulative breastfeeding duration"
                              ),
    level = case_when(
      grepl("d_R1menopause_lab3", term) ~ sub("d_R1menopause_lab3", "", term),
      grepl("age_meno", term) ~ sub("d_age_meno_tr_lab", "", term),
      grepl("hrt", term) ~ sub("d_R1hrt_tr_lab", "", term),
      grepl("parity", term) ~ sub("d_parity_lab", "", term),
      grepl("agebirth1", term) ~ sub("d_agebirth1_tr_lab", "", term),
      grepl("age_menarche", term) ~ sub("d_age_menarche_lab", "", term), 
      grepl("ocstatus2", term) ~ sub("d_ocstatus2_lab", "", term),
      grepl("bbd", term) ~ sub("d_bbd_lab", "", term),
      grepl("fambrca", term) ~ sub("d_fambrca_lab", "", term),
      grepl("R1alcohol", term) ~ sub("d_R1alcohol_units_lab", "", term),
      grepl("R1smokingstatus", term) ~ sub("d_R1smokingstatus_lab", "", term),
      grepl("bmi_entry", term) ~ sub("d_bmi_entry_lab", "", term),
      grepl("R1physmet_leis_who", term) ~ sub("d_R1physmet_leis_who_lab", "", term),
      grepl("md_qrt", term) ~ sub("d_md_qrt", "", term),
      grepl("bmi_20", term) ~ sub("d_bmi_20_lab", "", term),
      grepl("bf_dur", term) ~ sub("d_bf_dur_tr_lab", "", term)
    )
    )

# add subheading rows 

subheading_data <- md_forest_df %>%
  distinct(variable) %>%
  mutate(level = variable)

md_forest_df <- bind_rows(subheading_data, md_forest_df) %>%
  arrange(variable, factor(level, levels = unique(level)))



# create long format 
md_forest_long <- md_forest_df %>% 
  filter(!term %in% c("(Intercept)", "diagage", "d_R1toBC_y", "d_MDtoBC_y", "d_md_qrt2", "d_md_qrt3", "d_md_qrt4"),
         !is.na(level),
         level != "Mammographic breast density quartile",
         level != "Missing"
         ) %>% 
  select(level, estimate_mdl2, conf.low_mdl2, conf.high_mdl2, estimate_mdl4, conf.low_mdl4, conf.high_mdl4
         ) %>% 
  mutate(diff = if_else(is.na(estimate_mdl2), "-", as.character(round((estimate_mdl4 - estimate_mdl2), 2))) 
         ) %>% 
  pivot_longer(
    cols = c(estimate_mdl2, conf.low_mdl2, conf.high_mdl2, 
             estimate_mdl4, conf.low_mdl4, conf.high_mdl4
             ),
    names_to = c(".value", "mdl"),
    names_sep = "_"
  ) 
```

```{r, fig.width=8, fig.height=10}
#| echo: false

# 3. GENERATE FORESTPLOT

is_summary <- is.na(md_forest_long$estimate)


# develop forest

md_forest_long %>%
  group_by(mdl
           ) %>% 
  mutate(mean = estimate,
         lower = conf.low,
         upper = conf.high
         ) %>% 
  forestplot(labeltext = c(level, diff),
             xlog = FALSE,
             xlab = "OR",
             boxsize = 0.3,
             zero = 1,
             line.margin = 0.5,
             clip = c(0, 6),
             ticks = c(0, 1, 2, 3, 4, 5, 6),
             ci.vertices = T,
             ci.vertices.height = 0.2,
             legend = c("Model 2", "Model 4 (adjusted for MD)"),
             title = paste("Final model adjusted vs unadjusted for mamographic breast density quartiles"),
             is.summary = is_summary
             ) %>% 
  fp_set_style(box = c("grey50","deeppink2") %>%  lapply(function(x) gpar(fill = x, col = "#666666")),
               default = gpar(vertices = T),
               line = "grey40",
               summary = gpar(col = NA), # Set summary color to NA
               align = "lrrr",
               hrz_lines = gpar(col = NA), # Set line color to NA
               txt_gp = fpTxtGp(label = gpar(cex = 0.8),
                                    ticks = gpar(cex = 0.8),
                                    xlab = gpar(cex = 0.8),
                                )
               ) %>% 
  fp_decorate_graph(graph.pos = 2
                    ) %>% 
  fp_add_header(level = c("", "Risk factor"),
                diff = c("Absolute difference", "(Model 4 - Model 2)")
                ) %>% 
  fp_insert_row(level = "",
                position = 3,
                mean = NULL,
                lower = NULL,
                upper = NULL,
                is.summary = TRUE,
                boxsize = NA)
```

Note:

Some edits in the forest plot are still needed

\- ordering on the forest plot to be fixed - each row in the plot is also a row in the data behind the plot so not easily sorted by variable

\- BMI categories also jumped but reference is normal weight

\- scale even when specific ticks selected it still defaults to ticks by 0.5 points

# Supplementary material

## ST 1 - TC descriptives and association with detection mode

TC descriptives by detection mode, logistic regression of TC with detection mode (same principal as risk factor analysis but multiple exposures are TC)

::: {.callout-note style="color: orange"}
this will be in excel format in supplement
:::

```{r}
#| echo: false
#| warning: false

# summary Logistic Regression of Model 5  -----------------------------------------


# Adding TC trick variables
 
model <- df %>%
  glm(formula = d_dmode_n ~ 
      d_inv_status + d_grade_tr + d_morph4_tr_lab + d_pos_nodes_tr_lab 
      + d_tmsize_tr_lab + d_er_tr_lab + d_pr_tr_lab + d_her2_tr_lab, 
            family = binomial(link = "logit"))

# tidy up output
model5 <- tbl_regression(model, 
                         exponentiate = TRUE,
                         label = list(
                          d_inv_status ~ "Invasive status",
                          d_grade_tr ~ "Grade",
                          d_morph4_tr_lab ~ "Morphology type",
                          d_pos_nodes_tr_lab ~ "Positive nodes",
                          d_tmsize_tr_lab ~ "Tumour size",
                          d_er_tr_lab ~ "ER status",
                          d_pr_tr_lab ~ "PR status",
                          d_her2_tr_lab ~ "HER2 status"
                         )
                         ) 
  


# summary Logistic Regression of Model 6  -----------------------------------------



# remove columns with total and interval N - not simple with gtsummary, trying from scratch without adding the columns


# Create clean version of logistic output, without Ns

model <- df %>%
  glm(formula = d_dmode_n ~ 
      d_R1menopause_lab3 + d_age_meno_tr_lab 
      + d_R1hrt_sttyp_tr_lab
      + d_parity_lab + d_agebirth1_tr_lab + d_age_menarche_lab + d_ocstatus_lab 
      + d_bbd_lab + d_fambrca_lab
      + d_R1alcohol_units_lab + d_R1smokingstatus_lab 
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab + d_bmi_20_lab + d_bf_dur_tr_lab
      + diagage + d_R1toBC_y + yeardiag
      + d_inv_status + d_grade_tr + d_morph4_tr_lab + d_pos_nodes_tr_lab 
      + d_tmsize_tr_lab + d_er_tr_lab + d_pr_tr_lab + d_her2_tr_lab, 
            family = binomial(link = "logit"))

# tidy up output
model6 <- tbl_regression(model, 
                         exponentiate = TRUE,
                         label = list(
                          d_inv_status ~ "Invasive status",
                          d_grade_tr ~ "Grade",
                          d_morph4_tr_lab ~ "Morphology type",
                          d_pos_nodes_tr_lab ~ "Positive nodes",
                          d_tmsize_tr_lab ~ "Tumour size",
                          d_er_tr_lab ~ "ER status",
                          d_pr_tr_lab ~ "PR status",
                          d_her2_tr_lab ~ "HER2 status",
                          d_R1alcohol_units_lab ~ "Alcohol units",
                          d_R1smokingstatus_lab ~ "Smoking status",
                          d_bmi_entry_lab ~ "BMI at entry",
                          d_R1physmet_leis_who_lab ~ "Physical Activity (WHO cat)",
                          d_bbd_lab ~ "Benign breast disease",
                          d_fambrca_lab ~ "Family history of BC",
                          d_age_menarche_lab ~ "Age at menarche",
                          d_ocstatus_lab ~ "Oral contraceptives status",
                          d_R1menopause_lab3 ~ "Menopause",
                          d_age_meno_tr_lab ~ "Age at menopause",
                          d_R1hrt_sttyp_tr_lab ~ "HRT status",
                          d_parity_lab ~ "N of parous pregnancies",
                          d_agebirth1_tr_lab ~ "Age at first birth",
                          d_bf_dur_tr_lab ~ "Cummulative breastfeeding duration",
                          d_bmi_20_lab ~ "BMI at age 20y"
                         )
                         ) %>% 
  remove_row_type(variables = c(diagage, d_R1toBC_y, yeardiag), # remove covariates
                  type = "all")


# summary Logistic Regression of Model 7  -----------------------------------------



# remove columns with total and interval N - not simple with gtsummary, trying from scratch without adding the columns


# Create clean version of logistic output, without Ns

model <- df %>%
  glm(formula = d_dmode_n ~ d_R1menopause_lab3 + d_age_meno_tr_lab 
      + d_R1hrt_sttyp_tr_lab
      + d_parity_lab + d_agebirth1_tr_lab + d_age_menarche_lab + d_ocstatus_lab 
      + d_bbd_lab + d_fambrca_lab
      + d_R1alcohol_units_lab + d_R1smokingstatus_lab 
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab
      + diagage + d_R1toBC_y + yeardiag + d_bmi_20_lab + d_bf_dur_tr_lab
      + d_md_qrt + d_MDtoBC_lab
      + d_inv_status + d_grade_tr + d_morph4_tr_lab + d_pos_nodes_tr_lab 
      + d_tmsize_tr_lab + d_er_tr_lab + d_pr_tr_lab + d_her2_tr_lab,
            family = binomial(link = "logit"))

# tidy up output
model7 <- tbl_regression(model, 
                         exponentiate = TRUE,
                         label = list(
                          d_inv_status ~ "Invasive status",
                          d_grade_tr ~ "Grade",
                          d_morph4_tr_lab ~ "Morphology type",
                          d_pos_nodes_tr_lab ~ "Positive nodes",
                          d_tmsize_tr_lab ~ "Tumour size",
                          d_er_tr_lab ~ "ER status",
                          d_pr_tr_lab ~ "PR status",
                          d_her2_tr_lab ~ "HER2 status",
                          d_R1alcohol_units_lab ~ "Alcohol units",
                          d_R1smokingstatus_lab ~ "Smoking status",
                          d_bmi_entry_lab ~ "BMI at entry",
                          d_R1physmet_leis_who_lab ~ "Physical Activity (WHO cat)",
                          d_bbd_lab ~ "Benign breast disease",
                          d_fambrca_lab ~ "Family history of BC",
                          d_age_menarche_lab ~ "Age at menarche",
                          d_ocstatus_lab ~ "Oral contraceptives status",
                          d_R1menopause_lab3 ~ "Menopause",
                          d_age_meno_tr_lab ~ "Age at menopause",
                          d_R1hrt_sttyp_tr_lab ~ "HRT status",
                          d_parity_lab ~ "N of parous pregnancies",
                          d_agebirth1_tr_lab ~ "Age at first birth",
                          d_md_qrt ~ "Mammographic density",
                          d_bf_dur_tr_lab ~ "Cummulative breastfeeding duration",
                          d_bmi_20_lab ~ "BMI at age 20y"
                         )
                         ) %>% 
  remove_row_type(variables = c(diagage, d_R1toBC_y, yeardiag, d_MDtoBC_lab), # remove covariates
                  type = "all")

# Create descriptives by mode of detection ------------------------------------
# select(d_dmode_n, d_R1alcohol_units_lab, d_R1smokingstatus_lab, d_bmi_entry_lab,  d_R1physmet_leis_who_lab, d_bbd_lab, d_fambrca_lab, d_age_menarche_lab, d_ocstatus_lab, d_R1menopause_lab3, d_age_meno_tr_lab, d_R1hrt_tr_lab, d_parity_tr_lab, d_agebirth1_tr_lab)

# note: need to rename the non-trick variables according to their trick name to align with the model, which can then allow table merge. 

# include missing 
# create df
  
descr_tc_df <- df %>% 
  select(d_dmode, all_of(tumour_char), all_of(all_expl_variables), d_md_qrt, diagage, d_R1toBC_y 
         ) %>% 
  mutate(d_bmi_entry_lab = fct_relevel(d_bmi_entry_lab, "<18.5", "18.5-25", "25-30", "30+"),
         d_bmi_20_lab = fct_relevel(d_bmi_20_lab, "<18.5", "18.5-25", "25-30", "30+")
         ) %>% 
  mutate(d_grade_tr = d_grade_lab,
         d_morph4_tr_lab = d_morph4_lab,
         d_pos_nodes_tr_lab = d_pos_nodes_lab,
         d_tmsize_tr_lab = d_tumour_size_lab,
         d_er_tr_lab = d_er_status_lab,
         d_pr_tr_lab = d_pr_status_lab,
         d_her2_tr_lab = d_her2_status_lab,
         d_age_meno_tr_lab = d_age_menopause_lab,
         d_R1hrt_sttyp_tr_lab = d_R1hrt_sttyp_lab,
         d_agebirth1_tr_lab = d_agebirth1_lab,
         d_bf_dur_tr_lab = d_bf_dur_lab
         ) %>% 
  select(-d_grade_lab, -d_morph4_lab, -d_pos_nodes_lab, -d_tumour_size_lab, -d_er_status_lab, -d_pr_status_lab, -d_her2_status_lab, -diagage, -d_R1toBC_y, -d_age_menopause_lab, -d_R1hrt_sttyp_lab, -d_agebirth1_lab, -d_bf_dur_lab)  
  


# glimpse(descr_model2)
# str(test)



# create gs summary table
descr_tc <- descr_tc_df %>% 
  # reselect to ensure the correct order
  select(d_dmode,
         d_inv_status, 
         d_grade_tr, 
         d_morph4_tr_lab, 
         d_pos_nodes_tr_lab, 
         d_tmsize_tr_lab, 
         d_er_tr_lab, 
         d_pr_tr_lab, 
         d_her2_tr_lab, 
         d_age_menarche_lab,
         d_ocstatus_lab,
         d_parity_lab,
         d_agebirth1_tr_lab,
         d_bf_dur_tr_lab,
         d_R1menopause_lab3,
         d_age_meno_tr_lab,
         d_R1hrt_sttyp_tr_lab,
         d_bbd_lab,
         d_fambrca_lab,
         d_bmi_entry_lab,
         d_bmi_20_lab,
         d_R1physmet_leis_who_lab,
         d_R1alcohol_units_lab, 
         d_R1smokingstatus_lab,
         d_md_qrt) %>% 
  tbl_summary(by = d_dmode,
              statistic = list(
                all_continuous() ~ "{mean} ({sd})",
                all_categorical() ~ "{n} ({p})"
                ),
              digits = all_continuous() ~ 2,
              label = list(
                          d_inv_status ~ "Invasive status",
                          d_grade_tr ~ "Grade",
                          d_morph4_tr_lab ~ "Morphology type",
                          d_pos_nodes_tr_lab ~ "Positive nodes",
                          d_tmsize_tr_lab ~ "Tumour size",
                          d_er_tr_lab ~ "ER status",
                          d_pr_tr_lab ~ "PR status",
                          d_her2_tr_lab ~ "HER2 status",
                          d_R1alcohol_units_lab ~ "Alcohol units",
                          d_R1smokingstatus_lab ~ "Smoking status",
                          d_bmi_entry_lab ~ "BMI at entry",
                          d_R1physmet_leis_who_lab ~ "Physical Activity (WHO cat)",
                          d_bbd_lab ~ "Benign breast disease",
                          d_fambrca_lab ~ "Family history of BC",
                          d_age_menarche_lab ~ "Age at menarche",
                          d_ocstatus_lab ~ "Oral contraceptives status",
                          d_R1menopause_lab3 ~ "Menopause",
                          d_age_meno_tr_lab ~ "Age at menopause",
                          d_R1hrt_sttyp_tr_lab ~ "HRT status",
                          d_parity_lab ~ "N of parous pregnancies",
                          d_agebirth1_tr_lab ~ "Age at first birth",
                          d_md_qrt ~ "Mammographic density",
                          d_bf_dur_tr_lab ~ "Cummulative breastfeeding duration",
                          d_bmi_20_lab ~ "BMI at age 20y"
                         ),
              type = list(d_bbd_lab ~ "categorical",
                          d_fambrca_lab ~ "categorical")
              ) 


#descr_tc

  
# merge with analysis output -------------------------------------------------


ST1 <- tbl_merge(
  tbls = list(descr_tc, model5, model6, model7),
  tab_spanner = c("**Distribution by detection mode**", "**Model: TC only**", "**Model: TC + RF + time variables**", "**Model: TC + RF + MD + time variables**")
   ) 
   
  
   

ST1
```

## ST 2 - Differences in characteristics between those with and without mammographic density

this table needs additional formatting changes, labels, add p-value for difference between groups, add time from entry to dg and age at diagnosis

any other variables to add or remove?

::: {.callout-note style="color: red"}
## Question

Do we still need this table?
:::

```{r}
#| echo: false

ST2 <- df %>% 
  select(d_dmode, d_md_avail_lab, all_of(all_expl_variables), all_of(tumour_char), diagage, d_R1toBC_y, yeardiag_cat
         ) %>% 
  mutate(d_bmi_entry_lab = fct_relevel(d_bmi_entry_lab, "<18.5", "18.5-25", "25-30", "30+"),
         d_bmi_20_lab = fct_relevel(d_bmi_20_lab, "<18.5", "18.5-25", "25-30", "30+")
              ) %>%
  tbl_summary(by = d_md_avail_lab,
              statistic = list(
                all_continuous() ~ "{mean} ({sd})",
                all_categorical() ~ "{n} ({p})"
                ),
              digits = all_continuous() ~ 2,
              label = list(
                          d_R1alcohol_units_lab ~ "Alcohol units",
                          d_R1smokingstatus_lab ~ "Smoking status",
                          d_bmi_entry_lab ~ "BMI at entry",
                          d_R1physmet_leis_who_lab ~ "Physical Activity (WHO cat)",
                          d_bbd_lab ~ "Benign breast disease",
                          d_fambrca_lab ~ "Family history of BC",
                          d_age_menarche_lab ~ "Age at menarche",
                          d_ocstatus_lab ~ "Oral contraceptives status",
                          d_R1menopause_lab3 ~ "Menopause",
                          d_age_menopause_lab ~ "Age at menopause",
                          d_R1hrt_sttyp_lab ~ "HRT status",
                          d_parity_lab ~ "N of parous pregnancies",
                          d_agebirth1_lab ~ "Age at first birth",
                          d_inv_status ~ "Invasive status",
                          d_grade_lab ~ "Grade",
                          d_morph4_lab ~ "Morphology type",
                          d_pos_nodes_lab ~ "Positive nodes",
                          d_tumour_size_lab ~ "Tumour size",
                          d_er_status_lab ~ "ER status",
                          d_pr_status_lab ~ "PR status",
                          d_her2_status_lab ~ "HER2 status",
                          d_dmode ~ "Mode of detection",
                          yeardiag_cat ~ "Year of diagnosis",
                          d_bf_dur_lab ~ "Cummulative breastfeeding duration",
                          d_bmi_20_lab ~ "BMI at age 20y"
                ),
              type = list(d_bbd_lab ~ "categorical",
                          d_fambrca_lab ~ "categorical")
              ) %>% 
   remove_row_type(variables = c(diagage, d_R1toBC_y), # remove covariates
                  type = "all")

ST2
```

## Cross-tabulation between registry detection mode and in-house detection mode 

```{r}
# use row percentages and rename reg_sd levels to align with dmode

dmode_crosstab_tb <- df |> 
  mutate(
    reg_sd = factor(reg_sd, 
                    levels = c("N", "Y", "U"), 
                    labels = c("I", "SD", "Unknown"))
  ) |> 
  tbl_cross(
    row = reg_sd,
    col = d_dmode,
    percent = "row"
  ) |> 
  bold_labels()

dmode_crosstab_tb |> show_header_names()


dmode_crosstab_tb <- dmode_crosstab_tb |> 
  modify_header(
    label ~ "**Registry Detection Mode**"
  ) |> 
  modify_spanning_header(all_stat_cols() ~ "**In-house Detection Mode**" # Rename column variable
  ) 


dmode_crosstab_tb <- dmode_crosstab_tb |> 
  modify_table_body(
    ~ .x |> filter(label != "reg_sd")  # Remove rows where the label matches the old variable name
  )

dmode_crosstab_tb
```
