---
title: "Detection mode project - Analysis"
author: "Martina Brayley"
date: 05/06/2024
date-format: long
format: 
  html: 
    embed-resources: true
    toc: true
    toc-location: left
    number-sections: true
    theme: [flatly, style.css]
    page-layout: full
editor: visual
---

# Set up

The Master script links all parts of data preparation in 7 processing scripts. The code below excecutes the master file, running all 7 scripts within. The final product is the analytical dataset called an_df which combined all required variables from casummary, risk factors, mode of detection and mammographic density.

```{r}
#| include: false
#| echo: false

# execute master file 
source("C:/Users/MBrayley/OneDrive - The Institute of Cancer Research/Work/GitHub/BC_detection_mode/Martina/code/master.R")
```

Data summary - commented out as output too large

```{r}
#skim(an_df)
```

```{r}
#dfSummary(an_df)
```

# Preparation

Create lists of model variables as objects in the environment for easier manipulation.

Relevel reference groups within some exploratory variables.

Mammographic density is a special variable that will be used for subanalyses on cases with mammographic density only (to adjust for in the model)

```{r}
#| include: false
#| echo: false

## Store explanatory variables as an object -------------------------------

# select categorical variables that are not labelled

# exploratory vars 
all_expl_variables <- c( "d_R1alcohol_units_cat", "d_R1smokingstatus_cat", "d_bmi_entry_cat", "d_R1physmet_leis_who", "d_bbd", "d_fambrca", "d_age_menarche_cat", "d_ocstatus_cat", "d_R1menopause_cat3", "d_age_menopause_cat", "d_R1hrtstatus_cat", "d_parous_cat", "d_parity_cat", "d_agebirth1_cat")
  
lifestyle <- c("d_R1alcohol_status_cat", "d_R1alcohol_units_cat", "d_R1smokingstatus_cat", "d_bmi_entry_cat", "d_R1physmet_leis_who")
  
reproductive <- c("d_age_menarche_cat", "d_ocstatus_cat", "d_R1menopause_cat3", "d_age_menopause_cat", "d_R1hrtstatus_cat", "d_parous_cat", "d_parity_cat", "d_agebirth1_cat")
  
orther_rf <- c("d_bbd", "d_fambrca")

# mammo density excluded from the lists - special variable for subanalyses

# covariates 
all_covariates <- c("diagage", "d_R1toBC_y", "d_R1toMD_y", "d_MDtoBC_y", "d_inv_status", "d_grade_cat", "d_morph4", "d_pos_nodes_cat", "d_tumour_size", "d_er_status", "d_pr_status", "d_her2_status")  
  
tumour_char <- c("d_inv_status", "d_grade_cat", "d_morph4", "d_pos_nodes_cat", "d_tumour_size", "d_er_status", "d_pr_status", "d_her2_status")
  
time_vars <- c("diagage", "d_R1toBC_y", "d_R1toMD_y", "d_MDtoBC_y")


# with labels instead of categories --------------------------------------
# select categorical variables that are not labelled

# exploratory vars 
all_expl_variables <- c("d_R1alcohol_units_lab", "d_R1smokingstatus_lab", "d_bmi_entry_lab", "d_R1physmet_leis_who_lab", "d_bbd_lab", "d_fambrca_lab", "d_age_menarche_lab", "d_ocstatus_lab", "d_R1menopause_lab3", "d_age_menopause_lab", "d_R1hrtstatus_lab", "d_parous_lab", "d_parity_lab", "d_agebirth1_lab")
  
lifestyle <- c("d_R1alcohol_units_lab", "d_R1smokingstatus_lab", "d_bmi_entry_lab", "d_R1physmet_leis_who")
  
reproductive <- c("d_age_menarche_lab", "d_ocstatus_lab", "d_R1menopause_lab3", "d_age_menopause_lab", "d_R1hrtstatus_lab", "d_parous_lab", "d_parity_lab", "d_agebirth1_lab") # do we want to add breastfeeding? If so, it would also have be in the trick model
  
orther_rf <- c("d_bbd", "d_fambrca")

# mammo density excluded from the lists - special variable for subanalyses

# covariates 
all_covariates <- c("diagage", "d_R1toBC_y", "d_R1toMD_y", "d_MDtoBC_y", "d_inv_status", "d_grade_lab", "d_morph4", "d_pos_nodes_lab", "d_tumour_size", "d_er_status", "d_pr_status", "d_her2_status")  
  
tumour_char <- c("d_inv_status", "d_grade_lab", "d_morph4", "d_pos_nodes_lab", "d_tumour_size", "d_er_status", "d_pr_status", "d_her2_status")
  
time_vars <- c("diagage", "d_R1toBC_y", "d_R1toMD_y", "d_MDtoBC_y")
```

Ensure there are no missing values in outcome.

Decide on how to deal with missing values in other variables. When tried to filter out all rows with any missing values in explanatory variables the sample size was very low (only around 600) so complete case analysis may not be an option. In v3 of script 7 error codes were recoded to NA as it is automatically handled in analysis. This dataset should be used. If error codes are of interest, rerun v2 of script 7 in master file, which will produce a dataset with error codes.

```{r}

an_df %>% tabyl(d_dmode_n)

# remove rows with missing error codes in specified columns
# error_codes <- c(888, 777)
# 
# data <- an_df %>%
#   filter(!if_any(all_of(all_expl_variables), ~ . %in% error_codes))



# check numbers and percentages of missing values (coded out as output is too long)

# data %>% 
#   map(pct_miss)
# 
# data %>% 
#   map(n_miss)
```

# Univariate analysis

Repeat these steps for each exposure:

1.  Tabulation including missing
2.  Compute basic OR and chisqr test
3.  Univariate logistic regression

::: callout-important
Code under the empty heading has been excluded from the html file as not relevant
:::

```{r}
#| echo: false
#| include: false

# Example code for single exposure - the function below is derived on this


# # Alcohol status 
# 
# # cross tabulation
# an_df %$% tabpct(d_dmode_n, d_R1alcohol_units_lab, percent = "col", graph = F)
# # OR and chi-squared
# an_df %$% cc(d_dmode_n, d_R1alcohol_units_lab, graph = F)
# 
# # Logistic regression 
# 
# model <- an_df %>% 
#   glm(formula = d_dmode_n ~ d_R1alcohol_units_lab,
#             family = binomial(link = "logit")) 
# 
# # raw output display in log odds scale
# summary(model)
# 
# # using epiDisplay package to display odds ratios
# logistic.display(model)
# 
# # using tidy from broom package to convert output into a dataframe that can be further manipulated if needed
# tidy_model <- model %>% 
#   tidy(exponentiate = TRUE, conf.int = TRUE)
# tidy_model
# 
# 
# 
# 
# 
# 
# 
# 
# # age at first birth status - trying because of error and also has missing values
# 
# # cross tabulation
# data %$% tabpct(d_dmode_n, d_agebirth1_lab, percent = "col", graph = F)
# 
# data %>% tabyl(d_agebirth1_lab)
# 
# an_df %>% tabyl(d_agebirth1_lab)
# 
# # OR and chi-squared
# an_df %$% cc(d_dmode_n, d_agebirth1_lab, graph = F)
# 
# # Logistic regression 
# 
# model <- an_df %>% 
#   filter(d_agebirth1_lab != "Not known") %>% 
#   glm(formula = d_dmode_n ~ d_agebirth1_lab,
#             family = binomial(link = "logit")) 
# 
# # raw output display in log odds scale
# summary(model)
# 
# # using epiDisplay package to display odds ratios
# logistic.display(model)
# 
# # using tidy from broom package to convert output into a dataframe that can be further manipulated if needed
# tidy_model <- model %>% 
#   tidy(exponentiate = TRUE, conf.int = TRUE)
# tidy_model

```

## Define function for univariate analysis

```{r}
#| echo: false
#| include: false

# Univariate analysis function ----------------------------------------------------

# Define the function
perform_univariate_analysis <- function(data, outcome, exposure) {
  # Cross-tabulation with column percentages
  cat("\nCross-tabulation for", exposure, ":\n")
  tab_result <- tabpct(data[[outcome]], data[[exposure]], percent = "col", graph = FALSE)
  print(tab_result)
  
  # Odds Ratio and chi-squared
  cat("\nOdds Ratio and Chi-squared for", exposure, ":\n")
  tryCatch({
    or_chi_result <- cc(data[[outcome]], data[[exposure]], graph = FALSE)
    print(or_chi_result)
  }, error = function(e) {
    cat("An error occurred during OR and chi-squared calculation:", e$message, "\n")
  })
  
  # Logistic Regression
  model <- glm(formula = as.formula(paste(outcome, "~", exposure)),
               data = data,
               family = binomial(link = "logit"))
  
  # Display summary of logistic regression in log odds scale
  # cat("\nLogistic Regression Summary for", exposure, ":\n")
  # print(summary(model))
  
  # Display odds ratios using epiDisplay package
  cat("\nLogistic regression - Odds Ratios for", exposure, ":\n")
  logistic.display(model)
  
  # Capture output of logistic.display
  odds_ratios <- capture.output(logistic.display(model))
  # Print captured output
  cat(odds_ratios, sep = "\n")
}

#test for 1 exposure
# perform_univariate_analysis(data = an_df, outcome = "d_dmode_n", exposure = "d_R1alcohol_units_lab")

```

## Compute univariate analysis

### All cases

```{r}
#| echo: false
#| include: false

# loop for multiple exposures - Risk factors
exposures <- all_expl_variables

for (exposure in exposures) {
  cat("\nAnalyzing exposure:", exposure, "\n")
  perform_univariate_analysis(data = an_df, outcome = "d_dmode_n", exposure = exposure)
}
```

### Cases with Breast density

```{r}
#| echo: false
#| include: false
 
# loop for multiple exposures - Risk factors
exposures <- all_expl_variables

# filter data to include only cases with MD
data <- an_df %>% 
  filter(d_md_avail == 1)

for (exposure in exposures) {
  cat("\nAnalyzing exposure:", exposure, "\n")
  perform_univariate_analysis(data = data, outcome = "d_dmode_n", exposure = exposure)
}

rm(data)
```

## Univariate analysis tables

```{r}
#| echo: false
#| include: false

# set theme for gtsummary tables: 

theme_gtsummary_compact(set_theme = TRUE)
```

::: callout-note
Comparing all cases vs cases with MD only
:::

```{r}

# Univariate analysis with gtsummary

# all cases:
t1 <- an_df %>%
  mutate(d_bmi_entry_lab = fct_relevel(d_bmi_entry_lab, "18.5 to 25", "<18.5", "25 to 30", ">=30")
  ) %>%
  select(d_dmode_n, any_of(all_expl_variables)
         ) %>%
  tbl_uvregression(
    method = glm,
    y = d_dmode_n,
    method.args = list(family = binomial),
    exponentiate = TRUE,
    pvalue_fun = ~ style_pvalue(.x, digits = 2)
   ) %>%
  add_global_p() %>% # add global p-value
  add_nevent() %>% # add number of events of the outcome
  bold_labels() %>% 
  modify_header(stat_nevent = "**Interval N**",
                stat_n = "**Total N**")

#t1

 
# show_header_names(t1) # show header names to rename event N to Interval N


# MD cases only:
t2 <- an_df %>%
  filter(d_md_avail == 1
         ) %>% 
  mutate(d_bmi_entry_lab = fct_relevel(d_bmi_entry_lab, "18.5 to 25", "<18.5", "25 to 30", ">=30")
  ) %>%
  select(d_dmode_n, any_of(all_expl_variables)
         ) %>%
  tbl_uvregression(
    method = glm,
    y = d_dmode_n,
    method.args = list(family = binomial),
    exponentiate = TRUE,
    pvalue_fun = ~ style_pvalue(.x, digits = 2)
   ) %>%
  add_global_p() %>% # add global p-value
  add_nevent() %>% # add number of events of the outcome
  bold_labels() %>% 
  modify_header(stat_nevent = "**Interval N**",
                stat_n = "**Total N**")


# merge tables --------------------------------------------------------------- 


#univariate <- tbl_merge(tbls = list(t1, t2))


univariate <- tbl_merge(
  tbls = list(t1, t2),
  tab_spanner = c("**All cases**", "**Cases with MD**")
   ) 

univariate

  



```

# Test trick models

There are 3 groups of variables that require trick recoding (Tumour characteristics, parity related and menopause related). These variables are closely related and have 1 group within them that is the same in all of the variables. For example, if we look at the parity variables (binary parous, n of parous pregnancies, age at first birth), all 3 variables have a level of "non-parous" women, which all cannot be in the model. Therefore, to trick the model we recode the non-parous women in n of parous preg and age at first birth variables to the reference group. Although, it may seem counter inuitive, when the equation is written down including the indicator variables, the maths works.

When we run 2 models:

1.  a model on a subset of parous women only and with the remaining 2 parity related variables
2.  a model with all cases, the parous binary variable and the 2 recoded parity related variables (non parous women part of the reference group)

The estimates will be exactly the same. This needs to be tested for all 3 groups of these variables.

## Parity related variables

```{r}

# tabulate trick vars by parity to see if recoded properly -----------------------

## age at birth
an_df %>% tabyl(d_agebirth1_tr_lab, d_parous_lab) %>% 
  adorn_title()

## number of parous pregnancies
an_df %>% tabyl(d_parity_tr_lab, d_parous_lab) %>% 
  adorn_title()
# an_df %>% tabyl(d_parity_lab, d_parous_lab) %>% 
#   adorn_title()

# MODEL A - subset on parous women ------------------------------------------------

model <- an_df %>% 
  filter(d_parous_cat == 1) %>%
  glm(formula = d_dmode_n ~ d_parity_lab + d_agebirth1_tr_lab,
            family = binomial(link = "logit"))
#summary(model)
#logistic.display(model)

tbl_regression(model, exponentiate = TRUE) %>% 
  add_n(location = "label") %>% 
  add_nevent() %>% 
  modify_header(stat_nevent = "**Interval N**",
                stat_n = "**Total N**") 



# MODEL B - binary parity status in the model -------------------------------------

model <- an_df %>%
  glm(formula = d_dmode_n ~ d_parous_cat + d_parity_tr_lab + d_agebirth1_tr_lab,
            family = binomial(link = "logit"))
#summary(model)
#logistic.display(model)

tbl_regression(model, exponentiate = TRUE) %>% 
  add_n(location = "label") %>% 
  add_nevent() %>% 
  modify_header(stat_nevent = "**Interval N**",
                stat_n = "**Total N**")




```

::: {.callout-note style="color: green"}
Trick on parity variables **SUCCESSFULL**
:::

## Menopause related variables

```{r}

# tabulate trick vars by parity to see if recoded properly 

## age at menopause
an_df %>% tabyl(d_age_meno_tr_lab, d_R1menopause_lab3) %>% 
  adorn_title()

## HRT use 
an_df %>% tabyl(d_R1hrt_tr_lab, d_R1menopause_lab3) %>% 
  adorn_title()

# MODEL A - subset on postmenopausal  women ---------------------------------
model <- an_df %>% 
  filter(d_R1menopause_cat3 == 1) %>%
  glm(formula = d_dmode_n ~ d_age_meno_tr_lab + d_R1hrt_tr_lab,
            family = binomial(link = "logit"))
#summary(model)
#logistic.display(model)

tbl_regression(model, exponentiate = TRUE) %>% 
  add_n(location = "label") %>% 
  add_nevent() %>% 
  modify_header(stat_nevent = "**Interval N**",
                stat_n = "**Total N**")

# MODEL B - binary parity status in the model -------------------------------------

model <- an_df %>%
  glm(formula = d_dmode_n ~ d_R1menopause_lab3 + d_age_meno_tr_lab + d_R1hrt_tr_lab,
            family = binomial(link = "logit"))
#summary(model)
#logistic.display(model)

tbl_regression(model, exponentiate = TRUE) %>% 
  add_n(location = "label") %>% 
  add_nevent() %>% 
  modify_header(stat_nevent = "**Interval N**",
                stat_n = "**Total N**")
```

::: {.callout-note style="color: green"}
Trick on menopausae variables **SUCCESSFULL**
:::

## Tumour characteristics variables

All DCIS, including those with missing values in the variables below were coded as the reference group. This approach was done also by Louise. Assuming most DCIS would be missing so this seems correct.

```{r}
# tabulate trick vars by parity to see if recoded properly 

## grade
an_df %>% tabyl(d_grade_tr, d_inv_status_lab) %>% 
  adorn_title()

## morphology 
an_df %>% tabyl(d_morph4_tr, d_inv_status_lab) %>% 
  adorn_title()

## positive nodes
an_df %>% tabyl(d_pos_nodes_tr_lab, d_inv_status_lab) %>% 
  adorn_title()

## tumour size
an_df %>% tabyl(d_tmsize_tr, d_inv_status_lab) %>% 
  adorn_title()

## er status - reference positive
an_df %>% tabyl(d_er_tr_lab, d_inv_status_lab) %>% 
  adorn_title()

## pr status - reference positive
dev_an_df %>% tabyl(d_pr_tr_lab, d_inv_status_lab) %>% 
  adorn_title()

## her2 status - reference positive
an_df %>% tabyl(d_her2_tr_lab, d_inv_status_lab) %>% 
  adorn_title()



# MODEL A - subset on postmenopausal  women ---------------------------------

# note: keeping morhology as categorical without labels as it affects the outpu

model <- an_df %>% 
  filter(d_inv_status == 1) %>%
  glm(formula = d_dmode_n ~ d_grade_lab + d_morph4 + d_pos_nodes_lab + d_tumour_size + d_er_status + d_pr_status + d_her2_status,
            family = binomial(link = "logit"))
#summary(model)
#logistic.display(model)

tbl_regression(model, exponentiate = TRUE) %>% 
  add_n(location = "label") %>% 
  add_nevent() %>% 
  modify_header(stat_nevent = "**Interval N**",
                stat_n = "**Total N**")

# MODEL B - binary parity status in the model -------------------------------------

model <- an_df %>%
  glm(formula = d_dmode_n ~ d_inv_status + d_grade_tr + d_morph4_tr + d_pos_nodes_tr_lab + d_tmsize_tr_lab + d_er_tr_lab + d_pr_tr_lab + d_her2_tr_lab,
            family = binomial(link = "logit"))

#summary(model)
#logistic.display(model)

tbl_regression(model, exponentiate = TRUE) %>% 
  add_n(location = "label") %>% 
  add_nevent() %>% 
  modify_header(stat_nevent = "**Interval N**",
                stat_n = "**Total N**")
```

::: {.callout-note style="color: green"}
Trick on Tumour characteristics variables **SUCCESSFULL**
:::

## Conclusion of trick test

Trick test on all 3 sets of variables was successful. This means that we can use the trick variables in the multivariate analysis.

# Multivariate analysis

This is for the main analysis, subanalyses will be an independent section

The main analysis is comprised of 2 parts:

1.  Examining the relationship between risk factors and mode of detection
2.  Adjusting for mammographic density in a subset of cases with density available

## Part 1: Relationship between risk factors and mode of detection

An analysis with all selected exposures in the model (adjusted for each other). Also, will need to adjust for time and age variables. Then, adjust for tumour characteristics and explore the difference. If there is no or minimal difference the model adjusted for tumour characteristics will be presented in the supplement.

Therefore, part 1 of the multivariate analysis will include 3 models:

1.  Minimally adjusted - including all risk factor exposures
2.  Adjusted for time variables
3.  Fully adjusted - time variables and tumour characteristics.

**Model 1**: "d_R1alcohol_status_cat", "d_R1alcohol_units_cat", "d_R1smokingstatus_cat", "d_bmi_entry_cat", "d_R1physmet_leis_who", "d_bbd", "d_fambrca", "d_age_menarche_cat", "d_ocstatus_cat", "d_R1menopause_cat3", "d_age_menopause_cat", "d_R1hrtstatus_cat", "d_parous_cat", "d_parity_cat", "d_agebirth1_cat"

**Model 2**: Model 1 + "diagage", "d_R1toBC_y"

**Model 3**: subset of those with TC! Model 1 + "diagage", "d_R1toBC_y",

**Model 4**: Model 3 + "d_inv_status", "d_grade_cat", "d_morph4", "d_pos_nodes_cat", "d_tumour_size", "d_er_status", "d_pr_status", "d_her2_status"

### Code

```{r}

# Model 1 -----------------------------------------------------------
# include parity and meno trick variable instead of original variables

# only included alcohol status - check if we need to do the trick if including both
# relevel bmis so normal weight is the reference 

model <- an_df %>%
  mutate(d_bmi_entry_lab = relevel(d_bmi_entry_lab, ref = "18.5 to 25")
         ) %>% 
  glm(formula = d_dmode_n ~ d_R1alcohol_units_lab + d_R1smokingstatus_lab 
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab + d_bbd_lab + d_fambrca_lab
      + d_age_menarche_lab + d_ocstatus_lab 
      + d_R1menopause_cat3 + d_age_meno_tr_lab + d_R1hrt_tr_lab 
      + d_parous_lab + d_parity_tr_lab + d_agebirth1_tr_lab,
            family = binomial(link = "logit"))

# summary(model)
#logistic.display(model)

# # display result as a tidy model table with OR  
#  tidy_model <- model %>% 
#    tidy(exponentiate = TRUE, conf.int = TRUE) %>% 
#    mutate(OR = estimate) %>% 
#    select(-estimate) %>% 
#    select(term, OR, everything())
#  tidy_model 

model1 <- tbl_regression(model, exponentiate = TRUE) %>% 
  add_n(location = "level")  %>% 
  add_nevent(location = "level") %>% 
  add_global_p() %>% 
  modify_header(stat_nevent = "**Interval N**",
                stat_n = "**Total N**")
 
# model1
 
 # Model 2 ----------------------------------------------------------------
 # include parity and meno trick variable instead of original variables

# only included alcohol status - check if we need to do the trick if including both
# including "diagage", "d_R1toBC_y" - MD variables will be included in sub analysis

model <- an_df %>%
   mutate(d_bmi_entry_lab = relevel(d_bmi_entry_lab, ref = "18.5 to 25")
         ) %>%
  glm(formula = d_dmode_n ~ d_R1alcohol_units_lab + d_R1smokingstatus_lab 
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab + d_bbd_lab + d_fambrca_lab
      + d_age_menarche_lab + d_ocstatus_lab 
      + d_R1menopause_cat3 + d_age_meno_tr_lab + d_R1hrt_tr_lab 
      + d_parous_lab + d_parity_tr_lab + d_agebirth1_tr_lab
      + diagage + d_R1toBC_y,
            family = binomial(link = "logit"))

# summary(model)
# logistic.display(model)

# display result as a tidy model table with OR  
 # tidy_model <- model %>%
 #   tidy(exponentiate = TRUE, conf.int = TRUE) %>%
 #   mutate(OR = estimate) %>%
 #   select(-estimate) %>%
 #   select(term, OR, everything())
 # tidy_model

model2 <- tbl_regression(model, exponentiate = TRUE) %>% 
  add_n(location = "level")  %>% 
  add_nevent(location = "level") %>% 
  add_global_p() %>% 
  modify_header(stat_nevent = "**Interval N**",
                stat_n = "**Total N**")



# Model 3  ----------------------------------------------------------------
#same as model 2 but on the subset of those with TC

 # include parity and meno trick variable instead of original variables

# only included alcohol status - check if we need to do the trick if including both
# including "diagage", "d_R1toBC_y" - MD variables will be included in sub analysis

test <- an_df %>%
   drop_na(d_inv_status, d_grade_tr, d_morph4_tr_lab, d_pos_nodes_tr_lab, d_tmsize_tr_lab, d_er_tr_lab, d_pr_tr_lab, d_her2_tr_lab)
 
 
 model <- an_df %>%
   drop_na(d_inv_status, d_grade_tr, d_morph4_tr_lab, 
           d_pos_nodes_tr_lab, d_tmsize_tr_lab, 
           d_er_tr_lab, d_pr_tr_lab, d_her2_tr_lab
           ) %>% 
   mutate(d_bmi_entry_lab = relevel(d_bmi_entry_lab, ref = "18.5 to 25")
         ) %>%
  glm(formula = d_dmode_n ~ d_R1alcohol_units_lab + d_R1smokingstatus_lab 
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab + d_bbd_lab + d_fambrca_lab
      + d_age_menarche_lab + d_ocstatus_lab 
      + d_R1menopause_cat3 + d_age_meno_tr_lab + d_R1hrt_tr_lab 
      + d_parous_lab + d_parity_tr_lab + d_agebirth1_tr_lab
      + diagage + d_R1toBC_y,
            family = binomial(link = "logit"))

# summary(model)
# logistic.display(model)

# display result as a tidy model table with OR
# tidy_model <- model %>%
#   tidy(exponentiate = TRUE, conf.int = TRUE) %>%
#   mutate(OR = estimate) %>%
#   select(-estimate) %>%
#   select(term, OR, everything())
# tidy_model
 
model3 <- tbl_regression(model, exponentiate = TRUE) %>% 
  add_n(location = "level")  %>% 
  add_nevent(location = "level") %>% 
  add_global_p() %>% 
  modify_header(stat_nevent = "**Interval N**",
                stat_n = "**Total N**")

#model3

# Model 4 ---------------------------------------------------------------------

# Adding TC trick variables
 
model <- an_df %>%
   mutate(d_bmi_entry_lab = relevel(d_bmi_entry_lab, ref = "18.5 to 25")
         ) %>%
  glm(formula = d_dmode_n ~ d_R1alcohol_units_lab + d_R1smokingstatus_lab 
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab + d_bbd_lab + d_fambrca_lab
      + d_age_menarche_lab + d_ocstatus_lab 
      + d_R1menopause_cat3 + d_age_meno_tr_lab + d_R1hrt_tr_lab 
      + d_parous_lab + d_parity_tr_lab + d_agebirth1_tr_lab
      + diagage + d_R1toBC_y
      + d_inv_status + d_grade_tr + d_morph4_tr + d_pos_nodes_tr_lab 
      + d_tmsize_tr_lab + d_er_tr_lab + d_pr_tr_lab + d_her2_tr_lab, 
            family = binomial(link = "logit"))

# summary(model)
# logistic.display(model)
 

# display result as a tidy model table with OR
 # tidy_model <- model %>%
 #   tidy(exponentiate = TRUE, conf.int = TRUE) %>%
 #   mutate(OR = round(estimate, 2),
 #          std.error = round(std.error, 2),
 #          p.value = round(p.value, 4),
 #          conf.low = round(conf.low)
 #          ) %>%
 #   select(-estimate) %>%
 #   select(term, OR, everything())
 # tidy_model

model4 <- tbl_regression(model, exponentiate = TRUE) %>% 
  add_n(location = "level")  %>% 
  add_nevent(location = "level") %>% 
  add_global_p() %>% 
  modify_header(stat_nevent = "**Interval N**",
                stat_n = "**Total N**")




```

### Tables

These tables are for internal purposes only.

::: panel-tabset
## All cases

***Minimally adjusted vs time and age adjusted***

Model 1 - Minimally adjusted - only for other exposures

Model 2 - adjusted for time from entry to BC and age at diagnosis

```{r}
# create table for model 1 and 2 

model1and2 <- tbl_merge(
  tbls = list(model1, model2),
  tab_spanner = c("**Minimally adjusted**", "**Time and age adjusted**")
   ) 

model1and2
```

::: callout-note
## Summary

No changes in estimates after adjusting for age at diagnosis and time from entry to diagnosis
:::

## Cases with TC

**Adjusting for tumour characteristics**

Model 3 - model 2 (adjusted model) on subset of cases with TC

Model 4 - adjusting for TC

Model 3 and 4 have the same number of observations n = 857. After adjusting for tumour characteristics, there are some changes in the estimates (e.g. BMI - underweight group although CI large, BBD larger effect)

```{r}
model3and4 <- tbl_merge(
  tbls = list(model3, model4),
  tab_spanner = c("**Time and age adjusted**", "**Adjusted for TC**")
   ) 

model3and4

```

::: callout-note
## Summary

Relatively large changes in estimates for

-   BMI category \<18.5kg/m2 (underweight), although the confidence interval is very wide.

-   Benign breast disease - stronger effect after adjusting for tumour characteristics.
:::
:::

## Part 2: Adjusting for mammographic density

This part will require the same analysis as in part 1 on a **subset of cases with mamographic density**. Then expanding by a 4th model adjusting for mammographic density.

Not modelling TC here as the sample size would be significantly reduced.

**Model 5**: "d_R1alcohol_status_cat", "d_R1alcohol_units_cat", "d_R1smokingstatus_cat", "d_bmi_entry_cat", "d_R1physmet_leis_who", "d_bbd", "d_fambrca", "d_age_menarche_cat", "d_ocstatus_cat", "d_R1menopause_cat3", "d_age_menopause_cat", "d_R1hrtstatus_cat", "d_parous_cat", "d_parity_cat", "d_agebirth1_cat"

**Model 6**: Model 4 + "diagage", "d_R1toBC_y"

**Model 7**: Model 5 + "d_R1toMD_y", "d_MDtoBC_y" \[ decide whether to include all of these\]

**Model 8**: Model 6 +"d_md_qrt" \[as quartiles\]

### Code

```{r}


# Model 5 -----------------------------------------------------------
# include parity and meno trick variable instead of original variables

# only included alcohol status - check if we need to do the trick if including both
# relevel bmis so normal weight is the reference 

model <- an_df %>%
  filter(d_md_avail == 1 # include only those with md
         ) %>% 
  mutate(d_bmi_entry_lab = relevel(d_bmi_entry_lab, ref = "18.5 to 25") # relevel BMI
         ) %>% 
  glm(formula = d_dmode_n ~ d_R1alcohol_units_lab + d_R1smokingstatus_lab 
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab + d_bbd_lab + d_fambrca_lab
      + d_age_menarche_lab + d_ocstatus_lab 
      + d_R1menopause_cat3 + d_age_meno_tr_lab + d_R1hrt_tr_lab 
      + d_parous_lab + d_parity_tr_lab + d_agebirth1_tr_lab,
            family = binomial(link = "logit"))

# summary(model)
# logistic.display(model)

# display result as a tidy model table with OR  
 # tidy_model <- model %>% 
 #   tidy(exponentiate = TRUE, conf.int = TRUE) %>% 
 #   mutate(OR = estimate) %>% 
 #   select(-estimate) %>% 
 #   select(term, OR, everything())
 # tidy_model 

model5 <- tbl_regression(model, exponentiate = TRUE) %>% 
  add_n(location = "level")  %>% 
  add_nevent(location = "level") %>% 
  add_global_p() %>% 
  modify_header(stat_nevent = "**Interval N**",
                stat_n = "**Total N**")

 
 
 # Model 6 ----------------------------------------------------------------
 # include parity and meno trick variable instead of original variables

# only included alcohol status - check if we need to do the trick if including both
# including "diagage", "d_R1toBC_y" - MD variables will be included in sub analysis

model <- an_df %>%
   filter(d_md_avail == 1 # include only those with md
         ) %>%
   mutate(d_bmi_entry_lab = relevel(d_bmi_entry_lab, ref = "18.5 to 25")
         ) %>%
  glm(formula = d_dmode_n ~ d_R1alcohol_units_lab + d_R1smokingstatus_lab 
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab + d_bbd_lab + d_fambrca_lab
      + d_age_menarche_lab + d_ocstatus_lab 
      + d_R1menopause_cat3 + d_age_meno_tr_lab + d_R1hrt_tr_lab 
      + d_parous_lab + d_parity_tr_lab + d_agebirth1_tr_lab
      + diagage + d_R1toBC_y,
            family = binomial(link = "logit"))

# summary(model)
# logistic.display(model)

# # display result as a tidy model table with OR  
#  tidy_model <- model %>% 
#    tidy(exponentiate = TRUE, conf.int = TRUE) %>% 
#    mutate(OR = estimate) %>% 
#    select(-estimate) %>% 
#    select(term, OR, everything())
#  tidy_model 

model6 <- tbl_regression(model, exponentiate = TRUE) %>% 
  add_n(location = "level")  %>% 
  add_nevent(location = "level") %>% 
  add_global_p() %>% 
  modify_header(stat_nevent = "**Interval N**",
                stat_n = "**Total N**")
 # Model 7 ----------------------------------------------------------------
 # include parity and meno trick variable instead of original variables

# only included alcohol status - check if we need to do the trick if including both
# including "diagage", "d_R1toBC_y" and d_MDtoBC_y - not adjusted for R1toMD yet - discuss if needed

model <- an_df %>%
   filter(d_md_avail == 1 # include only those with md
         ) %>%
   mutate(d_bmi_entry_lab = relevel(d_bmi_entry_lab, ref = "18.5 to 25")
         ) %>%
  glm(formula = d_dmode_n ~ d_R1alcohol_units_lab + d_R1smokingstatus_lab 
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab + d_bbd_lab + d_fambrca_lab
      + d_age_menarche_lab + d_ocstatus_lab 
      + d_R1menopause_cat3 + d_age_meno_tr_lab + d_R1hrt_tr_lab 
      + d_parous_lab + d_parity_tr_lab + d_agebirth1_tr_lab
      + diagage + d_R1toBC_y + d_MDtoBC_y,
            family = binomial(link = "logit"))

# summary(model)
# logistic.display(model)

# # display result as a tidy model table with OR  
#  tidy_model <- model %>% 
#    tidy(exponentiate = TRUE, conf.int = TRUE) %>% 
#    mutate(OR = estimate) %>% 
#    select(-estimate) %>% 
#    select(term, OR, everything())
#  tidy_model

model7 <- tbl_regression(model, exponentiate = TRUE) %>% 
  add_n(location = "level")  %>% 
  add_nevent(location = "level") %>% 
  add_global_p() %>% 
  modify_header(stat_nevent = "**Interval N**",
                stat_n = "**Total N**")
 
# Model 8 ----------------------------------------------------------------
 # include parity and meno trick variable instead of original variables

# only included alcohol status - check if we need to do the trick if including both
# including "diagage", "d_R1toBC_y" - MD variables will be included in sub analysis

model <- an_df %>%
   filter(d_md_avail == 1 # include only those with md
         ) %>%
   mutate(d_bmi_entry_lab = relevel(d_bmi_entry_lab, ref = "18.5 to 25")
         ) %>%
  glm(formula = d_dmode_n ~ d_R1alcohol_status_lab + d_R1smokingstatus_lab 
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab + d_bbd_lab + d_fambrca_lab
      + d_age_menarche_lab + d_ocstatus_lab 
      + d_R1menopause_cat3 + d_age_meno_tr_lab + d_R1hrt_tr_lab 
      + d_parous_lab + d_parity_tr_lab + d_agebirth1_tr_lab
      + diagage + d_R1toBC_y + d_MDtoBC_y
      + d_md_qrt,
            family = binomial(link = "logit"))

# summary(model)
# logistic.display(model)

# # display result as a tidy model table with OR  
#  tidy_model <- model %>% 
#    tidy(exponentiate = TRUE, conf.int = TRUE) %>% 
#    mutate(OR = estimate) %>% 
#    select(-estimate) %>% 
#    select(term, OR, everything())
#  tidy_model

model8 <- tbl_regression(model, exponentiate = TRUE) %>% 
  add_n(location = "level")  %>% 
  add_nevent(location = "level") %>% 
  add_global_p() %>% 
  modify_header(stat_nevent = "**Interval N**",
                stat_n = "**Total N**")

```

### Tables 

This table is for internal purposes only.

```{r}
# combine results

model5to8 <- tbl_merge(
  tbls = list(model5, model6, model7, model8),
  tab_spanner = c("**Unadjusted**", "**Adjusted for DG age + entry to DG**", "**Adjusted for MDtoDG time**", "**Adjusted for MDtoDG time + MD**")
   ) 

model5to8
```

::: callout-note
## Summary

No significant changes after adjusting for mammographic density.
:::

## Part 3: Tumour characteristics only

For internal purpose or supplemental materials.

**Model 9**: TC only "d_inv_status", "d_grade_cat", "d_morph4", "d_pos_nodes_cat", "d_tumour_size", "d_er_status", "d_pr_status", "d_her2_status"

**Model 10**: "d_inv_status", "d_grade_cat", "d_morph4", "d_pos_nodes_cat", "d_tumour_size", "d_er_status", "d_pr_status", "d_her2_status", "diagage", "R1toBC"

### Code

```{r}

# Model 9 ------------------------------------------------------------------------

# Adding TC trick variables
 
model <- an_df %>%
  glm(formula = d_dmode_n ~ 
      d_inv_status + d_grade_tr + d_morph4_tr_lab + d_pos_nodes_tr_lab 
      + d_tmsize_tr_lab + d_er_tr_lab + d_pr_tr_lab + d_her2_tr_lab, 
            family = binomial(link = "logit"))

# summary(model)
# logistic.display(model)
 

# display result as a tidy model table with OR
 # tidy_model <- model %>%
 #   tidy(exponentiate = TRUE, conf.int = TRUE) %>%
 #   mutate(OR = round(estimate, 2),
 #          std.error = round(std.error, 2),
 #          p.value = round(p.value, 4),
 #          conf.low = round(conf.low)
 #          ) %>%
 #   select(-estimate) %>%
 #   select(term, OR, everything())
 # tidy_model

model9 <- tbl_regression(model, exponentiate = TRUE) %>% 
  add_n(location = "level")  %>% 
  add_nevent(location = "level") %>% 
  add_global_p() %>% 
  modify_header(stat_nevent = "**Interval N**",
                stat_n = "**Total N**")

# Model 9 ------------------------------------------------------------------------

# Adding TC trick variables
 
model <- an_df %>%
  glm(formula = d_dmode_n ~ 
      diagage + d_R1toBC_y
      + d_inv_status + d_grade_tr + d_morph4_tr_lab + d_pos_nodes_tr_lab 
      + d_tmsize_tr_lab + d_er_tr_lab + d_pr_tr_lab + d_her2_tr_lab, 
            family = binomial(link = "logit"))

# summary(model)
# logistic.display(model)
 

# display result as a tidy model table with OR
 # tidy_model <- model %>%
 #   tidy(exponentiate = TRUE, conf.int = TRUE) %>%
 #   mutate(OR = round(estimate, 2),
 #          std.error = round(std.error, 2),
 #          p.value = round(p.value, 4),
 #          conf.low = round(conf.low)
 #          ) %>%
 #   select(-estimate) %>%
 #   select(term, OR, everything())
 # tidy_model

model10 <- tbl_regression(model, exponentiate = TRUE) %>% 
  add_n(location = "level")  %>% 
  add_nevent(location = "level") %>% 
  add_global_p() %>% 
  modify_header(stat_nevent = "**Interval N**",
                stat_n = "**Total N**")

```

### Table

This table is for internal purpose only.

```{r}
# combine results 


model9and10 <- tbl_merge(
  tbls = list(model9, model10),
  tab_spanner = c("**Unadjusted**", "**Adjusted for DG age + entry to DG**")
   ) 

model9and10
```

::: callout-note
## Summary

Grade 3 of BC, positive nodes, tumour size of 21-50mm associated with being an interval case. Adjusting for age at diagnosis and time from entry to diagnosis did not change the estimates.
:::

# Draft tables for publication

This section is used for developing tables presenting analyses results for publication.

Table 1 and 2 are descriptive tables, which will be created in a separate Quarto file (Screening descriptives). Once the publication tables are finalised, new Quarto document will be produced with the final tables that can be then exported to different formats.

### Table structure: 

**Main tables**

Table 1 - general descriptives (population characteristics)

Table 2 - risk factor descriptives by density (+ figures)

Table 3 - Risk factor descriptives by detection mode with logistic regression results (unadjusted for TC) - all cases

Table 4 - subset on cases with MD. Same model as table 3 + adjusted for MD. Include mediation measures (% change)

**Supplements**

ST 1 - Descriptives for MD cases (same as table 1 but subset)

ST 2 - TC descriptives by detection mode with logistic regression results

ST 3 - Risk factor analysis adjusted for tumour characteristics (sensitivity analysis) - on subset of cases with TC

## Table 3 - Main analysis

Model 2 is the final selected model to present in the analysis. Model 2 is adjusted for all co-exposures + age at diagnosis and time from entry to BC diagnosis

```{r}

# summary Logistic Regression of Model 2  -----------------------------------------

# view existing output
model2

# remove columns with total and interval N - not simple with gtsummary, trying from scratch without adding the columns


# Create clean version of logistic output, without Ns
model <- an_df %>%
   mutate(d_bmi_entry_lab = relevel(d_bmi_entry_lab, ref = "18.5 to 25")
         ) %>%
  glm(formula = d_dmode_n ~ d_R1alcohol_units_lab + d_R1smokingstatus_lab 
      + d_bmi_entry_lab + d_R1physmet_leis_who_lab + d_bbd_lab + d_fambrca_lab
      + d_age_menarche_lab + d_ocstatus_lab 
      + d_R1menopause_cat3 + d_age_meno_tr_lab + d_R1hrt_tr_lab 
      + d_parous_lab + d_parity_tr_lab + d_agebirth1_tr_lab
      + diagage + d_R1toBC_y,
            family = binomial(link = "logit"))


model2 <- tbl_regression(model, exponentiate = TRUE) %>% 
  add_global_p() %>% 
  remove_row_type(variables = c(diagage, d_R1toBC_y), # remove covariates
                  type = "all")
  

model2





# Create descriptives by mode of detection ------------------------------------
# select(d_dmode_n, d_R1alcohol_units_lab, d_R1smokingstatus_lab, d_bmi_entry_lab,  d_R1physmet_leis_who_lab, d_bbd_lab, d_fambrca_lab, d_age_menarche_lab, d_ocstatus_lab, d_R1menopause_cat3, d_age_meno_tr_lab, d_R1hrt_tr_lab, d_parous_lab, d_parity_tr_lab, d_agebirth1_tr_lab)

descr_model2 <- an_df %>% 
  select(d_dmode, all_of(all_expl_variables))

names(descr_model2)




 an_df %>% 
  select(d_dmode, d_R1alcohol_units_lab, d_R1smokingstatus_lab, d_bmi_entry_lab, d_R1physmet_leis_who_lab, d_bbd_lab, d_fambrca_lab, d_age_menarche_lab, d_ocstatus_lab, d_R1menopause_cat3, d_age_meno_lab, d_R1hrt__lab, d_parous_lab, d_parity_tr_lab, d_agebirth1_tr_lab
         ) %>%
  tbl_summary(
    by = d_dmode,
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p})"
      ),
    digits = all_continuous() ~ 2,
    label = c(d_R1alcohol_units_lab ~ "Alcohol units",
              d_R1smokingstatus_lab ~ "Smoking status",
              d_bmi_entry_lab ~ "BMI",
              d_R1physmet_leis_who_lab ~ "Physical Activity (WHO cat)",
              d_bbd_lab ~ "Benign breast disease",
              d_fambrcaN_lab ~ "Family history of BC",
              d_age_menarche_lab ~ "Age at menarche",
              d_ocstatus_lab ~ "Oral contraceptives status",
              d_R1menopause_lab3 ~ "Menopause",
              d_age_menopause_lab ~ "Age at menopause",
              d_R1hrtstatus_lab ~ "HRT status",
              d_parous_lab ~ "Parous",
              d_parity_lab ~ "N of parous pregnancies",
              d_agebirth1_lab ~ "Age at first birth")
    
              )

  


descr_model2
label = c(d_R1alcohol_units_lab ~ "Alcohol units",
              d_R1smokingstatus_lab ~ "Smoking status",
              d_bmi_entry_lab ~ "BMI",
              d_R1physmet_leis_who_lab ~ "Physical Activity (WHO cat)",
              d_bbd_lab ~ "Benign breast disease",
              d_fambrcaN_lab ~ "Family history of BC",
              d_age_menarche_lab ~ "Age at menarche",
              d_ocstatus_lab ~ "Oral contraceptives status",
              d_R1menopause_lab3 ~ "Menopause",
              d_age_menopause_lab ~ "Age at menopause",
              d_R1hrtstatus_lab ~ "HRT status",
              d_parous_lab ~ "Parous",
              d_parity_lab ~ "N of parous pregnancies",
              d_agebirth1_lab ~ "Age at first birth"

# merge with analysis output -------------------------------------------------






# format the table -------------------------------------------------------------
```
