---
title: "Analysis"
author: "MB"
format: 
  html: 
    embed-resources: true
editor: visual
---

# Set up

The Master script links all parts of data preparation in 7 processing scripts. The code below excecutes the master file, running all 7 scripts within. The final product is the analytical dataset called an_df which combined all required variables from casummary, risk factors, mode of detection and mammographic density.

```{r}
#| include: false

# execute master file 
source("C:/Users/MBrayley/OneDrive - The Institute of Cancer Research/Work/GitHub/BC_detection_mode/Martina/code/master.R")
```

Data summary

```{r}
skim(an_df)
```

```{r}
dfSummary(an_df)
```

# 1. Preparation

Create lists of model variables as objects in the environment for easier manipulation.

Relevel reference groups within some exploratory variables.

Mammographic density is a special variable that will be used for subanalyses on cases with mammographic density only (to adjust for in the model)

```{r}
## Store explanatory variables as an object -------------------------------

# select categorical variables that are not labelled

# exploratory vars 
all_expl_variables <- c("d_R1alcohol_status_cat", "d_R1alcohol_units_cat", "d_R1smokingstatus_cat", "d_bmi_entry_cat", "d_R1physmet_leis_who", "d_bbd", "d_fambrca", "d_age_menarche_cat", "d_ocstatus_cat", "d_R1menopause_cat3", "d_age_menopause_cat", "d_R1hrtstatus_cat", "d_parous_cat", "d_parity_cat", "d_agebirth1_cat")
  
lifestyle <- c("d_R1alcohol_status_cat", "d_R1alcohol_units_cat", "d_R1smokingstatus_cat", "d_bmi_entry_cat", "d_R1physmet_leis_who")
  
reproductive <- c("d_age_menarche_cat", "d_ocstatus_cat", "d_R1menopause_cat3", "d_age_menopause_cat", "d_R1hrtstatus_cat", "d_parous_cat", "d_parity_cat", "d_agebirth1_cat")
  
orther_rf <- c("d_bbd", "d_fambrca")

# mammo density excluded from the lists - special variable for subanalyses

# covariates 
all_covariates <- c("diagage", "d_R1toBC_y", "d_R1toMD_y", "d_MDtoBC_y", "d_inv_status", "d_grade_cat", "d_morph7", "d_pos_nodes_cat", "d_tumour_size", "d_er_status", "d_pr_status", "d_her2_status")  
  
tumour_char <- c("d_inv_status", "d_grade_cat", "d_morph7", "d_pos_nodes_cat", "d_tumour_size", "d_er_status", "d_pr_status", "d_her2_status")
  
time_vars <- c("diagage", "d_R1toBC_y", "d_R1toMD_y", "d_MDtoBC_y")


# with labels instead of categories --------------------------------------
# select categorical variables that are not labelled

# exploratory vars 
all_expl_variables <- c("d_R1alcohol_status_lab", "d_R1alcohol_units_lab", "d_R1smokingstatus_lab", "d_bmi_entry_lab", "d_R1physmet_leis_who", "d_bbd", "d_fambrca", "d_age_menarche_lab", "d_ocstatus_lab", "d_R1menopause_lab3", "d_age_menopause_lab", "d_R1hrtstatus_lab", "d_parous_lab", "d_parity_lab", "d_agebirth1_lab")
  
lifestyle <- c("d_R1alcohol_status_lab", "d_R1alcohol_units_lab", "d_R1smokingstatus_lab", "d_bmi_entry_lab", "d_R1physmet_leis_who")
  
reproductive <- c("d_age_menarche_lab", "d_ocstatus_lab", "d_R1menopause_lab3", "d_age_menopause_lab", "d_R1hrtstatus_lab", "d_parous_lab", "d_parity_lab", "d_agebirth1_lab")
  
orther_rf <- c("d_bbd", "d_fambrca")

# mammo density excluded from the lists - special variable for subanalyses

# covariates 
all_covariates <- c("diagage", "d_R1toBC_y", "d_R1toMD_y", "d_MDtoBC_y", "d_inv_status", "d_grade_lab", "d_morph7", "d_pos_nodes_lab", "d_tumour_size", "d_er_status", "d_pr_status", "d_her2_status")  
  
tumour_char <- c("d_inv_status", "d_grade_lab", "d_morph7", "d_pos_nodes_lab", "d_tumour_size", "d_er_status", "d_pr_status", "d_her2_status")
  
time_vars <- c("diagage", "d_R1toBC_y", "d_R1toMD_y", "d_MDtoBC_y")
```

Ensure there are no missing values in outcome.

Decide on how to deal with missing values in other variables. When tried to filter out all rows with any missing values in explanatory variables the sample size was very low (only around 600) so complete case analysis is not an option.

```{r}

an_df %>% tabyl(d_dmode_n)

# remove rows with missing error codes in specified columns
# error_codes <- c(888, 777)
# 
# data <- an_df %>%
#   filter(!if_any(all_of(all_expl_variables), ~ . %in% error_codes))

```

# 2. Univariate analysis

Trick models - test model with

# test

```{r}

# option 1 
model <- an_df %>% 
  glm(formula = d_dmode_n ~ d_R1alcohol_status_lab,
            family = binomial(link = "logit"))
summary(model)

# create tidy model with original estimates 
tidy_model <- tidy(model)
conf_intervals <- confint(model)


# Get confidence intervals and convert to a data frame
conf_intervals <- confint(model)
conf_intervals_df <- as.data.frame(conf_intervals)
colnames(conf_intervals_df) <- c("conf.low", "conf.high")

# Ensure the term column is added to conf_intervals_df
conf_intervals_df$term <- rownames(conf_intervals_df)

# Combine the tidy model and confidence intervals
tidy_model <- tidy_model %>%
  left_join(conf_intervals_df, by = "term") %>%
  mutate(
    odds_ratio = exp(estimate),
    conf.low.odds = exp(conf.low),
    conf.high.odds = exp(conf.high)
  ) %>%
  select(term, estimate, std.error, conf.low, conf.high, statistic, p.value, odds_ratio, conf.low.odds, conf.high.odds)


tidy_model




# option 2 exponentiate in one

tidy_model <- an_df %>% 
  glm(formula = d_dmode_n ~ d_R1alcohol_status_lab,
            family = binomial(link = "logit")) %>% 
  tidy(exponentiate = TRUE, conf.int = TRUE) #%>%  # exponentiate and produce CIs
  #mutate(across(where(is.numeric), round, digits = 2))  # round all numeric columns

tidy_model




# option 3 to display both - works the best




model <- an_df %>% 
  glm(formula = d_dmode_n ~ d_parity_lab,
            family = binomial(link = "logit")) 
summary(model)




# compare these 2 models
model <- an_df %>% 
  filter(d_parous_cat == 1) %>%
  glm(formula = d_dmode_n ~ d_parity_lab,
            family = binomial(link = "logit")) 
summary(model)


model <- an_df %>% 
  glm(formula = d_dmode_n ~ d_parity_tr1 + d_parous_cat,
            family = binomial(link = "logit")) 
summary(model)

# -------------------------------------------------------------------








model <- an_df %>% 
  glm(formula = d_dmode_n ~ d_parity_tr1 + d_parous_cat,
            family = binomial(link = "logit")) 
summary(model)


an_df %>% 
  tabyl(d_parity_tr1)

an_df %>% tabyl(d_parity_tr_lab)







tidy_model <- model %>% 
  tidy(exponentiate = TRUE, conf.int = TRUE)
tidy_model








# example relevelling 

# model <- data %>%
#   mutate(d_R1alcohol_status_lab = relevel(d_R1alcohol_status_lab, ref = "Current")) %>%
#   glm(d_dmode_n ~ d_R1alcohol_status_lab, data = ., family = binomial)
# 
# summary(model)
```

### Strategy

Step 1. Tabulate all exposures with outcome (exposure, outcome)

## Alcohol

```{r}

an_df %$% tabpct(d_R1alcohol_status_cat, d_dmode_n, percent = "row", graph = F)
an_df %$% tabpct(d_R1alcohol_status_lab, d_dmode, percent = "row", graph = F)


an_df %>%  tabyl(d_dmode_n, d_dmode) 
an_df %>%  tabyl(d_R1alcohol_status_cat, d_R1alcohol_status_lab)
```
